### 2017-01-14 paper/paper.tex
paper: cleanup first half
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\documentclass[preprint]{sigplanconf-eurosys}
\usepackage{graphicx}

\title{Global Namespaces Scale as Well as Decoupled Namespaces}
\authorinfo{Paper ID: 165}
           {Total Pages: 12}
\date{}
\begin{document}
\maketitle

\begin{abstract}
Today's large and highly-parallel workloads are bottlenecked by metadata
services because many processes end up accessing the same shared resoure. In
HPC, state-of-the-art file systems are abandoning POSIX because the
sychronization and serializiation overheads are too costly -- and sometimes
even unnceccessary -- for some of their applications.  While the performance
benefits are plain for these users, other applications that rely on stronger
consistency must be re-written or deployed on a different system. We present
Cudele, a programmable file system that supports different degrees of
consistency and fault tolerance within the same namespace.  In this paper we
examine a general purpose file system and describe the overheads of the design
decisions that were made; namely strong consistency and fault tolerance. Then,
using Cudele, we relax those constraints on specific subtrees in the global
namespace and show improved performance without sacrificing the strong
consistency needs of other users.
\end{abstract}

\section{Introduction}

Today's client-server based file system metadata
services have scalability problems. It takes a lot of resources to server POSIX
metadata requests and applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs}. This is fine for small
workloads and file systems but as the system scales provisioning a metadata
server for every client is expensive and complicated.

Current hardware evolution and the rise of software-defined storage storage,
which uses techniqes like erasure coding, replication, and partitioning, have
ushered a new era of HPC computing; architectures are transitioning from
complex storage stacks with burst buffer, file system, object store, and tape
tiers to a two layer stack with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}. This trend exacerbates the metadata
scalability problem and has given rise to the serverless metadata services.

HPC workloads are so metadata intensive that new management techniques are
advocating reducing synchronization and serialization overheads by transferring
these responsibilities to the client. We call this approach decoupling the
namespace and the semantics of consistency differ amongst systems.

\begin{figure}[tb]
\centering
\includegraphics[width=75mm]{figures/subtree-policies.png}
\caption{Administrators can assign consistency and fault tolerance policies to
subtrees to get the benefits of some of the state-of-the-art HPC architectures.
}\label{fig:subtree-policies}
\end{figure}

We propose subtree policies, an interface that lets future programmers control
how the system manages different parts of the namespace.  For performance one
subtree can adopt weaker consistency semantics while another subtree can retain
the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
shows an example setup where a single global namespace has directories for
applications designed for different, state-of-the-art HPC architectures.  Our
system supports 3 forms of consistency and 2 forms of fault tolerance giving
the administrator a wide range of policies and optimizations depending on the
application's needs.

\section{Related Work}

Decoupling the namespaces has many advantages, including improved scalability,
higher resource utilization, and better performance.
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
are near-POSIX filesystems that give clients the ability to decouple subtrees
from the namespace so that the applications can execute metadata operations
without synchronization and serialization.  These operations are applied to a
local snapshot of the file system namespace and conflicts are resolved either
by the application or by an external service . Applications link into a
metadata server library to reduce resource utilization and code paths (e.g., no
daemons and less interprocess communication). Unfortunately, decoupling the
namespaces has costs: (1) merging metadata state back into the global namespace
is slow; (2) failures are local to the failing node; and (3) the systems are
not backwards compatible.

For (1), state-of-the-art systems manage consistency in non-traditional ways:
IndexFS maintains the global namespace but blocks operations from other clients
until the first client drops the lease, BatchFS does operations on a snapshot
of the namespace and merges batches of operations into the global namespace,
and DeltaFS never merges back into the global namespace. The merging for
BatchFS is done by an auxiliary metadata server running on the client and conflicts are
resolved by the application. Although DeltaFS never explicitly merges,
applications needing some degree of ground truth can either manage consistency
thesmelves on a read or add a bolt-on service to manage the consistency.

For (2), if the client fails and stays down, all metadata operations on the
decoupled namespace are lost. If the client recovers, the on-disk structures
(for BatchFS and DeltaFS this is the SSTables used in
TableFS~\cite{ren:atc2013-tablefs}) can be recovered. In other words, the
clients have state that cannot be recovered if the node stays failed and any
progress made will be lost. This scenario is a disaster for checkpoint-restart
where missed cycles may cause the checkpoint to bleed over into computation
time.

For (3), decoupled namespace approaches sacrifice POSIX going as far as
requiring the application to link against the systems they want to talk to. In
today's world of software defined caching, this can be a problem for large data
centers with many types and tiers of storage. Despite well-known  performance
problems POSIX and REST are the dominant APIs for data transfer.

Decoupling the namespace delays metadata consistency and sacrifices fault
tolerance. As shown in Table~\ref{table:namespaces}, metadata consistency is
provided by capabilities and fault tolerance is addressed with a journal.

\begin{table}[t!]
\centering
\begin{tabular}{l|l|l|l}
                  & decoupled~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}
                  & global~\cite{weil:sc2004-dyn-metadata, ren:sc2014-indexfs}
                  & implied~\cite{bent_plfs_2009} \\ \hline
  consistency     & eventual      & capabilities  & naming \\
  fault tol.      & local         & journal       & journal
\end{tabular}
\caption{Each technique for handling the namespace has its trade-offs.
Decoupled namespaces have the highest throughput because they have weaker
consistency and fault tolerance semantics.
\label{table:namespaces}}
\end{table}

\begin{figure}[tb]
\centering
\includegraphics[width=75mm]{figures/creates-motivation.png}
\caption{Create-heavy workloads (untar) incurr the highest disk, network, and CPU
utilization because the metadata server is managing consistency and fault
tolerance.}\label{fig:creates-motivation}
\end{figure}

\section{POSIX Overheads}

In our examiniation of the overheads of POSIX we benchmark and analyze CephFS,
the file system that uses the RADOS object store to store its data and
metadata. We choose CephFS because it is an open-source production quality
system. CephFS made one set of design decisions and we not asserting that the
design decisions that were made are superior but instead highlight the
effect those decisions have on performance.

To show how CephFS behaves under high metadata load we use a create-heavy
workload because these types of workloads incurr high CPU, network, and disk
usage.  Figure~\ref{fig:creates-motivation} shows a compilation of the Linux
kernel, which has a download, untar, compile, and remove phase. The untar phase
is characterized by many creates and has the highest resource usage, indicating
that it is stressing the consistency and journalling subsystems of the metadata
server. Also of note: a create-heavy workload does not help for caching indoes.

\subsection{Fault Tolerance}
\begin{figure*}[tb]%h
\centering
\includegraphics[width=180mm]{figures/throughput-journal.png}
\caption{Performance improves with larger journal segments because the metadata
server spends less time flushing the journal }\label{fig:throughput-journal}
\end{figure*}

%\begin{figure}[tb]%h
%\centering
%\includegraphics[width=80mm]{figures/throughput-journal-old.png}
%\caption{(a), (b), and (c) show increased activity halfway though the run; (d)
%shows that this activity is related to disk IO, indicating that the activity is
%the journal.  Despite the activity, we conclude that RADOS itself is not the
%bottleneck because our RADOS cluster can sustain 100MB/s write bandwidth.
%}\label{fig:throughput-journal-old}
%\end{figure}
%Figure~\ref{fig:throughput-cache} shows that journaling metadata updates into
%the object store has an overhead.  (a) shows that throughput degrades halfway
%through the run; (b) shows that cpu utilization spikes at this same point; (c)
%shows that outgoing network utilization spikes; and (d) shows that this
%activity spikes on the disks in RADOS.

% purpose of the journal
CephFS uses RADOS (1) as a metadata store for all information about files
including the hierarchical namespace and (2) as a staging area for the journal
of updates before they are applied to the metadata store. (1) is essentially a
cache that improves performance by serving metadata from memory and (2) is the
mechanisms for achieving fault tolerance.

% - sequential IO, trim redundant operations
Fault tolerance means that the client or server can fail and metadata will not
be lost.  CephFS addresses fault tolerance with a metadata journal that streams
into the resilient object store. Similar to LFS~\cite{} and WAFL~\cite{} the
metadata journal can grow to large sizes ensuring (1) sequential writes into
RADOS and (2) the ability for daemons to trim redundant or irrelevent journal
entries.

% Knobs in CephFS
The journal is striped over objects where multiple journal updates can reside
on the same object. There are two tunables for controlling the journal: the
segment size and the number of parallel segments that can be written in
parallel. The former is memory bound as larger segments take up more memory but
can reduce the time spent journaling and the latter is CPU bound.

% Effects on performance
Figure~\ref{fig:throughput-journal} shows that journaling metadata updates into
the object store has an overhead. Part (a) shows the runtime for different
journal segment sizes; the larger the segment size the bigger that the writes
into the object store are. The trade-off comes is in terms of memory because
larger segment sizes take up more space with their buffers. Parts (b), (c), and
(d) show the throughput over time for different segment sizes. Performance
suffers when time is spent journaling.

% Conclusion
Despite this overhead, we posit that the journal is sufficient to slow down
metadata throughput but not so much as to overwhelm RADOS because we measured
our peak bandwidth to be 100MB/s, which is the speed of our network link.

\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS, as far as
we can tell, when a client or server fails there is no recovery scheme. For
BatchFS, if a client fails when it is writing to the local log-structured
merged tree (implemented as an SSTable) then those batched metadata operations
on lost. For DeltaFS, if the client fails then on restart the computation does
the work again -- since the snapshots of the namespace are never globally
consistent, there is no ground truth the requires the failed namespace to
answer to anyone. On the server side, BatchFS and DeltaFS use IndexFS. Again,
IndexFS writes metadata to SSTables but it is not clear whether they ever
vacate memory, get written to disk, or are flushed to the object store.

\subsection{Strong Consistency}

\begin{figure*}[tb]%h
\centering
\includegraphics[width=180mm]{figures/throughput-cache-journal.png}
\caption{Journalling metadata updates has a bigger overhead than maintaining
the inode cache. For create-heavy workloads the inode cache offers no
performance benefits. }\label{fig:throughput-cache-journal}
\end{figure*}

\begin{figure*}[tb]%h
\centering
\includegraphics[width=180mm]{figures/inode-cache.png}
\caption{The inode cache improves metadata read performance but for our
create-heavy workload it is only an overhead. Most of the time maintaining the
cache is spent evicting and adding inodes.}\label{fig:inode-cache}
\end{figure*}

Access to POSIX metadata is strongly consistent, so reads and writes are
globally ordered. The synchronization and serialization machinery needed to
ensure that all clients see the same state has high overhead.  CephFS uses
capabilities to keep metadata strongly consistent. To reduce the number of RPCs
needed for consistency, clients can obtain capabilities for reading, reading
and updating, reads caching, writing, buffering writes, changing the file size,
and performing lazy IO.

\begin{figure*}[tb]%h
\centering
\includegraphics[width=180mm]{figures/throughput-droplease.png}
\caption{When a client create stream is ``isolated" then lookups resolve
locally but when a second client ``interferes" by creating in the same directory,
the directory inode capability is revoked forcing all clients to centralize
lookups at the metadata server.  \label{fig:throughput-droplease}}
\end{figure*}

% purpose of the inode ccahe - reduces RPCs (lookups for create, readdirs for
% stats)
To keep track of the read caching and write buffering capabilities, the clients
and metadata servers agree on the state of each inode using an inode cache.  If
a client has the directory inode cached it can do metadata writes (e.g.,
create) with a single RPC. If the client is not caching the directory inode
then it must do multiple RPCs to the metadata server to (1) determine if the
file exists and (2) do the actual create.  Unless the client immediately reads
all the inodes in the cache, the inode cache is less useful for create-heavy
workloads because the cached inodes are unused.

% benefits
The benefits of caching the directory inode when creating files is shown in
Figure~\ref{fig:throughput-droplease}(a).  If only one client is creating files
in a directory (``isolated" curve) then that client can lookup the existence of
new files locally before issuing a create request to the metadata server. If
another client starts creating files in the same directory (``interfere" curve)
then the directory inode transitions out of read caching and the first client
must send lookups to the metadata server. When other clients interfere the
request throughput is higher Figure~\ref{fig:throughput-droplease}(b) but the
runtime is slower because the isolated client scenario incurrs less requests.
Figures~\ref{fig:throughput-droplease}(c) and (d) plot the creates and lookups,
repsectively, over time; creates slow down and lookups dominate the request
load when the directory inode is shared in the interferring client scenario.

% TODO: what is the cost of trimming the cache?
% TODO: does CephFS still cache inodes when I turn off caching? Why is still keeping inodes in memory? Gah!
%\begin{figure}[tb]%h
%\centering
%\includegraphics[width=80mm]{figures/throughput-cache.png}
%\caption{ CephFS tries to maintain 100 thousand inodes in the cache;
%performance degrades when it cannot evict inodes fast enough to maintain this
%threshold. \label{fig:throughput-cache}}
%\end{figure}

% disadvantages
The drawbacks of the inode cache are shown in
Figure~\ref{fig:throughput-cache}.  Figure~\ref{fig:throughput-cache}(a) shows
the throughput in metadata requests per second for a single client creating 200
thousand files. Until time 950 the throughput is steady at just under 2000
ops/sec and the CPU utilization is at about 30\%. At time 950 seconds
throughput degrades which corresponds to more inodes in the cache
(Figure~\ref{fig:throughput-cache}(c)). The values in
Figure~\ref{fig:throughput-cache}(c) are:

\begin{itemize}
  \item inodes: total number of elemnts in the cache
  \item inodes pinned: 
  \item inodes pinned tail
  \item inodes with caps
  \item inodes top
  \item inodes bottom 
\end{itemize}

% how the inode cache works
CephFS tries to keep the cache at 100 thousand inodes so the degradation in
performance indicates that the metadata server cannot keep up with the
workload. Figure~\ref{fig:throughput-cache}(d) shows inodes getting added at
the same rate as before 950 seconds (ino\(+\)) but the rate at which inodes get
removed is less stable (ino\(-\)) indicating that cache eviction is taking more
of the metadata servers time.

\subsection{Consistency Overhead}
\begin{figure*}[tb]
\centering
\includegraphics[width=180mm]{figures/exp0-underloaded.png}
\caption{An underloaded metadata server adaquately services conflicting
clients; the create speeds are similar while the interferring operations are
limited by the cost of RPCs.}\label{fig:exp0-underloaded}
\end{figure*}

\begin{figure*}[tb]
\centering
\includegraphics[width=180mm]{figures/runtime-consistency-scale.png}
\caption{Scaling clients shows increased variability when another client
interferes; zooming in on runs with 7 clients we see that different types of
interferring operations have different effects on performance variability and
predictability.  }\label{fig:runtime-consistency-scale}
\end{figure*}

Figure~\ref{fig:exp0-underloaded} is a baseline showing that the metadata
server can service multiple clients when it is underloaded. One client creates
files in the same directory and another does a touch or stat 15 seconds into
the run. Figure~\ref{fig:exp0-underloaded}(a) shows the runtime of the client
doing the creates: "isolated" is when the create client is the only workload in
the cluster, "interfere stat" is the runtime of the create client when another
client does a stat, and "interfere touch" is the runtime of the create client
when another client does a touch. There is only a minor performance
degradation. Figure~\ref{fig:exp0-underloaded}(b) compares the baseline
the other client doing the touch or stat
%
%and Figure~\ref{fig:exp0-underloaded}
%has a throughput trace to show behavior.
%
%In Figure~\ref{fig:exp0-underloaded}(a), 
%creates only see a minor performance
%degradation while the interferring operations in
%Figure~\ref{fig:exp0-underloaded}(b) show a larger (yet not significant)
%discrepency. 
%
%
%
% isolated is when there is no interferrence from client 1.
To explain Figure~\ref{fig:exp0-underloaded}(a) we use
Figure~\ref{fig:exp0-behaviour}(b) to show why the runtime of creating 100
thousand files is unaffected. Compared to the throughput without an
interferring stat ("isolated" curve) the throughput of the metadata server with
an interferring stat ("inteferring stat" curve) is higher. This means the
metadata server is doing more operations suggesting that it can adequately
handle the demands of both workloads.

To explain Figure~\ref{fig:exp0-underloaded}(b) we use
Figure~\ref{fig:exp0-underloaded}(a) to show that operations are bounded by the
number of files; more files incurr more requests since both operations lookup
every file. The number of files shown in Figure~\ref{fig:exp0-behaviour}(a)
indicates that the local stat ("client 0" bar) is interacting with less files
than the remote stat ("client 1" bar). This graph also explains why the
isolated stats and touches take the longest from the remote clients; out of all
setups, the isolated operations are doing the most RPCs.  The difference
between local ("client 0 interfere" bar) and remote ("client 1 interfere" bar)
in Figure~\ref{fig:exp0-underloaded}(b) is the overhead of doing RPCs.
Isolated is the slowest because it requests the most files. On average the
local interferring client ("client 0" bar in
Figure~\ref{fig:exp0-behaviour}(a)) requests 40 thousand less files than the
remote interferring call ("client 1" bar in
Figure~\ref{fig:exp0-behaviour}(a)). This reflects how fast the local client
can get into the queue of requests, presumably because it bypasses capability
checks.

From this experiment we make 2 conclusions:
\begin{itemize}
  \item the metadata server is not overloaded because it can handle both workloads
  \item the metadata server needs to get global consistency from only 1 client
\end{itemize}

\textbf{Comparison to decoupled namespaces}: Decoupled namespaces merge batches
of metadata operations into the global namespaces when the job completes.  In
BatchFS the merge is delayed by the application using an API to switch between
asynchronous to sychronous mode. The merge itself is explicitly managed by the
application but future work looks at more automated methdologies. In DeltaFS
snapshots of the metadata subtrees stays on the client machines; there is no
ground truth and consistent namespaces are constructed and resolved at
application read time or when a 3rd party system (e.g., middleware, scheduler,
etc.) needs a view of the metadata.

\section{Methodology}

\subsection{Decoupled Namespaces in CephFS}

\begin{figure}[tb]
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\caption{Applications that can tolerate weaker consistency can decouple the
namespace, write updates to a local file, and delay metadata updates. The merge
and replay steps are optional if there is no need for a global namespace.
}\label{fig:decouple}
\end{figure}

%- How it works (timestamps)
Recall that CephFS uses RADOS (1) as a metadata store for all information about
files including the hierarchical namespace and (2) as a staging area for the
journal of updates before they are applied to the metadata store.  The journal
tool is used for disaster recovery and lets administrators view and modify the
journal of metadata updates; it can read the journal, erase events from the
journal, and apply the updates in the journal to the metadata store.  To apply
journal updates to the metadata store, the journal tool reads the journal
segments from RADOS objects and applies the update to the metadata store (which
are also stored as RADOS objects).

% General information about the journal
The journal segments are saved as objects in RADOS.  The journal has 4
pointers, described in `osdc/Journaler.h`:

\begin{itemize}
  \item write position: tail of the journal; points to the current session where we are appending events
  \item unused field: where someone is reading
  \item expire position: old journal segments
  \item trimmed position: where daemon is expiring old items
\end{itemize}

% How the journal tool works
Journal segments in RADDS have a header followed by serialized log events. The
log events are read by hopping over objects using the read offset and object
size pulled from the journal header.  After decoding them, we can examine the
metadata (1) about the event (e.g., type, timestamps, etc.) and (2) for inodes
that the event touches.

% The metablobs
The metadata for inodes that the event touches are called metadata blobs and
the ones associated with events are \textbf{unordered}; this layout makes
writing journal updates fast but the cost is realized when reading the metadata
blobs.  It makes sense to optimize for writing since reading only occurs on
failures. To reconstruct the namespace for the metadata blob, the journal tool
iterates over each metadata blob in the events and builds mappings of inodes to
directory entries (for directories) and parent inodes to child inodes/directory
entries.

% Journal import
The journal tool imports journals from binary files stored on disk.  First the
header of the dump is sanity checked and written into RADOS to the ``header"
object.  The ``header" object has metadata about the journal as well as the
locations of all the journal pointers (e.g., where the tail of the journal is,
where we are currently trimming, etc.).  Then the journal events are cleaned
(erasing trailing events that are not part of the header) and written as
objects into RADOS.  Note that while the journal is in RADOS, the metadata
servers do do not have the namespace reconstructed in memory so the metadata
cluster will not service requests relating to the journal of imported events.
To construct the namespace in the collective memory of the metadata servers we
need to first construct the namespace in RADOS. The journal tool can explicitly
do this by  applying the journal to the metadata store in RAODS. This will pull
the objects containing journal segments and replay them on the metadata store.
Finally, we delete the journal in RADOS and restart the metadata servers so
they rebuild their caches.

% Journal export
The journal tool exports journals to binary files stored on disk. First the
journal is scanned for the header and then journal is recovered. To recover the
journal the ``header" object is read off disk and then objects are probed in
order and starting from the write position saved in the header. Probing will
update the write position if it finds objects with data in them. There are two
types of objects:

1. journal objects (e.g., 200.000*)

2. metadata store objects (e.g., 1.inode)

We need to update the journal objects with the new events! How do we write out
new objects to RADOS (e.g., 200.000*)? These should be the corect sizes (e.g.,
4980 bytes) -- these log events that we are adding are not making them into
the journal objects... do we have to explicitly read them and then re-write
them?

TODO: pull the object, read it, append to it, make sure size changes. It looks
like we can only read these objects and apply doesn't actually touch them... it
only touches the metadata store objects (e.g., 1.inode*).

% TODO: figure out how import saves the journal on the metadata server
% - they aren't -- everything is stored in RADOS
% - we never touch the metadata servers!!!
% TODO: figure out how to read the header object
% TODO: figure out the mapping objects to journal segments (not 1x1)

% Journal export
When exporting a journal of events, the journal tool first scans the journal to
check for corruption. Then it recovers the journal by reading the ``header"
object out of RADOS.  After reading the header, the journal tool can pull
journal segments from RADOS because it knows how many objects to pull and how
far to seek within those objects.

%% The data structures
%The metadata that the event touches, including inodes, paths, and timestamps,
%are stored as metablobs. Each piece of metadata inside a metablob is called a
%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
%table client transactions (tid, version), renamed directory fragments (maps,
%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
%me (and you since you are reading this paper), this does not make any sense.
%
%Each directory fragment has an associated directory lump, which is just a bunch
%of metadata. The most interesting part of the dirlump is the fullbits array,
%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
%and then all the full bits, saving off the children and inode locations. The
%children tell us which dentry names an inode has and the inode locations map
%the parent inode to its child inode and dentry.

\subsubsection{Leveraging the Journal Tool}

%- How we leverage it
We leverage the journal tool's ability to reconstruct the namespace for the
metadata blob. First we scan journal and store all events in a map. Next we
create a new event and copy the root inode directory lump from another journal
event into the event's metadata blob. Finally, we add a new directory lump for
our new file or directory.

If
decoupled namespaces are turned on, the client writes metadata updates locally
and merges the updates with the journal tool. The client that decouples the
namespace operates without any consistency and any conflicts at the merge are
resolved in favor of this client. Updates by other clients ({\it i.e.} metadata
writes to the global namespace) are overwritten. Cudele adopts the following
process when an application decouples the namespace:

\begin{enumerate}
  \item metadata server exports a snapshot of the file system to a binary file
  \item client materializes the snapshot in memory
  \item client operates on the data structures in memory
  \item client exports the new snapshot to a binary file
  \item metadata server merges the snapshot into the global namespace
\end{enumerate}

Step 3 is the most complicated and requires understanding how the snapshot is
materialized in memory.

\subsubsection{Operating on Snapshots}

Our first implementation attempted to re-create journal events using the same
libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
to instantiate a Ceph inode and directory entry for the current file/dir and
its parent.  This is too hard because there are too many moving parts in the
metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
traverse up and down namespace, etc.). So when I tried add dentries and inodes
it was trying to traverse up/down and it would almost always segfault when it
was looking for something. These metablobs are supposed to be self container --
the problem is I do not know what is supposed to go *inside* them.

Our second idea was to copy the metadata blog and change just what we needed.
For example, we would save a binary dump of a generic \texttt{mkdir} event on
disk. When the application makes a directory, this dump would be loaded and the
fields would be changed before being written back to disk. Rather than
traversing up and down a namespace in memory of a metadata server, we should
traverse up and down the namespace *inside* the metadata blob. This
implementation requires disk IO and editing the log event is non-trivial for
two reasons:

\begin{itemize}

  \item methods do not edit events; they just write them

  \item the metadata that the event touches (e.g., the metablob) is unorganized
  on disk for performance -- it is trade-off for writing data faster serially and
  reconstructing information slowly since failure is not considered the norm

\end{itemize}

Faced with these challenges we landed on our final implementation: load the
snapshot into the the data structures used to examine and replay journals, edit
those data structures, and write them out to disk as binary.

\section{notes}
Linking clients into our custom libcephfs

Use namespace's recursive data structure to put policies on subtrees
- consistency: eventual vs. strong, global vs. local
  - e.g., BatchFS/DeltaFS: eventual, local
  - e.g., POSIX: strong, global
  - e.g., PLFS: no consistency
- fault tolerance: global vs. local
  - e.g., CephFS: global
  - e.g., BatchFS/DeltaFS: local

Experimental Setup
- Ceph: 9 OSDs, 1 MDS, 2 kernel client
- Workload limitations: blah

Workload: creates

Baseline: 200K creates in the same directory
- throughput: degrades at 950s
- CPU utilization: more at 950s
- inode cache: eviction dominate
- inodes +- to cache: eviction dominate
- per-disk throughput: RADOS not bottleneck

Experiment 1: Interference

\section{Experiments}

In this section, we quantify the costs of consistency and fault tolerance in
CephFS. If the components that ensure these semantics (i.e. capabilities and
journals, respectively) can be mitigated or delayed, then global namespaces
perform as well as decoupled namespaces.

\textbf{Experimental Setup}: we run our experiments on a 9 OSD, 3 MDS, 1 MON
Ceph cluster. The clients use the Ceph kernel client, which has been in the
mainline Linux kernel since TODO. We use the kernel client so that we can find
the true create speed of the server; our experiments show a low CPU utilziation
for the clients which indicates that we are stressing the servers more. We also
turn caching off becuase, as shown in figureX there is little difference, in
terms of performance between caching and not caching when using the kernel
client.

\subsection{Baseline}
Experiment 0: creates in the same directory
- setup: why we use caching, we use the kernel client, how we circumvent max fragment size

Experiment 0: creates with a stat
- Hypothesis: metadata read pauses creates and requires a snapshot in time
  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
- can we delay snapshot?

Experiment 1: creates with a readdir
- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
- client0: create 100k, client1: stat at 2 mins

Experiment 2: scale the number of files
- See if the open/close spike occurs 
- Try to see why open/close spike is allowed to happen
- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
- client0: create 100k, client1: touch at 2 mins

Experiment 3: see how fast the cache satisfies a read
- client0: create 100k, stat inodes
- client0: create 100k, client1: stat inodes

lient 0: creates, client 1 create(s)

\subsection{Journaling Overhead}

Journal to RADOS
Turn off journaling (large segment)
Journal to in-memory OSD

\subsection{Macrobenchmarks}
updatedb: http://lists.ceph.com/pipermail/ceph-users-ceph.com/2015-July/002768.html

\bibliography{paper}
\bibliographystyle{plain}
\end{document}

### 2017-01-31 paper/paper.tex
paper: finish related work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+% I work with the file system layer that sits on top of Ceph
+% - internal subsystems and interfaces that Ceph uses to provide a strongly consistent POSIX file system
+%   - e.g., load balancing mechanisms, for hotspots/flash crowds on directories
+%     caches for inodes, capabilities for guarding a shared resource, and a large journal for fault tolerance
+% - our programmability initiative talks about composing these internal interfaces for higher level services
+
+% 
+% Previous work: 
+% - master's thesis: scale-out vs. scale-up architectures 
+% - work in industry: file systems for the cloud, big data processing over
+% object storage, and automating workflows
+%
+% I work under Carlos at UC Santa Cruz and we study how to migrate file system
+% metadata to alleviate overloaded servers. The idea is to take the namespace
+% on one metadata server, chop it up into smaller subtrees, and farm those
+% subtrees out to other metadata servers.  It's very similar to IndexFS/ShardFS
+% but we try to leverage namespace locality by allowing variable sized
+% subtrees. We published a paper at SC'15 where we talk about a load balancing
+% API we built on CephFS that let's administrators control how to chop up
+% subtrees and when/where to move them (also got merged into Ceph).
+%
+% Go to conclusion
+%
+% Why I want to work with Brad
+% - new perspectives on file system designs, especially NON-POSIXy 
+% - work in a national lab: I know about theirarchitectures and problems but
+% I've never used an HPC system - real world applications and benchmarks; most
+% of my work has been accepted for novelty but not for fully benchmarking a
+% system
+% - lines up with my thesis
+%
+% What I want to work on:
+% - continue working on my thesis; this next phase of having a global namespace
+% but different consistency semantics within that namespace for data management
+%   - namespace: recursive data structure as a logical model for understanding the user's goals
+%   - merge: what is the proper way to merge deltafs/batchfs updates back into the global namespace
+%   - data management: deltaFS made the case that apps shouldn't use the fs to synchronize;
+%     we contend that you could use the fs as data management; remembering how to access your data
+%
+% What I have been working on 
+% - leverage internal services/subsystems in Ceph; avoids code duplication and
+% let's us test this over the same system 
+% - benchmarking the overheads of strong consistency/journalling - we've figured
+% out how to thrash capabilities and turn off the journal
 \begin{abstract}
+
 Today's large and highly-parallel workloads are bottlenecked by metadata
@@
 Cudele, a programmable file system that supports different degrees of
-consistency and fault tolerance within the same namespace.  In this paper we
-examine a general purpose file system and describe the overheads of the design
-decisions that were made; namely strong consistency and fault tolerance. Then,
-using Cudele, we relax those constraints on specific subtrees in the global
-namespace and show improved performance without sacrificing the strong
-consistency needs of other users.
+consistency and fault tolerance within the same namespace.  First, we take a
+POSIX compliant file system and relax the consistency constraints by
+implementing two metadata designs: delayed metadata update merge into the
+global namespace and client local views of metadata updates.  Second, we store
+the consistency and fault tolerance semantics in inodes so that subtrees within
+the same namespace can be optimized for different workloads; the inodes are
+programmable so that clients understand how to access the metadata in a
+subtree.  Third, we present a theoretical framework for a metadata service and
+a working implementation that provides the lowest common denominator needed to
+implement a wide range of consistency/fault tolerance semantics that can be
+benchmarked all on the same system.
+
 \end{abstract}
+
+% Questions
+% - do you still collaborate with CMU? Would you want all of us to work together?
+% - what projects do you want to work on?
+% - how closely would you want to work with Carlos?
+% - what system would you want to work on?
+
+% The main difference between CephFS and the CMU systems is that our approach
+% is a little more flexible.  We can partition the namespace in any way and can
+% co-locate items together if they are dependent on each other -- in theroy...
+% in practice it is very hard to figure out what the workload is trying to do.
+% Also, we can program the decision making based on different metrics, like
+% CPU, request rate, etc.
 
@@
 
-Today's client-server based file system metadata
-services have scalability problems. It takes a lot of resources to server POSIX
-metadata requests and applications perform better with dedicated metadata
+Today's client-server based file system metadata services have scalability
+problems. It takes a lot of resources to server POSIX metadata requests and
+applications perform better with dedicated metadata
 servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs}. This is fine for small
@@
 
-Decoupling the namespaces has many advantages, including improved scalability,
-higher resource utilization, and better performance.
-BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
-are near-POSIX filesystems that give clients the ability to decouple subtrees
-from the namespace so that the applications can execute metadata operations
-without synchronization and serialization.  These operations are applied to a
-local snapshot of the file system namespace and conflicts are resolved either
-by the application or by an external service . Applications link into a
-metadata server library to reduce resource utilization and code paths (e.g., no
-daemons and less interprocess communication). Unfortunately, decoupling the
-namespaces has costs: (1) merging metadata state back into the global namespace
-is slow; (2) failures are local to the failing node; and (3) the systems are
-not backwards compatible.
+% General
+The bottlenecks associated with accessing POSIX file system metadata are not limited
+to HPC workloads and the same challenges that plagued these systems for years are
+finding their way into the cloud. Workloads that deal with many small files
+({\it e.g.}, log processing and database
+queries~\cite{thusoo:sigmod2010-facebook-infrastructure}) and large numbers of
+simultaneous clients ({\it e.g.}, MapReduce
+jobs~\cite{mckusick:acm2010-gfs-evolution}), are subject to the scalability of
+the metadata service. The biggest challenge is that whenever a file
+is touched the client must access the file's metadata and maintaining a file
+system namespace imposes small, frequent accesses on the underlying storage
+system~\cite{roselli:atec2000-FS-workloads}.  Unfortunately, scaling file
+system metadata is a well-known problem and solutions for scaling data IO do
+not work for metadata IO~\cite{roselli:atec2000-FS-workloads,
+abad:techreport2012-fstrace, abad:ucc2012-mimesis,
+alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. There are two approaches
+for improving the performance of metadata access.
+
+\subsection{Metadata Load Balancing}
+
+% approaches to load balancing (strong = more resuorces per unit work, weak =
+% fixed resource per work unit)
+One approach for improving metadata performance and scalability is to alleviate
+overloaded servers by load balancing metadata IO across a cluster. Common
+techniques include partitioning metadata when there are many writes and
+replicating metadata when there are many reads. For example, IndexFS partitions
+directories and clients write to different partitions by grabbing leases and
+caching ancestor metadata for path traversal; it does well for strong scaling
+because servers can keep more inodes in the cache which results in less RPCs.
+Alternatively, ShardFS replicates directory state so servers do not need to
+contact peers for path traversal; it does well for read workloads because all
+file operations only require 1 RPC and for weak scaling because requests will
+never incurr extra RPCs due to a full cache.  CephFS employs both techniques to
+a lesser extent; directories can be replicated or sharded but the caching and
+replication policies do not change depending on the balancing technique.
+Despite the performance benefits these techniques add complexity and jeopardize
+the robustness and performance characteristics of the metadata service because
+the systems now need (1) policies to guide the migration decisions and (2)
+mechanisms to address inconsistent states across servers.
+
+% policies
+Setting policies for migrations is arguably more difficult than adding the
+migration mechanisms themselves.  For example, IndexFS and CephFS use the
+GIGA+~\cite{GIGA+} technique for partitioning directories at a predefined
+threshold and using lazy synchronization to redirect queries to the server that
+``owns" the targeted metadata.  Determining when to partition directories and
+when to migrate the directory fragments are policies that vary between systems:
+GIGA+ partitions directories when the size reaches a certain number of files
+and migrates directory fragments immediately; CephFS partitions directories
+when they reach a threshold size or when the write temperature reaches a
+certain value and migrates directory fragments when the hosting server has more
+load than the other servers in the metadata cluster. Another policy is when and
+how to replicate directory state; ShardFS replicates immediately and
+pessimistically while CephFS replicates only when the read temperature reaches
+a threshold.  There is a wide range of policies and it is difficult to address
+with tunables and hard-coded design decisions.
+
+% addressing inconsistency
+In addition to the policies, distributing metadata across a cluster requiress
+distributed transactions and cache coherence protocols to ensure strong
+consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
+directory state and uses optimistic concurrency control for conflicts; namely
+it does the operation and if there is a conflict at verification time it falls
+back to two-phase locking.  Another example is IndexFS's inode cache which
+reduces RPCs by caching ancestor paths -- the locality of this cache can be
+thrashed by random reads but performs well for metadata writes. For
+consistency, writes to directories in IndexFS block until the lease expires
+while writes to direcotries in ShardFS are slow for everyone as it either
+requires serialization or locking with many servers; reads in IndexFS are
+subject to cache locality while reads in ShardFS always resolve to 1 RPC.
+Another example of the overheads of addressing inconsistency is how CephFS
+maintains client sessions and inode caches for capabilities (which in turn make
+metadata access faster). When metadata is exchanged between metadata servers
+these sessions/caches must be flushed and new statistics exchanged with a
+scatter-gather process; this halts updates on the directories and blocks until
+the authoratitive metadata server responds.  These protocols are discusssed in
+more detail in the next section but their inclusion here is a testament to the
+complexity of migrating metadata.
+
+%For metadata writes it does a distributed transaction; monotonic writes with
+%concurrent clients fail and do pessimistic locking through a mediated lock
+%server to ensure strong consistency, non-monotonic writes grab locks at every
+%server. Zooming in on monotonic writes: if permissions increase it executes on
+%a primary then non-primary, if permissions decrease it executes on all
+%non-primary then on primary.
+
+The conclusion we have drawn from this related work is that metadata protocols
+have a bigger impact on performance and scalability than load balancing.  
+Understanding these protocols helps load balancing and gives us a better
+understanding of the metrics we should use to make migration decisions ({\it
+e.g.}, which operations reflect the state of the system), what types of
+requests cause the most load, and how an overloaded system reacts ({\it e.g.},
+increasing latencies, lower throughput, etc.).
+
+\subsection{Relaxing POSIX}
+
+% POSIX 
+POSIX workloads require strong consistency and many file systems improve
+performance by reducing the number of remote calls per operation ({\it i.e.}
+RPC amplification). As discussed in the previous section, caching with leases and
+replication are popular approaches to reducing the overheads of path traversals
+but their performance is subject to cache locality and the amount of available
+resources, respectively; for random workloads larger than the cache extra RPCs
+hurt performance~\cite{IndexFS, CephFS} and for write heavy workloads with more
+resources the RPCs for invalidations are harmful. Another approach to reducing
+RPCs is to use leases or capabilities.
+
+%IndexFS aggressively caches paths and
+%handles permissions by handing out leases for metadata writes; metadata may
+%only be modified when all leases have expired. 
+%IndexFS~\cite{ren:sc2014-indexfs} aggressively caches pathnames and their
+%permissions on the client servers. Modifications to metadata cached by clients
+%is delayed until all client leases have expired. This reduces the RPC
+%amplification to 1 when mutatating directory metadata ({\it e.g.,}
+%\texttt{mkdir}, \texttt{chmod}, etc.) because clients are not querying MDSs for
+%path traversals. The disadvantage of this approach is the high latency of
+%mutation operations (reads to filenames).  ShardFS~\cite{xiao:socc2015-shardfs}
+%replicates metadata (specifically directory lookup state) across the MDS
+%cluster, reducing the RPC amplification to 1 for file operations ({e.g.,}
+%\texttt{stat}, \texttt{chmod}, \texttt{chown}, etc.). Modifications to the
+%directory lookup state are done with optimistic concurrecny control and fall
+%back to retry if verification fails. The disadvantage of this appraoch is the
+%high number of RPCs for maintaining directory metadata mutations (writes to
+%directories).  CephFS also maintains an inode cache but its notion of leases
+%are much shorter, on the order of microseconds.
+
+% Non POSIX
+High performance computing has unique requirements for file systems ({e.g.},
+fast creates) and well-defined workloads (e.g., workflows) that make relaxing
+POSIX sensible.  One popular approach is to allow clients to ``lock'' parts of
+the namespace to improve performance and scalability by avoiding
+synchronization, false sharing, and serialization.  BatchFS assumes the
+application coordinates accesses to the namespace, so the clients can batch
+local operations and merge with a global namespace image lazily. Similarily,
+DeltaFS eliminates RPC traffic using subtree snapshots for non-conflicting
+workloads and middleware for conflicting workloads. MarFS gives administrators
+the ability to (1) lock ``project directories" and (2) allocate GPFS clusters
+for demanding directory workloads. TwoTiers eliminates high-latencies by
+storing metadata in a flash tier; apps lock the namespace so that metadata can
+be accessed more quickly.  Unfortunately, decoupling the namespaces has costs:
+(1) merging metadata state back into the global namespace is slow; (2) failures
+are local to the failing node; and (3) the systems are not backwards
+compatible.
+
+% NON POSIX
+%Decoupling the namespaces has many advantages, including improved scalability,
+%higher resource utilization, and better performance.  BatchFS and DeltaFS are
+%near-POSIX filesystems that give clients the ability to decouple subtrees from
+%the namespace so that the applications can execute metadata operations without
+%synchronization and serialization.  These operations are applied to a local
+%snapshot of the file system namespace and conflicts are resolved either by the
+%application or by an external service . Applications link into a metadata
+%server library to reduce resource utilization and code paths (e.g., no daemons
+%and less interprocess communication).
 
@@
 and DeltaFS never merges back into the global namespace. The merging for
-BatchFS is done by an auxiliary metadata server running on the client and conflicts are
-resolved by the application. Although DeltaFS never explicitly merges,
-applications needing some degree of ground truth can either manage consistency
-thesmelves on a read or add a bolt-on service to manage the consistency.
+BatchFS is done by an auxiliary metadata server running on the client and
+conflicts are resolved by the application. Although DeltaFS never explicitly
+merges, applications needing some degree of ground truth can either manage
+consistency thesmelves on a read or add a bolt-on service to manage the
+consistency.
 
@@
 decoupled namespace are lost. If the client recovers, the on-disk structures
-(for BatchFS and DeltaFS this is the SSTables used in
-TableFS~\cite{ren:atc2013-tablefs}) can be recovered. In other words, the
-clients have state that cannot be recovered if the node stays failed and any
-progress made will be lost. This scenario is a disaster for checkpoint-restart
-where missed cycles may cause the checkpoint to bleed over into computation
-time.
+(for BatchFS and DeltaFS this is the SSTables used in TableFS) can be
+recovered. In other words, the clients have state that cannot be recovered if
+the node stays failed and any progress will be lost. This scenario is a
+disaster for checkpoint-restart where missed cycles may cause the checkpoint to
+bleed over into computation time.
 
@@
 
-\begin{table}[t!]
-\centering
-\begin{tabular}{l|l|l|l}
-                  & decoupled~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}
-                  & global~\cite{weil:sc2004-dyn-metadata, ren:sc2014-indexfs}
-                  & implied~\cite{bent_plfs_2009} \\ \hline
-  consistency     & eventual      & capabilities  & naming \\
-  fault tol.      & local         & journal       & journal
-\end{tabular}
-\caption{Each technique for handling the namespace has its trade-offs.
-Decoupled namespaces have the highest throughput because they have weaker
-consistency and fault tolerance semantics.
-\label{table:namespaces}}
-\end{table}
+\section{POSIX Overheads}
 
@@
 \end{figure}
-
-\section{POSIX Overheads}
 
@@
 \subsection{Fault Tolerance}
+\begin{figure*}[tb] \centering
+\includegraphics[width=1\textwidth]{./figures/journal.png} 
+\caption{CephFS uses a journal to stage updates and tracks dirty metadata in
+the collective memory of the MDSs. Each MDS maintains its own journal, which is
+broken up into 4MB segments. These segments are pushed into RADOS and deleted
+when that particular segment is trimmed from the end of the log. In addition to
+journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
+\end{figure*}
+
 \begin{figure*}[tb]%h
@@
 % purpose of the journal
-CephFS uses RADOS (1) as a metadata store for all information about files
-including the hierarchical namespace and (2) as a staging area for the journal
-of updates before they are applied to the metadata store. (1) is essentially a
-cache that improves performance by serving metadata from memory and (2) is the
-mechanisms for achieving fault tolerance.
+As shown in Figure~\ref{fig:journal} CephFS uses RADOS (1) as a metadata store
+for all information about files including the hierarchical namespace and (2) as
+a staging area for the journal of updates before they are applied to the
+metadata store. (1) is essentially a cache that improves performance by serving
+metadata from memory and (2) is the mechanisms for achieving fault tolerance.
 
@@
 
-If
-decoupled namespaces are turned on, the client writes metadata updates locally
-and merges the updates with the journal tool. The client that decouples the
-namespace operates without any consistency and any conflicts at the merge are
-resolved in favor of this client. Updates by other clients ({\it i.e.} metadata
-writes to the global namespace) are overwritten. Cudele adopts the following
-process when an application decouples the namespace:
+If decoupled namespaces are turned on, the client writes metadata updates
+locally and merges the updates with the journal tool. The client that decouples
+the namespace operates without any consistency and any conflicts at the merge
+are resolved in favor of this client. Updates by other clients ({\it i.e.}
+metadata writes to the global namespace) are overwritten. Cudele adopts the
+following process when an application decouples the namespace:
 
@@
 
-\section{notes}
-Linking clients into our custom libcephfs
-
-Use namespace's recursive data structure to put policies on subtrees
-- consistency: eventual vs. strong, global vs. local
-  - e.g., BatchFS/DeltaFS: eventual, local
-  - e.g., POSIX: strong, global
-  - e.g., PLFS: no consistency
-- fault tolerance: global vs. local
-  - e.g., CephFS: global
-  - e.g., BatchFS/DeltaFS: local
-
-Experimental Setup
-- Ceph: 9 OSDs, 1 MDS, 2 kernel client
-- Workload limitations: blah
-
-Workload: creates
-
-Baseline: 200K creates in the same directory
-- throughput: degrades at 950s
-- CPU utilization: more at 950s
-- inode cache: eviction dominate
-- inodes +- to cache: eviction dominate
-- per-disk throughput: RADOS not bottleneck
-
-Experiment 1: Interference
-
-\section{Experiments}
+\section{Results}
 
@@
 
-\subsection{Baseline}
-Experiment 0: creates in the same directory
-- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
-
-Experiment 0: creates with a stat
-- Hypothesis: metadata read pauses creates and requires a snapshot in time
-  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
-- can we delay snapshot?
-
-Experiment 1: creates with a readdir
-- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
-- client0: create 100k, client1: stat at 2 mins
-
-Experiment 2: scale the number of files
-- See if the open/close spike occurs 
-- Try to see why open/close spike is allowed to happen
-- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
-- client0: create 100k, client1: touch at 2 mins
-
-Experiment 3: see how fast the cache satisfies a read
-- client0: create 100k, stat inodes
-- client0: create 100k, client1: stat inodes
-
-lient 0: creates, client 1 create(s)
+%\section{notes}
+%Linking clients into our custom libcephfs
+%
+%Use namespace's recursive data structure to put policies on subtrees
+%- consistency: eventual vs. strong, global vs. local
+%  - e.g., BatchFS/DeltaFS: eventual, local
+%  - e.g., POSIX: strong, global
+%  - e.g., PLFS: no consistency
+%- fault tolerance: global vs. local
+%  - e.g., CephFS: global
+%  - e.g., BatchFS/DeltaFS: local
+%
+%Experimental Setup
+%- Ceph: 9 OSDs, 1 MDS, 2 kernel client
+%- Workload limitations: blah
+%
+%Workload: creates
+%
+%Baseline: 200K creates in the same directory
+%- throughput: degrades at 950s
+%- CPU utilization: more at 950s
+%- inode cache: eviction dominate
+%- inodes +- to cache: eviction dominate
+%- per-disk throughput: RADOS not bottleneck
+%
+%Experiment 1: Interference
+%
+%\subsection{Baseline}
+%Experiment 0: creates in the same directory
+%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
+%
+%Experiment 0: creates with a stat
+%- Hypothesis: metadata read pauses creates and requires a snapshot in time
+%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
+%- can we delay snapshot?
+%
+%Experiment 1: creates with a readdir
+%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
+%- client0: create 100k, client1: stat at 2 mins
+%
+%Experiment 2: scale the number of files
+%- See if the open/close spike occurs 
+%- Try to see why open/close spike is allowed to happen
+%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
+%- client0: create 100k, client1: touch at 2 mins
+%
+%Experiment 3: see how fast the cache satisfies a read
+%- client0: create 100k, stat inodes
+%- client0: create 100k, client1: stat inodes
+%
+%lient 0: creates, client 1 create(s)
 

### 2017-02-02 experiments/fuse/batchfs/run.sh
benchmark: add batchfs results
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,28 @@
+#!/bin/bash
+
+set -ex
+
+# setup the docker command
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root"
+ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
+CEPH_ANSIBLE="$RUN -v `pwd`/site/roles/ceph-ansible:/root $ANSIBLE"
+SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
+
+# configure ceph and setup results directory
+cp site/* site/roles/ceph-ansible || true
+cp -r site/group_vars site/roles/ceph-ansible/
+cp site/hosts site/roles/ceph-ansible/hosts
+
+# cleanup and start ceph
+for i in "bogus"; do
+  for nfiles in 100000; do # 10000 1000 100 10; do
+    mkdir -p results/${nfiles}/logs || true
+    $SRL_ANSIBLE cleanup.yml
+    $CEPH_ANSIBLE ceph.yml cephfs.yml
+    $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
+     
+    ./ansible-playbook.sh -e nfiles=$nfiles ../workloads/journal.yml
+    ./ansible-playbook.sh -e site=$nfiles collect.yml
+  done
+  mv results results-run${i}
+done

### 2017-02-21 paper/paper.tex
paper: re-organize methodology
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \usepackage{graphicx}
+\usepackage{arydshln}
 
@@
 
+\begin{tabular}{ r | l | l | l }
+              & Decoupled & Global    & Implied   \\
+              & Namespace & Namespace & Namespace \\\hline
+  Example     & BatchFS & CephFS & PLFS \\
+              & DeltaFS & IndexFS &     \\
+  Consistency & eventual & strong & none? \\
+  Fault Tol.  & node local & journal & journal \\
+\end{tabular}
+
 \section{POSIX Overheads}
@@
 
+In this section, we quantify the costs of consistency and fault tolerance in
+CephFS. If the components that ensure these semantics (i.e. capabilities and
+journals, respectively) can be mitigated or delayed, then global namespaces
+perform as well as decoupled namespaces. We run our experiments on a 9 OSD, 3
+MDS, 1 MON Ceph cluster. The clients use the Ceph kernel client, which has been
+in the mainline Linux kernel since TODO. We use the kernel client so that we
+can find the true create speed of the server; our experiments show a low CPU
+utilziation for the clients which indicates that we are stressing the servers
+more. We also turn caching off becuase, as shown in figureX there is little
+difference, in terms of performance between caching and not caching when using
+the kernel client.
+
 \subsection{Fault Tolerance}
+\label{sec:fault-tolerance}
 \begin{figure*}[tb] \centering
@@
 
-From this experiment we make 2 conclusions:
-\begin{itemize}
-  \item the metadata server is not overloaded because it can handle both workloads
-  \item the metadata server needs to get global consistency from only 1 client
-\end{itemize}
+From this experiment we make 2 conclusions: (1) the metadata server is not
+overloaded because it can handle both workloads and (2) the metadata server
+needs to get global consistency from only 1 client
 
@@
 
-\section{Methodology}
-
-\subsection{Decoupled Namespaces in CephFS}
+\section{Methodology: Decoupled Namespaces}
 
 \begin{figure}[tb]
+\caption{Applications can decouple the namespace, write updates to a local
+journal, and delay metadata updates.  Table~\ref{table:spectrum} shows how
+these phases (represented by the arrows) can be combined to provide weaker
+consistency or fault tolerance semantics.  }\label{fig:decouple}
 \centering
 \includegraphics[width=90mm]{figures/fig-decouple.png}
-\caption{Applications that can tolerate weaker consistency can decouple the
-namespace, write updates to a local file, and delay metadata updates. The merge
-and replay steps are optional if there is no need for a global namespace.
-}\label{fig:decouple}
 \end{figure}
 
-%- How it works (timestamps)
-Recall that CephFS uses RADOS (1) as a metadata store for all information about
-files including the hierarchical namespace and (2) as a staging area for the
-journal of updates before they are applied to the metadata store.  The journal
-tool is used for disaster recovery and lets administrators view and modify the
-journal of metadata updates; it can read the journal, erase events from the
-journal, and apply the updates in the journal to the metadata store.  To apply
-journal updates to the metadata store, the journal tool reads the journal
-segments from RADOS objects and applies the update to the metadata store (which
-are also stored as RADOS objects).
+In this section we describe Cudele, our prototype system that lets future
+programmers compose mechanisms (Section~\S\ref{sec:cudeles-mechanisms}) to
+provide the necessary guarantees
+(Section~\S\ref{sec:setting-policies-with-cudele}) for their application.
+
+\subsection{Cudele's Mechanisms}
+\label{sec:cudeles-mechanisms}
+
+% describe the figure
+Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
+which entity they are performed by (gray boxes). The metadata store and journal
+are different ways of representing the namespace. The metadata store represents
+the namespace as a tree of directory fragments and is easier to read/traverse.
+On the other hand, the journal represent the namespace as a list of events. It
+a ``pile system"; writes are fast but reads are slow because state must be
+reconstructed.  Specifically, reads are slow because there is more data to
+read, it is unorganized, and many of the updates may be redundant.
+
+% describe the mechanisms
+Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Save, and
+Persist. ``RPCs" does round trip remote procedure calls to establish
+consistency; it is the default implementation for complying with POSIX in
+CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
+into the object store. ``Create" allows clients to append metadata events to an
+in-memory journal. ``Volatile apply" takes the in-memory journal on the client
+and applies it directly to the in-memory metadata store of the metadata server
+cluster. ``Save" takes the in-memory journal and writes it to the client's
+disk. ``Persist" saves the journal as a an object in the object store from the
+client.
+
+Next, we discuss how these mechanisms can be composed to get different
+consistency and fault tolerance semantics.
+
+%\begin{tabular}{ r | l }
+%  \(\Rightarrow\)   & Description \\\hline
+%  create            & events appended to in-memory journal \\
+%  v.apply           & journal volatily applied to in-memory metadata store \\
+%  save              & journal saved to client's disk \\
+%  persist           & journal saved in object store \\
+%  replay            & metadata servers materialize namespace \\
+%  RPCs              & round trip remote procedure calls \\
+%  stream            & metadata server streams journal into RADOS \\
+%\end{tabular}
+
+\subsection{Setting Policies with Cudele}
+\label{sec:setting-policies-with-cudele}
+
+\begin{table}[t]
+\begin{center}
+\caption{Cudele lets future programmers explore the spectrum of consistency and
+fault tolerance semantics shown below.\label{table:spectrum}}
+\begin{tabular}{ r | l | l | l }
+  consistency     \(\rightarrow\) &&& \\  
+  fault tolerance \(\downarrow\)  & none & eventual & global \\\hline
+  none   & create   & create   & RPCs    \\
+         &          & +v.apply &         \\\hdashline
+  local  & create   & create   & RPCs    \\
+         & +save    & +save    & +save   \\
+         &          & +v.apply &         \\\hdashline
+  global & create   & create   & RPCs    \\
+         & +persist & +persist & +stream \\
+         &          & +v.apply &         \\
+\end{tabular}
+\end{center}
+\end{table}
+
+% describe table
+The spectrum of consistency and fault tolerance guarantees that adminstrators
+can construct is shown in Table~\ref{table:spectrum}. The columns are the
+different consistency semantics and the rows cover the spectrum of fault
+tolerance guantees. For consistency: ``none" means the system does not handle
+merging updates into a global namespace and it is assumed that middleware or
+the application manages consistency lazily; ``eventual" merges updates either
+when the system has time (e.g., by a background daemon) or when the client is
+done writing; and updates in ``global" consistency are seen immediately by all
+clients. For fault tolerance, ``none" means that updates are volatile and will
+be lost on a crash of any component. Stronger guarantees are made with
+``local", which means updates will be retained if the client node recovers, and
+``global", where all updates are always recoverable.
+
+% which system they represent and which are impossible
+The cells in Table~\ref{table:spectrum} encompass many exisiting storage
+systems. POSIX systems like CephFS and IndexFS have global consistency and
+fault tolerance; DeltaFS has consistency and fault tolerance set to ``none";
+and BatchFS uses ``eventual" consistency and ``local" fault tolerance. These
+are just a few of the HPC examples.
+
+% how its done
+To compose the mechanisms administrators inject which steps (described in
+Section~\S\ref{sec:cudeles-mechanisms}) to run and which to use in parallel
+using a domain specific language. For example, to get the semantics of BatchFS,
+the administrator would inject the following pipeline:\\
+
+\noindent \texttt{create+save+v.apply}\\
+
+% what is impossible
+Although we can achieve all permutations of the different guarantees in
+Table~\ref{table:spectrum}, not all of them make much sense. For example, it
+makes little sense to do \texttt{creates+RPCs} since both steps do the same
+thing or \texttt{stream+save} since global fault tolerance is stronger and has
+more overhead than local fault tolerance.  Formulaically, valid steps
+include:\\
+
+\noindent \texttt{<create|RPCs> [+v.apply][+save][+persist][+stream]}
+
+\subsection{Cudele as a Programmable Storage System}
+
+% Programmability
+We use a programmable storage approach to expose internal subsystem
+abstractions as ``interfaces" for applications to create higher level
+functionality and services.
+
+\subsubsection{No Changes: Turning Knobs}
+
+RPCs is the default behavior of the storage system. Depending on the cache
+state and configuration each operation incurrs at least RPC. For example, if a
+client or metadata server is not caching the directory inode, all creates
+within that directory will result in a lookup and a create request. If the
+directory inode is cached then only the create needs to be sent.
+
+Re-using the debugging and testing features in the storage system, we can turn
+``stream" on and off. If it is off, then the metadata servers will not save
+journals in the object store and the daemons that apply the journal to the
+metadata store will never run. Performance numbers for enabling and disabling
+the stream phase are presented back in Section~\S\ref{sec:fault-tolerance}.
+
+\subsubsection{External Library: Reads/Writes Journal Events}
+
+% How it works
+For ``create", ``save", and ``persist" Cudele provides a library for clients to
+link into.  Recall that CephFS uses the object store (1) as a metadata store
+for all information about files including the hierarchical namespace and (2) as
+a staging area for the journal of updates before they are applied to the
+metadata store. As discussed previously, the metadata store is optimized for
+reading while the journal is optimized for writing.  Cudele re-uses the journal
+tool to read/write log events to memory and persistent storage.
+
+% creates
+The journal tool is used for disaster recovery and lets administrators view and
+modify the journal of metadata updates; it can read the journal, erase events
+from the journal, and apply the updates in the journal to the metadata store.
+To apply journal updates to the metadata store, the journal tool reads the
+journal segments from object store objects and applies the update to the
+metadata store (which are also stored as object store objects).  For create,
+clients write to an in-memory journal and updates are merged into the global
+namespace by replaying them onto the metadata store in the object store.
+
+% persist/save
+For save, clients write serialized log events to a file on local disk and for
+persist, clients store the file in the objects store. The overheads for save
+and persist is the write bandwidth of the local disk and object store,
+respectively.
+
+% Level of intrusion
+This required no changes to Ceph because the metadata server knows how to read
+the events we were writing into the object store.  The client writes metadata
+updates locally and merges the updates with the journal tool. The client that
+decouples the namespace operates without any consistency and any conflicts at
+the merge are resolved in favor of this client. Updates by other clients ({\it
+i.e.} metadata writes to the global namespace) are overwritten.  We leverage
+the journal tool's ability to reconstruct metadata events in memory. The client
+library is shown in Figure~\ref{fig:decouple}.  Cudele adopts the following
+process when an application decouples the namespace:
+
+\begin{enumerate}
+  \item ``\textbf{decoupled}": metadata server exports the journal events to a file
+  \item ``\textbf{transfer}": the file is pulled by the client from the metadata server
+  \item ``\textbf{snapshot}": client reads the file and materializes a snapshot of the
+        namespace in memory
+\end{enumerate}
+
+% Why we re-use stuff
+By using re-using the journal subsystem to implement the namespace decoupling,
+Cudele leverages the write/read optimized data structures, the formats for
+persisting events (similar to TableFS's SSTables), and the functions for
+replaying events onto the internal namespace data structures.
+
+%Step 3 is the most complicated and requires understanding how the snapshot is
+%materialized in memory. 
+%
+%\subsubsection{Operating on Snapshots} 
+%
+%Our first implementation attempted to re-create journal events using the same
+%libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
+%to instantiate a Ceph inode and directory entry for the current file/dir and
+%its parent.  This is too hard because there are too many moving parts in the
+%metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
+%traverse up and down namespace, etc.). So when I tried add dentries and inodes
+%it was trying to traverse up/down and it would almost always segfault when it
+%was looking for something. These metablobs are supposed to be self container --
+%the problem is I do not know what is supposed to go *inside* them. 
+%
+%Our second idea was to copy the metadata blog and change just what we needed.
+%For example, we would save a binary dump of a generic \texttt{mkdir} event on
+%disk. When the application makes a directory, this dump would be loaded and the
+%fields would be changed before being written back to disk. Rather than
+%traversing up and down a namespace in memory of a metadata server, we should
+%traverse up and down the namespace *inside* the metadata blob. This
+%implementation requires disk IO and editing the log event is non-trivial for
+%two reasons:
+%
+%\begin{itemize}
+%
+%  \item methods do not edit events; they just write them
+%
+%  \item the metadata that the event touches (e.g., the metablob) is unorganized
+%  on disk for performance -- it is trade-off for writing data faster serially and
+%  reconstructing information slowly since failure is not considered the norm
+%
+%\end{itemize}
+%
+%Faced with these challenges we landed on our final implementation: load the
+%snapshot into the the data structures used to examine and replay journals, edit
+%those data structures, and write them out to disk as binary.
+
+\subsubsection{Storage System Changes: Volatile Merge}
+
+%The metadata objects are located with naming schemes (200.000* for journal
+%objects and 1.inode for metadata storage objects).
+
+% How it works: socket for changing daemon's internal state (debugging, logging, behaviour)
+% 1. API for putting state into the daemon dynamically
+% 2. Documentation all the tunables
+% 3. Hooks directly into daemon code so we can use any parsing functionality in there
+
+\section{Implementation}
+\label{sec:implementation}
+
+% What we actually implemented
+Of these 6 mechanisms only 4 had to implemented. ``RPCs" and ``Stream" can be
+achieved with tunables in configuration files for the storage system (e.g., metadata cache size,
+logging on/off, etc.). 
+``Persist", ``Save", and ``Create" are implemented in
+separate programs and libraries and do not require modifications to the storage
+system. ``Volatile Apply" requires changes to the metadata server to inject
+updates back into the global namespace.
+
+% Outline of the section
+In this section, we describe how we decoupled the consistency and fault
+tolerance policies from their mechanisms to allow applications to choose the
+semantics they need. Each section corresponds ot the labeled arrows in
+Figure~\ref{fig:decouple}.
+
+\subsection{RPCs (Existing CephFS Implementation)}
+
+CephFS uses round-trip remote procedure calls (RPCs) for most of
+
+\subsection{Stream (Existing Implementation)}
 
@@
 
+\subsection{Create}
+
 % Journal import
@@
 order and starting from the write position saved in the header. Probing will
-update the write position if it finds objects with data in them. There are two
-types of objects:
-
-1. journal objects (e.g., 200.000*)
-
-2. metadata store objects (e.g., 1.inode)
-
-We need to update the journal objects with the new events! How do we write out
-new objects to RADOS (e.g., 200.000*)? These should be the corect sizes (e.g.,
-4980 bytes) -- these log events that we are adding are not making them into
-the journal objects... do we have to explicitly read them and then re-write
-them?
-
-TODO: pull the object, read it, append to it, make sure size changes. It looks
-like we can only read these objects and apply doesn't actually touch them... it
-only touches the metadata store objects (e.g., 1.inode*).
-
-% TODO: figure out how import saves the journal on the metadata server
-% - they aren't -- everything is stored in RADOS
-% - we never touch the metadata servers!!!
-% TODO: figure out how to read the header object
-% TODO: figure out the mapping objects to journal segments (not 1x1)
+update the write position if it finds objects with data in them.
 
@@
 
-\subsubsection{Leveraging the Journal Tool}
+%- How we leverage it
 
-%- How we leverage it
-We leverage the journal tool's ability to reconstruct the namespace for the
-metadata blob. First we scan journal and store all events in a map. Next we
-create a new event and copy the root inode directory lump from another journal
-event into the event's metadata blob. Finally, we add a new directory lump for
-our new file or directory.
-
-If decoupled namespaces are turned on, the client writes metadata updates
-locally and merges the updates with the journal tool. The client that decouples
-the namespace operates without any consistency and any conflicts at the merge
-are resolved in favor of this client. Updates by other clients ({\it i.e.}
-metadata writes to the global namespace) are overwritten. Cudele adopts the
-following process when an application decouples the namespace:
-
-\begin{enumerate}
-  \item metadata server exports a snapshot of the file system to a binary file
-  \item client materializes the snapshot in memory
-  \item client operates on the data structures in memory
-  \item client exports the new snapshot to a binary file
-  \item metadata server merges the snapshot into the global namespace
-\end{enumerate}
-
-Step 3 is the most complicated and requires understanding how the snapshot is
-materialized in memory.
-
-\subsubsection{Operating on Snapshots}
-
-Our first implementation attempted to re-create journal events using the same
-libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
-to instantiate a Ceph inode and directory entry for the current file/dir and
-its parent.  This is too hard because there are too many moving parts in the
-metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
-traverse up and down namespace, etc.). So when I tried add dentries and inodes
-it was trying to traverse up/down and it would almost always segfault when it
-was looking for something. These metablobs are supposed to be self container --
-the problem is I do not know what is supposed to go *inside* them.
-
-Our second idea was to copy the metadata blog and change just what we needed.
-For example, we would save a binary dump of a generic \texttt{mkdir} event on
-disk. When the application makes a directory, this dump would be loaded and the
-fields would be changed before being written back to disk. Rather than
-traversing up and down a namespace in memory of a metadata server, we should
-traverse up and down the namespace *inside* the metadata blob. This
-implementation requires disk IO and editing the log event is non-trivial for
-two reasons:
-
-\begin{itemize}
-
-  \item methods do not edit events; they just write them
-
-  \item the metadata that the event touches (e.g., the metablob) is unorganized
-  on disk for performance -- it is trade-off for writing data faster serially and
-  reconstructing information slowly since failure is not considered the norm
-
-\end{itemize}
-
-Faced with these challenges we landed on our final implementation: load the
-snapshot into the the data structures used to examine and replay journals, edit
-those data structures, and write them out to disk as binary.
+\subsection{Volatile Apply (v.apply)}
+\label{sec:vapply}
+\subsection{Save}
+\label{sec:save}
+\subsection{Persist}
+\label{sec:persist}
 
 \section{Results}
+\subsection{Microbenchmarks}
 
-In this section, we quantify the costs of consistency and fault tolerance in
-CephFS. If the components that ensure these semantics (i.e. capabilities and
-journals, respectively) can be mitigated or delayed, then global namespaces
-perform as well as decoupled namespaces.
-
-\textbf{Experimental Setup}: we run our experiments on a 9 OSD, 3 MDS, 1 MON
-Ceph cluster. The clients use the Ceph kernel client, which has been in the
-mainline Linux kernel since TODO. We use the kernel client so that we can find
-the true create speed of the server; our experiments show a low CPU utilziation
-for the clients which indicates that we are stressing the servers more. We also
-turn caching off becuase, as shown in figureX there is little difference, in
-terms of performance between caching and not caching when using the kernel
-client.
+\subsubsection{Per phase latencies}
+\begin{figure}[tb]
+\centering
+\includegraphics[width=90mm]{figures/global-v-decoupled.png}
+\caption{Compared to the \texttt{create} phase, saving and persisting
+updates (\texttt{create+save} and \texttt{create+save+persist}) experience only
+a 4.79\(\times\) and 8.66\(\times\) slowdown, in the worst case for 100K files.
+In contrast, maintaining \texttt{global} consistency is 905.70\(\times\)
+slower.  The disadvantage of decoupling the namespace is the merge phase where
+updates are applied to the metadata store (\texttt{create+apply}, resulting in a
+905.70\(\times\) slowdown for 100K files.}\label{fig:global-v-decoupled}
+\end{figure}
 

### 2017-02-22 paper/paper.tex
paper: add microbenchmarks
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\subsection{Cudele as a Programmable Storage System}
+\subsection{Subtree Policies}
+
+\section{Implementation}
+\label{sec:implementation}
+
+% Outline of the section
+Each section below corresponds ot the labeled arrows in
+Figure~\ref{fig:decouple}. This implementation decouples policies from
+mechanisms allowing applciations to choose the consistency and fault tolerance
+semantics they need.
 
 % Programmability
-We use a programmable storage approach to expose internal subsystem
-abstractions as ``interfaces" for applications to create higher level
-functionality and services.
-
-\subsubsection{No Changes: Turning Knobs}
+Of the 6 mechanisms in Figure~\ref{fig:decouple} 4 had to implemented and only
+1 required changes to the underlying storage system itself.  RPCs and stream
+can be achieved with tunables in configuration files for the storage system
+(e.g., metadata cache size, logging on/off, etc.).  Persist", save, and create
+are implemented as library and does not require modifications to the storage
+system. Volatile apply requires changes to the metadata server to inject
+updates back into the global namespace.
+
+\subsection{No Changes: RPCs, Stream}
 
@@
 Re-using the debugging and testing features in the storage system, we can turn
-``stream" on and off. If it is off, then the metadata servers will not save
+stream on and off. If it is off, then the metadata servers will not save
 journals in the object store and the daemons that apply the journal to the
@@
 
-\subsubsection{External Library: Reads/Writes Journal Events}
+% General information about the journal
+The journal segments are saved as objects in RADOS.  The journal has 4
+pointers, described in `osdc/Journaler.h`:
+
+\begin{itemize}
+  \item write position: tail of the journal; points to the current session where we are appending events
+  \item unused field: where someone is reading
+  \item expire position: old journal segments
+  \item trimmed position: where daemon is expiring old items
+\end{itemize}
+
+% How the journal tool works
+Journal segments in RADDS have a header followed by serialized log events. The
+log events are read by hopping over objects using the read offset and object
+size pulled from the journal header.  After decoding them, we can examine the
+metadata (1) about the event (e.g., type, timestamps, etc.) and (2) for inodes
+that the event touches.
+
+% The metablobs
+The metadata for inodes that the event touches are called metadata blobs and
+the ones associated with events are \textbf{unordered}; this layout makes
+writing journal updates fast but the cost is realized when reading the metadata
+blobs.  It makes sense to optimize for writing since reading only occurs on
+failures. To reconstruct the namespace for the metadata blob, the journal tool
+iterates over each metadata blob in the events and builds mappings of inodes to
+directory entries (for directories) and parent inodes to child inodes/directory
+entries.
+
+\subsection{External Library: Create, Save, Persist}
+%Reads/Writes Journal Events}
 
@@
 namespace by replaying them onto the metadata store in the object store.
-
-% persist/save
-For save, clients write serialized log events to a file on local disk and for
-persist, clients store the file in the objects store. The overheads for save
-and persist is the write bandwidth of the local disk and object store,
-respectively.
-
-% Level of intrusion
-This required no changes to Ceph because the metadata server knows how to read
-the events we were writing into the object store.  The client writes metadata
-updates locally and merges the updates with the journal tool. The client that
-decouples the namespace operates without any consistency and any conflicts at
-the merge are resolved in favor of this client. Updates by other clients ({\it
-i.e.} metadata writes to the global namespace) are overwritten.  We leverage
-the journal tool's ability to reconstruct metadata events in memory. The client
-library is shown in Figure~\ref{fig:decouple}.  Cudele adopts the following
-process when an application decouples the namespace:
-
-\begin{enumerate}
-  \item ``\textbf{decoupled}": metadata server exports the journal events to a file
-  \item ``\textbf{transfer}": the file is pulled by the client from the metadata server
-  \item ``\textbf{snapshot}": client reads the file and materializes a snapshot of the
-        namespace in memory
-\end{enumerate}
-
-% Why we re-use stuff
-By using re-using the journal subsystem to implement the namespace decoupling,
-Cudele leverages the write/read optimized data structures, the formats for
-persisting events (similar to TableFS's SSTables), and the functions for
-replaying events onto the internal namespace data structures.
-
-%Step 3 is the most complicated and requires understanding how the snapshot is
-%materialized in memory. 
-%
-%\subsubsection{Operating on Snapshots} 
-%
-%Our first implementation attempted to re-create journal events using the same
-%libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
-%to instantiate a Ceph inode and directory entry for the current file/dir and
-%its parent.  This is too hard because there are too many moving parts in the
-%metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
-%traverse up and down namespace, etc.). So when I tried add dentries and inodes
-%it was trying to traverse up/down and it would almost always segfault when it
-%was looking for something. These metablobs are supposed to be self container --
-%the problem is I do not know what is supposed to go *inside* them. 
-%
-%Our second idea was to copy the metadata blog and change just what we needed.
-%For example, we would save a binary dump of a generic \texttt{mkdir} event on
-%disk. When the application makes a directory, this dump would be loaded and the
-%fields would be changed before being written back to disk. Rather than
-%traversing up and down a namespace in memory of a metadata server, we should
-%traverse up and down the namespace *inside* the metadata blob. This
-%implementation requires disk IO and editing the log event is non-trivial for
-%two reasons:
-%
-%\begin{itemize}
-%
-%  \item methods do not edit events; they just write them
-%
-%  \item the metadata that the event touches (e.g., the metablob) is unorganized
-%  on disk for performance -- it is trade-off for writing data faster serially and
-%  reconstructing information slowly since failure is not considered the norm
-%
-%\end{itemize}
-%
-%Faced with these challenges we landed on our final implementation: load the
-%snapshot into the the data structures used to examine and replay journals, edit
-%those data structures, and write them out to disk as binary.
-
-\subsubsection{Storage System Changes: Volatile Merge}
-
-%The metadata objects are located with naming schemes (200.000* for journal
-%objects and 1.inode for metadata storage objects).
-
-% How it works: socket for changing daemon's internal state (debugging, logging, behaviour)
-% 1. API for putting state into the daemon dynamically
-% 2. Documentation all the tunables
-% 3. Hooks directly into daemon code so we can use any parsing functionality in there
-
-\section{Implementation}
-\label{sec:implementation}
-
-% What we actually implemented
-Of these 6 mechanisms only 4 had to implemented. ``RPCs" and ``Stream" can be
-achieved with tunables in configuration files for the storage system (e.g., metadata cache size,
-logging on/off, etc.). 
-``Persist", ``Save", and ``Create" are implemented in
-separate programs and libraries and do not require modifications to the storage
-system. ``Volatile Apply" requires changes to the metadata server to inject
-updates back into the global namespace.
-
-% Outline of the section
-In this section, we describe how we decoupled the consistency and fault
-tolerance policies from their mechanisms to allow applications to choose the
-semantics they need. Each section corresponds ot the labeled arrows in
-Figure~\ref{fig:decouple}.
-
-\subsection{RPCs (Existing CephFS Implementation)}
-
-CephFS uses round-trip remote procedure calls (RPCs) for most of
-
-\subsection{Stream (Existing Implementation)}
-
-% General information about the journal
-The journal segments are saved as objects in RADOS.  The journal has 4
-pointers, described in `osdc/Journaler.h`:
-
-\begin{itemize}
-  \item write position: tail of the journal; points to the current session where we are appending events
-  \item unused field: where someone is reading
-  \item expire position: old journal segments
-  \item trimmed position: where daemon is expiring old items
-\end{itemize}
-
-% How the journal tool works
-Journal segments in RADDS have a header followed by serialized log events. The
-log events are read by hopping over objects using the read offset and object
-size pulled from the journal header.  After decoding them, we can examine the
-metadata (1) about the event (e.g., type, timestamps, etc.) and (2) for inodes
-that the event touches.
-
-% The metablobs
-The metadata for inodes that the event touches are called metadata blobs and
-the ones associated with events are \textbf{unordered}; this layout makes
-writing journal updates fast but the cost is realized when reading the metadata
-blobs.  It makes sense to optimize for writing since reading only occurs on
-failures. To reconstruct the namespace for the metadata blob, the journal tool
-iterates over each metadata blob in the events and builds mappings of inodes to
-directory entries (for directories) and parent inodes to child inodes/directory
-entries.
-
-\subsection{Create}
 
@@
 
-%- How we leverage it
-
-\subsection{Volatile Apply (v.apply)}
-\label{sec:vapply}
-\subsection{Save}
-\label{sec:save}
-\subsection{Persist}
-\label{sec:persist}
+% persist/save
+For save, clients write serialized log events to a file on local disk and for
+persist, clients store the file in the objects store. The overheads for save
+and persist is the write bandwidth of the local disk and object store,
+respectively.
+
+% Level of intrusion
+This required no changes to Ceph because the metadata server knows how to read
+the events we were writing into the object store.  The client writes metadata
+updates locally and merges the updates with the journal tool. The client that
+decouples the namespace operates without any consistency and any conflicts at
+the merge are resolved in favor of this client. Updates by other clients ({\it
+i.e.} metadata writes to the global namespace) are overwritten.  We leverage
+the journal tool's ability to reconstruct metadata events in memory. The client
+library is shown in Figure~\ref{fig:decouple}.  Cudele adopts the following
+process when an application decouples the namespace:
+
+\begin{enumerate}
+  \item ``\textbf{decoupled}": metadata server exports the journal events to a file
+  \item ``\textbf{transfer}": the file is pulled by the client from the metadata server
+  \item ``\textbf{snapshot}": client reads the file and materializes a snapshot of the
+        namespace in memory
+\end{enumerate}
+
+% Why we re-use stuff
+By using re-using the journal subsystem to implement the namespace decoupling,
+Cudele leverages the write/read optimized data structures, the formats for
+persisting events (similar to TableFS's SSTables), and the functions for
+replaying events onto the internal namespace data structures.
+
+%Step 3 is the most complicated and requires understanding how the snapshot is
+%materialized in memory. 
+%
+%\subsubsection{Operating on Snapshots} 
+%
+%Our first implementation attempted to re-create journal events using the same
+%libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
+%to instantiate a Ceph inode and directory entry for the current file/dir and
+%its parent.  This is too hard because there are too many moving parts in the
+%metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
+%traverse up and down namespace, etc.). So when I tried add dentries and inodes
+%it was trying to traverse up/down and it would almost always segfault when it
+%was looking for something. These metablobs are supposed to be self container --
+%the problem is I do not know what is supposed to go *inside* them. 
+%
+%Our second idea was to copy the metadata blog and change just what we needed.
+%For example, we would save a binary dump of a generic \texttt{mkdir} event on
+%disk. When the application makes a directory, this dump would be loaded and the
+%fields would be changed before being written back to disk. Rather than
+%traversing up and down a namespace in memory of a metadata server, we should
+%traverse up and down the namespace *inside* the metadata blob. This
+%implementation requires disk IO and editing the log event is non-trivial for
+%two reasons:
+%
+%\begin{itemize}
+%
+%  \item methods do not edit events; they just write them
+%
+%  \item the metadata that the event touches (e.g., the metablob) is unorganized
+%  on disk for performance -- it is trade-off for writing data faster serially and
+%  reconstructing information slowly since failure is not considered the norm
+%
+%\end{itemize}
+%
+%Faced with these challenges we landed on our final implementation: load the
+%snapshot into the the data structures used to examine and replay journals, edit
+%those data structures, and write them out to disk as binary.
+
+\subsection{Storage System Changes: Volatile Merge}
+
+%The metadata objects are located with naming schemes (200.000* for journal
+%objects and 1.inode for metadata storage objects).
+
+% How it works: socket for changing daemon's internal state (debugging, logging, behaviour)
+% 1. API for putting state into the daemon dynamically
+% 2. Hooks directly into daemon code so we can use any parsing functionality in there
+% 3. Documentation all the tunables
+
+% 1. read journal of updates from file
+% 2. call replay (uses same code as when an MDS comes back) on each event
+% 3. skip inodes so MDS doesn't hand out those new inodes
 
@@
 \subsection{Microbenchmarks}
+\begin{figure*}[tb]
+\centering
+\includegraphics[width=180mm]{figures/results-microbenchmark.png}
+\caption{Here's a graph.}
+\end{figure*}
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=90mm]{figures/results-files.png}
+\caption{Here's a graph.}
+\end{figure}
 

### 2017-02-22 experiments/fuse/batchfs/run.sh
benchmark: finish microbenchmarks
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -8,21 +8,27 @@ ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
 CEPH_ANSIBLE="$RUN -v `pwd`/site/roles/ceph-ansible:/root $ANSIBLE"
 SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
 
-# configure ceph and setup results directory
-cp site/* site/roles/ceph-ansible || true
-cp -r site/group_vars site/roles/ceph-ansible/
-cp site/hosts site/roles/ceph-ansible/hosts
-
 # cleanup and start ceph
 for i in 0 1 2; do
   for nfiles in 10 100 1000 10000 100000; do
     mkdir -p results/${nfiles}/logs || true
-    $SRL_ANSIBLE cleanup.yml
-    $CEPH_ANSIBLE ceph.yml cephfs.yml
-    $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
-     
-    ./ansible-playbook.sh -e nfiles=$nfiles ../workloads/journal.yml
-    ./ansible-playbook.sh -e site=$nfiles collect.yml
+    for stream in "nostream" "stream"; do
+      # configure ceph and setup results directory
+      cp site/configs/${stream}.yml site/group_vars/all
+      cp site/* site/roles/ceph-ansible || true
+      cp -r site/group_vars site/roles/ceph-ansible/
+      cp site/hosts site/roles/ceph-ansible/hosts
+
+      # setup
+      $SRL_ANSIBLE cleanup.yml
+      $CEPH_ANSIBLE ceph.yml cephfs.yml
+      $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
+       
+      # benchmark
+      ./ansible-playbook.sh -e nfiles=$nfiles -e stream=$stream \
+        ../workloads/journal-rpcs.yml ../workloads/journal-vapply.yml
+      ./ansible-playbook.sh -e site=$nfiles collect.yml
+    done
   done
   mv results results-run${i}
 done

### 2017-03-08 paper/paper.tex
paper: switch template, refactor sections, add pretty code
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-01-31 paper/paper.tex
@@
-\documentclass[preprint]{sigplanconf-eurosys}
+% Template: http://www.acm.org/publications/proceedings-template
+\documentclass[sigconf]{acmart}
+\usepackage{booktabs} % For formal tables
+\usepackage{minted}
 \usepackage{graphicx}
+\usepackage{arydshln}
 
-\title{Global Namespaces Scale as Well as Decoupled Namespaces}
-\authorinfo{Paper ID: 165}
-           {Total Pages: 12}
-\date{}
+% Copyright
+%\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
+
+% DOI
+\acmDOI{10.475/123_4}
+
+% ISBN
+\acmISBN{123-4567-24-567/08/06}
+
+%Conference
+\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El
+  Paso, Texas USA} 
+\acmYear{1997}
+\copyrightyear{2016}
+
+\acmPrice{15.00}
+
 \begin{document}
+\title{Cudele: Programmable Consistency and Fault Tolerance}
+
+\author{Blind Author}
+\orcid{1234-5678-9012}
+\affiliation{%
+  \institution{Institute for Clarity in Documentation}
+  \streetaddress{P.O. Box 1212}
+  \city{Dublin} 
+  \state{Ohio} 
+  \postcode{43017-6221}
+}
+\email{blindauthor@corporation.com}
+
+\input{sections/0_abstract}
+
+%
+% The code below should be generated by the tool at
+% http://dl.acm.org/ccs.cfm
+% Please copy and paste the code instead of the example below. 
+%
+\begin{CCSXML}
+<ccs2012>
+ <concept>
+  <concept_id>10010520.10010553.10010562</concept_id>
+  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+  <concept_significance>500</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010575.10010755</concept_id>
+  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+  <concept_significance>300</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10010520.10010553.10010554</concept_id>
+  <concept_desc>Computer systems organization~Robotics</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+ <concept>
+  <concept_id>10003033.10003083.10003095</concept_id>
+  <concept_desc>Networks~Network reliability</concept_desc>
+  <concept_significance>100</concept_significance>
+ </concept>
+</ccs2012>  
+\end{CCSXML}
+
+\ccsdesc[500]{Computer systems organization~Embedded systems}
+\ccsdesc[300]{Computer systems organization~Redundancy}
+\ccsdesc{Computer systems organization~Robotics}
+\ccsdesc[100]{Networks~Network reliability}
+
+% We no longer use \terms command
+%\terms{Theory}
+
+\keywords{ACM proceedings, \LaTeX, text tagging}
+
 \maketitle
 
-% I work with the file system layer that sits on top of Ceph
-% - internal subsystems and interfaces that Ceph uses to provide a strongly consistent POSIX file system
-%   - e.g., load balancing mechanisms, for hotspots/flash crowds on directories
-%     caches for inodes, capabilities for guarding a shared resource, and a large journal for fault tolerance
-% - our programmability initiative talks about composing these internal interfaces for higher level services
+\input{sections/1_introduction}
+%\input{sections/2_related_work}
+\input{sections/3_posix_overheads}
+\input{sections/4_methodology}
+\input{sections/5_implementation}
+\input{sections/6_results}
 
-% 
-% Previous work: 
-% - master's thesis: scale-out vs. scale-up architectures 
-% - work in industry: file systems for the cloud, big data processing over
-% object storage, and automating workflows
-%
-% I work under Carlos at UC Santa Cruz and we study how to migrate file system
-% metadata to alleviate overloaded servers. The idea is to take the namespace
-% on one metadata server, chop it up into smaller subtrees, and farm those
-% subtrees out to other metadata servers.  It's very similar to IndexFS/ShardFS
-% but we try to leverage namespace locality by allowing variable sized
-% subtrees. We published a paper at SC'15 where we talk about a load balancing
-% API we built on CephFS that let's administrators control how to chop up
-% subtrees and when/where to move them (also got merged into Ceph).
-%
-% Go to conclusion
-%
-% Why I want to work with Brad
-% - new perspectives on file system designs, especially NON-POSIXy 
-% - work in a national lab: I know about theirarchitectures and problems but
-% I've never used an HPC system - real world applications and benchmarks; most
-% of my work has been accepted for novelty but not for fully benchmarking a
-% system
-% - lines up with my thesis
-%
-% What I want to work on:
-% - continue working on my thesis; this next phase of having a global namespace
-% but different consistency semantics within that namespace for data management
-%   - namespace: recursive data structure as a logical model for understanding the user's goals
-%   - merge: what is the proper way to merge deltafs/batchfs updates back into the global namespace
-%   - data management: deltaFS made the case that apps shouldn't use the fs to synchronize;
-%     we contend that you could use the fs as data management; remembering how to access your data
-%
-% What I have been working on 
-% - leverage internal services/subsystems in Ceph; avoids code duplication and
-% let's us test this over the same system 
-% - benchmarking the overheads of strong consistency/journalling - we've figured
-% out how to thrash capabilities and turn off the journal
-\begin{abstract}
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{sigproc}
 
-Today's large and highly-parallel workloads are bottlenecked by metadata
-services because many processes end up accessing the same shared resoure. In
-HPC, state-of-the-art file systems are abandoning POSIX because the
-sychronization and serializiation overheads are too costly -- and sometimes
-even unnceccessary -- for some of their applications.  While the performance
-benefits are plain for these users, other applications that rely on stronger
-consistency must be re-written or deployed on a different system.  We present
-Cudele, a programmable file system that supports different degrees of
-consistency and fault tolerance within the same namespace.  First, we take a
-POSIX compliant file system and relax the consistency constraints by
-implementing two metadata designs: delayed metadata update merge into the
-global namespace and client local views of metadata updates.  Second, we store
-the consistency and fault tolerance semantics in inodes so that subtrees within
-the same namespace can be optimized for different workloads; the inodes are
-programmable so that clients understand how to access the metadata in a
-subtree.  Third, we present a theoretical framework for a metadata service and
-a working implementation that provides the lowest common denominator needed to
-implement a wide range of consistency/fault tolerance semantics that can be
-benchmarked all on the same system.
-
-\end{abstract}
-
-% Questions
-% - do you still collaborate with CMU? Would you want all of us to work together?
-% - what projects do you want to work on?
-% - how closely would you want to work with Carlos?
-% - what system would you want to work on?
-
-% The main difference between CephFS and the CMU systems is that our approach
-% is a little more flexible.  We can partition the namespace in any way and can
-% co-locate items together if they are dependent on each other -- in theroy...
-% in practice it is very hard to figure out what the workload is trying to do.
-% Also, we can program the decision making based on different metrics, like
-% CPU, request rate, etc.
-
-\section{Introduction}
-
-Today's client-server based file system metadata services have scalability
-problems. It takes a lot of resources to server POSIX metadata requests and
-applications perform better with dedicated metadata
-servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs}. This is fine for small
-workloads and file systems but as the system scales provisioning a metadata
-server for every client is expensive and complicated.
-
-Current hardware evolution and the rise of software-defined storage storage,
-which uses techniqes like erasure coding, replication, and partitioning, have
-ushered a new era of HPC computing; architectures are transitioning from
-complex storage stacks with burst buffer, file system, object store, and tape
-tiers to a two layer stack with just a burst buffer and object
-store~\cite{bent:login16-hpc-trends}. This trend exacerbates the metadata
-scalability problem and has given rise to the serverless metadata services.
-
-HPC workloads are so metadata intensive that new management techniques are
-advocating reducing synchronization and serialization overheads by transferring
-these responsibilities to the client. We call this approach decoupling the
-namespace and the semantics of consistency differ amongst systems.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=75mm]{figures/subtree-policies.png}
-\caption{Administrators can assign consistency and fault tolerance policies to
-subtrees to get the benefits of some of the state-of-the-art HPC architectures.
-}\label{fig:subtree-policies}
-\end{figure}
-
-We propose subtree policies, an interface that lets future programmers control
-how the system manages different parts of the namespace.  For performance one
-subtree can adopt weaker consistency semantics while another subtree can retain
-the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
-shows an example setup where a single global namespace has directories for
-applications designed for different, state-of-the-art HPC architectures.  Our
-system supports 3 forms of consistency and 2 forms of fault tolerance giving
-the administrator a wide range of policies and optimizations depending on the
-application's needs.
-
-\section{Related Work}
-
-% General
-The bottlenecks associated with accessing POSIX file system metadata are not limited
-to HPC workloads and the same challenges that plagued these systems for years are
-finding their way into the cloud. Workloads that deal with many small files
-({\it e.g.}, log processing and database
-queries~\cite{thusoo:sigmod2010-facebook-infrastructure}) and large numbers of
-simultaneous clients ({\it e.g.}, MapReduce
-jobs~\cite{mckusick:acm2010-gfs-evolution}), are subject to the scalability of
-the metadata service. The biggest challenge is that whenever a file
-is touched the client must access the file's metadata and maintaining a file
-system namespace imposes small, frequent accesses on the underlying storage
-system~\cite{roselli:atec2000-FS-workloads}.  Unfortunately, scaling file
-system metadata is a well-known problem and solutions for scaling data IO do
-not work for metadata IO~\cite{roselli:atec2000-FS-workloads,
-abad:techreport2012-fstrace, abad:ucc2012-mimesis,
-alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. There are two approaches
-for improving the performance of metadata access.
-
-\subsection{Metadata Load Balancing}
-
-% approaches to load balancing (strong = more resuorces per unit work, weak =
-% fixed resource per work unit)
-One approach for improving metadata performance and scalability is to alleviate
-overloaded servers by load balancing metadata IO across a cluster. Common
-techniques include partitioning metadata when there are many writes and
-replicating metadata when there are many reads. For example, IndexFS partitions
-directories and clients write to different partitions by grabbing leases and
-caching ancestor metadata for path traversal; it does well for strong scaling
-because servers can keep more inodes in the cache which results in less RPCs.
-Alternatively, ShardFS replicates directory state so servers do not need to
-contact peers for path traversal; it does well for read workloads because all
-file operations only require 1 RPC and for weak scaling because requests will
-never incurr extra RPCs due to a full cache.  CephFS employs both techniques to
-a lesser extent; directories can be replicated or sharded but the caching and
-replication policies do not change depending on the balancing technique.
-Despite the performance benefits these techniques add complexity and jeopardize
-the robustness and performance characteristics of the metadata service because
-the systems now need (1) policies to guide the migration decisions and (2)
-mechanisms to address inconsistent states across servers.
-
-% policies
-Setting policies for migrations is arguably more difficult than adding the
-migration mechanisms themselves.  For example, IndexFS and CephFS use the
-GIGA+~\cite{GIGA+} technique for partitioning directories at a predefined
-threshold and using lazy synchronization to redirect queries to the server that
-``owns" the targeted metadata.  Determining when to partition directories and
-when to migrate the directory fragments are policies that vary between systems:
-GIGA+ partitions directories when the size reaches a certain number of files
-and migrates directory fragments immediately; CephFS partitions directories
-when they reach a threshold size or when the write temperature reaches a
-certain value and migrates directory fragments when the hosting server has more
-load than the other servers in the metadata cluster. Another policy is when and
-how to replicate directory state; ShardFS replicates immediately and
-pessimistically while CephFS replicates only when the read temperature reaches
-a threshold.  There is a wide range of policies and it is difficult to address
-with tunables and hard-coded design decisions.
-
-% addressing inconsistency
-In addition to the policies, distributing metadata across a cluster requiress
-distributed transactions and cache coherence protocols to ensure strong
-consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
-directory state and uses optimistic concurrency control for conflicts; namely
-it does the operation and if there is a conflict at verification time it falls
-back to two-phase locking.  Another example is IndexFS's inode cache which
-reduces RPCs by caching ancestor paths -- the locality of this cache can be
-thrashed by random reads but performs well for metadata writes. For
-consistency, writes to directories in IndexFS block until the lease expires
-while writes to direcotries in ShardFS are slow for everyone as it either
-requires serialization or locking with many servers; reads in IndexFS are
-subject to cache locality while reads in ShardFS always resolve to 1 RPC.
-Another example of the overheads of addressing inconsistency is how CephFS
-maintains client sessions and inode caches for capabilities (which in turn make
-metadata access faster). When metadata is exchanged between metadata servers
-these sessions/caches must be flushed and new statistics exchanged with a
-scatter-gather process; this halts updates on the directories and blocks until
-the authoratitive metadata server responds.  These protocols are discusssed in
-more detail in the next section but their inclusion here is a testament to the
-complexity of migrating metadata.
-
-%For metadata writes it does a distributed transaction; monotonic writes with
-%concurrent clients fail and do pessimistic locking through a mediated lock
-%server to ensure strong consistency, non-monotonic writes grab locks at every
-%server. Zooming in on monotonic writes: if permissions increase it executes on
-%a primary then non-primary, if permissions decrease it executes on all
-%non-primary then on primary.
-
-The conclusion we have drawn from this related work is that metadata protocols
-have a bigger impact on performance and scalability than load balancing.  
-Understanding these protocols helps load balancing and gives us a better
-understanding of the metrics we should use to make migration decisions ({\it
-e.g.}, which operations reflect the state of the system), what types of
-requests cause the most load, and how an overloaded system reacts ({\it e.g.},
-increasing latencies, lower throughput, etc.).
-
-\subsection{Relaxing POSIX}
-
-% POSIX 
-POSIX workloads require strong consistency and many file systems improve
-performance by reducing the number of remote calls per operation ({\it i.e.}
-RPC amplification). As discussed in the previous section, caching with leases and
-replication are popular approaches to reducing the overheads of path traversals
-but their performance is subject to cache locality and the amount of available
-resources, respectively; for random workloads larger than the cache extra RPCs
-hurt performance~\cite{IndexFS, CephFS} and for write heavy workloads with more
-resources the RPCs for invalidations are harmful. Another approach to reducing
-RPCs is to use leases or capabilities.
-
-%IndexFS aggressively caches paths and
-%handles permissions by handing out leases for metadata writes; metadata may
-%only be modified when all leases have expired. 
-%IndexFS~\cite{ren:sc2014-indexfs} aggressively caches pathnames and their
-%permissions on the client servers. Modifications to metadata cached by clients
-%is delayed until all client leases have expired. This reduces the RPC
-%amplification to 1 when mutatating directory metadata ({\it e.g.,}
-%\texttt{mkdir}, \texttt{chmod}, etc.) because clients are not querying MDSs for
-%path traversals. The disadvantage of this approach is the high latency of
-%mutation operations (reads to filenames).  ShardFS~\cite{xiao:socc2015-shardfs}
-%replicates metadata (specifically directory lookup state) across the MDS
-%cluster, reducing the RPC amplification to 1 for file operations ({e.g.,}
-%\texttt{stat}, \texttt{chmod}, \texttt{chown}, etc.). Modifications to the
-%directory lookup state are done with optimistic concurrecny control and fall
-%back to retry if verification fails. The disadvantage of this appraoch is the
-%high number of RPCs for maintaining directory metadata mutations (writes to
-%directories).  CephFS also maintains an inode cache but its notion of leases
-%are much shorter, on the order of microseconds.
-
-% Non POSIX
-High performance computing has unique requirements for file systems ({e.g.},
-fast creates) and well-defined workloads (e.g., workflows) that make relaxing
-POSIX sensible.  One popular approach is to allow clients to ``lock'' parts of
-the namespace to improve performance and scalability by avoiding
-synchronization, false sharing, and serialization.  BatchFS assumes the
-application coordinates accesses to the namespace, so the clients can batch
-local operations and merge with a global namespace image lazily. Similarily,
-DeltaFS eliminates RPC traffic using subtree snapshots for non-conflicting
-workloads and middleware for conflicting workloads. MarFS gives administrators
-the ability to (1) lock ``project directories" and (2) allocate GPFS clusters
-for demanding directory workloads. TwoTiers eliminates high-latencies by
-storing metadata in a flash tier; apps lock the namespace so that metadata can
-be accessed more quickly.  Unfortunately, decoupling the namespaces has costs:
-(1) merging metadata state back into the global namespace is slow; (2) failures
-are local to the failing node; and (3) the systems are not backwards
-compatible.
-
-% NON POSIX
-%Decoupling the namespaces has many advantages, including improved scalability,
-%higher resource utilization, and better performance.  BatchFS and DeltaFS are
-%near-POSIX filesystems that give clients the ability to decouple subtrees from
-%the namespace so that the applications can execute metadata operations without
-%synchronization and serialization.  These operations are applied to a local
-%snapshot of the file system namespace and conflicts are resolved either by the
-%application or by an external service . Applications link into a metadata
-%server library to reduce resource utilization and code paths (e.g., no daemons
-%and less interprocess communication).
-
-For (1), state-of-the-art systems manage consistency in non-traditional ways:
-IndexFS maintains the global namespace but blocks operations from other clients
-until the first client drops the lease, BatchFS does operations on a snapshot
-of the namespace and merges batches of operations into the global namespace,
-and DeltaFS never merges back into the global namespace. The merging for
-BatchFS is done by an auxiliary metadata server running on the client and
-conflicts are resolved by the application. Although DeltaFS never explicitly
-merges, applications needing some degree of ground truth can either manage
-consistency thesmelves on a read or add a bolt-on service to manage the
-consistency.
-
-For (2), if the client fails and stays down, all metadata operations on the
-decoupled namespace are lost. If the client recovers, the on-disk structures
-(for BatchFS and DeltaFS this is the SSTables used in TableFS) can be
-recovered. In other words, the clients have state that cannot be recovered if
-the node stays failed and any progress will be lost. This scenario is a
-disaster for checkpoint-restart where missed cycles may cause the checkpoint to
-bleed over into computation time.
-
-For (3), decoupled namespace approaches sacrifice POSIX going as far as
-requiring the application to link against the systems they want to talk to. In
-today's world of software defined caching, this can be a problem for large data
-centers with many types and tiers of storage. Despite well-known performance
-problems POSIX and REST are the dominant APIs for data transfer.
-
-Decoupling the namespace delays metadata consistency and sacrifices fault
-tolerance. As shown in Table~\ref{table:namespaces}, metadata consistency is
-provided by capabilities and fault tolerance is addressed with a journal.
-
-\section{POSIX Overheads}
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=75mm]{figures/creates-motivation.png}
-\caption{Create-heavy workloads (untar) incurr the highest disk, network, and CPU
-utilization because the metadata server is managing consistency and fault
-tolerance.}\label{fig:creates-motivation}
-\end{figure}
-
-In our examiniation of the overheads of POSIX we benchmark and analyze CephFS,
-the file system that uses the RADOS object store to store its data and
-metadata. We choose CephFS because it is an open-source production quality
-system. CephFS made one set of design decisions and we not asserting that the
-design decisions that were made are superior but instead highlight the
-effect those decisions have on performance.
-
-To show how CephFS behaves under high metadata load we use a create-heavy
-workload because these types of workloads incurr high CPU, network, and disk
-usage.  Figure~\ref{fig:creates-motivation} shows a compilation of the Linux
-kernel, which has a download, untar, compile, and remove phase. The untar phase
-is characterized by many creates and has the highest resource usage, indicating
-that it is stressing the consistency and journalling subsystems of the metadata
-server. Also of note: a create-heavy workload does not help for caching indoes.
-
-\subsection{Fault Tolerance}
-\begin{figure*}[tb] \centering
-\includegraphics[width=1\textwidth]{./figures/journal.png} 
-\caption{CephFS uses a journal to stage updates and tracks dirty metadata in
-the collective memory of the MDSs. Each MDS maintains its own journal, which is
-broken up into 4MB segments. These segments are pushed into RADOS and deleted
-when that particular segment is trimmed from the end of the log. In addition to
-journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
-\end{figure*}
-
-\begin{figure*}[tb]%h
-\centering
-\includegraphics[width=180mm]{figures/throughput-journal.png}
-\caption{Performance improves with larger journal segments because the metadata
-server spends less time flushing the journal }\label{fig:throughput-journal}
-\end{figure*}
-
-%\begin{figure}[tb]%h
-%\centering
-%\includegraphics[width=80mm]{figures/throughput-journal-old.png}
-%\caption{(a), (b), and (c) show increased activity halfway though the run; (d)
-%shows that this activity is related to disk IO, indicating that the activity is
-%the journal.  Despite the activity, we conclude that RADOS itself is not the
-%bottleneck because our RADOS cluster can sustain 100MB/s write bandwidth.
-%}\label{fig:throughput-journal-old}
-%\end{figure}
-%Figure~\ref{fig:throughput-cache} shows that journaling metadata updates into
-%the object store has an overhead.  (a) shows that throughput degrades halfway
-%through the run; (b) shows that cpu utilization spikes at this same point; (c)
-%shows that outgoing network utilization spikes; and (d) shows that this
-%activity spikes on the disks in RADOS.
-
-% purpose of the journal
-As shown in Figure~\ref{fig:journal} CephFS uses RADOS (1) as a metadata store
-for all information about files including the hierarchical namespace and (2) as
-a staging area for the journal of updates before they are applied to the
-metadata store. (1) is essentially a cache that improves performance by serving
-metadata from memory and (2) is the mechanisms for achieving fault tolerance.
-
-% - sequential IO, trim redundant operations
-Fault tolerance means that the client or server can fail and metadata will not
-be lost.  CephFS addresses fault tolerance with a metadata journal that streams
-into the resilient object store. Similar to LFS~\cite{} and WAFL~\cite{} the
-metadata journal can grow to large sizes ensuring (1) sequential writes into
-RADOS and (2) the ability for daemons to trim redundant or irrelevent journal
-entries.
-
-% Knobs in CephFS
-The journal is striped over objects where multiple journal updates can reside
-on the same object. There are two tunables for controlling the journal: the
-segment size and the number of parallel segments that can be written in
-parallel. The former is memory bound as larger segments take up more memory but
-can reduce the time spent journaling and the latter is CPU bound.
-
-% Effects on performance
-Figure~\ref{fig:throughput-journal} shows that journaling metadata updates into
-the object store has an overhead. Part (a) shows the runtime for different
-journal segment sizes; the larger the segment size the bigger that the writes
-into the object store are. The trade-off comes is in terms of memory because
-larger segment sizes take up more space with their buffers. Parts (b), (c), and
-(d) show the throughput over time for different segment sizes. Performance
-suffers when time is spent journaling.
-
-% Conclusion
-Despite this overhead, we posit that the journal is sufficient to slow down
-metadata throughput but not so much as to overwhelm RADOS because we measured
-our peak bandwidth to be 100MB/s, which is the speed of our network link.
-
-\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS, as far as
-we can tell, when a client or server fails there is no recovery scheme. For
-BatchFS, if a client fails when it is writing to the local log-structured
-merged tree (implemented as an SSTable) then those batched metadata operations
-on lost. For DeltaFS, if the client fails then on restart the computation does
-the work again -- since the snapshots of the namespace are never globally
-consistent, there is no ground truth the requires the failed namespace to
-answer to anyone. On the server side, BatchFS and DeltaFS use IndexFS. Again,
-IndexFS writes metadata to SSTables but it is not clear whether they ever
-vacate memory, get written to disk, or are flushed to the object store.
-
-\subsection{Strong Consistency}
-
-\begin{figure*}[tb]%h
-\centering
-\includegraphics[width=180mm]{figures/throughput-cache-journal.png}
-\caption{Journalling metadata updates has a bigger overhead than maintaining
-the inode cache. For create-heavy workloads the inode cache offers no
-performance benefits. }\label{fig:throughput-cache-journal}
-\end{figure*}
-
-\begin{figure*}[tb]%h
-\centering
-\includegraphics[width=180mm]{figures/inode-cache.png}
-\caption{The inode cache improves metadata read performance but for our
-create-heavy workload it is only an overhead. Most of the time maintaining the
-cache is spent evicting and adding inodes.}\label{fig:inode-cache}
-\end{figure*}
-
-Access to POSIX metadata is strongly consistent, so reads and writes are
-globally ordered. The synchronization and serialization machinery needed to
-ensure that all clients see the same state has high overhead.  CephFS uses
-capabilities to keep metadata strongly consistent. To reduce the number of RPCs
-needed for consistency, clients can obtain capabilities for reading, reading
-and updating, reads caching, writing, buffering writes, changing the file size,
-and performing lazy IO.
-
-\begin{figure*}[tb]%h
-\centering
-\includegraphics[width=180mm]{figures/throughput-droplease.png}
-\caption{When a client create stream is ``isolated" then lookups resolve
-locally but when a second client ``interferes" by creating in the same directory,
-the directory inode capability is revoked forcing all clients to centralize
-lookups at the metadata server.  \label{fig:throughput-droplease}}
-\end{figure*}
-
-% purpose of the inode ccahe - reduces RPCs (lookups for create, readdirs for
-% stats)
-To keep track of the read caching and write buffering capabilities, the clients
-and metadata servers agree on the state of each inode using an inode cache.  If
-a client has the directory inode cached it can do metadata writes (e.g.,
-create) with a single RPC. If the client is not caching the directory inode
-then it must do multiple RPCs to the metadata server to (1) determine if the
-file exists and (2) do the actual create.  Unless the client immediately reads
-all the inodes in the cache, the inode cache is less useful for create-heavy
-workloads because the cached inodes are unused.
-
-% benefits
-The benefits of caching the directory inode when creating files is shown in
-Figure~\ref{fig:throughput-droplease}(a).  If only one client is creating files
-in a directory (``isolated" curve) then that client can lookup the existence of
-new files locally before issuing a create request to the metadata server. If
-another client starts creating files in the same directory (``interfere" curve)
-then the directory inode transitions out of read caching and the first client
-must send lookups to the metadata server. When other clients interfere the
-request throughput is higher Figure~\ref{fig:throughput-droplease}(b) but the
-runtime is slower because the isolated client scenario incurrs less requests.
-Figures~\ref{fig:throughput-droplease}(c) and (d) plot the creates and lookups,
-repsectively, over time; creates slow down and lookups dominate the request
-load when the directory inode is shared in the interferring client scenario.
-
-% TODO: what is the cost of trimming the cache?
-% TODO: does CephFS still cache inodes when I turn off caching? Why is still keeping inodes in memory? Gah!
-%\begin{figure}[tb]%h
-%\centering
-%\includegraphics[width=80mm]{figures/throughput-cache.png}
-%\caption{ CephFS tries to maintain 100 thousand inodes in the cache;
-%performance degrades when it cannot evict inodes fast enough to maintain this
-%threshold. \label{fig:throughput-cache}}
-%\end{figure}
-
-% disadvantages
-The drawbacks of the inode cache are shown in
-Figure~\ref{fig:throughput-cache}.  Figure~\ref{fig:throughput-cache}(a) shows
-the throughput in metadata requests per second for a single client creating 200
-thousand files. Until time 950 the throughput is steady at just under 2000
-ops/sec and the CPU utilization is at about 30\%. At time 950 seconds
-throughput degrades which corresponds to more inodes in the cache
-(Figure~\ref{fig:throughput-cache}(c)). The values in
-Figure~\ref{fig:throughput-cache}(c) are:
-
-\begin{itemize}
-  \item inodes: total number of elemnts in the cache
-  \item inodes pinned: 
-  \item inodes pinned tail
-  \item inodes with caps
-  \item inodes top
-  \item inodes bottom 
-\end{itemize}
-
-% how the inode cache works
-CephFS tries to keep the cache at 100 thousand inodes so the degradation in
-performance indicates that the metadata server cannot keep up with the
-workload. Figure~\ref{fig:throughput-cache}(d) shows inodes getting added at
-the same rate as before 950 seconds (ino\(+\)) but the rate at which inodes get
-removed is less stable (ino\(-\)) indicating that cache eviction is taking more
-of the metadata servers time.
-
-\subsection{Consistency Overhead}
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=180mm]{figures/exp0-underloaded.png}
-\caption{An underloaded metadata server adaquately services conflicting
-clients; the create speeds are similar while the interferring operations are
-limited by the cost of RPCs.}\label{fig:exp0-underloaded}
-\end{figure*}
-
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=180mm]{figures/runtime-consistency-scale.png}
-\caption{Scaling clients shows increased variability when another client
-interferes; zooming in on runs with 7 clients we see that different types of
-interferring operations have different effects on performance variability and
-predictability.  }\label{fig:runtime-consistency-scale}
-\end{figure*}
-
-Figure~\ref{fig:exp0-underloaded} is a baseline showing that the metadata
-server can service multiple clients when it is underloaded. One client creates
-files in the same directory and another does a touch or stat 15 seconds into
-the run. Figure~\ref{fig:exp0-underloaded}(a) shows the runtime of the client
-doing the creates: "isolated" is when the create client is the only workload in
-the cluster, "interfere stat" is the runtime of the create client when another
-client does a stat, and "interfere touch" is the runtime of the create client
-when another client does a touch. There is only a minor performance
-degradation. Figure~\ref{fig:exp0-underloaded}(b) compares the baseline
-the other client doing the touch or stat
-%
-%and Figure~\ref{fig:exp0-underloaded}
-%has a throughput trace to show behavior.
-%
-%In Figure~\ref{fig:exp0-underloaded}(a), 
-%creates only see a minor performance
-%degradation while the interferring operations in
-%Figure~\ref{fig:exp0-underloaded}(b) show a larger (yet not significant)
-%discrepency. 
-%
-%
-%
-% isolated is when there is no interferrence from client 1.
-To explain Figure~\ref{fig:exp0-underloaded}(a) we use
-Figure~\ref{fig:exp0-behaviour}(b) to show why the runtime of creating 100
-thousand files is unaffected. Compared to the throughput without an
-interferring stat ("isolated" curve) the throughput of the metadata server with
-an interferring stat ("inteferring stat" curve) is higher. This means the
-metadata server is doing more operations suggesting that it can adequately
-handle the demands of both workloads.
-
-To explain Figure~\ref{fig:exp0-underloaded}(b) we use
-Figure~\ref{fig:exp0-underloaded}(a) to show that operations are bounded by the
-number of files; more files incurr more requests since both operations lookup
-every file. The number of files shown in Figure~\ref{fig:exp0-behaviour}(a)
-indicates that the local stat ("client 0" bar) is interacting with less files
-than the remote stat ("client 1" bar). This graph also explains why the
-isolated stats and touches take the longest from the remote clients; out of all
-setups, the isolated operations are doing the most RPCs.  The difference
-between local ("client 0 interfere" bar) and remote ("client 1 interfere" bar)
-in Figure~\ref{fig:exp0-underloaded}(b) is the overhead of doing RPCs.
-Isolated is the slowest because it requests the most files. On average the
-local interferring client ("client 0" bar in
-Figure~\ref{fig:exp0-behaviour}(a)) requests 40 thousand less files than the
-remote interferring call ("client 1" bar in
-Figure~\ref{fig:exp0-behaviour}(a)). This reflects how fast the local client
-can get into the queue of requests, presumably because it bypasses capability
-checks.
-
-From this experiment we make 2 conclusions:
-\begin{itemize}
-  \item the metadata server is not overloaded because it can handle both workloads
-  \item the metadata server needs to get global consistency from only 1 client
-\end{itemize}
-
-\textbf{Comparison to decoupled namespaces}: Decoupled namespaces merge batches
-of metadata operations into the global namespaces when the job completes.  In
-BatchFS the merge is delayed by the application using an API to switch between
-asynchronous to sychronous mode. The merge itself is explicitly managed by the
-application but future work looks at more automated methdologies. In DeltaFS
-snapshots of the metadata subtrees stays on the client machines; there is no
-ground truth and consistent namespaces are constructed and resolved at
-application read time or when a 3rd party system (e.g., middleware, scheduler,
-etc.) needs a view of the metadata.
-
-\section{Methodology}
-
-\subsection{Decoupled Namespaces in CephFS}
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=90mm]{figures/fig-decouple.png}
-\caption{Applications that can tolerate weaker consistency can decouple the
-namespace, write updates to a local file, and delay metadata updates. The merge
-and replay steps are optional if there is no need for a global namespace.
-}\label{fig:decouple}
-\end{figure}
-
-%- How it works (timestamps)
-Recall that CephFS uses RADOS (1) as a metadata store for all information about
-files including the hierarchical namespace and (2) as a staging area for the
-journal of updates before they are applied to the metadata store.  The journal
-tool is used for disaster recovery and lets administrators view and modify the
-journal of metadata updates; it can read the journal, erase events from the
-journal, and apply the updates in the journal to the metadata store.  To apply
-journal updates to the metadata store, the journal tool reads the journal
-segments from RADOS objects and applies the update to the metadata store (which
-are also stored as RADOS objects).
-
-% General information about the journal
-The journal segments are saved as objects in RADOS.  The journal has 4
-pointers, described in `osdc/Journaler.h`:
-
-\begin{itemize}
-  \item write position: tail of the journal; points to the current session where we are appending events
-  \item unused field: where someone is reading
-  \item expire position: old journal segments
-  \item trimmed position: where daemon is expiring old items
-\end{itemize}
-
-% How the journal tool works
-Journal segments in RADDS have a header followed by serialized log events. The
-log events are read by hopping over objects using the read offset and object
-size pulled from the journal header.  After decoding them, we can examine the
-metadata (1) about the event (e.g., type, timestamps, etc.) and (2) for inodes
-that the event touches.
-
-% The metablobs
-The metadata for inodes that the event touches are called metadata blobs and
-the ones associated with events are \textbf{unordered}; this layout makes
-writing journal updates fast but the cost is realized when reading the metadata
-blobs.  It makes sense to optimize for writing since reading only occurs on
-failures. To reconstruct the namespace for the metadata blob, the journal tool
-iterates over each metadata blob in the events and builds mappings of inodes to
-directory entries (for directories) and parent inodes to child inodes/directory
-entries.
-
-% Journal import
-The journal tool imports journals from binary files stored on disk.  First the
-header of the dump is sanity checked and written into RADOS to the ``header"
-object.  The ``header" object has metadata about the journal as well as the
-locations of all the journal pointers (e.g., where the tail of the journal is,
-where we are currently trimming, etc.).  Then the journal events are cleaned
-(erasing trailing events that are not part of the header) and written as
-objects into RADOS.  Note that while the journal is in RADOS, the metadata
-servers do do not have the namespace reconstructed in memory so the metadata
-cluster will not service requests relating to the journal of imported events.
-To construct the namespace in the collective memory of the metadata servers we
-need to first construct the namespace in RADOS. The journal tool can explicitly
-do this by  applying the journal to the metadata store in RAODS. This will pull
-the objects containing journal segments and replay them on the metadata store.
-Finally, we delete the journal in RADOS and restart the metadata servers so
-they rebuild their caches.
-
-% Journal export
-The journal tool exports journals to binary files stored on disk. First the
-journal is scanned for the header and then journal is recovered. To recover the
-journal the ``header" object is read off disk and then objects are probed in
-order and starting from the write position saved in the header. Probing will
-update the write position if it finds objects with data in them. There are two
-types of objects:
-
-1. journal objects (e.g., 200.000*)
-
-2. metadata store objects (e.g., 1.inode)
-
-We need to update the journal objects with the new events! How do we write out
-new objects to RADOS (e.g., 200.000*)? These should be the corect sizes (e.g.,
-4980 bytes) -- these log events that we are adding are not making them into
-the journal objects... do we have to explicitly read them and then re-write
-them?
-
-TODO: pull the object, read it, append to it, make sure size changes. It looks
-like we can only read these objects and apply doesn't actually touch them... it
-only touches the metadata store objects (e.g., 1.inode*).
-
-% TODO: figure out how import saves the journal on the metadata server
-% - they aren't -- everything is stored in RADOS
-% - we never touch the metadata servers!!!
-% TODO: figure out how to read the header object
-% TODO: figure out the mapping objects to journal segments (not 1x1)
-
-% Journal export
-When exporting a journal of events, the journal tool first scans the journal to
-check for corruption. Then it recovers the journal by reading the ``header"
-object out of RADOS.  After reading the header, the journal tool can pull
-journal segments from RADOS because it knows how many objects to pull and how
-far to seek within those objects.
-
-%% The data structures
-%The metadata that the event touches, including inodes, paths, and timestamps,
-%are stored as metablobs. Each piece of metadata inside a metablob is called a
-%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
-%table client transactions (tid, version), renamed directory fragments (maps,
-%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
-%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
-%me (and you since you are reading this paper), this does not make any sense.
-%
-%Each directory fragment has an associated directory lump, which is just a bunch
-%of metadata. The most interesting part of the dirlump is the fullbits array,
-%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
-%and then all the full bits, saving off the children and inode locations. The
-%children tell us which dentry names an inode has and the inode locations map
-%the parent inode to its child inode and dentry.
-
-\subsubsection{Leveraging the Journal Tool}
-
-%- How we leverage it
-We leverage the journal tool's ability to reconstruct the namespace for the
-metadata blob. First we scan journal and store all events in a map. Next we
-create a new event and copy the root inode directory lump from another journal
-event into the event's metadata blob. Finally, we add a new directory lump for
-our new file or directory.
-
-If decoupled namespaces are turned on, the client writes metadata updates
-locally and merges the updates with the journal tool. The client that decouples
-the namespace operates without any consistency and any conflicts at the merge
-are resolved in favor of this client. Updates by other clients ({\it i.e.}
-metadata writes to the global namespace) are overwritten. Cudele adopts the
-following process when an application decouples the namespace:
-
-\begin{enumerate}
-  \item metadata server exports a snapshot of the file system to a binary file
-  \item client materializes the snapshot in memory
-  \item client operates on the data structures in memory
-  \item client exports the new snapshot to a binary file
-  \item metadata server merges the snapshot into the global namespace
-\end{enumerate}
-
-Step 3 is the most complicated and requires understanding how the snapshot is
-materialized in memory.
-
-\subsubsection{Operating on Snapshots}
-
-Our first implementation attempted to re-create journal events using the same
-libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
-to instantiate a Ceph inode and directory entry for the current file/dir and
-its parent.  This is too hard because there are too many moving parts in the
-metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
-traverse up and down namespace, etc.). So when I tried add dentries and inodes
-it was trying to traverse up/down and it would almost always segfault when it
-was looking for something. These metablobs are supposed to be self container --
-the problem is I do not know what is supposed to go *inside* them.
-
-Our second idea was to copy the metadata blog and change just what we needed.
-For example, we would save a binary dump of a generic \texttt{mkdir} event on
-disk. When the application makes a directory, this dump would be loaded and the
-fields would be changed before being written back to disk. Rather than
-traversing up and down a namespace in memory of a metadata server, we should
-traverse up and down the namespace *inside* the metadata blob. This
-implementation requires disk IO and editing the log event is non-trivial for
-two reasons:
-
-\begin{itemize}
-
-  \item methods do not edit events; they just write them
-
-  \item the metadata that the event touches (e.g., the metablob) is unorganized
-  on disk for performance -- it is trade-off for writing data faster serially and
-  reconstructing information slowly since failure is not considered the norm
-
-\end{itemize}
-
-Faced with these challenges we landed on our final implementation: load the
-snapshot into the the data structures used to examine and replay journals, edit
-those data structures, and write them out to disk as binary.
-
-\section{Results}
-
-In this section, we quantify the costs of consistency and fault tolerance in
-CephFS. If the components that ensure these semantics (i.e. capabilities and
-journals, respectively) can be mitigated or delayed, then global namespaces
-perform as well as decoupled namespaces.
-
-\textbf{Experimental Setup}: we run our experiments on a 9 OSD, 3 MDS, 1 MON
-Ceph cluster. The clients use the Ceph kernel client, which has been in the
-mainline Linux kernel since TODO. We use the kernel client so that we can find
-the true create speed of the server; our experiments show a low CPU utilziation
-for the clients which indicates that we are stressing the servers more. We also
-turn caching off becuase, as shown in figureX there is little difference, in
-terms of performance between caching and not caching when using the kernel
-client.
-
-%\section{notes}
-%Linking clients into our custom libcephfs
-%
-%Use namespace's recursive data structure to put policies on subtrees
-%- consistency: eventual vs. strong, global vs. local
-%  - e.g., BatchFS/DeltaFS: eventual, local
-%  - e.g., POSIX: strong, global
-%  - e.g., PLFS: no consistency
-%- fault tolerance: global vs. local
-%  - e.g., CephFS: global
-%  - e.g., BatchFS/DeltaFS: local
-%
-%Experimental Setup
-%- Ceph: 9 OSDs, 1 MDS, 2 kernel client
-%- Workload limitations: blah
-%
-%Workload: creates
-%
-%Baseline: 200K creates in the same directory
-%- throughput: degrades at 950s
-%- CPU utilization: more at 950s
-%- inode cache: eviction dominate
-%- inodes +- to cache: eviction dominate
-%- per-disk throughput: RADOS not bottleneck
-%
-%Experiment 1: Interference
-%
-%\subsection{Baseline}
-%Experiment 0: creates in the same directory
-%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
-%
-%Experiment 0: creates with a stat
-%- Hypothesis: metadata read pauses creates and requires a snapshot in time
-%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
-%- can we delay snapshot?
-%
-%Experiment 1: creates with a readdir
-%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
-%- client0: create 100k, client1: stat at 2 mins
-%
-%Experiment 2: scale the number of files
-%- See if the open/close spike occurs 
-%- Try to see why open/close spike is allowed to happen
-%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
-%- client0: create 100k, client1: touch at 2 mins
-%
-%Experiment 3: see how fast the cache satisfies a read
-%- client0: create 100k, stat inodes
-%- client0: create 100k, client1: stat inodes
-%
-%lient 0: creates, client 1 create(s)
-
-\subsection{Journaling Overhead}
-
-Journal to RADOS
-Turn off journaling (large segment)
-Journal to in-memory OSD
-
-\subsection{Macrobenchmarks}
-updatedb: http://lists.ceph.com/pipermail/ceph-users-ceph.com/2015-July/002768.html
-
-\bibliography{paper}
-\bibliographystyle{plain}
 \end{document}

### 2017-03-08 paper/sections/1_introduction.tex
paper: switch template, refactor sections, add pretty code
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Introduction}

Today's client-server based file system metadata services have scalability
problems. It takes a lot of resources to server POSIX metadata requests and
applications perform better with dedicated metadata
servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs}. This is fine for small
workloads and file systems but as the system scales provisioning a metadata
server for every client is expensive and complicated.

Current hardware evolution and the rise of software-defined storage storage,
which uses techniqes like erasure coding, replication, and partitioning, have
ushered a new era of HPC computing; architectures are transitioning from
complex storage stacks with burst buffer, file system, object store, and tape
tiers to a two layer stack with just a burst buffer and object
store~\cite{bent:login16-hpc-trends}. This trend exacerbates the metadata
scalability problem and has given rise to the serverless metadata services.
%\begin{listing}
\begin{minted}{python}
def hello():
\end{minted}
%\caption{JSON example} 
%\label{json-example}
%\end{listing}

HPC workloads are so metadata intensive that new management techniques are
advocating reducing synchronization and serialization overheads by transferring
these responsibilities to the client. We call this approach decoupling the
namespace and the semantics of consistency differ amongst systems.

\begin{figure}[tb]
\centering
\includegraphics[width=75mm]{figures/subtree-policies.png}
\caption{Administrators can assign consistency and fault tolerance policies to
subtrees to get the benefits of some of the state-of-the-art HPC architectures.
}\label{fig:subtree-policies}
\end{figure}

We propose subtree policies, an interface that lets future programmers control
how the system manages different parts of the namespace.  For performance one
subtree can adopt weaker consistency semantics while another subtree can retain
the rigidity of POSIX's strong consistency. Figure~\ref{fig:subtree-policies}
shows an example setup where a single global namespace has directories for
applications designed for different, state-of-the-art HPC architectures.  Our
system supports 3 forms of consistency and 2 forms of fault tolerance giving
the administrator a wide range of policies and optimizations depending on the
application's needs.

### 2017-03-08 paper/sections/4_methodology.tex
paper: switch template, refactor sections, add pretty code
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Methodology: Decoupled Namespaces}

\begin{figure}[tb]
\caption{Applications can decouple the namespace, write updates to a local
journal, and delay metadata updates.  Table~\ref{table:spectrum} shows how
these phases (represented by the arrows) can be combined to provide weaker
consistency or fault tolerance semantics.  }\label{fig:decouple}
\centering
\includegraphics[width=90mm]{figures/fig-decouple.png}
\end{figure}

In this section we describe Cudele, our prototype system that lets future
programmers compose mechanisms (Section~\S\ref{sec:cudeles-mechanisms}) to
provide the necessary guarantees
(Section~\S\ref{sec:setting-policies-with-cudele}) for their application.

\subsection{Cudele's Mechanisms}
\label{sec:cudeles-mechanisms}

% describe the figure
Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
which entity they are performed by (gray boxes). The metadata store and journal
are different ways of representing the namespace. The metadata store represents
the namespace as a tree of directory fragments and is easier to read/traverse.
On the other hand, the journal represent the namespace as a list of events. It
a ``pile system"; writes are fast but reads are slow because state must be
reconstructed.  Specifically, reads are slow because there is more data to
read, it is unorganized, and many of the updates may be redundant.

% describe the mechanisms
Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Save, and
Persist. ``RPCs" does round trip remote procedure calls to establish
consistency; it is the default implementation for complying with POSIX in
CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
into the object store. ``Create" allows clients to append metadata events to an
in-memory journal. ``Volatile apply" takes the in-memory journal on the client
and applies it directly to the in-memory metadata store of the metadata server
cluster. ``Save" takes the in-memory journal and writes it to the client's
disk. ``Persist" saves the journal as a an object in the object store from the
client.

Next, we discuss how these mechanisms can be composed to get different
consistency and fault tolerance semantics.

%\begin{tabular}{ r | l }
%  \(\Rightarrow\)   & Description \\\hline
%  create            & events appended to in-memory journal \\
%  v\_apply           & journal volatily applied to in-memory metadata store \\
%  save              & journal saved to client's disk \\
%  persist           & journal saved in object store \\
%  replay            & metadata servers materialize namespace \\
%  RPCs              & round trip remote procedure calls \\
%  stream            & metadata server streams journal into RADOS \\
%\end{tabular}

\subsection{Setting Policies with Cudele}
\label{sec:setting-policies-with-cudele}

\begin{table}[t]
\begin{center}
\caption{Future programmers can explore the consistency (C) and
durability (D) spectrums by composing Cudele mechanisms. The consistency
and durability properties are not guaranteed until all mechanisms in the cell
are complete (i.e. the compositions should be considered atomic) and there are
no guarantees while tranisitioning between policies. \label{table:spectrum}}
\begin{tabular}{ l | l | l | l }
  C \(\rightarrow\) &&& \\  
  D \(\downarrow\)  & none              & eventual        & strong  \\\hline
  none                       & create            & create          & RPCs    \\
                             &                   & +volatile apply &         \\\hdashline
  local                      & create            & create          & RPCs    \\
                             & +local persist    & +local persist  & +local persist \\
                             &                   & +volatile apply &         \\\hdashline
  global                     & create            & create          & RPCs    \\
                             & +global persist   & +global persist & +stream \\
                             &                   & +volatile apply &         \\
\end{tabular}
\end{center}
\end{table}

% describe table
The spectrum of consistency and fault tolerance guarantees that adminstrators
can construct is shown in Table~\ref{table:spectrum}. The columns are the
different consistency semantics and the rows cover the spectrum of fault
tolerance guantees. For consistency: ``none" means the system does not handle
merging updates into a global namespace and it is assumed that middleware or
the application manages consistency lazily; ``eventual" merges updates either
when the system has time (e.g., by a background daemon) or when the client is
done writing; and updates in ``global" consistency are seen immediately by all
clients. For fault tolerance, ``none" means that updates are volatile and will
be lost on a crash of any component. Stronger guarantees are made with
``local", which means updates will be retained if the client node recovers, and
``global", where all updates are always recoverable.

% which system they represent and which are impossible
The cells in Table~\ref{table:spectrum} encompass many exisiting storage
systems. POSIX systems like CephFS and IndexFS have global consistency and
fault tolerance; DeltaFS has consistency and fault tolerance set to ``none";
and BatchFS uses ``eventual" consistency and ``local" fault tolerance. These
are just a few of the HPC examples.

% how its done
To compose the mechanisms administrators inject which steps (described in
Section~\S\ref{sec:cudeles-mechanisms}) to run and which to use in parallel
using a domain specific language. For example, to get the semantics of BatchFS,
the administrator would inject the following pipeline:\\

\noindent \texttt{create+save+volatile apply}\\

% what is impossible
Although we can achieve all permutations of the different guarantees in
Table~\ref{table:spectrum}, not all of them make much sense. For example, it
makes little sense to do \texttt{creates+RPCs} since both steps do the same
thing or \texttt{stream+save} since global fault tolerance is stronger and has
more overhead than local fault tolerance.  Formulaically, valid steps
include:\\

\noindent \texttt{<create|RPCs> [+v\_apply][+save][+persist][+stream]}

\subsection{Cudele Namespace API: Per-Subtree Policies}

% file type
To assign consistency and fault tolerance to the subtrees we store policies in
the directory inode. This approach uses the File Type interface from the
Malacology programmable store system~\cite{} and it tells clients how to access
the underlying data or metadata. The underlying implementation stores
executable code in the inode that calls the different Cudele mechanisms. Of
course, there are many security and access control aspects of this approach but
that is beyond the scope of this paper.

% the interface
The interface for setting the subtree policies is with \{path,
<block|overwrite>, pre-allocated inodes\} tuples. For example:

\texttt{(msevilla/mydir, policies.yml)}

would decouple the path \texttt{msevilla/mydir} and would apply the policies in
\texttt{policies.txt}. The policies file supports the following values:

\begin{itemize}

  \item \texttt{allocated\_inodes}: the number of inodes to allocate to the
  decoupled namespace (default 100)

  \item \texttt{interfere\_callback}: how to handle a request from another
  client targeted at the now decoupled subtree (default \texttt{block})

  \item \texttt{consistency\_callback}: which consistency model to use (default
  \texttt{RPCs})

  \item \texttt{durability\_callback}: which durability model to use (default
  \texttt{stream})

\end{itemize}

Given these default values decoupling the namespace with an empty policies file
would give the application 100 inodes but the subtree would behave like the
existing CephFS implementation. Table~\ref{} shows how one would use the Cudele
API to implement policies from related work.

%

For \texttt{block}, any requests to this part of the namespace return with
``Device is busy", which will spare the MDS from wasting resources for updates
that may get overwritten. If the application does not mind losing updates, for
example it wants approximations for results that take too long to compute, it
can select \texttt{overwrite}. In this case, metadata will be written and the
computation from the decoupled namespace will take priority because the resutls
are more accurate.

% // limit consistency
% if metadata write (create) -- try to reduce error bars for many creates and runtime of one touch)
%   if !RPCs
%     if overwrite
%       do nothing and let the RPC go
%     else
%       return -EBUSY
%
% // limit fault tolerance
% if !stream
%   return immediately on mdlog

### 2017-03-08 paper/sections/0_abstract.tex
paper: switch template, refactor sections, add pretty code
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{abstract}
Today's large and highly-parallel workloads are bottlenecked by metadata
services because many processes end up accessing the same shared resoure. In
HPC, state-of-the-art file systems are abandoning POSIX because the
sychronization and serializiation overheads are too costly -- and sometimes
even unnceccessary -- for some of their applications.  While the performance
benefits are plain for these users, other applications that rely on stronger
consistency must be re-written or deployed on a different system.  We present
Cudele, a programmable file system that supports different degrees of
consistency and fault tolerance within the same namespace.  First, we take a
POSIX compliant file system and relax the consistency constraints by
implementing two metadata designs: delayed metadata update merge into the
global namespace and client local views of metadata updates.  Second, we store
the consistency and fault tolerance semantics in inodes so that subtrees within
the same namespace can be optimized for different workloads; the inodes are
programmable so that clients understand how to access the metadata in a
subtree.  Third, we present a theoretical framework for a metadata service and
a working implementation that provides the lowest common denominator needed to
implement a wide range of consistency/fault tolerance semantics that can be
benchmarked all on the same system.
\end{abstract}

### 2017-03-08 paper/sections/2_related_work.tex
paper: switch template, refactor sections, add pretty code
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Related Work}

% General
The bottlenecks associated with accessing POSIX file system metadata are not limited
to HPC workloads and the same challenges that plagued these systems for years are
finding their way into the cloud. Workloads that deal with many small files
({\it e.g.}, log processing and database
queries~\cite{thusoo:sigmod2010-facebook-infrastructure}) and large numbers of
simultaneous clients ({\it e.g.}, MapReduce
jobs~\cite{mckusick:acm2010-gfs-evolution}), are subject to the scalability of
the metadata service. The biggest challenge is that whenever a file
is touched the client must access the file's metadata and maintaining a file
system namespace imposes small, frequent accesses on the underlying storage
system~\cite{roselli:atec2000-FS-workloads}.  Unfortunately, scaling file
system metadata is a well-known problem and solutions for scaling data IO do
not work for metadata IO~\cite{roselli:atec2000-FS-workloads,
abad:techreport2012-fstrace, abad:ucc2012-mimesis,
alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. There are two approaches
for improving the performance of metadata access.

\subsection{Metadata Load Balancing}

% approaches to load balancing (strong = more resuorces per unit work, weak =
% fixed resource per work unit)
One approach for improving metadata performance and scalability is to alleviate
overloaded servers by load balancing metadata IO across a cluster. Common
techniques include partitioning metadata when there are many writes and
replicating metadata when there are many reads. For example, IndexFS partitions
directories and clients write to different partitions by grabbing leases and
caching ancestor metadata for path traversal; it does well for strong scaling
because servers can keep more inodes in the cache which results in less RPCs.
Alternatively, ShardFS replicates directory state so servers do not need to
contact peers for path traversal; it does well for read workloads because all
file operations only require 1 RPC and for weak scaling because requests will
never incurr extra RPCs due to a full cache.  CephFS employs both techniques to
a lesser extent; directories can be replicated or sharded but the caching and
replication policies do not change depending on the balancing technique.
Despite the performance benefits these techniques add complexity and jeopardize
the robustness and performance characteristics of the metadata service because
the systems now need (1) policies to guide the migration decisions and (2)
mechanisms to address inconsistent states across servers.

% policies
Setting policies for migrations is arguably more difficult than adding the
migration mechanisms themselves.  For example, IndexFS and CephFS use the
GIGA+~\cite{GIGA+} technique for partitioning directories at a predefined
threshold and using lazy synchronization to redirect queries to the server that
``owns" the targeted metadata.  Determining when to partition directories and
when to migrate the directory fragments are policies that vary between systems:
GIGA+ partitions directories when the size reaches a certain number of files
and migrates directory fragments immediately; CephFS partitions directories
when they reach a threshold size or when the write temperature reaches a
certain value and migrates directory fragments when the hosting server has more
load than the other servers in the metadata cluster. Another policy is when and
how to replicate directory state; ShardFS replicates immediately and
pessimistically while CephFS replicates only when the read temperature reaches
a threshold.  There is a wide range of policies and it is difficult to address
with tunables and hard-coded design decisions.

% addressing inconsistency
In addition to the policies, distributing metadata across a cluster requiress
distributed transactions and cache coherence protocols to ensure strong
consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
directory state and uses optimistic concurrency control for conflicts; namely
it does the operation and if there is a conflict at verification time it falls
back to two-phase locking.  Another example is IndexFS's inode cache which
reduces RPCs by caching ancestor paths -- the locality of this cache can be
thrashed by random reads but performs well for metadata writes. For
consistency, writes to directories in IndexFS block until the lease expires
while writes to direcotries in ShardFS are slow for everyone as it either
requires serialization or locking with many servers; reads in IndexFS are
subject to cache locality while reads in ShardFS always resolve to 1 RPC.
Another example of the overheads of addressing inconsistency is how CephFS
maintains client sessions and inode caches for capabilities (which in turn make
metadata access faster). When metadata is exchanged between metadata servers
these sessions/caches must be flushed and new statistics exchanged with a
scatter-gather process; this halts updates on the directories and blocks until
the authoratitive metadata server responds.  These protocols are discusssed in
more detail in the next section but their inclusion here is a testament to the
complexity of migrating metadata.

%For metadata writes it does a distributed transaction; monotonic writes with
%concurrent clients fail and do pessimistic locking through a mediated lock
%server to ensure strong consistency, non-monotonic writes grab locks at every
%server. Zooming in on monotonic writes: if permissions increase it executes on
%a primary then non-primary, if permissions decrease it executes on all
%non-primary then on primary.

The conclusion we have drawn from this related work is that metadata protocols
have a bigger impact on performance and scalability than load balancing.  
Understanding these protocols helps load balancing and gives us a better
understanding of the metrics we should use to make migration decisions ({\it
e.g.}, which operations reflect the state of the system), what types of
requests cause the most load, and how an overloaded system reacts ({\it e.g.},
increasing latencies, lower throughput, etc.).

\subsection{Relaxing POSIX}

% POSIX 
POSIX workloads require strong consistency and many file systems improve
performance by reducing the number of remote calls per operation ({\it i.e.}
RPC amplification). As discussed in the previous section, caching with leases and
replication are popular approaches to reducing the overheads of path traversals
but their performance is subject to cache locality and the amount of available
resources, respectively; for random workloads larger than the cache extra RPCs
hurt performance~\cite{IndexFS, CephFS} and for write heavy workloads with more
resources the RPCs for invalidations are harmful. Another approach to reducing
RPCs is to use leases or capabilities.

%IndexFS aggressively caches paths and
%handles permissions by handing out leases for metadata writes; metadata may
%only be modified when all leases have expired. 
%IndexFS~\cite{ren:sc2014-indexfs} aggressively caches pathnames and their
%permissions on the client servers. Modifications to metadata cached by clients
%is delayed until all client leases have expired. This reduces the RPC
%amplification to 1 when mutatating directory metadata ({\it e.g.,}
%\texttt{mkdir}, \texttt{chmod}, etc.) because clients are not querying MDSs for
%path traversals. The disadvantage of this approach is the high latency of
%mutation operations (reads to filenames).  ShardFS~\cite{xiao:socc2015-shardfs}
%replicates metadata (specifically directory lookup state) across the MDS
%cluster, reducing the RPC amplification to 1 for file operations ({e.g.,}
%\texttt{stat}, \texttt{chmod}, \texttt{chown}, etc.). Modifications to the
%directory lookup state are done with optimistic concurrecny control and fall
%back to retry if verification fails. The disadvantage of this appraoch is the
%high number of RPCs for maintaining directory metadata mutations (writes to
%directories).  CephFS also maintains an inode cache but its notion of leases
%are much shorter, on the order of microseconds.

% Non POSIX
High performance computing has unique requirements for file systems ({e.g.},
fast creates) and well-defined workloads (e.g., workflows) that make relaxing
POSIX sensible.  One popular approach is to allow clients to ``lock'' parts of
the namespace to improve performance and scalability by avoiding
synchronization, false sharing, and serialization.  BatchFS assumes the
application coordinates accesses to the namespace, so the clients can batch
local operations and merge with a global namespace image lazily. Similarily,
DeltaFS eliminates RPC traffic using subtree snapshots for non-conflicting
workloads and middleware for conflicting workloads. MarFS gives administrators
the ability to (1) lock ``project directories" and (2) allocate GPFS clusters
for demanding directory workloads. TwoTiers eliminates high-latencies by
storing metadata in a flash tier; apps lock the namespace so that metadata can
be accessed more quickly.  Unfortunately, decoupling the namespaces has costs:
(1) merging metadata state back into the global namespace is slow; (2) failures
are local to the failing node; and (3) the systems are not backwards
compatible.

% NON POSIX
%Decoupling the namespaces has many advantages, including improved scalability,
%higher resource utilization, and better performance.  BatchFS and DeltaFS are
%near-POSIX filesystems that give clients the ability to decouple subtrees from
%the namespace so that the applications can execute metadata operations without
%synchronization and serialization.  These operations are applied to a local
%snapshot of the file system namespace and conflicts are resolved either by the
%application or by an external service . Applications link into a metadata
%server library to reduce resource utilization and code paths (e.g., no daemons
%and less interprocess communication).

For (1), state-of-the-art systems manage consistency in non-traditional ways:
IndexFS maintains the global namespace but blocks operations from other clients
until the first client drops the lease, BatchFS does operations on a snapshot
of the namespace and merges batches of operations into the global namespace,
and DeltaFS never merges back into the global namespace. The merging for
BatchFS is done by an auxiliary metadata server running on the client and
conflicts are resolved by the application. Although DeltaFS never explicitly
merges, applications needing some degree of ground truth can either manage
consistency thesmelves on a read or add a bolt-on service to manage the
consistency.

For (2), if the client fails and stays down, all metadata operations on the
decoupled namespace are lost. If the client recovers, the on-disk structures
(for BatchFS and DeltaFS this is the SSTables used in TableFS) can be
recovered. In other words, the clients have state that cannot be recovered if
the node stays failed and any progress will be lost. This scenario is a
disaster for checkpoint-restart where missed cycles may cause the checkpoint to
bleed over into computation time.

For (3), decoupled namespace approaches sacrifice POSIX going as far as
requiring the application to link against the systems they want to talk to. In
today's world of software defined caching, this can be a problem for large data
centers with many types and tiers of storage. Despite well-known performance
problems POSIX and REST are the dominant APIs for data transfer.

Decoupling the namespace delays metadata consistency and sacrifices fault
tolerance. As shown in Table~\ref{table:namespaces}, metadata consistency is
provided by capabilities and fault tolerance is addressed with a journal.

\begin{tabular}{ r | l | l | l }
              & Decoupled & Global    & Implied   \\
              & Namespace & Namespace & Namespace \\\hline
  Example     & BatchFS & CephFS & PLFS \\
              & DeltaFS & IndexFS &     \\
  Consistency & eventual & strong & none? \\
  Fault Tol.  & node local & journal & journal \\
\end{tabular}

### 2017-03-08 paper/sections/3_posix_overheads.tex
paper: switch template, refactor sections, add pretty code
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{POSIX Overheads}

\begin{figure}[tb]
\centering
\includegraphics[width=75mm]{figures/creates-motivation.png}
\caption{Create-heavy workloads (untar) incurr the highest disk, network, and CPU
utilization because the metadata server is managing consistency and fault
tolerance.}\label{fig:creates-motivation}
\end{figure}

In our examiniation of the overheads of POSIX we benchmark and analyze CephFS,
the file system that uses the RADOS object store to store its data and
metadata. We choose CephFS because it is an open-source production quality
system. CephFS made one set of design decisions and we not asserting that the
design decisions that were made are superior but instead highlight the
effect those decisions have on performance.

To show how CephFS behaves under high metadata load we use a create-heavy
workload because these types of workloads incurr high CPU, network, and disk
usage.  Figure~\ref{fig:creates-motivation} shows a compilation of the Linux
kernel, which has a download, untar, compile, and remove phase. The untar phase
is characterized by many creates and has the highest resource usage, indicating
that it is stressing the consistency and journalling subsystems of the metadata
server. Also of note: a create-heavy workload does not help for caching indoes.

In this section, we quantify the costs of consistency and fault tolerance in
CephFS. If the components that ensure these semantics (i.e. capabilities and
journals, respectively) can be mitigated or delayed, then global namespaces
perform as well as decoupled namespaces. We run our experiments on a 9 OSD, 3
MDS, 1 MON Ceph cluster. The clients use the Ceph kernel client, which has been
in the mainline Linux kernel since TODO. We use the kernel client so that we
can find the true create speed of the server; our experiments show a low CPU
utilziation for the clients which indicates that we are stressing the servers
more. We also turn caching off becuase, as shown in figureX there is little
difference, in terms of performance between caching and not caching when using
the kernel client.

\subsection{Fault Tolerance}
\label{sec:fault-tolerance}
\begin{figure*}[tb] \centering
\includegraphics[width=1\textwidth]{./figures/journal.png} 
\caption{CephFS uses a journal to stage updates and tracks dirty metadata in
the collective memory of the MDSs. Each MDS maintains its own journal, which is
broken up into 4MB segments. These segments are pushed into RADOS and deleted
when that particular segment is trimmed from the end of the log. In addition to
journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
\end{figure*}

\begin{figure*}[tb]%h
\centering
\includegraphics[width=180mm]{figures/throughput-journal.png}
\caption{Performance improves with larger journal segments because the metadata
server spends less time flushing the journal }\label{fig:throughput-journal}
\end{figure*}

%\begin{figure}[tb]%h
%\centering
%\includegraphics[width=80mm]{figures/throughput-journal-old.png}
%\caption{(a), (b), and (c) show increased activity halfway though the run; (d)
%shows that this activity is related to disk IO, indicating that the activity is
%the journal.  Despite the activity, we conclude that RADOS itself is not the
%bottleneck because our RADOS cluster can sustain 100MB/s write bandwidth.
%}\label{fig:throughput-journal-old}
%\end{figure}
%Figure~\ref{fig:throughput-cache} shows that journaling metadata updates into
%the object store has an overhead.  (a) shows that throughput degrades halfway
%through the run; (b) shows that cpu utilization spikes at this same point; (c)
%shows that outgoing network utilization spikes; and (d) shows that this
%activity spikes on the disks in RADOS.

% purpose of the journal
As shown in Figure~\ref{fig:journal} CephFS uses RADOS (1) as a metadata store
for all information about files including the hierarchical namespace and (2) as
a staging area for the journal of updates before they are applied to the
metadata store. (1) is essentially a cache that improves performance by serving
metadata from memory and (2) is the mechanisms for achieving fault tolerance.

% - sequential IO, trim redundant operations
Fault tolerance means that the client or server can fail and metadata will not
be lost.  CephFS addresses fault tolerance with a metadata journal that streams
into the resilient object store. Similar to LFS~\cite{} and WAFL~\cite{} the
metadata journal can grow to large sizes ensuring (1) sequential writes into
RADOS and (2) the ability for daemons to trim redundant or irrelevent journal
entries.

% Knobs in CephFS
The journal is striped over objects where multiple journal updates can reside
on the same object. There are two tunables for controlling the journal: the
segment size and the number of parallel segments that can be written in
parallel. The former is memory bound as larger segments take up more memory but
can reduce the time spent journaling and the latter is CPU bound.

% Effects on performance
Figure~\ref{fig:throughput-journal} shows that journaling metadata updates into
the object store has an overhead. Part (a) shows the runtime for different
journal segment sizes; the larger the segment size the bigger that the writes
into the object store are. The trade-off comes is in terms of memory because
larger segment sizes take up more space with their buffers. Parts (b), (c), and
(d) show the throughput over time for different segment sizes. Performance
suffers when time is spent journaling.

% Conclusion
Despite this overhead, we posit that the journal is sufficient to slow down
metadata throughput but not so much as to overwhelm RADOS because we measured
our peak bandwidth to be 100MB/s, which is the speed of our network link.

\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS, as far as
we can tell, when a client or server fails there is no recovery scheme. For
BatchFS, if a client fails when it is writing to the local log-structured
merged tree (implemented as an SSTable) then those batched metadata operations
on lost. For DeltaFS, if the client fails then on restart the computation does
the work again -- since the snapshots of the namespace are never globally
consistent, there is no ground truth the requires the failed namespace to
answer to anyone. On the server side, BatchFS and DeltaFS use IndexFS. Again,
IndexFS writes metadata to SSTables but it is not clear whether they ever
vacate memory, get written to disk, or are flushed to the object store.

\subsection{Strong Consistency}

\begin{figure*}[tb]%h
\centering
\includegraphics[width=180mm]{figures/throughput-cache-journal.png}
\caption{Journalling metadata updates has a bigger overhead than maintaining
the inode cache. For create-heavy workloads the inode cache offers no
performance benefits. }\label{fig:throughput-cache-journal}
\end{figure*}

\begin{figure*}[tb]%h
\centering
\includegraphics[width=180mm]{figures/inode-cache.png}
\caption{The inode cache improves metadata read performance but for our
create-heavy workload it is only an overhead. Most of the time maintaining the
cache is spent evicting and adding inodes.}\label{fig:inode-cache}
\end{figure*}

Access to POSIX metadata is strongly consistent, so reads and writes are
globally ordered. The synchronization and serialization machinery needed to
ensure that all clients see the same state has high overhead.  CephFS uses
capabilities to keep metadata strongly consistent. To reduce the number of RPCs
needed for consistency, clients can obtain capabilities for reading, reading
and updating, reads caching, writing, buffering writes, changing the file size,
and performing lazy IO.

\begin{figure*}[tb]%h
\centering
\includegraphics[width=180mm]{figures/throughput-droplease.png}
\caption{When a client create stream is ``isolated" then lookups resolve
locally but when a second client ``interferes" by creating in the same directory,
the directory inode capability is revoked forcing all clients to centralize
lookups at the metadata server.  \label{fig:throughput-droplease}}
\end{figure*}

% purpose of the inode ccahe - reduces RPCs (lookups for create, readdirs for
% stats)
To keep track of the read caching and write buffering capabilities, the clients
and metadata servers agree on the state of each inode using an inode cache.  If
a client has the directory inode cached it can do metadata writes (e.g.,
create) with a single RPC. If the client is not caching the directory inode
then it must do multiple RPCs to the metadata server to (1) determine if the
file exists and (2) do the actual create.  Unless the client immediately reads
all the inodes in the cache, the inode cache is less useful for create-heavy
workloads because the cached inodes are unused.

% benefits
The benefits of caching the directory inode when creating files is shown in
Figure~\ref{fig:throughput-droplease}(a).  If only one client is creating files
in a directory (``isolated" curve) then that client can lookup the existence of
new files locally before issuing a create request to the metadata server. If
another client starts creating files in the same directory (``interfere" curve)
then the directory inode transitions out of read caching and the first client
must send lookups to the metadata server. When other clients interfere the
request throughput is higher Figure~\ref{fig:throughput-droplease}(b) but the
runtime is slower because the isolated client scenario incurrs less requests.
Figures~\ref{fig:throughput-droplease}(c) and (d) plot the creates and lookups,
repsectively, over time; creates slow down and lookups dominate the request
load when the directory inode is shared in the interferring client scenario.

% TODO: what is the cost of trimming the cache?
% TODO: does CephFS still cache inodes when I turn off caching? Why is still keeping inodes in memory? Gah!
%\begin{figure}[tb]%h
%\centering
%\includegraphics[width=80mm]{figures/throughput-cache.png}
%\caption{ CephFS tries to maintain 100 thousand inodes in the cache;
%performance degrades when it cannot evict inodes fast enough to maintain this
%threshold. \label{fig:throughput-cache}}
%\end{figure}

% disadvantages
The drawbacks of the inode cache are shown in
Figure~\ref{fig:throughput-cache}.  Figure~\ref{fig:throughput-cache}(a) shows
the throughput in metadata requests per second for a single client creating 200
thousand files. Until time 950 the throughput is steady at just under 2000
ops/sec and the CPU utilization is at about 30\%. At time 950 seconds
throughput degrades which corresponds to more inodes in the cache
(Figure~\ref{fig:throughput-cache}(c)). The values in
Figure~\ref{fig:throughput-cache}(c) are:

\begin{itemize}
  \item inodes: total number of elemnts in the cache
  \item inodes pinned: 
  \item inodes pinned tail
  \item inodes with caps
  \item inodes top
  \item inodes bottom 
\end{itemize}

% how the inode cache works
CephFS tries to keep the cache at 100 thousand inodes so the degradation in
performance indicates that the metadata server cannot keep up with the
workload. Figure~\ref{fig:throughput-cache}(d) shows inodes getting added at
the same rate as before 950 seconds (ino\(+\)) but the rate at which inodes get
removed is less stable (ino\(-\)) indicating that cache eviction is taking more
of the metadata servers time.

\subsection{Consistency Overhead}
\begin{figure*}[tb]
\centering
\includegraphics[width=180mm]{figures/exp0-underloaded.png}
\caption{An underloaded metadata server adaquately services conflicting
clients; the create speeds are similar while the interferring operations are
limited by the cost of RPCs.}\label{fig:exp0-underloaded}
\end{figure*}

\begin{figure*}[tb]
\centering
\includegraphics[width=180mm]{figures/runtime-consistency-scale.png}
\caption{Scaling clients shows increased variability when another client
interferes; zooming in on runs with 7 clients we see that different types of
interferring operations have different effects on performance variability and
predictability.  }\label{fig:runtime-consistency-scale}
\end{figure*}

Figure~\ref{fig:exp0-underloaded} is a baseline showing that the metadata
server can service multiple clients when it is underloaded. One client creates
files in the same directory and another does a touch or stat 15 seconds into
the run. Figure~\ref{fig:exp0-underloaded}(a) shows the runtime of the client
doing the creates: "isolated" is when the create client is the only workload in
the cluster, "interfere stat" is the runtime of the create client when another
client does a stat, and "interfere touch" is the runtime of the create client
when another client does a touch. There is only a minor performance
degradation. Figure~\ref{fig:exp0-underloaded}(b) compares the baseline
the other client doing the touch or stat
%
%and Figure~\ref{fig:exp0-underloaded}
%has a throughput trace to show behavior.
%
%In Figure~\ref{fig:exp0-underloaded}(a), 
%creates only see a minor performance
%degradation while the interferring operations in
%Figure~\ref{fig:exp0-underloaded}(b) show a larger (yet not significant)
%discrepency. 
%
%
%
% isolated is when there is no interferrence from client 1.
To explain Figure~\ref{fig:exp0-underloaded}(a) we use
Figure~\ref{fig:exp0-behaviour}(b) to show why the runtime of creating 100
thousand files is unaffected. Compared to the throughput without an
interferring stat ("isolated" curve) the throughput of the metadata server with
an interferring stat ("inteferring stat" curve) is higher. This means the
metadata server is doing more operations suggesting that it can adequately
handle the demands of both workloads.

To explain Figure~\ref{fig:exp0-underloaded}(b) we use
Figure~\ref{fig:exp0-underloaded}(a) to show that operations are bounded by the
number of files; more files incurr more requests since both operations lookup
every file. The number of files shown in Figure~\ref{fig:exp0-behaviour}(a)
indicates that the local stat ("client 0" bar) is interacting with less files
than the remote stat ("client 1" bar). This graph also explains why the
isolated stats and touches take the longest from the remote clients; out of all
setups, the isolated operations are doing the most RPCs.  The difference
between local ("client 0 interfere" bar) and remote ("client 1 interfere" bar)
in Figure~\ref{fig:exp0-underloaded}(b) is the overhead of doing RPCs.
Isolated is the slowest because it requests the most files. On average the
local interferring client ("client 0" bar in
Figure~\ref{fig:exp0-behaviour}(a)) requests 40 thousand less files than the
remote interferring call ("client 1" bar in
Figure~\ref{fig:exp0-behaviour}(a)). This reflects how fast the local client
can get into the queue of requests, presumably because it bypasses capability
checks.

From this experiment we make 2 conclusions: (1) the metadata server is not
overloaded because it can handle both workloads and (2) the metadata server
needs to get global consistency from only 1 client

\textbf{Comparison to decoupled namespaces}: Decoupled namespaces merge batches
of metadata operations into the global namespaces when the job completes.  In
BatchFS the merge is delayed by the application using an API to switch between
asynchronous to sychronous mode. The merge itself is explicitly managed by the
application but future work looks at more automated methdologies. In DeltaFS
snapshots of the metadata subtrees stays on the client machines; there is no
ground truth and consistent namespaces are constructed and resolved at
application read time or when a 3rd party system (e.g., middleware, scheduler,
etc.) needs a view of the metadata.

### 2017-03-08 paper/sections/5_implementation.tex
paper: switch template, refactor sections, add pretty code
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Implementation}
\label{sec:implementation}

% Outline of the section
Each section below corresponds ot the labeled arrows in
Figure~\ref{fig:decouple}. This implementation decouples policies from
mechanisms allowing applciations to choose the consistency and fault tolerance
semantics they need.

% Programmability
Of the 6 mechanisms in Figure~\ref{fig:decouple} 4 had to implemented and only
1 required changes to the underlying storage system itself.  RPCs and stream
can be achieved with tunables in configuration files for the storage system
(e.g., metadata cache size, logging on/off, etc.).  Persist", save, and create
are implemented as library and does not require modifications to the storage
system. Volatile apply requires changes to the metadata server to inject
updates back into the global namespace.

\subsection{No Changes: RPCs, Stream}

RPCs is the default behavior of the storage system. Depending on the cache
state and configuration each operation incurrs at least RPC. For example, if a
client or metadata server is not caching the directory inode, all creates
within that directory will result in a lookup and a create request. If the
directory inode is cached then only the create needs to be sent.

Re-using the debugging and testing features in the storage system, we can turn
stream on and off. If it is off, then the metadata servers will not save
journals in the object store and the daemons that apply the journal to the
metadata store will never run. Performance numbers for enabling and disabling
the stream phase are presented back in Section~\S\ref{sec:fault-tolerance}.

% General information about the journal
The journal segments are saved as objects in RADOS.  The journal has 4
pointers, described in `osdc/Journaler.h`:

\begin{itemize}
  \item write position: tail of the journal; points to the current session where we are appending events
  \item unused field: where someone is reading
  \item expire position: old journal segments
  \item trimmed position: where daemon is expiring old items
\end{itemize}

% How the journal tool works
Journal segments in RADDS have a header followed by serialized log events. The
log events are read by hopping over objects using the read offset and object
size pulled from the journal header.  After decoding them, we can examine the
metadata (1) about the event (e.g., type, timestamps, etc.) and (2) for inodes
that the event touches.

% The metablobs
The metadata for inodes that the event touches are called metadata blobs and
the ones associated with events are \textbf{unordered}; this layout makes
writing journal updates fast but the cost is realized when reading the metadata
blobs.  It makes sense to optimize for writing since reading only occurs on
failures. To reconstruct the namespace for the metadata blob, the journal tool
iterates over each metadata blob in the events and builds mappings of inodes to
directory entries (for directories) and parent inodes to child inodes/directory
entries.

\subsection{External Library: Create, Save, Persist}
\label{sec:external-library}
%Reads/Writes Journal Events}

% How it works
For ``create", ``save", and ``persist" Cudele provides a library for clients to
link into.  Recall that CephFS uses the object store (1) as a metadata store
for all information about files including the hierarchical namespace and (2) as
a staging area for the journal of updates before they are applied to the
metadata store. As discussed previously, the metadata store is optimized for
reading while the journal is optimized for writing.  Cudele re-uses the journal
tool to read/write log events to memory and persistent storage.

% creates
The journal tool is used for disaster recovery and lets administrators view and
modify the journal of metadata updates; it can read the journal, erase events
from the journal, and apply the updates in the journal to the metadata store.
To apply journal updates to the metadata store, the journal tool reads the
journal segments from object store objects and applies the update to the
metadata store (which are also stored as object store objects).  For create,
clients write to an in-memory journal and updates are merged into the global
namespace by replaying them onto the metadata store in the object store.

% Journal import
The journal tool imports journals from binary files stored on disk.  First the
header of the dump is sanity checked and written into RADOS to the ``header"
object.  The ``header" object has metadata about the journal as well as the
locations of all the journal pointers (e.g., where the tail of the journal is,
where we are currently trimming, etc.).  Then the journal events are cleaned
(erasing trailing events that are not part of the header) and written as
objects into RADOS.  Note that while the journal is in RADOS, the metadata
servers do do not have the namespace reconstructed in memory so the metadata
cluster will not service requests relating to the journal of imported events.
To construct the namespace in the collective memory of the metadata servers we
need to first construct the namespace in RADOS. The journal tool can explicitly
do this by  applying the journal to the metadata store in RAODS. This will pull
the objects containing journal segments and replay them on the metadata store.
Finally, we delete the journal in RADOS and restart the metadata servers so
they rebuild their caches.

% Journal export
The journal tool exports journals to binary files stored on disk. First the
journal is scanned for the header and then journal is recovered. To recover the
journal the ``header" object is read off disk and then objects are probed in
order and starting from the write position saved in the header. Probing will
update the write position if it finds objects with data in them.

% Journal export
When exporting a journal of events, the journal tool first scans the journal to
check for corruption. Then it recovers the journal by reading the ``header"
object out of RADOS.  After reading the header, the journal tool can pull
journal segments from RADOS because it knows how many objects to pull and how
far to seek within those objects.

%% The data structures
%The metadata that the event touches, including inodes, paths, and timestamps,
%are stored as metablobs. Each piece of metadata inside a metablob is called a
%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
%table client transactions (tid, version), renamed directory fragments (maps,
%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
%me (and you since you are reading this paper), this does not make any sense.
%
%Each directory fragment has an associated directory lump, which is just a bunch
%of metadata. The most interesting part of the dirlump is the fullbits array,
%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
%and then all the full bits, saving off the children and inode locations. The
%children tell us which dentry names an inode has and the inode locations map
%the parent inode to its child inode and dentry.

% persist/save
For save, clients write serialized log events to a file on local disk and for
persist, clients store the file in the objects store. The overheads for save
and persist is the write bandwidth of the local disk and object store,
respectively.

% Level of intrusion
This required no changes to Ceph because the metadata server knows how to read
the events we were writing into the object store.  The client writes metadata
updates locally and merges the updates with the journal tool. The client that
decouples the namespace operates without any consistency and any conflicts at
the merge are resolved in favor of this client. Updates by other clients ({\it
i.e.} metadata writes to the global namespace) are overwritten.  We leverage
the journal tool's ability to reconstruct metadata events in memory. The client
library is shown in Figure~\ref{fig:decouple}.  Cudele adopts the following
process when an application decouples the namespace:

\begin{enumerate}
  \item ``\textbf{decoupled}": metadata server exports the journal events to a file
  \item ``\textbf{transfer}": the file is pulled by the client from the metadata server
  \item ``\textbf{snapshot}": client reads the file and materializes a snapshot of the
        namespace in memory
\end{enumerate}

% Why we re-use stuff
By using re-using the journal subsystem to implement the namespace decoupling,
Cudele leverages the write/read optimized data structures, the formats for
persisting events (similar to TableFS's SSTables), and the functions for
replaying events onto the internal namespace data structures.

%Step 3 is the most complicated and requires understanding how the snapshot is
%materialized in memory. 
%
%\subsubsection{Operating on Snapshots} 
%
%Our first implementation attempted to re-create journal events using the same
%libraries that the metadata server uses. To construct a \texttt{mkdir} we tried
%to instantiate a Ceph inode and directory entry for the current file/dir and
%its parent.  This is too hard because there are too many moving parts in the
%metadata server (e.g., a mdlog class, stuff in memory, assumption that we can
%traverse up and down namespace, etc.). So when I tried add dentries and inodes
%it was trying to traverse up/down and it would almost always segfault when it
%was looking for something. These metablobs are supposed to be self container --
%the problem is I do not know what is supposed to go *inside* them. 
%
%Our second idea was to copy the metadata blog and change just what we needed.
%For example, we would save a binary dump of a generic \texttt{mkdir} event on
%disk. When the application makes a directory, this dump would be loaded and the
%fields would be changed before being written back to disk. Rather than
%traversing up and down a namespace in memory of a metadata server, we should
%traverse up and down the namespace *inside* the metadata blob. This
%implementation requires disk IO and editing the log event is non-trivial for
%two reasons:
%
%\begin{itemize}
%
%  \item methods do not edit events; they just write them
%
%  \item the metadata that the event touches (e.g., the metablob) is unorganized
%  on disk for performance -- it is trade-off for writing data faster serially and
%  reconstructing information slowly since failure is not considered the norm
%
%\end{itemize}
%
%Faced with these challenges we landed on our final implementation: load the
%snapshot into the the data structures used to examine and replay journals, edit
%those data structures, and write them out to disk as binary.

\subsection{Storage System Changes: Volatile Apply}

% - how it makes no gaurantees
The volatile apply mechanism ({\it i.e.} the ``v\_apply" arrow in
Figure~\ref{fig:decouple}) takes an in-memory journal on the client and applies
the updates directly to the in-memory namespace maintained by the metadata
servers. We say volatile because Cudele makes no consistency or fault
tolerance guarantees. If a concurrent update from a client occurs there is no
rule for resolving conflicts and if the client or metadata server crashes there
may be no way to recover. Relaxing these constraints gives volatile apply the
best performance.

% difference between apply and volatile apply
In contrast, apply uses the the object store to merge the journal of updates
from the client to the metadata server. Apply is safer than volatile apply but
has a performance overhead because objects in the metadata store need to be
read from and written back to the object store.  Apply was already implemented
by the journal tool: it reads the journal from the object store, adds events to
the journal, and writes the metadata updates out to the metadata store in the
object store. Cudele's volatile apply executes ``save" (described
in~\S\ref{sec:external-library}), transfers the file of metadata updates to the
metadata server, and then merges the updates directly into the in-memory
metadata store of the metadata cluster.

% Why so slow?
Creating many files in the same directory would touch the same object but the
existing implementation results in this object being repeatedly pushed/pulled.

% - how it actually works
% - bugs I fixed in the last couple of commits

%The metadata objects are located with naming schemes (200.000* for journal
%objects and 1.inode for metadata storage objects).

% How it works: socket for changing daemon's internal state (debugging, logging, behaviour)
% 1. API for putting state into the daemon dynamically
% 2. Hooks directly into daemon code so we can use any parsing functionality in there
% 3. Documentation all the tunables

% 1. read journal of updates from file
% 2. call replay (uses same code as when an MDS comes back) on each event
% 3. skip inodes so MDS doesn't hand out those new inodes

### 2017-03-08 paper/sections/6_results.tex
paper: switch template, refactor sections, add pretty code
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Results}
\subsection{Microbenchmarks}
\begin{figure*}[tb]
\centering
\includegraphics[width=180mm]{figures/results-microbenchmark.png}
\caption{Here's a graph.}
\end{figure*}

\begin{figure}[tb]
\centering
\includegraphics[width=90mm]{figures/results-files.png}
\caption{Here's a graph.}
\end{figure}

\subsubsection{Per phase latencies}
\begin{figure}[tb]
\centering
\includegraphics[width=90mm]{figures/global-v-decoupled.png}
\caption{Compared to the \texttt{create} phase, saving and persisting
updates (\texttt{create+save} and \texttt{create+save+persist}) experience only
a 4.79\(\times\) and 8.66\(\times\) slowdown, in the worst case for 100K files.
In contrast, maintaining \texttt{global} consistency is 905.70\(\times\)
slower.  The disadvantage of decoupling the namespace is the merge phase where
updates are applied to the metadata store (\texttt{create+apply}, resulting in a
905.70\(\times\) slowdown for 100K files.}\label{fig:global-v-decoupled}
\end{figure}

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: eventual vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: eventual, local
%  - e.g., POSIX: strong, global
%  - e.g., PLFS: no consistency
%- fault tolerance: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 MDS, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)

\subsection{Journaling Overhead}

Journal to RADOS
Turn off journaling (large segment)
Journal to in-memory OSD

\subsection{Macrobenchmarks}
updatedb: http://lists.ceph.com/pipermail/ceph-users-ceph.com/2015-July/002768.html

### 2017-03-09 images/ceph/build.sh
image: build newest   - mds: support locked decoupled namespaces   - bug fixes for inode number
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -15,8 +15,10 @@ docker pull ceph/daemon:tag-build-master-jewel-ubuntu-14.04
 docker tag ceph/daemon:tag-build-master-jewel-ubuntu-14.04 ceph/daemon:jewel
 
 #-e SHA1_OR_REF="ec12610bb1358fb2eab74cb7b0483c2613bae38f" \
+#-e SHA1_OR_REF="1fdfa94459026b70e8925c0a530005456c526260" \
+#-e SHA1_OR_REF="f4d39ca86e29555a45a4cff3f03862081bcc1f84" \
 dmake \
-  -e SHA1_OR_REF="1fdfa94459026b70e8925c0a530005456c526260" \
+  -e SHA1_OR_REF="11d3bc22ceb151c23c54edaadab2d9d0ec9659fe" \
   -e GIT_URL="https://github.com/michaelsevilla/ceph.git" \
   -e BUILD_THREADS=`grep processor /proc/cpuinfo | wc -l` \
   -e CONFIGURE_FLAGS="-DWITH_TESTS=OFF" \
@@ -25,6 +27,8 @@ dmake \
 cd -
 
 #docker tag ec12610bb1358fb2eab74cb7b0483c2613bae38f-base ceph/daemon:ec12610
-docker tag ceph-1fdfa94459026b70e8925c0a530005456c526260 ceph/daemon:1fdfa94
+#docker tag ceph-1fdfa94459026b70e8925c0a530005456c526260 ceph/daemon:1fdfa94
+#docker tag ceph-f4d39ca86e29555a45a4cff3f03862081bcc1f84 ceph/daemon:f4d39ca
+docker tag ceph-11d3bc22ceb151c23c54edaadab2d9d0ec9659fe ceph/daemon:11d3bc2
 docker build -t tmp .
-docker tag tmp piha.soe.ucsc.edu:5000/ceph/daemon:1fdfa94
+docker tag tmp piha.soe.ucsc.edu:5000/ceph/daemon:11d3bc2

### 2017-03-18 paper/sections/4_methodology.tex
paper: add hotstorage draft
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \begin{tabular}{ l | l | l | l }
-  C \(\rightarrow\) &&& \\  
-  D \(\downarrow\)  & none              & eventual        & strong  \\\hline
+  Consistency \(\rightarrow\) &&& \\  
+  Durability \(\downarrow\)  & invisible         & eventual        & strong  \\\hline
   none                       & create            & create          & RPCs    \\
@@
   local                      & create            & create          & RPCs    \\
-                             & +local persist    & +local persist  & +local persist \\
-                             &                   & +volatile apply &         \\\hdashline
+                             & +local persist    & +local persist  & +local  \\
+                             &                   & +volatile apply &  persist\\\hdashline
   global                     & create            & create          & RPCs    \\
@@
 
-\subsection{Cudele Namespace API: Per-Subtree Policies}
+\subsection{Cudele Namespace API:\\Per-Subtree Policies}
 
@@
 
+\begin{listing}
+\begin{minted}[frame=single,
+               framesep=3mm,
+               xleftmargin=21pt,
+               tabsize=4]{js}
+{     
+    "allocated_inodes": "100000"
+    "interfere_policy": "block"
+    "consistency": "create"
+    "durability": "local_persist"
+}
+\end{minted}
+\caption{Implementing DeltaFS with Cudele.}
+\label{src:deltafs}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[frame=single,
+               framesep=3mm,
+               xleftmargin=21pt,
+               tabsize=4]{js}
+{     
+    "allocated_inodes": "100000"
+    "interfere_policy": "block"
+    "consistency": "create+volatile_apply"
+    "durability": "local_persist"
+}
+\end{minted}
+\caption{Implementing BatchFS with Cudele.}
+\label{src:batchfs}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[frame=single,
+               framesep=3mm,
+               xleftmargin=21pt,
+               tabsize=4]{js}
+{     
+    "allocated_inodes": "-"
+    "interfere_policy": "-"
+    "consistency": "RPCs"
+    "durability": "stream"
+}
+\end{minted}
+\caption{Existing CephFS on Cudele.}
+\label{src:batchfs}
+\end{listing}
+
 % // limit consistency

### 2017-03-18 paper/build.sh
paper: cleanup garbage
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.synctex.gz >> /dev/null 2>&1
+rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.synctex.gz build.log >> /dev/null 2>&1
 set -ex
 
 # Build the paper
@@ -22,6 +22,9 @@ if [ $ERR != "0" ] ; then
   cat build.log
   exit 1
 fi
+
+set +e
+rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.synctex.gz build.log >> /dev/null 2>&1
 
 echo "SUCCESS"
 exit 0

### 2017-03-20 paper/sections/0_abstract.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \begin{abstract}
-Today's large and highly-parallel workloads are bottlenecked by metadata
-services because many processes end up accessing the same shared resoure. In
-HPC, state-of-the-art file systems are abandoning POSIX because the
-sychronization and serializiation overheads are too costly -- and sometimes
-even unnceccessary -- for some of their applications.  While the performance
-benefits are plain for these users, other applications that rely on stronger
-consistency must be re-written or deployed on a different system.  We present
-Cudele, a programmable file system that supports different degrees of
-consistency and fault tolerance within the same namespace.  First, we take a
-POSIX compliant file system and relax the consistency constraints by
-implementing two metadata designs: delayed metadata update merge into the
-global namespace and client local views of metadata updates.  Second, we store
-the consistency and fault tolerance semantics in inodes so that subtrees within
-the same namespace can be optimized for different workloads; the inodes are
-programmable so that clients understand how to access the metadata in a
-subtree.  Third, we present a theoretical framework for a metadata service and
-a working implementation that provides the lowest common denominator needed to
-implement a wide range of consistency/fault tolerance semantics that can be
-benchmarked all on the same system.
+
+HPC file systems are abandoning POSIX because the synchronization and
+serialization overheads are too costly -- and often uneccessary -- for their
+applications.  We provide an API for the client to (1) merge updates into the
+global namespace and (2) assign policies to subtrees so that other clients
+cannot interfere with the decoupled namespace.  This allows administrators to
+optimize subtrees within the same namespace for different workloads. We draw
+conclusions about the performance impact of previously unexplored
+consistency/durability metadata designs and show that consistency can cause a
+104\(\times\) slow down while merging updates (7\(\times\) slow down) and
+maintaining durability (\(10\times\) slow down) have a more reasonable cost.
+
 \end{abstract}

### 2017-03-20 paper/sections/1_introduction.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+% What is the problem?
 Today's client-server based file system metadata services have scalability
-problems. It takes a lot of resources to server POSIX metadata requests and
+problems. It takes a lot of resources to service POSIX metadata requests so
 applications perform better with dedicated metadata
-servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs}. This is fine for small
-workloads and file systems but as the system scales provisioning a metadata
-server for every client is expensive and complicated.
+servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs}. Unfortunately,
+provisioning a metadata server for every client is expensive and complicated.
 
+% How does this fit in current landscape?
 Current hardware evolution and the rise of software-defined storage storage,
 which uses techniqes like erasure coding, replication, and partitioning, have
-ushered a new era of HPC computing; architectures are transitioning from
+ushered in a new era of HPC computing; architectures are transitioning from
 complex storage stacks with burst buffer, file system, object store, and tape
@@
 store~\cite{bent:login16-hpc-trends}. This trend exacerbates the metadata
-scalability problem and has given rise to the serverless metadata services.
-%\begin{listing}
-\begin{minted}{python}
-def hello():
-\end{minted}
-%\caption{JSON example} 
-%\label{json-example}
-%\end{listing}
+scalability problem
 
-HPC workloads are so metadata intensive that new management techniques are
-advocating reducing synchronization and serialization overheads by transferring
-these responsibilities to the client. We call this approach decoupling the
-namespace and the semantics of consistency differ amongst systems.
+% What is HPC doing?
+To address this trend, HPC research has advocated client side processing and an
+emphasis on serverless metadata services.  The techniques propose new metadata
+management techniques that reduce synchronization and serialization overheads
+by transferring these responsibilities to the client. We call this approach
+``decoupling the namespace" and the semantics of consistency differ amongst
+systems.  While the performance benefits are obvious for these users,
+applications that rely on stronger consistency or durability guarantees must be
+re-written or deployed on a different system.
 
@@
 
+% What did we do
 We propose subtree policies, an interface that lets future programmers control
@@
 shows an example setup where a single global namespace has directories for
-applications designed for different, state-of-the-art HPC architectures.  Our
-system supports 3 forms of consistency and 2 forms of fault tolerance giving
-the administrator a wide range of policies and optimizations depending on the
-application's needs.
+applications designed for different, state-of-the-art HPC architectures.  We
+present Cudele, a prototype programmable file system that supports different
+degrees of consistency and durability by exposing mechanisms used within the
+file system as a client library.  Cudele supports 3 forms of consistency
+(invisible, eventual, and strong) and 3 degrees of durability (none, local, and
+global) giving the administrator a wide range of policies and optimizations
+that can be custom fit to an application. Our contributions:
+
+\begin{enumerate}
+
+  \item a prototype that lets administrators program a range of
+  consistency and durability semantics (9 permutations), allowing them to custom
+  fit the storage system to the application.
+
+  \item an API for assigning and programming policies to subtrees in the file
+  system namespace.
+
+  \item an analysis of recently proposed research systems compared against
+  previously unexplored metadata designs.
+
+\end{enumerate}
+
+% Results
+Our results confirm the assertions of ``clean-state" research systems that
+decouple namespaces; specifically that the technique drastically improve
+performance (104\(\times\) speed up) but we go a step further by quantifying
+the costs of merging updates (7\(\times\) slow down) and maintaining durability
+(\(10\times\) slow down). We also show the effect of having a metadata specific
+file format in systems that are based on in-memory data structures.
+Section~\ref{sec:related-work} places Cudele in the context of other related
+work. Section~\ref{sec:posix-overheads} quantifies the cost of POSIX
+consistency and system-defined durability and
+Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
+prototype and API. Section~\ref{sec:implementation} describes Cudele's
+mechanisms and shows how re-using internal subsystems results in an
+implementation of less than 500 lines of code. The evaluation in
+Section~\ref{sec:evaluation} quantifies the overheads and performance gains of
+explored and previously unexplored metadata designs.

### 2017-03-21 experiments/prep-viz.sh
graph: start viz script
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,40 @@
+#!/bin/bash
+
+#set -x
+
+echo "-- untar the journal"
+
+DIR="kernel/journal/results-done"
+SUBDIRS=`ls $DIR`
+MDS="issdm-12"
+for i in $SUBDIRS; do
+  cd $DIR/$i
+  echo "DIR=" `pwd`
+  for j in `ls *.tar.gz`; do
+    #tar xzf $j
+    if [ $j == $MDS.tar.gz ]; then
+      echo "... parsing MDS=$j"    
+      for m in \
+         "tmp/graphite/whisper/$MDS/mds_mem/ino.wsp" \
+         "tmp/graphite/whisper/$MDS/mds_mem/ino+.wsp" \
+         "tmp/graphite/whisper/$MDS/mds_mem/ino-.wsp" \
+         "tmp/graphite/whisper/$MDS/mds/inodes.wsp" \
+         "tmp/graphite/whisper/$MDS/mds/inodes_bottom.wsp" \
+         "tmp/graphite/whisper/$MDS/mds/inodes_with_caps.wsp" \
+         "tmp/graphite/whisper/$MDS/mds/inodes_pin_tail.wsp" \
+         "tmp/graphite/whisper/$MDS/mds/inodes_pinned.wsp" \
+         "tmp/graphite/whisper/$MDS/mds/inodes_top.wsp" \
+         "tmp/graphite/whisper/$MDS/mds_server/handle_client_request_tput.wsp" \
+         "tmp/graphite/whisper/$MDS/cputotals/user.wsp" \
+         "tmp/graphite/whisper/$MDS/cputotals/sys.wsp" \
+        ; do
+        docker run -v `pwd`/tmp:/tmp --entrypoint=whisper-dump.py --rm \
+          michaelsevilla/graphite ${m} > `basename ${m}`.out
+        docker run -v `pwd`/:/tmp --entrypoint=/bin/bash --rm \
+          michaelsevilla/graphite -c "sed -i \"s/:/,/g\" tmp/`basename $m`.out"
+        echo "... ... `basename $m`"
+      done
+    fi
+  done
+  cd - >> /dev/null
+done

### 2017-03-22 paper/sections/3_posix_overheads.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: popperize graphs and fix links
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \section{POSIX Overheads}
+\label{sec:posix-overheads}
 
@@
 \label{sec:fault-tolerance}
-\begin{figure*}[tb] \centering
-\includegraphics[width=1\textwidth]{./figures/journal.png} 
+
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
+      \caption{}
+      \label{fig:phy-design}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/behavior-journal.png}
+      \caption{}
+      \label{fig:batching}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/slowdown-cache.png}
+      \caption{}
+      \label{fig:batching-outlier}
+  \end{subfigure}
+  \caption{(a) relative performance differences are different after storage
+  software upgrade. (b) total throughput with and without batching. (c)
+  identifying and handling an outlier maintains the benefits
+  of batching without the performance degradation of unnecessarily large I/O
+  requests.}
+\end{figure*}
+
+\begin{figure}[tb] \centering
+\includegraphics[width=1\linewidth]{./graphs/behavior-cache.png} 
+\caption{The inode cache improves metadata read performance but for our
+create-heavy workload it is only an overhead. Most of the time maintaining the
+cache is spent evicting and adding inodes.}\label{fig:inode-cache}
+\end{figure}
+
+\begin{figure}[tb] \centering
+\includegraphics[width=1\linewidth]{./figures/journal.png} 
 \caption{CephFS uses a journal to stage updates and tracks dirty metadata in
@@
 journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[tb]%h
-\centering
-\includegraphics[width=180mm]{figures/throughput-journal.png}
-\caption{Performance improves with larger journal segments because the metadata
-server spends less time flushing the journal }\label{fig:throughput-journal}
-\end{figure*}
+%\begin{figure*}[tb]%h
+%\centering
+%\includegraphics[width=180mm]{figures/throughput-journal.png}
+%\caption{Performance improves with larger journal segments because the metadata
+%server spends less time flushing the journal }\label{fig:throughput-journal}
+%\end{figure*}
 
@@
 
-\begin{figure*}[tb]%h
-\centering
-\includegraphics[width=180mm]{figures/throughput-cache-journal.png}
-\caption{Journalling metadata updates has a bigger overhead than maintaining
-the inode cache. For create-heavy workloads the inode cache offers no
-performance benefits. }\label{fig:throughput-cache-journal}
-\end{figure*}
-
-\begin{figure*}[tb]%h
-\centering
-\includegraphics[width=180mm]{figures/inode-cache.png}
-\caption{The inode cache improves metadata read performance but for our
-create-heavy workload it is only an overhead. Most of the time maintaining the
-cache is spent evicting and adding inodes.}\label{fig:inode-cache}
-\end{figure*}
-
 Access to POSIX metadata is strongly consistent, so reads and writes are
@@
 
-\begin{figure*}[tb]%h
+\begin{figure*}[t]
   \centering
-\includegraphics[width=180mm]{figures/throughput-droplease.png}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/behavior-interfere.png}
+      \caption{}
+      \label{fig:phy-design}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere.png}
+      \caption{}
+      \label{fig:batching}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere-types.png}
+      \caption{}
+      \label{fig:batching-outlier}
+  \end{subfigure}
   \caption{When a client create stream is ``isolated" then lookups resolve
-locally but when a second client ``interferes" by creating in the same directory,
-the directory inode capability is revoked forcing all clients to centralize
-lookups at the metadata server.  \label{fig:throughput-droplease}}
+  locally but when a second client ``interferes" by creating in the same
+  directory, the directory inode capability is revoked forcing all clients to
+  centralize lookups at the metadata server.  An underloaded metadata server
+  adaquately services conflicting clients; the create speeds are similar while
+  the interferring operations are limited by the cost of RPCs.  Scaling clients
+  shows increased variability when another client interferes; zooming in on runs
+  with 7 clients we see that different types of interferring operations have
+  different effects on performance variability and predictability.
+  \label{fig:throughput-droplease}}
 \end{figure*}
@@
 \subsection{Consistency Overhead}
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=180mm]{figures/exp0-underloaded.png}
-\caption{An underloaded metadata server adaquately services conflicting
-clients; the create speeds are similar while the interferring operations are
-limited by the cost of RPCs.}\label{fig:exp0-underloaded}
-\end{figure*}
-
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=180mm]{figures/runtime-consistency-scale.png}
-\caption{Scaling clients shows increased variability when another client
-interferes; zooming in on runs with 7 clients we see that different types of
-interferring operations have different effects on performance variability and
-predictability.  }\label{fig:runtime-consistency-scale}
-\end{figure*}
 

### 2017-03-22 paper/sections/6_results.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: popperize graphs and fix links
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
-\section{Results}
+\section{Evaluation}
+\label{sec:evaluation}
+
 \subsection{Microbenchmarks}
-\begin{figure*}[tb]
-\centering
-\includegraphics[width=180mm]{figures/results-microbenchmark.png}
-\caption{Here's a graph.}
-\end{figure*}
+%\begin{figure*}[tb]
+%\centering
+%\includegraphics[width=180mm]{graphs/slowdown.png}
+%\caption{Compared to the \texttt{create} phase, saving and persisting
+%updates (\texttt{create+save} and \texttt{create+save+persist}) experience only
+%a 4.79\(\times\) and 8.66\(\times\) slowdown, in the worst case for 100K files.
+%In contrast, maintaining \texttt{global} consistency is 905.70\(\times\)
+%slower.  The disadvantage of decoupling the namespace is the merge phase where
+%updates are applied to the metadata store (\texttt{create+apply}, resulting in a
+%905.70\(\times\) slowdown for 100K files.}\label{fig:global-v-decoupled}
+%\end{figure*}
 
@@
 \centering
-\includegraphics[width=90mm]{figures/results-files.png}
-\caption{Here's a graph.}
+\includegraphics[width=75mm]{graphs/slowdown-mechanisms.png}
+\caption{}
+\end{figure}
+\begin{figure}[tb]
+\centering
+\includegraphics[width=75mm]{graphs/slowdown-related-work.png}
+\caption{}
+\end{figure}
+\begin{figure}[tb]
+\centering
+\includegraphics[width=75mm]{graphs/slowdown-strong-v-eventual.png}
+\caption{}
 \end{figure}
@@
 \subsubsection{Per phase latencies}
-\begin{figure}[tb]
-\centering
-\includegraphics[width=90mm]{figures/global-v-decoupled.png}
-\caption{Compared to the \texttt{create} phase, saving and persisting
-updates (\texttt{create+save} and \texttt{create+save+persist}) experience only
-a 4.79\(\times\) and 8.66\(\times\) slowdown, in the worst case for 100K files.
-In contrast, maintaining \texttt{global} consistency is 905.70\(\times\)
-slower.  The disadvantage of decoupling the namespace is the merge phase where
-updates are applied to the metadata store (\texttt{create+apply}, resulting in a
-905.70\(\times\) slowdown for 100K files.}\label{fig:global-v-decoupled}
-\end{figure}
 
@@
 
-\subsection{Journaling Overhead}
-
-Journal to RADOS
-Turn off journaling (large segment)
-Journal to in-memory OSD
+\subsection{Isolation from EBUSY}
+\subsection{Isolation from Global Writes, Cost of Overwrites}
+\subsection{Partial Directory Listings}
 

### 2017-03-22 experiments/prep-viz.sh

@@ -1,40 +1,0 @@
-#!/bin/bash
-
-#set -x
-
-echo "-- untar the journal"
-
-DIR="kernel/journal/results-done"
-SUBDIRS=`ls $DIR`
-MDS="issdm-12"
-for i in $SUBDIRS; do
-  cd $DIR/$i
-  echo "DIR=" `pwd`
-  for j in `ls *.tar.gz`; do
-    #tar xzf $j
-    if [ $j == $MDS.tar.gz ]; then
-      echo "... parsing MDS=$j"    
-      for m in \
-         "tmp/graphite/whisper/$MDS/mds_mem/ino.wsp" \
-         "tmp/graphite/whisper/$MDS/mds_mem/ino+.wsp" \
-         "tmp/graphite/whisper/$MDS/mds_mem/ino-.wsp" \
-         "tmp/graphite/whisper/$MDS/mds/inodes.wsp" \
-         "tmp/graphite/whisper/$MDS/mds/inodes_bottom.wsp" \
-         "tmp/graphite/whisper/$MDS/mds/inodes_with_caps.wsp" \
-         "tmp/graphite/whisper/$MDS/mds/inodes_pin_tail.wsp" \
-         "tmp/graphite/whisper/$MDS/mds/inodes_pinned.wsp" \
-         "tmp/graphite/whisper/$MDS/mds/inodes_top.wsp" \
-         "tmp/graphite/whisper/$MDS/mds_server/handle_client_request_tput.wsp" \
-         "tmp/graphite/whisper/$MDS/cputotals/user.wsp" \
-         "tmp/graphite/whisper/$MDS/cputotals/sys.wsp" \
-        ; do
-        docker run -v `pwd`/tmp:/tmp --entrypoint=whisper-dump.py --rm \
-          michaelsevilla/graphite ${m} > `basename ${m}`.out
-        docker run -v `pwd`/:/tmp --entrypoint=/bin/bash --rm \
-          michaelsevilla/graphite -c "sed -i \"s/:/,/g\" tmp/`basename $m`.out"
-        echo "... ... `basename $m`"
-      done
-    fi
-  done
-  cd - >> /dev/null
-done

### 2017-03-22 viz.sh

@@ -1,0 +1,82 @@
+#!/bin/bash
+
+#echo "-- untar the journal"
+#DIR="experiments/kernel/journal/results-done"
+#SUBDIRS=`ls $DIR`
+#MDS="issdm-12"
+#for i in $SUBDIRS; do
+#  cd $DIR/$i
+#  echo "DIR=" `pwd`
+#  for j in `ls *.tar.gz`; do
+#    tar xzf $j
+#    if [ $j == $MDS.tar.gz ]; then
+#      echo "... parsing MDS=$j"    
+#      for m in \
+#         "tmp/graphite/whisper/$MDS/mds_mem/ino.wsp" \
+#         "tmp/graphite/whisper/$MDS/mds_mem/ino+.wsp" \
+#         "tmp/graphite/whisper/$MDS/mds_mem/ino-.wsp" \
+#         "tmp/graphite/whisper/$MDS/mds/inodes.wsp" \
+#         "tmp/graphite/whisper/$MDS/mds/inodes_bottom.wsp" \
+#         "tmp/graphite/whisper/$MDS/mds/inodes_with_caps.wsp" \
+#         "tmp/graphite/whisper/$MDS/mds/inodes_pin_tail.wsp" \
+#         "tmp/graphite/whisper/$MDS/mds/inodes_pinned.wsp" \
+#         "tmp/graphite/whisper/$MDS/mds/inodes_top.wsp" \
+#         "tmp/graphite/whisper/$MDS/mds_server/handle_client_request_tput.wsp" \
+#         "tmp/graphite/whisper/$MDS/cputotals/user.wsp" \
+#         "tmp/graphite/whisper/$MDS/cputotals/sys.wsp" \
+#        ; do
+#        docker run -v `pwd`/tmp:/tmp --entrypoint=whisper-dump.py --rm \
+#          michaelsevilla/graphite ${m} > `basename ${m}`.out
+#        docker run -v `pwd`/:/tmp --entrypoint=/bin/bash --rm \
+#          michaelsevilla/graphite -c "sed -i \"s/:/,/g\" tmp/`basename $m`.out"
+#        echo "... ... `basename $m`"
+#      done
+#    fi
+#  done
+#  cd - >> /dev/null
+#done
+#
+#DIR="experiments/kernel/journal/results/journal180-cache"
+#SUBDIRS=`ls $DIR`
+#MDS="issdm-12"
+#cd $DIR
+#echo "DIR=" `pwd`
+#for j in `ls *.tar.gz`; do
+#  tar xzf $j
[…]

### 2017-03-23 paper/sections/3_posix_overheads.tex
paper: edit section 3
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \centering
-\includegraphics[width=75mm]{figures/creates-motivation.png}
+\includegraphics[width=1.0\linewidth]{figures/creates-motivation.png}
 \caption{Create-heavy workloads (untar) incurr the highest disk, network, and CPU
@@
 In our examiniation of the overheads of POSIX we benchmark and analyze CephFS,
-the file system that uses the RADOS object store to store its data and
-metadata. We choose CephFS because it is an open-source production quality
-system. CephFS made one set of design decisions and we not asserting that the
-design decisions that were made are superior but instead highlight the
-effect those decisions have on performance.
+the file system that uses the Ceph's object store ({\it i.e.} RADOS) to store
+its data and metadata. We choose CephFS because it is an open-source production
+quality system. This file system is an implementation of one set of design
+decisions; we are not asserting that those decisions are superior but instead
+highlight the effect that those decisions have on performance.
 

### 2017-03-23 paper/sections/6_results.tex
paper: edit section 3
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \label{sec:evaluation}
+
+We evaluate Cudele on a 15 node cluster, partitioned into 8 object storage
+servers, 3 metadata servers, and 2 monitor servers. The object storage servers
+double as clients which is fine because clients are CPU and memory bound while
+object storage servers are disk IO bound. All daemons run as a single process
+which is the default setting for Ceph and the nodes have 2 dual core 2GHz
+processors with 8GB of RAM. There are three daemons per object storage server
+(one for each disk formatted with XFS) and they share an SSD for the journal.
+The nodes are running Ubuntu 12.04.4, kernel version 3.2.0-63 but all
+experiments run in Docker containers; this makes it easier to tear down and
+re-initialize ({\it e.g.}, dropping the kernel cache) for the cluster between
+experiments.
+
+This paper adheres to The Popper
+Convention\footnote{\url{http://falsifiable.us}}~\cite{jimenez_popper_2016}, so
+experiments presented here are available in the repository for this
+article\footnote{\url{https://github.com/michaelsevilla/cudele-popper/tree/}}.
+Experiments can be examined in more detail, or even re-run, by visiting the
+\texttt{[source]} link next to each figure. That link points to a Jupyter
+notebook that shows the analysis and source code for that graph and its parent
+folder contains all the associated artifacts.
 
@@
 \centering
-\includegraphics[width=75mm]{graphs/slowdown-mechanisms.png}
+\includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
 \caption{}
@@
 \centering
-\includegraphics[width=75mm]{graphs/slowdown-related-work.png}
+\includegraphics[width=1.0\linewidth]{graphs/slowdown-related-work.png}
 \caption{}
@@
 \centering
-\includegraphics[width=75mm]{graphs/slowdown-strong-v-eventual.png}
+\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-eventual.png}
 \caption{}

### 2017-03-25 paper/sections/2_related_work.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: cleanup section 2 and prune section 3
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \section{Related Work} 
+\label{sec:related-work}
 
@@
 migration mechanisms themselves.  For example, IndexFS and CephFS use the
-GIGA+~\cite{GIGA+} technique for partitioning directories at a predefined
+GIGA+~\cite{patil:fast2011-giga} technique for partitioning directories at a predefined
 threshold and using lazy synchronization to redirect queries to the server that
@@
 resources, respectively; for random workloads larger than the cache extra RPCs
-hurt performance~\cite{IndexFS, CephFS} and for write heavy workloads with more
+hurt performance~\cite{ren:sc2014-indexfs, weil:sc2004-dyn-metadata} and for write heavy workloads with more
 resources the RPCs for invalidations are harmful. Another approach to reducing
@@
 
-Decoupling the namespace delays metadata consistency and sacrifices fault
-tolerance. As shown in Table~\ref{table:namespaces}, metadata consistency is
-provided by capabilities and fault tolerance is addressed with a journal.
-
-\begin{tabular}{ r | l | l | l }
-              & Decoupled & Global    & Implied   \\
-              & Namespace & Namespace & Namespace \\\hline
-  Example     & BatchFS & CephFS & PLFS \\
-              & DeltaFS & IndexFS &     \\
-  Consistency & eventual & strong & none? \\
-  Fault Tol.  & node local & journal & journal \\
+Decoupling the namespace delays metadata consistency and sacrifices durability. 
+As shown in Table~\ref{table:namespaces}, metadata consistency is
+provided by capabilities and metadata durability is addressed with a journal.
+\begin{table}
+\begin{tabular}{ r | l | l }
+              & Decoupled & Global    \\
+              & Namespace & Namespace \\\hline
+  Example     & BatchFS~\cite{zheng:pdsw2014-batchfs} & CephFS~\cite{weil:sc2004-dyn-metadata} \\
+              & DeltaFS~\cite{zheng:pdsw2015-deltafs} & IndexFS~\cite{ren:sc2014-indexfs}      \\
+  Consistency & eventual & strong     \\
+  Durability  & node local & journal  \\
 \end{tabular}
+\caption{State-of-the-art systems in HPC improve file system metadata
+performance by relaxing consistency and durability
+guarantees.\label{table:namespaces}}
+\end{table}

### 2017-03-25 paper/sections/3_posix_overheads.tex
paper: cleanup section 2 and prune section 3
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=1.0\linewidth]{figures/creates-motivation.png}
-\caption{Create-heavy workloads (untar) incurr the highest disk, network, and CPU
-utilization because the metadata server is managing consistency and fault
-tolerance.}\label{fig:creates-motivation}
+\caption{Create-heavy workloads (untar) incurr the highest disk, network, and
+CPU utilization because of the consistency and durability demands of CephFS
+.}\label{fig:creates-motivation}
 \end{figure}
@@
 quality system. This file system is an implementation of one set of design
-decisions; we are not asserting that those decisions are superior but instead
-highlight the effect that those decisions have on performance.
+decisions and our goal is to highlight the effect that those decisions have on
+performance.
 
-To show how CephFS behaves under high metadata load we use a create-heavy
-workload because these types of workloads incurr high CPU, network, and disk
-usage.  Figure~\ref{fig:creates-motivation} shows a compilation of the Linux
-kernel, which has a download, untar, compile, and remove phase. The untar phase
-is characterized by many creates and has the highest resource usage, indicating
-that it is stressing the consistency and journalling subsystems of the metadata
-server. Also of note: a create-heavy workload does not help for caching indoes.
+To show how the file system behaves under high metadata load we use a
+create-heavy workload. Create-heavy workloads are studied the most in HPC
+research because of the checkpoint/restart use case but they also happen to
+stress the underyling storage system the most.
+Figure~\ref{fig:creates-motivation} shows the resource utilization of compiling
+the Linux kernel.  The untar phase, which is characterized by many creates, has
+the highest resource usage which suggests that it is stressing the consistency
+and journalling subsystems of the metadata server the most. Traditional file
+system techniques for improving performance, such as caching inodes, do not
+help for create-heavy workloads.
 
-In this section, we quantify the costs of consistency and fault tolerance in
-CephFS. If the components that ensure these semantics (i.e. capabilities and
-journals, respectively) can be mitigated or delayed, then global namespaces
-perform as well as decoupled namespaces. We run our experiments on a 9 OSD, 3
-MDS, 1 MON Ceph cluster. The clients use the Ceph kernel client, which has been
-in the mainline Linux kernel since TODO. We use the kernel client so that we
-can find the true create speed of the server; our experiments show a low CPU
-utilziation for the clients which indicates that we are stressing the servers
-more. We also turn caching off becuase, as shown in figureX there is little
-difference, in terms of performance between caching and not caching when using
-the kernel client.
+In this section, we quantify the costs of strong consistency and global
+durability in CephFS.  We use the kernel client so that we can find the true
+create speed of the server; our experiments show a low CPU utilziation for the
+clients which indicates that we are stressing the servers more.
 
-\subsection{Fault Tolerance}
+\subsection{Durability}
 \label{sec:fault-tolerance}
-
-\begin{figure*}[t]
-  \centering
-  \begin{subfigure}[b]{.3\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
-      \caption{}
-      \label{fig:phy-design}
-  \end{subfigure}
-  \begin{subfigure}[b]{.3\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/behavior-journal.png}
-      \caption{}
-      \label{fig:batching}
-  \end{subfigure}
-  \begin{subfigure}[b]{.3\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/slowdown-cache.png}
-      \caption{}
-      \label{fig:batching-outlier}
-  \end{subfigure}
-  \caption{(a) relative performance differences are different after storage
-  software upgrade. (b) total throughput with and without batching. (c)
-  identifying and handling an outlier maintains the benefits
-  of batching without the performance degradation of unnecessarily large I/O
-  requests.}
-\end{figure*}
-
-\begin{figure}[tb] \centering
-\includegraphics[width=1\linewidth]{./graphs/behavior-cache.png} 
-\caption{The inode cache improves metadata read performance but for our
-create-heavy workload it is only an overhead. Most of the time maintaining the
-cache is spent evicting and adding inodes.}\label{fig:inode-cache}
-\end{figure}
 
@@
 \end{figure}
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
+      \caption{Journal Segment Size} \label{fig:slowdown-journal-a}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/behavior-journal.png}
+      \caption{180MB Segment vs. No Journal}
+      \label{fig:slowdown-journal-b}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/slowdown-cache.png}
+      \caption{Inode Cache}
+      \label{fig:slowdown-journal-c}
+  \end{subfigure}
+  \caption{The overhead of the file system metadata journal. The segment size
+  is the threshold that the metadata server starts
+  trimming.\label{fig:slowdown-journal}}
+\end{figure*}
 
-%\begin{figure*}[tb]%h
-%\centering
-%\includegraphics[width=180mm]{figures/throughput-journal.png}
-%\caption{Performance improves with larger journal segments because the metadata
-%server spends less time flushing the journal }\label{fig:throughput-journal}
-%\end{figure*}
+% what is durability
+While durability is not specified by POSIX, users expect that their metadata is
+safe. We define three types of durability: global, local, and none.  Global
+durability means that the client or server can fail at any time and metadata
+will not be lost. Local durability means that metadata can be lost if the
+client or server fails and stays failed. None means that metadata is volatile
+and that the system provides no guarantees when clients or servers fail.\\
 
-%\begin{figure}[tb]%h
-%\centering
-%\includegraphics[width=80mm]{figures/throughput-journal-old.png}
-%\caption{(a), (b), and (c) show increased activity halfway though the run; (d)
-%shows that this activity is related to disk IO, indicating that the activity is
-%the journal.  Despite the activity, we conclude that RADOS itself is not the
-%bottleneck because our RADOS cluster can sustain 100MB/s write bandwidth.
-%}\label{fig:throughput-journal-old}
-%\end{figure}
-%Figure~\ref{fig:throughput-cache} shows that journaling metadata updates into
-%the object store has an overhead.  (a) shows that throughput degrades halfway
-%through the run; (b) shows that cpu utilization spikes at this same point; (c)
-%shows that outgoing network utilization spikes; and (d) shows that this
-%activity spikes on the disks in RADOS.
+% - sequential IO, trim redundant operations
+\noindent\textbf{CephFS Design}: a journal of metadata updates that streams into
+the resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
+WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal can grow to large sizes
+ensuring (1) sequential writes into RADOS and (2) the ability for daemons to
+trim redundant or irrelevent journal entries.  The journal is striped over
+objects where multiple journal updates can reside on the same object. There are
+two tunables for controlling the journal: the segment size and the number of
+parallel segments that can be written in parallel. The former is memory bound
+as larger segments take up more memory but can reduce the time spent journaling
+and the latter is CPU bound.
 
 % purpose of the journal
-As shown in Figure~\ref{fig:journal} CephFS uses RADOS (1) as a metadata store
-for all information about files including the hierarchical namespace and (2) as
-a staging area for the journal of updates before they are applied to the
-metadata store. (1) is essentially a cache that improves performance by serving
-metadata from memory and (2) is the mechanisms for achieving fault tolerance.
-
-% - sequential IO, trim redundant operations
-Fault tolerance means that the client or server can fail and metadata will not
-be lost.  CephFS addresses fault tolerance with a metadata journal that streams
-into the resilient object store. Similar to LFS~\cite{} and WAFL~\cite{} the
-metadata journal can grow to large sizes ensuring (1) sequential writes into
-RADOS and (2) the ability for daemons to trim redundant or irrelevent journal
-entries.
-
-% Knobs in CephFS
-The journal is striped over objects where multiple journal updates can reside
-on the same object. There are two tunables for controlling the journal: the
-segment size and the number of parallel segments that can be written in
-parallel. The former is memory bound as larger segments take up more memory but
-can reduce the time spent journaling and the latter is CPU bound.
+As shown in Figure~\ref{fig:journal}, in addition to the metadata journal,
+CephFS also represents metadata in RADOS as a metadata store, where directories
+and their file inodes are stored as objects.  The metadata server applies the
+updates in the journal to the metadata store when the journal reaches a certain
+size. The metadata store is optimized for recovery ({\it i.e.} reading) while
+the metadata journal is write-optimized.
 
 % Effects on performance
-Figure~\ref{fig:throughput-journal} shows that journaling metadata updates into
-the object store has an overhead. Part (a) shows the runtime for different
-journal segment sizes; the larger the segment size the bigger that the writes
-into the object store are. The trade-off comes is in terms of memory because
-larger segment sizes take up more space with their buffers. Parts (b), (c), and
-(d) show the throughput over time for different segment sizes. Performance
-suffers when time is spent journaling.
+Figure~\ref{fig:slowdown-journal} shows that journaling metadata updates into
+the object store has an overhead. Figure~\ref{fig:slowdown-journal-a} shows the
+effect of journaling of different journal segment sizes; the larger the segment
+size the bigger that the writes into the object store are. The trade-off for
+better performance is memory consumption because larger segment sizes take up
+more space with their buffers. Figure~\ref{fig:slowdown-journal-b} shows how
+the metadata server periodically stops serving requests to flush ({\it i.e.}
+apply journal updates to) the metadata store.  The journal overhead is
+sufficient enough to slow down metadata throughput but not so much as to
+overwhelm the bandidth of the object store. We measured our peak bandwidth to
+be 100MB/s, which is the speed of our network link.\\
 
-% Conclusion
-Despite this overhead, we posit that the journal is sufficient to slow down
-metadata throughput but not so much as to overwhelm RADOS because we measured
-our peak bandwidth to be 100MB/s, which is the speed of our network link.
-
-\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS, as far as
-we can tell, when a client or server fails there is no recovery scheme. For
-BatchFS, if a client fails when it is writing to the local log-structured
-merged tree (implemented as an SSTable) then those batched metadata operations
-on lost. For DeltaFS, if the client fails then on restart the computation does
-the work again -- since the snapshots of the namespace are never globally
-consistent, there is no ground truth the requires the failed namespace to
-answer to anyone. On the server side, BatchFS and DeltaFS use IndexFS. Again,
-IndexFS writes metadata to SSTables but it is not clear whether they ever
-vacate memory, get written to disk, or are flushed to the object store.
+\noindent\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS,
+as far as we can tell, when a client or server fails there is no recovery
+scheme. For BatchFS, if a client fails when it is writing to the local
+log-structured merged tree (implemented as an SSTable) then those batched
+metadata operations are lost. For DeltaFS, if the client fails then on restart
+the computation does the work again -- since the snapshots of the namespace are
+never globally consistent and there is no ground truth.  On the server side,
+BatchFS and DeltaFS use IndexFS. Again, IndexFS writes metadata to SSTables but
+it is not clear whether they ever vacate memory, get written to disk, or are
+flushed to the object store.
 
 \subsection{Strong Consistency} 
+\label{sec:strong-consistency}
 
-Access to POSIX metadata is strongly consistent, so reads and writes are
-globally ordered. The synchronization and serialization machinery needed to
-ensure that all clients see the same state has high overhead.  CephFS uses
-capabilities to keep metadata strongly consistent. To reduce the number of RPCs
-needed for consistency, clients can obtain capabilities for reading, reading
-and updating, reads caching, writing, buffering writes, changing the file size,
-and performing lazy IO.
+Access to metadata in a POSIX-compliant file system is strongly consistent, so
+reads and writes are globally ordered.  The synchronization and serialization
+machinery needed to ensure that all clients see the same state has high
+overhead.\\
+
+\noindent\textbf{CephFS Design}: capabilities keep metadata strongly
+consistent. To reduce the number of RPCs needed for consistency, clients can
+obtain capabilities for reading, reading and updating, caching reads, writing,
+buffering writes, changing the file size, and performing lazy IO.
 
@@
       \includegraphics[width=1.0\linewidth]{graphs/behavior-interfere.png}
-      \caption{}
-      \label{fig:phy-design}
+      \caption{Inteference forces \texttt{lookup()}s}
+      \label{fig:interfere-a}
   \end{subfigure}
@@
       \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere.png}
-      \caption{}
-      \label{fig:batching}
+      \caption{Interference hurts performance}
+      \label{fig:interfere-b}
   \end{subfigure}
@@
       \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere-types.png}
-      \caption{}
-      \label{fig:batching-outlier}
+      \caption{Different interferring requests}
+      \label{fig:interfere-c}
   \end{subfigure}
@@
   directory, the directory inode capability is revoked forcing all clients to
-  centralize lookups at the metadata server.  An underloaded metadata server
-  adaquately services conflicting clients; the create speeds are similar while
-  the interferring operations are limited by the cost of RPCs.  Scaling clients
-  shows increased variability when another client interferes; zooming in on runs
-  with 7 clients we see that different types of interferring operations have
-  different effects on performance variability and predictability.
-  \label{fig:throughput-droplease}}
+  centralize lookups at the metadata server.  
+  \label{fig:interfere}}
 \end{figure*}
 
-% purpose of the inode ccahe - reduces RPCs (lookups for create, readdirs for
-% stats)
+% inode cache - reduces RPCs (lookups for create, readdirs for stats)
 To keep track of the read caching and write buffering capabilities, the clients
 and metadata servers agree on the state of each inode using an inode cache.  If
-a client has the directory inode cached it can do metadata writes (e.g.,
+a client has the directory inode cached it can do metadata writes ({\it e.g.},
 create) with a single RPC. If the client is not caching the directory inode
-then it must do multiple RPCs to the metadata server to (1) determine if the
-file exists and (2) do the actual create.  Unless the client immediately reads
-all the inodes in the cache, the inode cache is less useful for create-heavy
-workloads because the cached inodes are unused.
+then it must do an extra RPC to determine if the file exists.  Unless the
+client immediately reads all the inodes in the cache ({\it i.e.} \texttt{ls
+-alR}), the inode cache is less useful for create-heavy workloads.
 
-% benefits
+% benefits PROBLEM -- IS THIS THE METADATA PROTOCOL OR JUST THE OVERLOADEDNESS?
 The benefits of caching the directory inode when creating files is shown in
-Figure~\ref{fig:throughput-droplease}(a).  If only one client is creating files
-in a directory (``isolated" curve) then that client can lookup the existence of
-new files locally before issuing a create request to the metadata server. If
-another client starts creating files in the same directory (``interfere" curve)
-then the directory inode transitions out of read caching and the first client
-must send lookups to the metadata server. When other clients interfere the
-request throughput is higher Figure~\ref{fig:throughput-droplease}(b) but the
-runtime is slower because the isolated client scenario incurrs less requests.
-Figures~\ref{fig:throughput-droplease}(c) and (d) plot the creates and lookups,
-repsectively, over time; creates slow down and lookups dominate the request
-load when the directory inode is shared in the interferring client scenario.
+Figure~\ref{fig:interfere-a}.  If only one client is creating files in a
+directory (``isolated" curve on \(y1\) axis) then that client can lookup the
+existence of new files locally before issuing a create request to the metadata
+server. If another client starts creating files in the same directory
+(``interfere" curve on \(y1\) axis) then the directory inode transitions out of
+read caching and the first client must send \texttt{lookup()}s to the metadata
+server (``interfere" curve on \(y2\) axis). These extra requests increase the
+throughput of the ``interfere" curve because the metadata server can handle the
+extra load but the overall performance decreases.  This degradation is shown in
+Figure~\ref{fig:interfere-b}, where we scale the number of clients and show
+increased runtime and variability. The performance is compared to a single
+isolated client and he error bar is the standard deviation of the runtime of
+all clients.  For the ``interfere" bars, each client creates files in private
+directories and at 30 seconds we launch another process that creates files in
+those directories.  For less than 7 clients, the runtime and deviation is worse
+when clients interfere. At 7 clients, the metadata server is overloaded so the
+directory caching in the ``isolated" bars has no benefit. Zooming in on runs
+with 7 clients in Figure~\ref{fig:interfere-c} we see how different
+interferring operations affect performance, again when compared to the
+performance of a single isolate client.  ``create" has little effect but
+``stat" and ``create many" cause noticeable slowdowns because of the extra work
+they impose on the metadata server.\\
 
@@
 % TODO: does CephFS still cache inodes when I turn off caching? Why is still keeping inodes in memory? Gah!
-%\begin{figure}[tb]%h
-%\centering
-%\includegraphics[width=80mm]{figures/throughput-cache.png}
-%\caption{ CephFS tries to maintain 100 thousand inodes in the cache;
-%performance degrades when it cannot evict inodes fast enough to maintain this
-%threshold. \label{fig:throughput-cache}}
-%\end{figure}
 
-% disadvantages
-The drawbacks of the inode cache are shown in
-Figure~\ref{fig:throughput-cache}.  Figure~\ref{fig:throughput-cache}(a) shows
-the throughput in metadata requests per second for a single client creating 200
-thousand files. Until time 950 the throughput is steady at just under 2000
-ops/sec and the CPU utilization is at about 30\%. At time 950 seconds
-throughput degrades which corresponds to more inodes in the cache
-(Figure~\ref{fig:throughput-cache}(c)). The values in
-Figure~\ref{fig:throughput-cache}(c) are:
-
-\begin{itemize}
-  \item inodes: total number of elemnts in the cache
-  \item inodes pinned: 
-  \item inodes pinned tail
-  \item inodes with caps
-  \item inodes top
-  \item inodes bottom 
-\end{itemize}
-
-% how the inode cache works
-CephFS tries to keep the cache at 100 thousand inodes so the degradation in
-performance indicates that the metadata server cannot keep up with the
-workload. Figure~\ref{fig:throughput-cache}(d) shows inodes getting added at
-the same rate as before 950 seconds (ino\(+\)) but the rate at which inodes get
-removed is less stable (ino\(-\)) indicating that cache eviction is taking more
-of the metadata servers time.
-
-\subsection{Consistency Overhead}
-
-Figure~\ref{fig:exp0-underloaded} is a baseline showing that the metadata
-server can service multiple clients when it is underloaded. One client creates
-files in the same directory and another does a touch or stat 15 seconds into
-the run. Figure~\ref{fig:exp0-underloaded}(a) shows the runtime of the client
-doing the creates: "isolated" is when the create client is the only workload in
-the cluster, "interfere stat" is the runtime of the create client when another
-client does a stat, and "interfere touch" is the runtime of the create client
-when another client does a touch. There is only a minor performance
-degradation. Figure~\ref{fig:exp0-underloaded}(b) compares the baseline
-the other client doing the touch or stat
-%
-%and Figure~\ref{fig:exp0-underloaded}
-%has a throughput trace to show behavior.
-%
-%In Figure~\ref{fig:exp0-underloaded}(a), 
-%creates only see a minor performance
-%degradation while the interferring operations in
-%Figure~\ref{fig:exp0-underloaded}(b) show a larger (yet not significant)
-%discrepency. 
-%
-%
-%
-% isolated is when there is no interferrence from client 1.
-To explain Figure~\ref{fig:exp0-underloaded}(a) we use
-Figure~\ref{fig:exp0-behaviour}(b) to show why the runtime of creating 100
-thousand files is unaffected. Compared to the throughput without an
-interferring stat ("isolated" curve) the throughput of the metadata server with
-an interferring stat ("inteferring stat" curve) is higher. This means the
-metadata server is doing more operations suggesting that it can adequately
-handle the demands of both workloads.
-
-To explain Figure~\ref{fig:exp0-underloaded}(b) we use
-Figure~\ref{fig:exp0-underloaded}(a) to show that operations are bounded by the
-number of files; more files incurr more requests since both operations lookup
-every file. The number of files shown in Figure~\ref{fig:exp0-behaviour}(a)
-indicates that the local stat ("client 0" bar) is interacting with less files
-than the remote stat ("client 1" bar). This graph also explains why the
-isolated stats and touches take the longest from the remote clients; out of all
-setups, the isolated operations are doing the most RPCs.  The difference
-between local ("client 0 interfere" bar) and remote ("client 1 interfere" bar)
-in Figure~\ref{fig:exp0-underloaded}(b) is the overhead of doing RPCs.
-Isolated is the slowest because it requests the most files. On average the
-local interferring client ("client 0" bar in
-Figure~\ref{fig:exp0-behaviour}(a)) requests 40 thousand less files than the
-remote interferring call ("client 1" bar in
-Figure~\ref{fig:exp0-behaviour}(a)). This reflects how fast the local client
-can get into the queue of requests, presumably because it bypasses capability
-checks.
-
-From this experiment we make 2 conclusions: (1) the metadata server is not
-overloaded because it can handle both workloads and (2) the metadata server
-needs to get global consistency from only 1 client
-
-\textbf{Comparison to decoupled namespaces}: Decoupled namespaces merge batches
-of metadata operations into the global namespaces when the job completes.  In
-BatchFS the merge is delayed by the application using an API to switch between
-asynchronous to sychronous mode. The merge itself is explicitly managed by the
-application but future work looks at more automated methdologies. In DeltaFS
-snapshots of the metadata subtrees stays on the client machines; there is no
-ground truth and consistent namespaces are constructed and resolved at
-application read time or when a 3rd party system (e.g., middleware, scheduler,
-etc.) needs a view of the metadata.
+\noindent\textbf{Comparison to decoupled namespaces}: Decoupled namespaces
+merge batches of metadata operations into the global namespaces when the job
+completes.  In BatchFS the merge is delayed by the application using an API to
+switch between asynchronous to sychronous mode. The merge itself is explicitly
+managed by the application but future work looks at more automated
+methdologies. In DeltaFS snapshots of the metadata subtrees stays on the client
+machines; there is no ground truth and consistent namespaces are constructed
+and resolved at application read time or when a 3rd party system ({\it e.g.},
+middleware, scheduler, etc.) needs a view of the metadata. As a result, all the
+overheads of maintaining consistency that we showed above are delayed until the
+merge phase.

### 2017-03-27 paper/sections/4_methodology.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: write first half of results section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \section{Methodology: Decoupled Namespaces}
+\label{sec:methodology-decoupled-namespaces}
 
@@
 \begin{tabular}{ l | l | l | l }
-  Consistency \(\rightarrow\) &&& \\  
-  Durability \(\downarrow\)  & invisible         & eventual        & strong  \\\hline
+  C \(\rightarrow\) &&& \\  
+  D \(\downarrow\)  & invisible         & eventual        & strong  \\\hline
   none                       & create            & create          & RPCs    \\

### 2017-03-27 paper/sections/6_results.tex
paper: write first half of results section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 experiments presented here are available in the repository for this
-article\footnote{\url{https://github.com/michaelsevilla/cudele-popper/tree/}}.
+article\footnote{\url{https://github.com/michaelsevilla/cudele-popper/}}.
 Experiments can be examined in more detail, or even re-run, by visiting the
 \texttt{[source]} link next to each figure. That link points to a Jupyter
-notebook that shows the analysis and source code for that graph and its parent
-folder contains all the associated artifacts.
+notebook that shows the analysis and source code for that graph, which points
+to an experiment and its artifacts.
 
-\subsection{Microbenchmarks}
 %\begin{figure*}[tb]
@@
 
+\subsection{Cudele Mechanism Performance}
+
 \begin{figure}[tb]
@@
 \includegraphics[width=1.0\linewidth]{graphs/slowdown-mechanisms.png}
-\caption{}
+\caption{
+[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
+The performance of the Cudele mechanisms normalized to the runtime of the
+create mechanism. The runtime of the create mechanism is the time it takes to
+write 100 file creates to the client's in-memory journal of metadata
+updates.\label{fig:slowdown-mechanisms}}
 \end{figure}
+
 \begin{figure}[tb]
 \centering
-\includegraphics[width=1.0\linewidth]{graphs/slowdown-related-work.png}
-\caption{}
+\includegraphics[width=1.0\linewidth]{graphs/behavior-journal-size.png}
+\caption{
+[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
+In-memory client journal scales with the number of
+updates.\label{fig:behavior-journal-size}}
 \end{figure}
+
+Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the Cudele
+mechanisms, normalized to the time it takes to write 100K file create updates
+to the client's in-memory journal ({\it i.e.} the create mechanism). Bars above
+\(10^0\) are slower than the create mechanism and bars below are faster.
+``Stream" is an approximation of the overhead and is calculated by subtracting
+the runtime of the job with the journal turned off from the runtime with the
+journal turned on. All the slowdowns and speedups reported are for the 100K
+file create job, the largest workoad we tested.
+
+% RPCs vs. apply: calls to metadata server vs. RADOS
+\subsubsection{Overhead of RPCs} ``RPCs" is \(66\times\) slower than ``volatile
+apply" because sending individual metadata updates over the network is costly.
+While ``RPCs" sends a request for every file create, ``nonvolatile apply"
+writes all the updates to the in-memory journal and applies them to the
+in-memory data structures in the metadata server. While communicating the
+decoupled namespace directly to the metadata server is faster, communicating
+through the object store (``nonvolatile apply") is \(10\times\) slower.
+
+% TODO: why is apply so slow.  
+% apply: no CephFS changes, pulls/pushes same RADOS obj.
+% v_apply vs. apply/persist: communicating through RADOS
+\subsubsection{Overhead of ``nonvolatile apply"} The cost of ``nonvolatile
+apply" is much larger than all the other mechanisms.  That mechanism was not
+implemented as part of Cudele -- it was a debugging and recovery tool packaged
+with CephFS. It works by iterating over the updates in the journal and pulling
+all objects that {\it may} be affected by the update.  This means that two
+objects are repeatedly pulled, updated, and pushed: the object that houses the
+experiment directory and the object that contains the root directory ({\it
+i.e.} \texttt{/}).  The cost of communcating through the object store is shown
+by comparing the runtime of ``volatile apply" + ``global persist" to
+``nonvolatile apply". These two operations end up with the same final metadata
+state but using ``nonvolatile apply" is clearly inferior.
+
+% persist vs. save: one disk vs. many
+\subsubsection{Parallelism of the Object Store} Comparing ``local" and ``global
+persist" demonstrates the bandwidth advantages of storing the journal in a
+distributed object store. As the journal size increases, the ``global persist"
+performance is \(1.5\times\) faster because the object store is leveraging the
+collective bandwidth of the disks in the cluster. This benefit comes from the
+object store itself but should be acknowledged when making decisions for the
+application; the size of the object store can help mitigate the overheads of
+globally persisting metadata updates.
+
+\subsubsection{Journal Size} Figure~\ref{fig:behavior-journal-size} shows the
+amount of storage per journal update (\(y\) axis) for the range of file creates
+we tested (\(x\) axis). The increase in file size is linear with the number of
+metadata creates and suggests that updates for a million files would be
+\(2.5\text{KB}*1\text{ million files} = 2.38\text{GB}\). Transfer times for
+files this large on an HPC network are reasonable.
+
+\subsection{Eventual vs. Strong Consistency}
 \begin{figure}[tb]
@@
 \includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-eventual.png}
-\caption{}
+\caption{The RPC per metadata update of ``Strong Consistency" has a large
+overhead compared to the decoupled namespace strategy of ``Eventual
+Consistency".\label{fig:slowdown-strong-eventual}}
 \end{figure}
 
-\subsubsection{Per phase latencies}
+Figure~\ref{fig:slowdown-strong-eventual} shows the runtimes of eventual and
+strong consistency based systems implemented on Cudele, normalized to the
+runtime of the create mechanism (again, just creating files in the client's
+in-memory journal). We scaled the number of files up to 100K which is the
+maximum size of a directory by default in CephFS. We use the following
+compositions from the mechanisms in Table~\ref{table:spectrum}:
+
+\begin{itemize}
+  \item Strong Consistency
+  \begin{itemize}
+    \item[]  \texttt{RPCs} \(+\) \texttt{stream}
+  \end{itemize}
+  \item Strong Consistency, No Fault Tolerance
+  \begin{itemize}
+    \item[] \texttt{RPCs}
+  \end{itemize}
+  \item Eventual Consistency
+  \begin{itemize}
+    \item[] \texttt{creates} \(+\) \texttt{local persist}
+  \end{itemize}
+  \item Eventual Consistency, No Fault Tolerance
+  \begin{itemize}
+    \item[] \texttt{creates} \(+\) \texttt{local persist} \(+\) \texttt{volatile apply}
+  \end{itemize}
+\end{itemize}
+
+% Final file system metadata states are equivalent No guarantees while
+% transitioning mechanisms
+We compare these semantics because the final metadata states are equivalent.
+Cudele makes no guarantees during execution of the mechanisms or when
+transitioning semantics -- the semantics are guaranteed {\it once the mechanism
+completes}. So if servers fail during a mechanism, metadata or data may be
+lost.
+
+% DeltaFS vs. BatchFS vs. POSIX: decoupled is faster
+% (5-7x) vs. (90-104x) slower
+% Does not scale?
+% Strong Consistency > 10X slower than Eventual Consistency 
+\subsubsection{Speedups of Decoupled Namespaces} Eventual consistency uses the
+decoupled namespace strategy and shows up to a \(20\times\) speedup over the
+traditional namespaces that use RPCs. Compared to the baseline the slowdown is
+\(5-7\times\) for Strong Consistency, which emulates BatchFS and
+\(90-104\times\) for Eventual Consistency, which emulates DeltaFS.
+
+% POSIX (no stream) vs. POSIX: cost of durability < consistency
+\subsubsection{Durability \(<<\) Consistency} The \(1.15\times\) overhead of
+``Strong Consistency" compared to ``Strong Consistency, No Fault Tolerance" for
+100K files is neglible. It suggests that the overhead of consistency is much
+larger than the overhead of durability. This conclusion should be stronger as
+we scale the number of files because the cost of streaming the journal into the
+object store is constant. We omit the same analysis for ``Eventual Consistency"
+because the runtimes are so short that the normalized slowdowns are misleading.
+
+% Eventual Consistency benefit not due to metadata format
+\subsubsection{Metadata Formats} Because the metadata formats are the same for
+all schemes we argue that the performance gain for decoupled namespaces comes
+from relaxing the consistency guarantees and not from the metadata formats, as
+was argued in previous work outlining the benefit of
+SSTables~\cite{ren:atc2013-tablefs, ren:sc2014-indexfs}.
 
@@
 \subsection{Partial Directory Listings}
-
-\subsection{Macrobenchmarks}
-updatedb: http://lists.ceph.com/pipermail/ceph-users-ceph.com/2015-July/002768.html

### 2017-03-28 paper/sections/1_introduction.tex
paper: remove fault tolerance and add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=75mm]{figures/subtree-policies.png}
-\caption{Administrators can assign consistency and fault tolerance policies to
+\caption{Administrators can assign consistency and durability policies to
 subtrees to get the benefits of some of the state-of-the-art HPC architectures.

### 2017-03-28 paper/sections/4_methodology.tex
paper: remove fault tolerance and add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 journal, and delay metadata updates.  Table~\ref{table:spectrum} shows how
-these phases (represented by the arrows) can be combined to provide weaker
-consistency or fault tolerance semantics.  }\label{fig:decouple}
+these mechanisms (represented by the arrows) can be combined to provide weaker
+consistency or durability semantics.  }\label{fig:decouple}
 \centering
@@
 Next, we discuss how these mechanisms can be composed to get different
-consistency and fault tolerance semantics.
+consistency and durability semantics.
 
@@
 % describe table
-The spectrum of consistency and fault tolerance guarantees that adminstrators
+The spectrum of consistency and durability guarantees that adminstrators
 can construct is shown in Table~\ref{table:spectrum}. The columns are the
-different consistency semantics and the rows cover the spectrum of fault
-tolerance guantees. For consistency: ``none" means the system does not handle
+different consistency semantics and the rows cover the spectrum of durability
+guarantees. For consistency: ``none" means the system does not handle
 merging updates into a global namespace and it is assumed that middleware or
@@
 done writing; and updates in ``global" consistency are seen immediately by all
-clients. For fault tolerance, ``none" means that updates are volatile and will
+clients. For durability, ``none" means that updates are volatile and will
 be lost on a crash of any component. Stronger guarantees are made with
@@
 systems. POSIX systems like CephFS and IndexFS have global consistency and
-fault tolerance; DeltaFS has consistency and fault tolerance set to ``none";
-and BatchFS uses ``eventual" consistency and ``local" fault tolerance. These
+durability; DeltaFS has consistency and durability set to ``none";
+and BatchFS uses ``eventual" consistency and ``local" durability. These
 are just a few of the HPC examples.
@@
 makes little sense to do \texttt{creates+RPCs} since both steps do the same
-thing or \texttt{stream+save} since global fault tolerance is stronger and has
-more overhead than local fault tolerance.  Formulaically, valid steps
+thing or \texttt{stream+save} since global durability is stronger and has
+more overhead than local durability.  Formulaically, valid steps
 include:\\
@@
 % file type
-To assign consistency and fault tolerance to the subtrees we store policies in
+To assign consistency and durability to the subtrees we store policies in
 the directory inode. This approach uses the File Type interface from the
@@
 
+\subsection{Programmability Approach}
+
+% what is programmability
+A programmable storage system exposes internal subsystem as building blocks for
+higher level services. This `dirty-slate' approach limits redundant code and
+leverages the robustness of the underlying storage system. Cudele uses this
+approach to great success and requires only:
+
+\begin{itemize}
+
+  \item 354 lines of library code
+
+  \item 219 lines of non-destructive metadata server code, which is not used
+  unless it is turned on
+
+  \item 4 lines of destructive client/server code to check whether a namespace
+  is decoupled
+
+\end{itemize}
+
+% how it works in CephFS, how I used it
+\subsubsection{Metadata Store}
+
+\subsubsection{Journal}
+
+\subsubsection{Journal Tool}
+The journal tool is used for disaster recovery and lets administrators view and
+modify the journal of metadata updates; it can read the journal, erase events
+from the journal, and apply the updates in the journal to the metadata store.
+To apply journal updates to the metadata store, the journal tool reads the
+journal segments from object store objects and applies the update to the
+metadata store (which are also stored as object store objects).
+
+% Journal import
+The journal tool imports journals from binary files stored on disk.  First the
+header of the dump is sanity checked and written into RADOS to the ``header"
+object.  The ``header" object has metadata about the journal as well as the
+locations of all the journal pointers (e.g., where the tail of the journal is,
+where we are currently trimming, etc.).  Then the journal events are cleaned
+(erasing trailing events that are not part of the header) and written as
+objects into RADOS.  Note that while the journal is in RADOS, the metadata
+servers do do not have the namespace reconstructed in memory so the metadata
+cluster will not service requests relating to the journal of imported events.
+To construct the namespace in the collective memory of the metadata servers we
+need to first construct the namespace in RADOS. The journal tool can explicitly
+do this by  applying the journal to the metadata store in RAODS. This will pull
+the objects containing journal segments and replay them on the metadata store.
+Finally, we delete the journal in RADOS and restart the metadata servers so
+they rebuild their caches.
+
+% Journal export
+The journal tool exports journals to binary files stored on disk. First the
+journal is scanned for the header and then journal is recovered. To recover the
+journal the ``header" object is read off disk and then objects are probed in
+order and starting from the write position saved in the header. Probing will
+update the write position if it finds objects with data in them.
+
+% Journal export
+When exporting a journal of events, the journal tool first scans the journal to
+check for corruption. Then it recovers the journal by reading the ``header"
+object out of RADOS.  After reading the header, the journal tool can pull
+journal segments from RADOS because it knows how many objects to pull and how
+far to seek within those objects.
+
+%% The data structures
+%The metadata that the event touches, including inodes, paths, and timestamps,
+%are stored as metablobs. Each piece of metadata inside a metablob is called a
+%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
+%table client transactions (tid, version), renamed directory fragments (maps,
+%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
+%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
+%me (and you since you are reading this paper), this does not make any sense.
+%
+%Each directory fragment has an associated directory lump, which is just a bunch
+%of metadata. The most interesting part of the dirlump is the fullbits array,
+%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
+%and then all the full bits, saving off the children and inode locations. The
+%children tell us which dentry names an inode has and the inode locations map
+%the parent inode to its child inode and dentry.
+
+\subsubsection{Inode Cache}
+
+\subsubsection{Large Inodes}
+
+% lead into implementation
+
 % // limit consistency
@@
 %
-% // limit fault tolerance
+% // limit durability
 % if !stream

### 2017-03-28 paper/sections/5_implementation.tex
paper: remove fault tolerance and add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 Figure~\ref{fig:decouple}. This implementation decouples policies from
-mechanisms allowing applciations to choose the consistency and fault tolerance
+mechanisms allowing applciations to choose the consistency and durability
 semantics they need.
@@
 metadata store will never run. Performance numbers for enabling and disabling
-the stream phase are presented back in Section~\S\ref{sec:fault-tolerance}.
+the stream mechanism are presented back in Section~\S\ref{sec:durability}.
 
@@
 % creates
-The journal tool is used for disaster recovery and lets administrators view and
-modify the journal of metadata updates; it can read the journal, erase events
-from the journal, and apply the updates in the journal to the metadata store.
-To apply journal updates to the metadata store, the journal tool reads the
-journal segments from object store objects and applies the update to the
-metadata store (which are also stored as object store objects).  For create,
+For create,
 clients write to an in-memory journal and updates are merged into the global
 namespace by replaying them onto the metadata store in the object store.
-
-% Journal import
-The journal tool imports journals from binary files stored on disk.  First the
-header of the dump is sanity checked and written into RADOS to the ``header"
-object.  The ``header" object has metadata about the journal as well as the
-locations of all the journal pointers (e.g., where the tail of the journal is,
-where we are currently trimming, etc.).  Then the journal events are cleaned
-(erasing trailing events that are not part of the header) and written as
-objects into RADOS.  Note that while the journal is in RADOS, the metadata
-servers do do not have the namespace reconstructed in memory so the metadata
-cluster will not service requests relating to the journal of imported events.
-To construct the namespace in the collective memory of the metadata servers we
-need to first construct the namespace in RADOS. The journal tool can explicitly
-do this by  applying the journal to the metadata store in RAODS. This will pull
-the objects containing journal segments and replay them on the metadata store.
-Finally, we delete the journal in RADOS and restart the metadata servers so
-they rebuild their caches.
-
-% Journal export
-The journal tool exports journals to binary files stored on disk. First the
-journal is scanned for the header and then journal is recovered. To recover the
-journal the ``header" object is read off disk and then objects are probed in
-order and starting from the write position saved in the header. Probing will
-update the write position if it finds objects with data in them.
-
-% Journal export
-When exporting a journal of events, the journal tool first scans the journal to
-check for corruption. Then it recovers the journal by reading the ``header"
-object out of RADOS.  After reading the header, the journal tool can pull
-journal segments from RADOS because it knows how many objects to pull and how
-far to seek within those objects.
-
-%% The data structures
-%The metadata that the event touches, including inodes, paths, and timestamps,
-%are stored as metablobs. Each piece of metadata inside a metablob is called a
-%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
-%table client transactions (tid, version), renamed directory fragments (maps,
-%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
-%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
-%me (and you since you are reading this paper), this does not make any sense.
-%
-%Each directory fragment has an associated directory lump, which is just a bunch
-%of metadata. The most interesting part of the dirlump is the fullbits array,
-%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
-%and then all the full bits, saving off the children and inode locations. The
-%children tell us which dentry names an inode has and the inode locations map
-%the parent inode to its child inode and dentry.
 
@@
 the updates directly to the in-memory namespace maintained by the metadata
-servers. We say volatile because Cudele makes no consistency or fault
-tolerance guarantees. If a concurrent update from a client occurs there is no
+servers. We say volatile because Cudele makes no consistency or durability
+guarantees. If a concurrent update from a client occurs there is no
 rule for resolving conflicts and if the client or metadata server crashes there

### 2017-03-28 paper/sections/6_results.tex
paper: remove fault tolerance and add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 %  - e.g., PLFS: no consistency
-%- fault tolerance: global vs. local
+%- durability: global vs. local
 %  - e.g., CephFS: global
@@
 \subsection{Isolation from EBUSY}
+
+% setup: 2cs sep dirs, 2cs DN, 1c malicious write
+In this section, we show the isolation benefits of subtree policies.  We repeat
+the experiment from Figure~\ref{fig:interfere-b}, where clients write to their
+own private directories and another client interferes at 30 seconds.  We also
+have 2 clients write to decoupled namespaces and merge their updates 90
+seconds. The
+
+\subsubsection{Benefits of Isolated Subtree Policies}
+
+\subsubsection{Cost of Merging}
+\subsubsection{Scaling Concurrent Merges}
+
 \subsection{Isolation from Global Writes, Cost of Overwrites}

### 2017-03-28 paper/sections/3_posix_overheads.tex
paper: remove fault tolerance and add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: cleanup terminology in section 4
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \subsection{Durability}
-\label{sec:fault-tolerance}
+\label{sec:durability}
 
@@
 \caption{CephFS uses a journal to stage updates and tracks dirty metadata in
-the collective memory of the MDSs. Each MDS maintains its own journal, which is
+the collective memory of the metadata servers. Each metadata server maintains its own journal, which is
 broken up into 4MB segments. These segments are pushed into RADOS and deleted

### 2017-03-28 paper/sections/4_methodology.tex
paper: cleanup terminology in section 4
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 the namespace as a tree of directory fragments and is easier to read/traverse.
-On the other hand, the journal represent the namespace as a list of events. It
+The journal represent the namespace as a list of events. It is
 a ``pile system"; writes are fast but reads are slow because state must be
@@
 % describe the mechanisms
-Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Save, and
+Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
 Persist. ``RPCs" does round trip remote procedure calls to establish
@@
 and applies it directly to the in-memory metadata store of the metadata server
-cluster. ``Save" takes the in-memory journal and writes it to the client's
-disk. ``Persist" saves the journal as a an object in the object store from the
+cluster. ``Local Persist" takes the in-memory journal and writes it to the client's
+disk. ``Global Persist" saves the journal as a an object in the object store from the
 client.
@@
 and durability properties are not guaranteed until all mechanisms in the cell
-are complete (i.e. the compositions should be considered atomic) and there are
+are complete ({\it i.e.} the compositions should be considered atomic) and there are
 no guarantees while tranisitioning between policies. \label{table:spectrum}}
@@
 % describe table
-The spectrum of consistency and durability guarantees that adminstrators
-can construct is shown in Table~\ref{table:spectrum}. The columns are the
-different consistency semantics and the rows cover the spectrum of durability
-guarantees. For consistency: ``none" means the system does not handle
-merging updates into a global namespace and it is assumed that middleware or
-the application manages consistency lazily; ``eventual" merges updates either
-when the system has time (e.g., by a background daemon) or when the client is
-done writing; and updates in ``global" consistency are seen immediately by all
-clients. For durability, ``none" means that updates are volatile and will
-be lost on a crash of any component. Stronger guarantees are made with
-``local", which means updates will be retained if the client node recovers, and
-``global", where all updates are always recoverable.
+The spectrum of consistency and durability guarantees that adminstrators can
+construct is shown in Table~\ref{table:spectrum}. The columns are the different
+consistency semantics and the rows cover the spectrum of durability guarantees.
+For consistency: ``invisible" means the system does not handle merging updates
+into a global namespace and it is assumed that middleware or the application
+manages consistency lazily; ``eventual" merges updates at some time in the
+future ({\it e.g.}, when the system has time, when number of updates reaches a
+certain threshold, when the client is done writing, etc.); ; and updates in
+``global" consistency are seen immediately by all clients. For durability,
+``none" means that updates are volatile and will be lost on a failure. Stronger
+guarantees are made with ``local", which means updates will be retained if the
+client node recovers, and ``global", where all updates are always recoverable.
 
 % which system they represent and which are impossible
-The cells in Table~\ref{table:spectrum} encompass many exisiting storage
-systems. POSIX systems like CephFS and IndexFS have global consistency and
-durability; DeltaFS has consistency and durability set to ``none";
-and BatchFS uses ``eventual" consistency and ``local" durability. These
-are just a few of the HPC examples.
+Existing, state-of-the-art systems in HPC can be represented by the cells in
+Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
+have global consistency and durability; DeltaFS uses ``invisible" consistency
+and ``local" durability and BatchFS uses ``eventual" consistency and ``local"
+durability. These are just a few of the HPC examples.
 
@@
 Although we can achieve all permutations of the different guarantees in
-Table~\ref{table:spectrum}, not all of them make much sense. For example, it
+Table~\ref{table:spectrum}, not all of them make sense. For example, it
 makes little sense to do \texttt{creates+RPCs} since both steps do the same
-thing or \texttt{stream+save} since global durability is stronger and has
-more overhead than local durability.  Formulaically, valid steps
+thing or \texttt{stream+save} since ``global" durability is stronger and has
+more overhead than ``local" durability.  Formulaically, valid steps
 include:\\
@@
 \subsection{Cudele Namespace API:\\Per-Subtree Policies}
-
-% file type
-To assign consistency and durability to the subtrees we store policies in
-the directory inode. This approach uses the File Type interface from the
-Malacology programmable store system~\cite{} and it tells clients how to access
-the underlying data or metadata. The underlying implementation stores
-executable code in the inode that calls the different Cudele mechanisms. Of
-course, there are many security and access control aspects of this approach but
-that is beyond the scope of this paper.
+\label{sec:cudele-namespace-api-per-subtree-policies}
 
@@
 The interface for setting the subtree policies is with \{path,
-<block|overwrite>, pre-allocated inodes\} tuples. For example:
+\(<\)block\(|\)overwrite\(>\), pre-allocated inodes\} tuples. For example:
 
@@
 would give the application 100 inodes but the subtree would behave like the
-existing CephFS implementation. Table~\ref{} shows how one would use the Cudele
+existing CephFS implementation. Table~\ref{table:spectrum} shows how one would use the Cudele
 API to implement policies from related work.
@@
 
-For \texttt{block}, any requests to this part of the namespace return with
-``Device is busy", which will spare the MDS from wasting resources for updates
+For \texttt{block}, any requests to this part of the namespace returns with
+``Device is busy", which will spare the metadata server from wasting resources for updates
 that may get overwritten. If the application does not mind losing updates, for
@@
 can select \texttt{overwrite}. In this case, metadata will be written and the
-computation from the decoupled namespace will take priority because the resutls
+computation from the decoupled namespace will take priority at merge time because the results
 are more accurate.
@@
 leverages the robustness of the underlying storage system. Cudele uses this
-approach to great success and requires only:
+approach and re-uses some of the building blocks from the Malacology
+programmable storage system~\ref{sevilla:eurosys17} to great success and
+requires only:
 
@@
 
+% file type
+To assign consistency and durability to the subtrees we store policies in
+the directory inode. This approach uses the File Type interface from the
+Malacology programmable store system~\cite{sevilla:eurosys17} and it tells clients how to access
+the underlying data or metadata. The underlying implementation stores
+executable code in the inode that calls the different Cudele mechanisms. Of
+course, there are many security and access control aspects of this approach but
+that is beyond the scope of this paper.
+
 % lead into implementation

### 2017-03-28 paper/sections/4_methodology.tex
paper: add programmability section
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 which entity they are performed by (gray boxes). The metadata store and journal
-are different ways of representing the namespace. The metadata store represents
-the namespace as a tree of directory fragments and is easier to read/traverse.
-The journal represent the namespace as a list of events. It is
-a ``pile system"; writes are fast but reads are slow because state must be
-reconstructed.  Specifically, reads are slow because there is more data to
-read, it is unorganized, and many of the updates may be redundant.
-
-% describe the mechanisms
-Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
+are different ways of representing the namespace.  Cudele presents 6
+mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
 Persist. ``RPCs" does round trip remote procedure calls to establish
@@
 and applies it directly to the in-memory metadata store of the metadata server
-cluster. ``Local Persist" takes the in-memory journal and writes it to the client's
-disk. ``Global Persist" saves the journal as a an object in the object store from the
-client.
-
-Next, we discuss how these mechanisms can be composed to get different
-consistency and durability semantics.
+cluster. ``Local Persist" takes the in-memory journal and writes it to the
+client's disk. ``Global Persist" saves the journal as a an object in the object
+store from the client.  Next, we discuss how these mechanisms can be composed
+to get different consistency and durability semantics.
 
@@
 approach and re-uses some of the building blocks from the Malacology
-programmable storage system~\ref{sevilla:eurosys17} to great success and
+programmable storage system~\cite{sevilla:eurosys17} to great success and
 requires only:
@@
 \subsubsection{Metadata Store}
+\label{sec:metadata-store}
+
+% how it works in CephFS
+In CephFS, the metadata store is a data structure that represents the file
+system namespace. This data structure is stored in two places: in memory ({\it
+i.e.} in the collective memory of the metadata server cluster) and as objects
+in the object store. In the object store, directories and their inodes are
+stored together in objects to improve the performance of scans.  The metadata
+store data structure is structured as a tree of directory fragments making it
+easier to read and traverse.\\
+
+% how i used it
+\noindent\textbf{Cudele}: uses the metadata store format to write objects
+formatted like the journal into the object store. By writing with the same
+format, the metadata servers can read and use the recovery code to materialize
+the updates from a client's decoupled namespace ({\it i.e.} merge).
 
 \subsubsection{Journal}
+\label{sec:journal}
+
+The journal is the second way that CephFS represents the file system namespace;
+it is a log of metadata updates that can materialize the namespace when the
+updates are replayed onto the metadata store. The journal is a ``pile system";
+writes are fast but reads are slow because state must be reconstructed.
+Specifically, reads are slow because there is more state to read, it is
+unorganized, and many of the updates may be redundant.\\
+
+\noindent\textbf{Cudele}: uses the journal subsystems within the metadata
+server to replay updates onto the in-memory metadata store.  When the clients
+are ready to merge their updates back into the global namespace, they pass a
+binary file of  metadata updates to the metadata server. The metadata server
+uses its recovery code to ``replay" the updates.
 
 \subsubsection{Journal Tool}
+\label{sec:journal-tool}
+
 The journal tool is used for disaster recovery and lets administrators view and
-modify the journal of metadata updates; it can read the journal, erase events
-from the journal, and apply the updates in the journal to the metadata store.
-To apply journal updates to the metadata store, the journal tool reads the
-journal segments from object store objects and applies the update to the
-metadata store (which are also stored as object store objects).
+modify the journal. It can read the journal, export the journal as a file,
+erase events, and apply updates to the metadata store.  To apply journal
+updates to the metadata store, the journal tool reads the journal from object
+store objects and replays the updates on the metadata store in the object
+store.\\
+
+\noindent\textbf{Cudele}: using the journal tool, clients can materialize the
+journal in memory and append events to their own copy. When they are done they
+export their journal to a binary file. Clients also use the journal tool to
+save their updates to disk or to the object store.
 
 % Journal import
-The journal tool imports journals from binary files stored on disk.  First the
-header of the dump is sanity checked and written into RADOS to the ``header"
-object.  The ``header" object has metadata about the journal as well as the
-locations of all the journal pointers (e.g., where the tail of the journal is,
-where we are currently trimming, etc.).  Then the journal events are cleaned
-(erasing trailing events that are not part of the header) and written as
-objects into RADOS.  Note that while the journal is in RADOS, the metadata
-servers do do not have the namespace reconstructed in memory so the metadata
-cluster will not service requests relating to the journal of imported events.
-To construct the namespace in the collective memory of the metadata servers we
-need to first construct the namespace in RADOS. The journal tool can explicitly
-do this by  applying the journal to the metadata store in RAODS. This will pull
-the objects containing journal segments and replay them on the metadata store.
-Finally, we delete the journal in RADOS and restart the metadata servers so
-they rebuild their caches.
+%The journal tool imports journals from binary files stored on disk.  First the
+%header of the dump is sanity checked and written into RADOS to the ``header"
+%object.  The ``header" object has metadata about the journal as well as the
+%locations of all the journal pointers (e.g., where the tail of the journal is,
+%where we are currently trimming, etc.).  Then the journal events are cleaned
+%(erasing trailing events that are not part of the header) and written as
+%objects into RADOS.  Note that while the journal is in RADOS, the metadata
+%servers do do not have the namespace reconstructed in memory so the metadata
+%cluster will not service requests relating to the journal of imported events.
+%To construct the namespace in the collective memory of the metadata servers we
+%need to first construct the namespace in RADOS. The journal tool can explicitly
+%do this by  applying the journal to the metadata store in RAODS. This will pull
+%the objects containing journal segments and replay them on the metadata store.
+%Finally, we delete the journal in RADOS and restart the metadata servers so
+%they rebuild their caches.
 
 % Journal export
-The journal tool exports journals to binary files stored on disk. First the
-journal is scanned for the header and then journal is recovered. To recover the
-journal the ``header" object is read off disk and then objects are probed in
-order and starting from the write position saved in the header. Probing will
-update the write position if it finds objects with data in them.
+%First the
+%journal is scanned for the header and then journal is recovered. To recover the
+%journal the ``header" object is read off disk and then objects are probed in
+%order and starting from the write position saved in the header. Probing will
+%update the write position if it finds objects with data in them.
 
 % Journal export
-When exporting a journal of events, the journal tool first scans the journal to
-check for corruption. Then it recovers the journal by reading the ``header"
-object out of RADOS.  After reading the header, the journal tool can pull
-journal segments from RADOS because it knows how many objects to pull and how
-far to seek within those objects.
+%When exporting a journal of events, the journal tool first scans the journal to
+%check for corruption. Then it recovers the journal by reading the ``header"
+%object out of RADOS.  After reading the header, the journal tool can pull
+%journal segments from RADOS because it knows how many objects to pull and how
+%far to seek within those objects.
 
@@
 \subsubsection{Inode Cache}
+\label{sec:inode-cache}
+
+In CephFS, the inode cache reduces the number of RPCs between clients and
+metadata servers. Without contention clients can resolve metadata reads locally
+thus reducing operations like \texttt{lookup()}. The size of the inode cache is
+configurable so as not to saturate the memory on the metadata server -- inodes
+in CephFS are about 1400
+bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
+The inode cache has code for manipulating inode numbers, such as pre-allocating
+inodes to clients.\\
+
+\noindent\textbf{Cudele}: uses the internal inode cache code to allocate inodes
+to clients that decouple parts of the namespace and to skip inodes used by the
+client at merge time.
 
 \subsubsection{Large Inodes}
+\label{sec:large-inodes}
 
-% file type
-To assign consistency and durability to the subtrees we store policies in
-the directory inode. This approach uses the File Type interface from the
-Malacology programmable store system~\cite{sevilla:eurosys17} and it tells clients how to access
-the underlying data or metadata. The underlying implementation stores
-executable code in the inode that calls the different Cudele mechanisms. Of
-course, there are many security and access control aspects of this approach but
-that is beyond the scope of this paper.
+In CephFS, inodes store policies for policies, like how the file is striped
+across the object store or for managing subtrees for load balancing. These
+policies adhere to logical partitionings of metadata or data, like Ceph pools and file
+system namespace subtrees. To implement this, the namespace data structure has
+the ability to recursively apply policies to subtrees and to isolate subtrees
+from each other.  \\
+
+\noindent\textbf{Cudele}: uses the large inodes to store consistency and
+durability policies in the directory inode. This approach uses the File Type
+interface from the Malacology programmable store
+system~\cite{sevilla:eurosys17} and it tells clients how to access the
+underlying metadata. The underlying implementation stores executable code in
+the inode that calls the different Cudele mechanisms. Of course, there are many
+security and access control aspects of this approach but that is beyond the
+scope of this paper.
 

### 2017-03-28 paper/sections/5_implementation.tex
paper: cleanup terminology in section 4
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: prune implementation section and make it readable
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % Outline of the section
-Each section below corresponds ot the labeled arrows in
+Each section below corresponds to the labeled arrows in
 Figure~\ref{fig:decouple}. This implementation decouples policies from
-mechanisms allowing applciations to choose the consistency and durability
+mechanisms allowing applications to choose the consistency and durability
 semantics they need.
@@
 Of the 6 mechanisms in Figure~\ref{fig:decouple} 4 had to implemented and only
-1 required changes to the underlying storage system itself.  RPCs and stream
-can be achieved with tunables in configuration files for the storage system
-(e.g., metadata cache size, logging on/off, etc.).  Persist", save, and create
-are implemented as library and does not require modifications to the storage
-system. Volatile apply requires changes to the metadata server to inject
-updates back into the global namespace.
+1 required changes to the underlying storage system itself.  ``RPCs" and
+``Stream" can be achieved with configuration settings 
+.  ``Global Persist", ``Local Persist",
+``Nonvolatile Apply", and ``Create" are implemented as a library and they do
+not require modifications to the storage system. ``Volatile Apply" requires
+changes to the metadata server to inject updates back into the global
+namespace.
 
 \subsection{No Changes: RPCs, Stream}
+\label{sec:no-changes-rpcs-stream}
 
-RPCs is the default behavior of the storage system. Depending on the cache
-state and configuration each operation incurrs at least RPC. For example, if a
-client or metadata server is not caching the directory inode, all creates
-within that directory will result in a lookup and a create request. If the
-directory inode is cached then only the create needs to be sent.
-
-Re-using the debugging and testing features in the storage system, we can turn
-stream on and off. If it is off, then the metadata servers will not save
+RPCs is the default behavior of CephFS.  Using existing configuration settings
+({\it e.g.}, metadata cache size, logging on/off, etc.) in Ceph we can turn
+``Stream" on and off.  If it is off, then the metadata servers will not save
 journals in the object store and the daemons that apply the journal to the
-metadata store will never run. Performance numbers for enabling and disabling
-the stream mechanism are presented back in Section~\S\ref{sec:durability}.
+metadata store will never run.
 
 % General information about the journal
-The journal segments are saved as objects in RADOS.  The journal has 4
-pointers, described in `osdc/Journaler.h`:
+%The journal segments are saved as objects in RADOS.  The journal has 4
+%pointers, described in `osdc/Journaler.h`:
+%
+%\begin{itemize}
+%  \item write position: tail of the journal; points to the current session where we are appending events
+%  \item unused field: where someone is reading
+%  \item expire position: old journal segments
+%  \item trimmed position: where daemon is expiring old items
+%\end{itemize}
+%
+%% How the journal tool works
+%Journal segments in RADDS have a header followed by serialized log events. The
+%log events are read by hopping over objects using the read offset and object
+%size pulled from the journal header.  After decoding them, we can examine the
+%metadata (1) about the event (e.g., type, timestamps, etc.) and (2) for inodes
+%that the event touches.
+%
+%% The metablobs
+%The metadata for inodes that the event touches are called metadata blobs and
+%the ones associated with events are \textbf{unordered}; this layout makes
+%writing journal updates fast but the cost is realized when reading the metadata
+%blobs.  It makes sense to optimize for writing since reading only occurs on
+%failures. To reconstruct the namespace for the metadata blob, the journal tool
+%iterates over each metadata blob in the events and builds mappings of inodes to
+%directory entries (for directories) and parent inodes to child inodes/directory
+%entries.
 
-\begin{itemize}
-  \item write position: tail of the journal; points to the current session where we are appending events
-  \item unused field: where someone is reading
-  \item expire position: old journal segments
-  \item trimmed position: where daemon is expiring old items
-\end{itemize}
-
-% How the journal tool works
-Journal segments in RADDS have a header followed by serialized log events. The
-log events are read by hopping over objects using the read offset and object
-size pulled from the journal header.  After decoding them, we can examine the
-metadata (1) about the event (e.g., type, timestamps, etc.) and (2) for inodes
-that the event touches.
-
-% The metablobs
-The metadata for inodes that the event touches are called metadata blobs and
-the ones associated with events are \textbf{unordered}; this layout makes
-writing journal updates fast but the cost is realized when reading the metadata
-blobs.  It makes sense to optimize for writing since reading only occurs on
-failures. To reconstruct the namespace for the metadata blob, the journal tool
-iterates over each metadata blob in the events and builds mappings of inodes to
-directory entries (for directories) and parent inodes to child inodes/directory
-entries.
-
-\subsection{External Library: Create, Save, Persist}
-\label{sec:external-library}
+\subsection{Library: Create, Local/Global Persist, Nonvolatile Apply}
+\label{sec:library-create-local-global-persist}
 %Reads/Writes Journal Events}
@@
 % How it works
-For ``create", ``save", and ``persist" Cudele provides a library for clients to
-link into.  Recall that CephFS uses the object store (1) as a metadata store
-for all information about files including the hierarchical namespace and (2) as
-a staging area for the journal of updates before they are applied to the
-metadata store. As discussed previously, the metadata store is optimized for
-reading while the journal is optimized for writing.  Cudele re-uses the journal
-tool to read/write log events to memory and persistent storage.
-
-% creates
-For create,
-clients write to an in-memory journal and updates are merged into the global
-namespace by replaying them onto the metadata store in the object store.
-
-% persist/save
-For save, clients write serialized log events to a file on local disk and for
-persist, clients store the file in the objects store. The overheads for save
-and persist is the write bandwidth of the local disk and object store,
-respectively.
+For these mechanisms, Cudele provides a library for clients to link into.  For
+``Create", clients write to an in-memory journal and updates are merged into
+the global namespace by replaying them onto the metadata store either in the
+metadata servers or the object store.  For ``Local Persist", clients write
+serialized log events to a file on local disk and for ``Global Persist",
+clients push the journal into the objects store. The overheads for both ``Local
+Persist" and ``Global Persist" is the write bandwidth of the local disk and
+object store, respectively. Finally, for ``Nonvolatile Apply" the journal
+updates are replayed onto the metadata store in the object store and the
+metadata servers are restarted.
 
 % Level of intrusion
-This required no changes to Ceph because the metadata server knows how to read
-the events we were writing into the object store.  The client writes metadata
-updates locally and merges the updates with the journal tool. The client that
-decouples the namespace operates without any consistency and any conflicts at
-the merge are resolved in favor of this client. Updates by other clients ({\it
-i.e.} metadata writes to the global namespace) are overwritten.  We leverage
-the journal tool's ability to reconstruct metadata events in memory. The client
-library is shown in Figure~\ref{fig:decouple}.  Cudele adopts the following
-process when an application decouples the namespace:
-
-\begin{enumerate}
-  \item ``\textbf{decoupled}": metadata server exports the journal events to a file
-  \item ``\textbf{transfer}": the file is pulled by the client from the metadata server
-  \item ``\textbf{snapshot}": client reads the file and materializes a snapshot of the
-        namespace in memory
-\end{enumerate}
+These implementations required no changes to Ceph because the metadata servers
+know how to read the events the library is writing into the object store.  By
+re-using the journal subsystem to implement the namespace decoupling, Cudele
+leverages the write/read optimized data structures, the formats for persisting
+events (similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the
+functions for replaying events onto the internal namespace data structures.
 
 % Why we re-use stuff
-By using re-using the journal subsystem to implement the namespace decoupling,
-Cudele leverages the write/read optimized data structures, the formats for
-persisting events (similar to TableFS's SSTables), and the functions for
-replaying events onto the internal namespace data structures.
-
 %Step 3 is the most complicated and requires understanding how the snapshot is
@@
 % - how it makes no gaurantees
-The volatile apply mechanism ({\it i.e.} the ``v\_apply" arrow in
-Figure~\ref{fig:decouple}) takes an in-memory journal on the client and applies
-the updates directly to the in-memory namespace maintained by the metadata
-servers. We say volatile because Cudele makes no consistency or durability
-guarantees. If a concurrent update from a client occurs there is no
-rule for resolving conflicts and if the client or metadata server crashes there
-may be no way to recover. Relaxing these constraints gives volatile apply the
-best performance.
+The ``Volatile Apply" mechanism takes an in-memory journal on the client and
+applies the updates directly to the in-memory namespace maintained by the
+metadata servers. We say volatile because -- in exchange for peak performance
+-- Cudele makes no consistency or durability guarantees while the mechanism is
+executing.  If a concurrent update from a client occurs there is no rule for
+resolving conflicts and if the client or metadata server crashes there may be
+no way to recover.
 
 % difference between apply and volatile apply
-In contrast, apply uses the the object store to merge the journal of updates
-from the client to the metadata server. Apply is safer than volatile apply but
+In contrast, ``Nonvolatile Apply" uses the the object store to merge the journal of updates
+from the client to the metadata server. ``Nonvolatile Apply" is safer but
 has a performance overhead because objects in the metadata store need to be
-read from and written back to the object store.  Apply was already implemented
+read from and written back to the object store.  ``Volatile Apply" was already implemented
 by the journal tool: it reads the journal from the object store, adds events to
 the journal, and writes the metadata updates out to the metadata store in the
-object store. Cudele's volatile apply executes ``save" (described
-in~\S\ref{sec:external-library}), transfers the file of metadata updates to the
-metadata server, and then merges the updates directly into the in-memory
-metadata store of the metadata cluster.
-
-% Why so slow?
-Creating many files in the same directory would touch the same object but the
-existing implementation results in this object being repeatedly pushed/pulled.
+object store.
 
@@
 % 1. read journal of updates from file
-% 2. call replay (uses same code as when an MDS comes back) on each event
-% 3. skip inodes so MDS doesn't hand out those new inodes
+% 2. call replay (uses same code as when an metadata server comes back) on each event
+% 3. skip inodes so metadata server doesn't hand out those new inodes

### 2017-03-29 paper/sections/0_abstract.tex
paper: wordsmith and push related work backwards
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 serialization overheads are too costly -- and often uneccessary -- for their
-applications.  We provide an API for the client to (1) merge updates into the
-global namespace and (2) assign policies to subtrees so that other clients
-cannot interfere with the decoupled namespace.  This allows administrators to
-optimize subtrees within the same namespace for different workloads. We draw
-conclusions about the performance impact of previously unexplored
-consistency/durability metadata designs and show that consistency can cause a
-104\(\times\) slow down while merging updates (7\(\times\) slow down) and
-maintaining durability (\(10\times\) slow down) have a more reasonable cost.
+applications.  We provide an API for the client to (1) decouple subtrees from
+the global namespace, (2) merge metadata updates back into the global namespace
+and (3) assign policies to subtrees so that other clients cannot interfere with
+the decoupled namespace.  This allows administrators to optimize subtrees
+within the same namespace for different workloads. We draw conclusions about
+the performance impact of previously unexplored consistency/durability metadata
+designs and show that consistency can cause a 104\(\times\) slow down while
+merging updates (7\(\times\) slow down) and maintaining durability
+(\(10\times\) slow down) have a more reasonable cost.
 

### 2017-03-29 paper/sections/1_introduction.tex
paper: wordsmith and push related work backwards
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 applications perform better with dedicated metadata
-servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs}. Unfortunately,
-provisioning a metadata server for every client is expensive and complicated.
-
-% How does this fit in current landscape?
-Current hardware evolution and the rise of software-defined storage storage,
-which uses techniqes like erasure coding, replication, and partitioning, have
-ushered in a new era of HPC computing; architectures are transitioning from
+servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
+metadata server for every client is expensive and complicated. This problem is
+exacerbated by current trends in HPC.  Architectures are transitioning from
 complex storage stacks with burst buffer, file system, object store, and tape
-tiers to a two layer stack with just a burst buffer and object
-store~\cite{bent:login16-hpc-trends}. This trend exacerbates the metadata
-scalability problem
+tiers to a more simplified stack with just a burst buffer and object
+store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
+instead of data movement.
 
 % What is HPC doing?
-To address this trend, HPC research has advocated client side processing and an
-emphasis on serverless metadata services.  The techniques propose new metadata
-management techniques that reduce synchronization and serialization overheads
-by transferring these responsibilities to the client. We call this approach
-``decoupling the namespace" and the semantics of consistency differ amongst
-systems.  While the performance benefits are obvious for these users,
-applications that rely on stronger consistency or durability guarantees must be
-re-written or deployed on a different system.
+HPC's unique requirements for file systems ({\it e.g.}, fast creates) and well
+defined workloads ({\it e.g.}, workloads) make relaxing POSIX sensible. Without
+the constraints of strong consistency and durability, HPC systems can now do
+more client side processing with an emphasis on serverless metadata services.
+One popular approach is to ``decouple the namespace", where clients lock the
+subtree they want exclusive access to so other clients cannot interfere. This
+delayed merge ({\it i.e.} a form of eventual consistency) and relaxed
+durability improves performance and scalability by avoiding the costs of RPCs,
+synchronization, false sharing, and serialization.  While the performance
+benefits are obvious for these users, applications that rely on stronger
+consistency or durability guarantees must be re-written or deployed on a
+different system. The consistency and durability semantics for these systems is
+shown in Table~\ref{table:namespaces}.
 
@@
 
+\begin{table}
+\begin{tabular}{ r | l | l }
+              & Decoupled & Global    \\
+              & Namespace & Namespace \\\hline
+  Example     & BatchFS~\cite{zheng:pdsw2014-batchfs} & CephFS~\cite{weil:sc2004-dyn-metadata} \\
+              & DeltaFS~\cite{zheng:pdsw2015-deltafs} & IndexFS~\cite{ren:sc2014-indexfs}      \\
+  Consistency & eventual & strong     \\
+  Durability  & node local & global  \\
+\end{tabular}
+\caption{State-of-the-art systems in HPC improve file system metadata
+performance by relaxing consistency and durability
+guarantees.\label{table:namespaces}}
+\end{table}
+
 % What did we do
 We propose subtree policies, an interface that lets future programmers control
-how the system manages different parts of the namespace.  For performance one
+how the storage system manages different parts of the file system namespace.  For performance one
 subtree can adopt weaker consistency semantics while another subtree can retain
@@
 
-  \item an API for assigning and programming policies to subtrees in the file
-  system namespace.
+  \item an API for programming consistency/durability policies and assigning
+  them to subtrees in the file system namespace.
 
-  \item an analysis of recently proposed research systems compared against
+  \item a comparison of the strategies used in recently proposed research systems against
   previously unexplored metadata designs.

### 2017-03-29 paper/sections/2_related_work.tex
paper: cleanup terminology in section 4
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: wordsmith and push related work backwards
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 techniques include partitioning metadata when there are many writes and
-replicating metadata when there are many reads. For example, IndexFS partitions
+replicating metadata when there are many reads. For example, IndexFS~\cite{ren:sc2014-indexfs} partitions
 directories and clients write to different partitions by grabbing leases and
@@
 pessimistically while CephFS replicates only when the read temperature reaches
-a threshold.  There is a wide range of policies and it is difficult to address
-with tunables and hard-coded design decisions.
+a threshold.  There is a wide range of policies and it is difficult to maneuver
+tunables and hard-coded design decisions.
 
@@
 consistency, writes to directories in IndexFS block until the lease expires
-while writes to direcotries in ShardFS are slow for everyone as it either
+while writes to directories in ShardFS are slow for everyone as it either
 requires serialization or locking with many servers; reads in IndexFS are
@@
 %amplification to 1 when mutatating directory metadata ({\it e.g.,}
-%\texttt{mkdir}, \texttt{chmod}, etc.) because clients are not querying MDSs for
+%\texttt{mkdir}, \texttt{chmod}, etc.) because clients are not querying metadata servers for
 %path traversals. The disadvantage of this approach is the high latency of
 %mutation operations (reads to filenames).  ShardFS~\cite{xiao:socc2015-shardfs}
-%replicates metadata (specifically directory lookup state) across the MDS
+%replicates metadata (specifically directory lookup state) across the metadata server
 %cluster, reducing the RPC amplification to 1 for file operations ({e.g.,}
@@
 % Non POSIX
-High performance computing has unique requirements for file systems ({e.g.},
-fast creates) and well-defined workloads (e.g., workflows) that make relaxing
-POSIX sensible.  One popular approach is to allow clients to ``lock'' parts of
-the namespace to improve performance and scalability by avoiding
-synchronization, false sharing, and serialization.  BatchFS assumes the
-application coordinates accesses to the namespace, so the clients can batch
-local operations and merge with a global namespace image lazily. Similarily,
-DeltaFS eliminates RPC traffic using subtree snapshots for non-conflicting
-workloads and middleware for conflicting workloads. MarFS gives administrators
-the ability to (1) lock ``project directories" and (2) allocate GPFS clusters
-for demanding directory workloads. TwoTiers eliminates high-latencies by
-storing metadata in a flash tier; apps lock the namespace so that metadata can
-be accessed more quickly.  Unfortunately, decoupling the namespaces has costs:
-(1) merging metadata state back into the global namespace is slow; (2) failures
-are local to the failing node; and (3) the systems are not backwards
-compatible.
+High performance computing has unique requirements for file systems ({\it
+e.g.}, fast creates) and well-defined workloads ({\it e.g.}, workflows) that
+make relaxing POSIX sensible.  BatchFS assumes the application coordinates
+accesses to the namespace, so the clients can batch local operations and merge
+with a global namespace image lazily. Similarily, DeltaFS eliminates RPC
+traffic using subtree snapshots for non-conflicting workloads and middleware
+for conflicting workloads. MarFS gives administrators the ability to lock
+``project directories" and allocate GPFS clusters for demanding metadata
+workloads. TwoTiers eliminates high-latencies by storing metadata in a flash
+tier; apps lock the namespace so that metadata can be accessed more quickly.
+Unfortunately, decoupling the namespace has costs: (1) merging metadata state
+back into the global namespace is slow; (2) failures are local to the failing
+node; and (3) the systems are not backwards compatible.
 
@@
 problems POSIX and REST are the dominant APIs for data transfer.
-
-Decoupling the namespace delays metadata consistency and sacrifices durability. 
-As shown in Table~\ref{table:namespaces}, metadata consistency is
-provided by capabilities and metadata durability is addressed with a journal.
-\begin{table}
-\begin{tabular}{ r | l | l }
-              & Decoupled & Global    \\
-              & Namespace & Namespace \\\hline
-  Example     & BatchFS~\cite{zheng:pdsw2014-batchfs} & CephFS~\cite{weil:sc2004-dyn-metadata} \\
-              & DeltaFS~\cite{zheng:pdsw2015-deltafs} & IndexFS~\cite{ren:sc2014-indexfs}      \\
-  Consistency & eventual & strong     \\
-  Durability  & node local & journal  \\
-\end{tabular}
-\caption{State-of-the-art systems in HPC improve file system metadata
-performance by relaxing consistency and durability
-guarantees.\label{table:namespaces}}
-\end{table}

### 2017-03-30 images/ceph/build.sh
images: update to version that excludes certain clients
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -13,12 +13,14 @@ wget https://raw.githubusercontent.com/systemslab/docker-cephdev/master/aliases.
 rm aliases.sh
 docker pull ceph/daemon:tag-build-master-jewel-ubuntu-14.04
 docker tag ceph/daemon:tag-build-master-jewel-ubuntu-14.04 ceph/daemon:jewel
+docker pull cephbuilder/ceph:latest
 
 #-e SHA1_OR_REF="ec12610bb1358fb2eab74cb7b0483c2613bae38f" \
 #-e SHA1_OR_REF="1fdfa94459026b70e8925c0a530005456c526260" \
 #-e SHA1_OR_REF="f4d39ca86e29555a45a4cff3f03862081bcc1f84" \
+#-e SHA1_OR_REF="11d3bc22ceb151c23c54edaadab2d9d0ec9659fe" \
 dmake \
-  -e SHA1_OR_REF="11d3bc22ceb151c23c54edaadab2d9d0ec9659fe" \
+  -e SHA1_OR_REF="c8198b367959e5e7ccb362987ae541a526ccaf59" \
   -e GIT_URL="https://github.com/michaelsevilla/ceph.git" \
   -e BUILD_THREADS=`grep processor /proc/cpuinfo | wc -l` \
   -e CONFIGURE_FLAGS="-DWITH_TESTS=OFF" \
@@ -29,6 +31,7 @@ cd -
 #docker tag ec12610bb1358fb2eab74cb7b0483c2613bae38f-base ceph/daemon:ec12610
 #docker tag ceph-1fdfa94459026b70e8925c0a530005456c526260 ceph/daemon:1fdfa94
 #docker tag ceph-f4d39ca86e29555a45a4cff3f03862081bcc1f84 ceph/daemon:f4d39ca
-docker tag ceph-11d3bc22ceb151c23c54edaadab2d9d0ec9659fe ceph/daemon:11d3bc2
+#docker tag ceph-11d3bc22ceb151c23c54edaadab2d9d0ec9659fe ceph/daemon:11d3bc2
+docker tag ceph-c8198b367959e5e7ccb362987ae541a526ccaf59 ceph/daemon:c8198b3
 docker build -t tmp .
-docker tag tmp piha.soe.ucsc.edu:5000/ceph/daemon:11d3bc2
+docker tag tmp piha.soe.ucsc.edu:5000/ceph/daemon:c8198b3

### 2017-03-30 paper/paper.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: popperize graphs and fix links
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: edit section 3
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: wordsmith and push related work backwards
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: jeff intro comments v0
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % Template: http://www.acm.org/publications/proceedings-template
-\documentclass[sigconf]{acmart}
+\documentclass[sigconf, preprint]{acmart}
 \usepackage{booktabs} % For formal tables
@@
 \usepackage{arydshln}
+\usepackage{subcaption}
 
-% Copyright
-%\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-\acmDOI{10.475/123_4}
-
-% ISBN
-\acmISBN{123-4567-24-567/08/06}
-
-%Conference
-\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El
-  Paso, Texas USA} 
-\acmYear{1997}
-\copyrightyear{2016}
-
-\acmPrice{15.00}
+%% Copyright
+%%\setcopyright{none}
+%%\setcopyright{acmcopyright}
+%%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%%\setcopyright{usgov}
+%%\setcopyright{usgovmixed}
+%%\setcopyright{cagov}
+%%\setcopyright{cagovmixed}
+%
+%
+%% DOI
+%\acmDOI{10.475/123_4}
+%
+%% ISBN
+%\acmISBN{123-4567-24-567/08/06}
+%
+%%Conference
+%\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El
+%  Paso, Texas USA} 
+%\acmYear{1997}
+%\copyrightyear{2016}
+%
+%\acmPrice{15.00}
 
 \begin{document}
-\title{Cudele: Programmable Consistency and Fault Tolerance}
+\title{Cudele: A File System with Programmable Consistency and Durability}
 
@@
 
+%%
+%% The code below should be generated by the tool at
+%% http://dl.acm.org/ccs.cfm
+%% Please copy and paste the code instead of the example below. 
+%%
+%\begin{CCSXML}
+%<ccs2012>
+% <concept>
+%  <concept_id>10010520.10010553.10010562</concept_id>
+%  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+%  <concept_significance>500</concept_significance>
+% </concept>
+% <concept>
+%  <concept_id>10010520.10010575.10010755</concept_id>
+%  <concept_desc>Computer systems organization~Redundancy</concept_desc>
+%  <concept_significance>300</concept_significance>
+% </concept>
+% <concept>
+%  <concept_id>10010520.10010553.10010554</concept_id>
+%  <concept_desc>Computer systems organization~Robotics</concept_desc>
+%  <concept_significance>100</concept_significance>
+% </concept>
+% <concept>
+%  <concept_id>10003033.10003083.10003095</concept_id>
+%  <concept_desc>Networks~Network reliability</concept_desc>
+%  <concept_significance>100</concept_significance>
+% </concept>
+%</ccs2012>  
+%\end{CCSXML}
 %
-% The code below should be generated by the tool at
-% http://dl.acm.org/ccs.cfm
-% Please copy and paste the code instead of the example below. 
+%\ccsdesc[500]{Computer systems organization~Embedded systems}
+%\ccsdesc[300]{Computer systems organization~Redundancy}
+%\ccsdesc{Computer systems organization~Robotics}
+%\ccsdesc[100]{Networks~Network reliability}
 %
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>  
-\end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
-
-% We no longer use \terms command
-%\terms{Theory}
-
-\keywords{ACM proceedings, \LaTeX, text tagging}
+%% We no longer use \terms command
+%%\terms{Theory}
+%
+%\keywords{ACM proceedings, \LaTeX, text tagging}
 
@@
 \input{sections/1_introduction}
-%\input{sections/2_related_work}
 \input{sections/3_posix_overheads}
@@
 \input{sections/6_results}
+\input{sections/2_related_work}
 
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sigproc}
+\bibliography{paper}
 

### 2017-03-30 paper/sections/0_abstract.tex
paper: jeff intro comments v0
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-03-20 paper/sections/0_abstract.tex
@@
 
-HPC file systems are abandoning POSIX because the synchronization and
-serialization overheads are too costly -- and often uneccessary -- for their
-applications.  We provide an API for the client to (1) merge updates into the
-global namespace and (2) assign policies to subtrees so that other clients
-cannot interfere with the decoupled namespace.  This allows administrators to
-optimize subtrees within the same namespace for different workloads. We draw
-conclusions about the performance impact of previously unexplored
-consistency/durability metadata designs and show that consistency can cause a
-104\(\times\) slow down while merging updates (7\(\times\) slow down) and
-maintaining durability (\(10\times\) slow down) have a more reasonable cost.
+HPC developers debate abandoning POSIX because the synchronization and
+serialization overheads of providing strong consistency and durability are too
+costly -- and often uneccessary -- for their applications.  Unfortunately,
+designing near-POSIX file systems excludes applications that rely on strong
+consistency or durability, forcing developers to re-write their applications or
+deploy them on a different system.  We present a file system and API that
+allows clients to specify their consistency/durability requirements and assign
+them to subtrees in the namespace, allowing administrators to optimize subtrees
+within the same namespace for different workloads.  We draw conclusions about
+the performance impact of unexplored consistency/durability metadata designs
+and show that strong consistency can cause a 104\(\times\) slow down while merging
+updates (7\(\times\) slow down) and maintaining durability (\(10\times\) slow
+down) have more reasonable costs.
 

### 2017-03-30 paper/sections/1_introduction.tex
paper: jeff intro comments v0
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % What is the problem?
-Today's client-server based file system metadata services have scalability
-problems. It takes a lot of resources to service POSIX metadata requests so
-applications perform better with dedicated metadata
+File system metadata services in HPC have scalability problems. It has been
+shown that HPC workloads, especially checkpoint-restart jobs, are resource
+intensive primarily due to contention for the same directories and inodes ({\it
+e.g.}, path traversal). Applications perform better with dedicated metadata
 servers~\cite{sevilla:sc15-mantle, ren:sc2014-indexfs} but provisioning a
-metadata server for every client is expensive and complicated. This problem is
-exacerbated by current trends in HPC.  Architectures are transitioning from
-complex storage stacks with burst buffer, file system, object store, and tape
-tiers to a more simplified stack with just a burst buffer and object
+metadata server for every client is unreasonable. This problem is exacerbated
+by current trends in HPC, where architectures are transitioning from complex
+storage stacks with burst buffer, file system, object store, and tape tiers to
+more simplified stackis with just a burst buffer and object
 store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
@@
 HPC's unique requirements for file systems ({\it e.g.}, fast creates) and well
-defined workloads ({\it e.g.}, workloads) make relaxing POSIX sensible. Without
-the constraints of strong consistency and durability, HPC systems can now do
-more client side processing with an emphasis on serverless metadata services.
-One popular approach is to ``decouple the namespace", where clients lock the
-subtree they want exclusive access to so other clients cannot interfere. This
-delayed merge ({\it i.e.} a form of eventual consistency) and relaxed
-durability improves performance and scalability by avoiding the costs of RPCs,
-synchronization, false sharing, and serialization.  While the performance
-benefits are obvious for these users, applications that rely on stronger
-consistency or durability guarantees must be re-written or deployed on a
-different system. The consistency and durability semantics for these systems is
-shown in Table~\ref{table:namespaces}.
+defined workloads ({\it e.g.}, workflows) have given rise to file systems with
+weaker consistency constraints than POSIX and less durability than programmers
+are accustomed to.  Without these constraints, HPC systems can do more
+client side processing with an emphasis on serverless metadata services.
+
+One popular approach for relaxing consistency and durability is to ``decouple
+the namespace", where clients lock the subtree they want exclusive access to so
+other clients cannot interfere. This delayed merge ({\it i.e.} a form of
+eventual consistency) and relaxed durability improves performance and
+scalability by avoiding the costs of RPCs, synchronization, false sharing, and
+serialization.  While the performance benefits are obvious for these users,
+applications that rely on stronger consistency or durability guarantees must be
+re-written or deployed on a different system. The consistency and durability
+semantics for these systems is shown in Table~\ref{table:namespaces}.
 

### 2017-03-31 paper/sections/3_posix_overheads.tex
paper: fix creates motivation graph
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \centering
-\includegraphics[width=1.0\linewidth]{figures/creates-motivation.png}
+\includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
 \caption{Create-heavy workloads (untar) incurr the highest disk, network, and
 CPU utilization because of the consistency and durability demands of CephFS
-.}\label{fig:creates-motivation}
+.}\label{fig:overhead-creates}
 \end{figure}
@@
 stress the underyling storage system the most.
-Figure~\ref{fig:creates-motivation} shows the resource utilization of compiling
+Figure~\ref{fig:overhead-creates} shows the resource utilization of compiling
 the Linux kernel.  The untar phase, which is characterized by many creates, has

### 2017-03-31 viz.sh

@@ -1,62 +1,62 @@
 #!/bin/bash
 
-#echo "-- untar the journal"
-#DIR="experiments/kernel/journal/results-done"
-#SUBDIRS=`ls $DIR`
-#MDS="issdm-12"
-#for i in $SUBDIRS; do
-#  cd $DIR/$i
-#  echo "DIR=" `pwd`
-#  for j in `ls *.tar.gz`; do
-#    tar xzf $j
-#    if [ $j == $MDS.tar.gz ]; then
-#      echo "... parsing MDS=$j"    
-#      for m in \
-#         "tmp/graphite/whisper/$MDS/mds_mem/ino.wsp" \
-#         "tmp/graphite/whisper/$MDS/mds_mem/ino+.wsp" \
-#         "tmp/graphite/whisper/$MDS/mds_mem/ino-.wsp" \
-#         "tmp/graphite/whisper/$MDS/mds/inodes.wsp" \
-#         "tmp/graphite/whisper/$MDS/mds/inodes_bottom.wsp" \
-#         "tmp/graphite/whisper/$MDS/mds/inodes_with_caps.wsp" \
-#         "tmp/graphite/whisper/$MDS/mds/inodes_pin_tail.wsp" \
-#         "tmp/graphite/whisper/$MDS/mds/inodes_pinned.wsp" \
-#         "tmp/graphite/whisper/$MDS/mds/inodes_top.wsp" \
-#         "tmp/graphite/whisper/$MDS/mds_server/handle_client_request_tput.wsp" \
-#         "tmp/graphite/whisper/$MDS/cputotals/user.wsp" \
-#         "tmp/graphite/whisper/$MDS/cputotals/sys.wsp" \
-#        ; do
-#        docker run -v `pwd`/tmp:/tmp --entrypoint=whisper-dump.py --rm \
-#          michaelsevilla/graphite ${m} > `basename ${m}`.out
-#        docker run -v `pwd`/:/tmp --entrypoint=/bin/bash --rm \
-#          michaelsevilla/graphite -c "sed -i \"s/:/,/g\" tmp/`basename $m`.out"
-#        echo "... ... `basename $m`"
-#      done
-#    fi
-#  done
-#  cd - >> /dev/null
-#done
-#
-#DIR="experiments/kernel/journal/results/journal180-cache"
-#SUBDIRS=`ls $DIR`
-#MDS="issdm-12"
-#cd $DIR
-#echo "DIR=" `pwd`
-#for j in `ls *.tar.gz`; do
-#  tar xzf $j
[…]

### 2017-03-31 experiments/fuse/scale-clients-isolated/run.sh
benchmark: add isolated clients
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,53 @@
+#!/bin/bash
+
+set -ex
+
+# setup the docker command
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root"
+ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
+CEPH_ANSIBLE="$RUN -v `pwd`/site/roles/ceph-ansible:/root $ANSIBLE"
+SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
+
+#for nclients in 2 3 4 5 6 7 8; do
+#  site="nojournal-cache"
+#  for run in `seq 0 2`; do
+#    for job in "creates" "creates-touchstream" "creates-stat" "creates-touch"; do
+
+for nclients in 2 3 4 5 6 7; do
+  site="nojournal-cache"
+  for run in 0 1 2; do
+    for job in "creates-touchstream"; do
+      for decoupled in 0 1; do
+        # configure ceph and setup results directory
+        results="$job-$nclients-decoupled-$decoupled/run$run"
+        mkdir -p results/$results/logs || true
+        cp site/* site/roles/ceph-ansible || true
+        cp site_confs/${site}.yml site/group_vars/all
+        cp -r site/group_vars site/roles/ceph-ansible/
+        cp inventory/${nclients}client site/hosts
+        cp inventory/${nclients}client site/roles/ceph-ansible/hosts
+  
+        # cleanup and start ceph
+        for k in `seq 0 3`; do
+          $SRL_ANSIBLE cleanup.yml || true
+        done
+        $SRL_ANSIBLE cleanup.yml
+        $CEPH_ANSIBLE ceph.yml #cephfs.yml
+        $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
+  
+        ## warmup and get baseline
+        #for i in `seq 0 2`; do
+        #  ./ansible-playbook.sh -e site=$results -e nfiles=100000 ../workloads/warmup.yml
+        #done
+  
+        # run job
+        ./ansible-playbook.sh -e site=$results -e nfiles=10 ../workloads/setup_mdtest.yml
+        if [ $decoupled -eq 1 ]; then
+          ./ansible-playbook.sh -e site=$results -e nfiles=10 ../workloads/decouple.yml
+        fi
[…]

### 2017-03-31 paper/sections/0_abstract.tex
paper: noah's intro comments and new figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-HPC developers debate abandoning POSIX because the synchronization and
+HPC developers are abandoning POSIX because the synchronization and
 serialization overheads of providing strong consistency and durability are too
@@
 deploy them on a different system.  We present a file system and API that
-allows clients to specify their consistency/durability requirements and assign
+lets clients to specify their consistency/durability requirements and assign
 them to subtrees in the namespace, allowing administrators to optimize subtrees

### 2017-03-31 paper/sections/1_introduction.tex
paper: noah's intro comments and new figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 File system metadata services in HPC have scalability problems. It has been
-shown that HPC workloads, especially checkpoint-restart jobs, are resource
-intensive primarily due to contention for the same directories and inodes ({\it
+shown that HPC workloads are metadata resource intensive because administrative
+tasks, like checkpointing~\cite{bent_plfs_2009} or scanning the file
+system~\cite{zheng:pdsw2014-batchfs}, on large data sets leads to contention
+for the same directories and inodes ({\it
 e.g.}, path traversal). Applications perform better with dedicated metadata
@@
 storage stacks with burst buffer, file system, object store, and tape tiers to
-more simplified stackis with just a burst buffer and object
+more simplified stacks with just a burst buffer and object
 store~\cite{bent:login16-hpc-trends}; this puts more pressure on data access
-instead of data movement.
-
-% What is HPC doing?
-HPC's unique requirements for file systems ({\it e.g.}, fast creates) and well
-defined workloads ({\it e.g.}, workflows) have given rise to file systems with
-weaker consistency constraints than POSIX and less durability than programmers
-are accustomed to.  Without these constraints, HPC systems can do more
-client side processing with an emphasis on serverless metadata services.
-
-One popular approach for relaxing consistency and durability is to ``decouple
-the namespace", where clients lock the subtree they want exclusive access to so
-other clients cannot interfere. This delayed merge ({\it i.e.} a form of
-eventual consistency) and relaxed durability improves performance and
-scalability by avoiding the costs of RPCs, synchronization, false sharing, and
-serialization.  While the performance benefits are obvious for these users,
-applications that rely on stronger consistency or durability guarantees must be
-re-written or deployed on a different system. The consistency and durability
-semantics for these systems is shown in Table~\ref{table:namespaces}.
+because more requests end up hitting the same layer and latencies cannot be 
+hidden while data migrates across tiers.
 
@@
 \centering
-\includegraphics[width=75mm]{figures/subtree-policies.png}
+\includegraphics[width=0.5\textwidth]{figures/subtree-policies.png}
 \caption{Administrators can assign consistency and durability policies to
@@
 
-\begin{table}
-\begin{tabular}{ r | l | l }
-              & Decoupled & Global    \\
-              & Namespace & Namespace \\\hline
-  Example     & BatchFS~\cite{zheng:pdsw2014-batchfs} & CephFS~\cite{weil:sc2004-dyn-metadata} \\
-              & DeltaFS~\cite{zheng:pdsw2015-deltafs} & IndexFS~\cite{ren:sc2014-indexfs}      \\
-  Consistency & eventual & strong     \\
-  Durability  & node local & global  \\
-\end{tabular}
-\caption{State-of-the-art systems in HPC improve file system metadata
-performance by relaxing consistency and durability
-guarantees.\label{table:namespaces}}
-\end{table}
+% What is HPC doing?
+To address this, developers are relaxing the consistency and durability
+semantics in the file system because weaker guarantees are sufficient for their
+applications. For example, batch style jobs often do not need the strong
+consistency that the file system provides, so
+BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
+do more client side processing and merge updates when the job is done. HPC
+developers are turning to this non-POSIX solution because their applications
+are well-understood ({\it e.g.}, well-defined read/write phases,
+synchronization only needed during certain phases, workflows describing
+computation, etc.) and because they wreak havoc on file systems designed for
+general-purpose workloads ({\it e.g.}, checkpoint-restart's N-N and N-1 create
+patterns).
+
+% One example
+One popular approach for relaxing consistency and durability is to ``decouple
+the namespace", where clients lock the subtree they want exclusive access to as
+a way to tell the file system that the subtree is important or may cause
+resource contention in the near-future~\cite{grider:pdsw2015-marfs,
+zheng:pdsw2015-deltafs, zheng:pdsw2014-batchfs, ren:sc2014-indexfs,
+bent:slides-twotiers}. Then the file system can change its internal structure
+to optimize performance. For example, the file system could enter a mode that
+prevents other clients from interferring with the decoupled directory.  This
+delayed merge ({\it i.e.} a form of eventual consistency) and relaxed
+durability improves performance and scalability by avoiding the costs of RPCs,
+synchronization, false sharing, and serialization.  The consistency and
+durability semantics for these systems is shown in the table in
+Figure~\ref{fig:subtree-policies} . While the performance benefits are obvious
+for these users, applications that rely on the file system's guarantees must be
+deployed on an entirely different system or re-written to coordinate strong
+consistency/durability themselves.
+
+%\begin{table}
+%\begin{tabular}{ r | l }
+%  Subtree         & Example \\\hline
+%  (1)   & \{Index, Batch\}FS~\cite{ren:sc2014-indexfs, zheng:pdsw2014-batchfs} \\
+%  (2)   & \{Index, Ceph\}FS~\cite{ren:sc2014-indexfs, weil:sc2004-dyn-metadata} \\
+%  (3)   & RAMDisk \\
+%  (4)   & DeltaFS~\cite{zheng:pdsw2015-deltafs} \\
+%\end{tabular}
+%
+%\caption{State-of-the-art systems in HPC improve file system metadata
+%performance by relaxing consistency and durability guarantees. Note that
+%IndexFS also supports weak consistency with bulk inserts.
+%\label{table:namespaces}} \end{table}
 
@@
 We propose subtree policies, an interface that lets future programmers control
-how the storage system manages different parts of the file system namespace.  For performance one
+the consistency and durability for subtrees in the file system namespace. For performance, one
 subtree can adopt weaker consistency semantics while another subtree can retain
@@
 file system as a client library.  Cudele supports 3 forms of consistency
-(invisible, eventual, and strong) and 3 degrees of durability (none, local, and
+(invisible, weak, and strong) and 3 degrees of durability (none, local, and
 global) giving the administrator a wide range of policies and optimizations

### 2017-03-31 paper/sections/4_methodology.tex
paper: prune implementation section and make it readable
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: noah's intro comments and new figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
   C \(\rightarrow\) &&& \\  
-  D \(\downarrow\)  	     & invisible         & eventual        & strong  \\\hline
+  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
   none                       & create            & create          & RPCs    \\
@@
 into a global namespace and it is assumed that middleware or the application
-manages consistency lazily; ``eventual" merges updates at some time in the
+manages consistency lazily; ``weak" merges updates at some time in the
 future ({\it e.g.}, when the system has time, when number of updates reaches a
@@
 have global consistency and durability; DeltaFS uses ``invisible" consistency
-and ``local" durability and BatchFS uses ``eventual" consistency and ``local"
+and ``local" durability and BatchFS uses ``weak" consistency and ``local"
 durability. These are just a few of the HPC examples.
@@
 metadata servers. Without contention clients can resolve metadata reads locally
-thus reducing operations like \texttt{lookup()}. The size of the inode cache is
-configurable so as not to saturate the memory on the metadata server -- inodes
-in CephFS are about 1400
+thus reducing operations like \texttt{lookup()}.  For example, if a client or
+metadata server is not caching the directory inode, all creates within that
+directory will result in a lookup and a create request. If the directory inode
+is cached then only the create needs to be sent.  The size of the inode cache
+is configurable so as not to saturate the memory on the metadata server --
+inodes in CephFS are about 1400
 bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.

### 2017-03-31 paper/sections/6_results.tex
paper: cleanup terminology in section 4
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: prune implementation section and make it readable
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: noah's intro comments and new figure 1
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \centering
-\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-eventual.png}
+\includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-weak.png}
 \caption{The RPC per metadata update of ``Strong Consistency" has a large
 overhead compared to the decoupled namespace strategy of ``Eventual
-Consistency".\label{fig:slowdown-strong-eventual}}
+Consistency".\label{fig:slowdown-strong-weak}}
 \end{figure}
 
-Figure~\ref{fig:slowdown-strong-eventual} shows the runtimes of eventual and
+Figure~\ref{fig:slowdown-strong-weak} shows the runtimes of weak and
 strong consistency based systems implemented on Cudele, normalized to the
@@
 %Use namespace's recursive data structure to put policies on subtrees
-%- consistency: eventual vs. strong, global vs. local
-%  - e.g., BatchFS/DeltaFS: eventual, local
+%- consistency: weak vs. strong, global vs. local
+%  - e.g., BatchFS/DeltaFS: weak, local
 %  - e.g., POSIX: strong, global
@@
 %Experimental Setup
-%- Ceph: 9 OSDs, 1 MDS, 2 kernel client
+%- Ceph: 9 OSDs, 1 metadata server, 2 kernel client
 %- Workload limitations: blah
@@
 \subsection{Isolation from Global Writes, Cost of Overwrites}
+\subsection{Cost of Nonvolatile Apply}
+% Why so slow?
+Creating many files in the same directory would touch the same object but the
+existing implementation results in this object being repeatedly pushed/pulled.
+
 \subsection{Partial Directory Listings}

### 2017-03-31 paper/sections/1_introduction.tex
paper: noah section 2 comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=0.5\textwidth]{figures/subtree-policies.png}
-\caption{Administrators can assign consistency and durability policies to
-subtrees to get the benefits of some of the state-of-the-art HPC architectures.
+\caption{Administrators can assign weaker consistency and durability policies
+to subtrees to get the same performance benefits of state-of-the-art HPC
+architectures. Applications that need stronger guarantees can still reside in
+the same namespace.
 }\label{fig:subtree-policies}

### 2017-03-31 paper/sections/3_posix_overheads.tex
paper: noah section 2 comments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \caption{Create-heavy workloads (untar) incurr the highest disk, network, and
-CPU utilization because of the consistency and durability demands of CephFS
-.}\label{fig:overhead-creates}
+CPU utilization because of the consistency and durability demands of
+CephFS.}\label{fig:overhead-creates}
 \end{figure}
@@
 the highest resource usage which suggests that it is stressing the consistency
-and journalling subsystems of the metadata server the most. Traditional file
+and journaling subsystems of the metadata server the most. Traditional file
 system techniques for improving performance, such as caching inodes, do not
@@
 In this section, we quantify the costs of strong consistency and global
-durability in CephFS.  We use the kernel client so that we can find the true
-create speed of the server; our experiments show a low CPU utilziation for the
-clients which indicates that we are stressing the servers more.
+durability in CephFS. At the end of each subsection we compare the approach to
+``decoupled namespaces", the technique in related work that detaches subtrees
+from the global namespace to relax consistency/durability guarantees.  We use
+the kernel client so that we can find the true create speed of the server; our
+experiments show a low CPU utilziation for the clients which indicates that we
+are stressing the servers more.
 
@@
 % what is durability
-While durability is not specified by POSIX, users expect that their metadata is
-safe. We define three types of durability: global, local, and none.  Global
+While durability is not specified by POSIX, users expect that files they create
+or modify survive failures.  We define three types of durability: global,
+local, and none.  Global
 durability means that the client or server can fail at any time and metadata
 will not be lost. Local durability means that metadata can be lost if the
-client or server fails and stays failed. None means that metadata is volatile
-and that the system provides no guarantees when clients or servers fail.\\
+client or server stays down after a failure. None means that metadata is volatile
+and that the system provides no guarantees when clients or servers fail. 
+None is different than local durability because regardless of the type failure,
+metadata will be lost when components die in a None configuration.\\
 
@@
 the resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
-WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal can grow to large sizes
-ensuring (1) sequential writes into RADOS and (2) the ability for daemons to
+WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to grow to large which
+ensures (1) sequential writes into the object store and (2) the ability for daemons to
 trim redundant or irrelevent journal entries.  The journal is striped over
@@
 two tunables for controlling the journal: the segment size and the number of
-parallel segments that can be written in parallel. The former is memory bound
-as larger segments take up more memory but can reduce the time spent journaling
-and the latter is CPU bound.
+parallel segments that can be written in parallel. Unless the journal saturates 
+memory or CPU resources, larger values for these tunables results in better
+performance.
 
@@
 never globally consistent and there is no ground truth.  On the server side,
-BatchFS and DeltaFS use IndexFS. Again, IndexFS writes metadata to SSTables but
-it is not clear whether they ever vacate memory, get written to disk, or are
-flushed to the object store.
+BatchFS and DeltaFS use IndexFS. IndexFS writes metadata to SSTables, which
+initially reside in memory but are later to flushed to the underlying
+distributed file system.
 
@@
 Access to metadata in a POSIX-compliant file system is strongly consistent, so
-reads and writes are globally ordered.  The synchronization and serialization
-machinery needed to ensure that all clients see the same state has high
-overhead.\\
+reads and writes to the same inode or directory are globally ordered.  The
+synchronization and serialization machinery needed to ensure that all clients
+see the same state has high overhead.\\
 
@@
 increased runtime and variability. The performance is compared to a single
-isolated client and he error bar is the standard deviation of the runtime of
+isolated client and the error bar is the standard deviation of the runtime of
 all clients.  For the ``interfere" bars, each client creates files in private

### 2017-04-01 paper/sections/4_methodology.tex
paper: integrate noah section 3 feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-03-28 paper/sections/4_methodology.tex
@@
-\section{Methodology: Decoupled Namespaces}
+\section{Methodology: Global Namespace with Per-Subtree Consistency/Durability}
 \label{sec:methodology-decoupled-namespaces}
@@
 \begin{figure}[tb]
-\caption{Applications can decouple the namespace, write updates to a local
-journal, and delay metadata updates.  Table~\ref{table:spectrum} shows how
-these mechanisms (represented by the arrows) can be combined to provide weaker
-consistency or durability semantics.  }\label{fig:decouple}
 \centering
 \includegraphics[width=90mm]{figures/fig-decouple.png}
+\caption{Applications decouple the namespace, write updates to a local journal,
+and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
 \end{figure}
 
-In this section we describe Cudele, our prototype system that lets future
-programmers compose mechanisms (Section~\S\ref{sec:cudeles-mechanisms}) to
-provide the necessary guarantees
-(Section~\S\ref{sec:setting-policies-with-cudele}) for their application.
+In this section we describe CudeleFS, our file system that lets administrators
+assign consistency and durability semantics to subtrees in the global
+namespace. A \textbf{mechanism} is an abstraction and basic building block for
+constructing consistency and durability guarantees. CudeleFS exposes these
+mechanisms and the administrator composes them together to construct
+\textbf{policies}. These policies are assigned to subtrees and they dictate how
+the file system should handle operations within that subtree.  Below, we
+describe the mechanisms, the policies, and the API for assigning policies to
+subtrees.
 
@@
 Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
-which entity they are performed by (gray boxes). The metadata store and journal
-are different ways of representing the namespace. The metadata store represents
-the namespace as a tree of directory fragments and is easier to read/traverse.
-The journal represent the namespace as a list of events. It is
-a ``pile system"; writes are fast but reads are slow because state must be
-reconstructed.  Specifically, reads are slow because there is more data to
-read, it is unorganized, and many of the updates may be redundant.
+which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
+description of what each mechanism does.  Of the 6 mechanisms in
+Figure~\ref{fig:decouple} only 4 had to implemented and just 1 required changes to
+the underlying storage system itself.
 
-% describe the mechanisms
-Cudele presents 6 mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
-Persist. ``RPCs" does round trip remote procedure calls to establish
-consistency; it is the default implementation for complying with POSIX in
-CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
-into the object store. ``Create" allows clients to append metadata events to an
-in-memory journal. ``Volatile apply" takes the in-memory journal on the client
-and applies it directly to the in-memory metadata store of the metadata server
-cluster. ``Local Persist" takes the in-memory journal and writes it to the client's
-disk. ``Global Persist" saves the journal as a an object in the object store from the
-client.
+\subsubsection{No Changes} ``RPCs" and ``Stream" are part of the default CephFS
+implementation; they are used to get strong consistency and global durability,
+respectively.  Using existing configuration settings in Ceph we can turn
+``Stream" on and off.  If it is off, then the metadata servers will not save
+journals in the object store and the daemons that apply the journal to the
+metadata store will never run.
 
-Next, we discuss how these mechanisms can be composed to get different
-consistency and durability semantics.
+\subsubsection{Library} for ``Create", ``Nonvolatile Apply", ``Local/Global
+Persist", CudeleFS provides a library for clients to link into and all
+operations are performed by the client.  Decoupled clients use the ``Create"
+mechanism to append metadata updates to a local, in-memory journal.  For
+``Local Persist", clients write serialized log events to a file on local disk
+and for ``Global Persist", clients push the journal into the objects store. The
+overheads for both ``Local Persist" and ``Global Persist" is the write
+bandwidth of the local disk and object store, respectively.  ``Nonvolatile
+Apply" replays the client's journal onto the metadata cluster's metadata store.
+The client's in-memory journal is written into the object store and the
+metadata servers are restarted. When the metadata servers re-initialize, they
+notice new journal updates in the object store and replay the events onto their
+in-memory metadata stores.  These implementations required no changes to CephFS
+because the metadata servers know how to read the events the library is writing
+into the object store.  By re-using the journal subsystem to implement the
+namespace decoupling, Cudele leverages the write/read optimized data
+structures, the formats for persisting events (similar to TableFS's
+SSTables~\cite{ren:atc2013-tablefs}), and the functions for replaying events
+onto the internal namespace data structures.
 
-%\begin{tabular}{ r | l }
-%  \(\Rightarrow\)   & Description \\\hline
-%  create            & events appended to in-memory journal \\
-%  v\_apply           & journal volatily applied to in-memory metadata store \\
-%  save              & journal saved to client's disk \\
-%  persist           & journal saved in object store \\
-%  replay            & metadata servers materialize namespace \\
-%  RPCs              & round trip remote procedure calls \\
-%  stream            & metadata server streams journal into RADOS \\
-%\end{tabular}
+% - how it makes no gaurantees
+\subsubsection{Changes to Metadata Server} the ``Volatile Apply" mechanism
+takes an in-memory journal on the client and applies the updates directly to
+the in-memory namespace maintained by the metadata servers. We say volatile
+because -- in exchange for peak performance -- Cudele makes no consistency or
+durability guarantees while ``Volatile Apply" is executing.  If a concurrent
+update from a client occurs there is no rule for resolving conflicts and if the
+client or metadata server crashes there may be no way to recover.
 
-\subsection{Setting Policies with Cudele}
+% difference between apply and volatile apply
+In contrast, ``Nonvolatile Apply" uses the the object store to apply the
+journal of updates from the client onto the metadata servers' metadata store.
+``Nonvolatile Apply" is safer but has a performance overhead because objects in
+the metadata store need to be read from and written back to the object store.
+
+%The metadata store and journal
+%are different ways of representing the namespace.  Cudele presents 6
+%mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
+%Persist. ``RPCs" does round trip remote procedure calls to establish
+%consistency; it is the default implementation for complying with POSIX in
+%CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
+%into the object store. ``Create" allows clients to append metadata events to an
+%in-memory journal. ``Volatile apply" 
+% ``Local Persist" takes the in-memory journal and writes it to the
+%client's disk. ``Global Persist" saves the journal as a an object in the object
+%store from the client.  Next, we discuss how these mechanisms can be composed
+%to get different consistency and durability semantics.
+
+\begin{table}
+\begin{tabular}{ r | l }
+  Mechanism         & Description \\\hline
+  RPCs              & round trip remote procedure calls \\
+  Stream            & stream journal into object store \\
+  Create            & events appended to in-memory journal \\
+  Volatile Apply    & apply to metadata store in obj store \\
+  Nonvolatile Apply & apply to metadata store in memory \\
+  Local Persist     & journal saved to client's disk \\
+  Global Persist    & journal saved in object store \\
+\end{tabular}
+\caption{Cudele's mechanisms are composed together to form consistency and
+durability semantics.\label{table:mechanisms}} 
+\end{table}
+
+\subsection{Defining Policies in Cudele}
 \label{sec:setting-policies-with-cudele}
@@
 \caption{Future programmers can explore the consistency (C) and
-durability (D) spectrums by composing Cudele mechanisms. The consistency
-and durability properties are not guaranteed until all mechanisms in the cell
-are complete ({\it i.e.} the compositions should be considered atomic) and there are
-no guarantees while tranisitioning between policies. \label{table:spectrum}}
+durability (D) spectrums by composing Cudele mechanisms. 
+\label{table:spectrum}}
 \begin{tabular}{ l | l | l | l }
   C \(\rightarrow\) &&& \\  
-  D \(\downarrow\)  	     & invisible         & eventual        & strong  \\\hline
+  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
   none                       & create            & create          & RPCs    \\
@@
 into a global namespace and it is assumed that middleware or the application
-manages consistency lazily; ``eventual" merges updates at some time in the
+manages consistency lazily; ``weak" merges updates at some time in the
 future ({\it e.g.}, when the system has time, when number of updates reaches a
-certain threshold, when the client is done writing, etc.); ; and updates in
-``global" consistency are seen immediately by all clients. For durability,
+certain threshold, when the client is done writing, etc.); and updates in
+``strong" consistency are seen immediately by all clients. For durability,
 ``none" means that updates are volatile and will be lost on a failure. Stronger
@@
 have global consistency and durability; DeltaFS uses ``invisible" consistency
-and ``local" durability and BatchFS uses ``eventual" consistency and ``local"
-durability. These are just a few of the HPC examples.
-
-% how its done
-To compose the mechanisms administrators inject which steps (described in
-Section~\S\ref{sec:cudeles-mechanisms}) to run and which to use in parallel
-using a domain specific language. For example, to get the semantics of BatchFS,
-the administrator would inject the following pipeline:\\
-
-\noindent \texttt{create+save+volatile apply}\\
+and ``local" durability and BatchFS uses ``weak" consistency and ``local"
+durability.  To compose the mechanisms administrators inject which mechanisms
+(described in Section~\S\ref{sec:cudeles-mechanisms}) to run and which to use
+in parallel using a domain specific language.
 
@@
 Table~\ref{table:spectrum}, not all of them make sense. For example, it
-makes little sense to do \texttt{creates+RPCs} since both steps do the same
+makes little sense to do \texttt{creates+RPCs} since both mechanisms do the same
 thing or \texttt{stream+save} since ``global" durability is stronger and has
-more overhead than ``local" durability.  Formulaically, valid steps
-include:\\
+more overhead than ``local" durability.
 
-\noindent \texttt{<create|RPCs> [+v\_apply][+save][+persist][+stream]}
+% talk of eventual consistency
+The consistency and durability properties in Table~\ref{table:spectrum} are not
+guaranteed until all mechanisms in the cell are complete. The compositions
+should be considered atomic and there are no guarantees while transitioning
+between policies. For example, updates are not deemed globally consistent until
+they are safely saved in the object store. If a failure occurs during ``global
+persist" or if we inject a new policy that chanages a subtree from ``local
+persist" to ``global persist", CudeleFS no guarantee until the mechanisms are
+complete.
 
-\subsection{Cudele Namespace API:\\Per-Subtree Policies}
-\label{sec:cudele-namespace-api-per-subtree-policies}
+\subsection{Cudele Namespace API}
+\label{sec:cudele-namespace-api}
 
@@
 would decouple the path \texttt{msevilla/mydir} and would apply the policies in
-\texttt{policies.txt}. The policies file supports the following values:
+\texttt{policies.yml}. The policies file supports the following values:
 
@@
   \item \texttt{interfere\_callback}: how to handle a request from another
-  client targeted at the now decoupled subtree (default \texttt{block})
+  client targeted at the now decoupled subtree (default \texttt{overwrite})
 
@@
 
+% description
+For \texttt{block}, any requests to this part of the namespace returns with
+``Device is busy", which will spare the metadata server from wasting resources
+for updates that may get overwritten. If the application does not mind losing
+updates, for example it wants approximations for results that take too long to
+compute, it can select \texttt{overwrite}. In this case, metadata will be
+written and the computation from the decoupled namespace will take priority at
+merge time because the results are more accurate.
+
+% examples
 Given these default values decoupling the namespace with an empty policies file
 would give the application 100 inodes but the subtree would behave like the
-existing CephFS implementation. Table~\ref{table:spectrum} shows how one would use the Cudele
-API to implement policies from related work.
-
-%
-
-For \texttt{block}, any requests to this part of the namespace returns with
-``Device is busy", which will spare the metadata server from wasting resources for updates
-that may get overwritten. If the application does not mind losing updates, for
-example it wants approximations for results that take too long to compute, it
-can select \texttt{overwrite}. In this case, metadata will be written and the
-computation from the decoupled namespace will take priority at merge time because the results
-are more accurate.
+existing CephFS implementation. To implement DeltaFS on CudeleFS, the user
+would use the configuration from Listing~\ref{src:deltafs}. BatchFS's merging
+back into the global namespace can be achieved with the configuration in
+Listing~\ref{src:batchfs}.
 
@@
 \end{minted}
-\caption{Implementing DeltaFS with Cudele.}
+\caption{Implementing DeltaFS with CudeleFS.}
 \label{src:deltafs}
@@
 \end{minted}
-\caption{Implementing BatchFS with Cudele.}
+\caption{Implementing BatchFS with CudeleFS.}
 \label{src:batchfs}
@@
 
-\begin{listing}
-\begin{minted}[frame=single,
-               framesep=3mm,
-               xleftmargin=21pt,
-               tabsize=4]{js}
-{     
-    "allocated_inodes": "-"
-    "interfere_policy": "-"
-    "consistency": "RPCs"
-    "durability": "stream"
-}
-\end{minted}
-\caption{Existing CephFS on Cudele.}
-\label{src:batchfs}
-\end{listing}
-
-\subsection{Programmability Approach}
-
-% what is programmability
-A programmable storage system exposes internal subsystem as building blocks for
-higher level services. This `dirty-slate' approach limits redundant code and
-leverages the robustness of the underlying storage system. Cudele uses this
-approach and re-uses some of the building blocks from the Malacology
-programmable storage system~\ref{sevilla:eurosys17} to great success and
-requires only:
-
-\begin{itemize}
-
-  \item 354 lines of library code
-
-  \item 219 lines of non-destructive metadata server code, which is not used
-  unless it is turned on
-
-  \item 4 lines of destructive client/server code to check whether a namespace
-  is decoupled
-
-\end{itemize}
-
-% how it works in CephFS, how I used it
-\subsubsection{Metadata Store}
-
-\subsubsection{Journal}
-
-\subsubsection{Journal Tool}
-The journal tool is used for disaster recovery and lets administrators view and
-modify the journal of metadata updates; it can read the journal, erase events
-from the journal, and apply the updates in the journal to the metadata store.
-To apply journal updates to the metadata store, the journal tool reads the
-journal segments from object store objects and applies the update to the
-metadata store (which are also stored as object store objects).
-
-% Journal import
-The journal tool imports journals from binary files stored on disk.  First the
-header of the dump is sanity checked and written into RADOS to the ``header"
-object.  The ``header" object has metadata about the journal as well as the
-locations of all the journal pointers (e.g., where the tail of the journal is,
-where we are currently trimming, etc.).  Then the journal events are cleaned
-(erasing trailing events that are not part of the header) and written as
-objects into RADOS.  Note that while the journal is in RADOS, the metadata
-servers do do not have the namespace reconstructed in memory so the metadata
-cluster will not service requests relating to the journal of imported events.
-To construct the namespace in the collective memory of the metadata servers we
-need to first construct the namespace in RADOS. The journal tool can explicitly
-do this by  applying the journal to the metadata store in RAODS. This will pull
-the objects containing journal segments and replay them on the metadata store.
-Finally, we delete the journal in RADOS and restart the metadata servers so
-they rebuild their caches.
-
-% Journal export
-The journal tool exports journals to binary files stored on disk. First the
-journal is scanned for the header and then journal is recovered. To recover the
-journal the ``header" object is read off disk and then objects are probed in
-order and starting from the write position saved in the header. Probing will
-update the write position if it finds objects with data in them.
-
-% Journal export
-When exporting a journal of events, the journal tool first scans the journal to
-check for corruption. Then it recovers the journal by reading the ``header"
-object out of RADOS.  After reading the header, the journal tool can pull
-journal segments from RADOS because it knows how many objects to pull and how
-far to seek within those objects.
-
-%% The data structures
-%The metadata that the event touches, including inodes, paths, and timestamps,
-%are stored as metablobs. Each piece of metadata inside a metablob is called a
-%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
-%table client transactions (tid, version), renamed directory fragments (maps,
-%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
-%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
-%me (and you since you are reading this paper), this does not make any sense.
-%
-%Each directory fragment has an associated directory lump, which is just a bunch
-%of metadata. The most interesting part of the dirlump is the fullbits array,
-%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
-%and then all the full bits, saving off the children and inode locations. The
-%children tell us which dentry names an inode has and the inode locations map
-%the parent inode to its child inode and dentry.
-
-\subsubsection{Inode Cache}
-
-\subsubsection{Large Inodes}
-
-% file type
-To assign consistency and durability to the subtrees we store policies in
-the directory inode. This approach uses the File Type interface from the
-Malacology programmable store system~\cite{sevilla:eurosys17} and it tells clients how to access
-the underlying data or metadata. The underlying implementation stores
-executable code in the inode that calls the different Cudele mechanisms. Of
-course, there are many security and access control aspects of this approach but
-that is beyond the scope of this paper.
-
-% lead into implementation
-
-% // limit consistency
-% if metadata write (create) -- try to reduce error bars for many creates and runtime of one touch)
-%   if !RPCs
-%     if overwrite
-%       do nothing and let the RPC go
-%     else
-%       return -EBUSY
-%
-% // limit durability
-% if !stream
-%   return immediately on mdlog
+%\begin{listing}
+%\begin{minted}[frame=single,
+%               framesep=3mm,
+%               xleftmargin=21pt,
+%               tabsize=4]{js}
+%{     
+%    "allocated_inodes": "-"
+%    "interfere_policy": "-"
+%    "consistency": "RPCs"
+%    "durability": "stream"
+%}
+%\end{minted}
+%\caption{Existing CephFS on Cudele.}
+%\label{src:batchfs}
+%\end{listing}

### 2017-04-01 paper/sections/5_implementation.tex
paper: integrate noah section 3 feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-% Outline of the section
-Each section below corresponds to the labeled arrows in
-Figure~\ref{fig:decouple}. This implementation decouples policies from
-mechanisms allowing applications to choose the consistency and durability
-semantics they need.
+A programmable storage system exposes internal subsystem as building blocks for
+higher level services. This `dirty-slate' approach limits redundant code and
+leverages the robustness of the underlying storage system. Cudele uses this
+approach and re-uses some of the building blocks from the Malacology
+programmable storage system~\cite{sevilla:eurosys17} to great success and
+requires only:
 
-% Programmability
-Of the 6 mechanisms in Figure~\ref{fig:decouple} 4 had to implemented and only
-1 required changes to the underlying storage system itself.  ``RPCs" and
-``Stream" can be achieved with configuration settings 
-.  ``Global Persist", ``Local Persist",
-``Nonvolatile Apply", and ``Create" are implemented as a library and they do
-not require modifications to the storage system. ``Volatile Apply" requires
-changes to the metadata server to inject updates back into the global
-namespace.
+\begin{itemize}
 
-\subsection{No Changes: RPCs, Stream}
-\label{sec:no-changes-rpcs-stream}
+  \item 354 lines of library code
 
-RPCs is the default behavior of CephFS.  Using existing configuration settings
-({\it e.g.}, metadata cache size, logging on/off, etc.) in Ceph we can turn
-``Stream" on and off.  If it is off, then the metadata servers will not save
-journals in the object store and the daemons that apply the journal to the
-metadata store will never run.
+  \item 219 lines of non-destructive metadata server code, which is not used
+  unless it is turned on
+
+  \item 4 lines of destructive client/server code to check whether a namespace
+  is decoupled
+
+\end{itemize}
+
+% how it works in CephFS, how I used it
+\subsection{Metadata Store}
+\label{sec:metadata-store}
+
+% how it works in CephFS
+In CephFS, the metadata store is a data structure that represents the file
+system namespace. This data structure is stored in two places: in memory ({\it
+i.e.} in the collective memory of the metadata server cluster) and as objects
+in the object store. In the object store, directories and their inodes are
+stored together in objects to improve the performance of scans.  The metadata
+store data structure is structured as a tree of directory fragments making it
+easier to read and traverse.\\
+
+% how i used it
+\noindent\textbf{Cudele}: uses the metadata store format to write objects
+formatted like the journal into the object store. By writing with the same
+format, the metadata servers can read and use the recovery code to materialize
+the updates from a client's decoupled namespace ({\it i.e.} merge).
+
+\subsection{Journal}
+\label{sec:journal}
+
+The journal is the second way that CephFS represents the file system namespace;
+it is a log of metadata updates that can materialize the namespace when the
+updates are replayed onto the metadata store. The journal is a ``pile system";
+writes are fast but reads are slow because state must be reconstructed.
+Specifically, reads are slow because there is more state to read, it is
+unorganized, and many of the updates may be redundant.\\
+
+\noindent\textbf{Cudele}: uses the journal subsystems within the metadata
+server to replay updates onto the in-memory metadata store.  When the clients
+are ready to merge their updates back into the global namespace, they pass a
+binary file of  metadata updates to the metadata server. The metadata server
+uses its recovery code to ``replay" the updates.
+
+\subsection{Journal Tool}
+\label{sec:journal-tool}
+
+The journal tool is used for disaster recovery and lets administrators view and
+modify the journal. It can read the journal, export the journal as a file,
+erase events, and apply updates to the metadata store.  To apply journal
+updates to the metadata store, the journal tool reads the journal from object
+store objects and replays the updates on the metadata store in the object
+store.\\
+
+\noindent\textbf{Cudele}: using the journal tool, clients can materialize the
+journal in memory and append events to their own copy. When they are done they
+export their journal to a binary file. Clients also use the journal tool to
+save their updates to disk or to the object store.
+
+% Journal import
+%The journal tool imports journals from binary files stored on disk.  First the
+%header of the dump is sanity checked and written into RADOS to the ``header"
+%object.  The ``header" object has metadata about the journal as well as the
+%locations of all the journal pointers (e.g., where the tail of the journal is,
+%where we are currently trimming, etc.).  Then the journal events are cleaned
+%(erasing trailing events that are not part of the header) and written as
+%objects into RADOS.  Note that while the journal is in RADOS, the metadata
+%servers do do not have the namespace reconstructed in memory so the metadata
+%cluster will not service requests relating to the journal of imported events.
+%To construct the namespace in the collective memory of the metadata servers we
+%need to first construct the namespace in RADOS. The journal tool can explicitly
+%do this by  applying the journal to the metadata store in RAODS. This will pull
+%the objects containing journal segments and replay them on the metadata store.
+%Finally, we delete the journal in RADOS and restart the metadata servers so
+%they rebuild their caches.
+
+% Journal export
+%First the
+%journal is scanned for the header and then journal is recovered. To recover the
+%journal the ``header" object is read off disk and then objects are probed in
+%order and starting from the write position saved in the header. Probing will
+%update the write position if it finds objects with data in them.
+
+% Journal export
+%When exporting a journal of events, the journal tool first scans the journal to
+%check for corruption. Then it recovers the journal by reading the ``header"
+%object out of RADOS.  After reading the header, the journal tool can pull
+%journal segments from RADOS because it knows how many objects to pull and how
+%far to seek within those objects.
+
+%% The data structures
+%The metadata that the event touches, including inodes, paths, and timestamps,
+%are stored as metablobs. Each piece of metadata inside a metablob is called a
+%dirlump. A dump has a section for lumps (dirfrag, dirlump), roots (dentries),
+%table client transactions (tid, version), renamed directory fragments (maps,
+%versions, alloc/preallocated inodes), inodes starting a truncate, inodes
+%finishing a truncate, destroyed inodes, and client requests. Unfortunately for
+%me (and you since you are reading this paper), this does not make any sense.
+%
+%Each directory fragment has an associated directory lump, which is just a bunch
+%of metadata. The most interesting part of the dirlump is the fullbits array,
+%which has a dentry OR inode. To walk the tree, iterate over all the dirlumps
+%and then all the full bits, saving off the children and inode locations. The
+%children tell us which dentry names an inode has and the inode locations map
+%the parent inode to its child inode and dentry.
+
+\subsection{Inode Cache}
+\label{sec:inode-cache}
+
+In CephFS, the inode cache reduces the number of RPCs between clients and
+metadata servers. Without contention clients can resolve metadata reads locally
+thus reducing operations like \texttt{lookup()}.  For example, if a client or
+metadata server is not caching the directory inode, all creates within that
+directory will result in a lookup and a create request. If the directory inode
+is cached then only the create needs to be sent.  The size of the inode cache
+is configurable so as not to saturate the memory on the metadata server --
+inodes in CephFS are about 1400
+bytes\footnote{http://docs.ceph.com/docs/jewel/dev/mds\_internals/data-structures/}.
+The inode cache has code for manipulating inode numbers, such as pre-allocating
+inodes to clients.\\
+
+\noindent\textbf{Cudele}: uses the internal inode cache code to allocate inodes
+to clients that decouple parts of the namespace and to skip inodes used by the
+client at merge time.
+
+\subsection{Large Inodes}
+\label{sec:large-inodes}
+
+In CephFS, inodes store policies for policies, like how the file is striped
+across the object store or for managing subtrees for load balancing. These
+policies adhere to logical partitionings of metadata or data, like Ceph pools and file
+system namespace subtrees. To implement this, the namespace data structure has
+the ability to recursively apply policies to subtrees and to isolate subtrees
+from each other.  \\
+
+\noindent\textbf{Cudele}: uses the large inodes to store consistency and
+durability policies in the directory inode. This approach uses the File Type
+interface from the Malacology programmable store
+system~\cite{sevilla:eurosys17} and it tells clients how to access the
+underlying metadata. The underlying implementation stores executable code in
+the inode that calls the different Cudele mechanisms. Of course, there are many
+security and access control aspects of this approach but that is beyond the
+scope of this paper.
 
@@
 
-\subsection{Library: Create, Local/Global Persist, Nonvolatile Apply}
-\label{sec:library-create-local-global-persist}
 %Reads/Writes Journal Events}
@@
 % How it works
-For these mechanisms, Cudele provides a library for clients to link into.  For
-``Create", clients write to an in-memory journal and updates are merged into
-the global namespace by replaying them onto the metadata store either in the
-metadata servers or the object store.  For ``Local Persist", clients write
-serialized log events to a file on local disk and for ``Global Persist",
-clients push the journal into the objects store. The overheads for both ``Local
-Persist" and ``Global Persist" is the write bandwidth of the local disk and
-object store, respectively. Finally, for ``Nonvolatile Apply" the journal
-updates are replayed onto the metadata store in the object store and the
-metadata servers are restarted.
-
-% Level of intrusion
-These implementations required no changes to Ceph because the metadata servers
-know how to read the events the library is writing into the object store.  By
-re-using the journal subsystem to implement the namespace decoupling, Cudele
-leverages the write/read optimized data structures, the formats for persisting
-events (similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the
-functions for replaying events onto the internal namespace data structures.
-
 % Why we re-use stuff
@@
 %
-%\subsubsection{Operating on Snapshots} 
+%\subsection{Operating on Snapshots} 
 %
@@
 
-\subsection{Storage System Changes: Volatile Apply}
-
-% - how it makes no gaurantees
-The ``Volatile Apply" mechanism takes an in-memory journal on the client and
-applies the updates directly to the in-memory namespace maintained by the
-metadata servers. We say volatile because -- in exchange for peak performance
--- Cudele makes no consistency or durability guarantees while the mechanism is
-executing.  If a concurrent update from a client occurs there is no rule for
-resolving conflicts and if the client or metadata server crashes there may be
-no way to recover.
-
-% difference between apply and volatile apply
-In contrast, ``Nonvolatile Apply" uses the the object store to merge the journal of updates
-from the client to the metadata server. ``Nonvolatile Apply" is safer but
-has a performance overhead because objects in the metadata store need to be
-read from and written back to the object store.  ``Volatile Apply" was already implemented
-by the journal tool: it reads the journal from the object store, adds events to
-the journal, and writes the metadata updates out to the metadata store in the
-object store.
-
 % - how it actually works

### 2017-04-01 paper/sections/5_implementation.tex
paper: wordsmith section 5 and add conclusion/discussion
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-A programmable storage system exposes internal subsystem as building blocks for
-higher level services. This `dirty-slate' approach limits redundant code and
-leverages the robustness of the underlying storage system. Cudele uses this
-approach and re-uses some of the building blocks from the Malacology
-programmable storage system~\cite{sevilla:eurosys17} to great success and
-requires only:
-
-\begin{itemize}
-
-  \item 354 lines of library code
-
-  \item 219 lines of non-destructive metadata server code, which is not used
-  unless it is turned on
-
-  \item 4 lines of destructive client/server code to check whether a namespace
-  is decoupled
-
-\end{itemize}
+We use a ``programmable storage" approach~\cite{sevilla:eurosys17} to design
+CudeleFS; namely, we try to leverage components inside Ceph to inherit the
+robustness and correctness of the internal subsystems. In this section, we
+first describe a Ceph internal subsystem or component and then we show how we
+use it in CudeleFS.
 

### 2017-04-01 paper/sections/6_results.tex
paper: wordsmith section 5 and add conclusion/discussion
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\subsection{Eventual vs. Strong Consistency}
+\subsection{Weak vs. Strong Consistency}
 \begin{figure}[tb]
@@
 \caption{The RPC per metadata update of ``Strong Consistency" has a large
-overhead compared to the decoupled namespace strategy of ``Eventual
+overhead compared to the decoupled namespace strategy of ``Weak
 Consistency".\label{fig:slowdown-strong-weak}}
@@
   \end{itemize}
-  \item Eventual Consistency
+  \item Weak Consistency
   \begin{itemize}
@@
   \end{itemize}
-  \item Eventual Consistency, No Fault Tolerance
+  \item Weak Consistency, No Fault Tolerance
   \begin{itemize}
@@
 % Does not scale?
-% Strong Consistency > 10X slower than Eventual Consistency 
-\subsubsection{Speedups of Decoupled Namespaces} Eventual consistency uses the
+% Strong Consistency > 10X slower than Weak Consistency 
+\subsubsection{Speedups of Decoupled Namespaces} Weak consistency uses the
 decoupled namespace strategy and shows up to a \(20\times\) speedup over the
@@
 \(5-7\times\) for Strong Consistency, which emulates BatchFS and
-\(90-104\times\) for Eventual Consistency, which emulates DeltaFS.
+\(90-104\times\) for Weak Consistency, which emulates DeltaFS.
 
@@
 we scale the number of files because the cost of streaming the journal into the
-object store is constant. We omit the same analysis for ``Eventual Consistency"
+object store is constant. We omit the same analysis for ``Weak Consistency"
 because the runtimes are so short that the normalized slowdowns are misleading.
 
-% Eventual Consistency benefit not due to metadata format
+% Weak Consistency benefit not due to metadata format
 \subsubsection{Metadata Formats} Because the metadata formats are the same for
@@
 
-\subsection{Isolation from EBUSY}
-
+\subsection{Benefits of Isolated Subtree Policies}
 % setup: 2cs sep dirs, 2cs DN, 1c malicious write
@@
 
-\subsubsection{Benefits of Isolated Subtree Policies}
+\subsection{Scaling Concurrent Merges}
 
-\subsubsection{Cost of Merging}
-\subsubsection{Scaling Concurrent Merges}
+\subsection{Partial Directory Listings}
 
-\subsection{Isolation from Global Writes, Cost of Overwrites}
 \subsection{Cost of Nonvolatile Apply}
@@
 existing implementation results in this object being repeatedly pushed/pulled.
-
-\subsection{Partial Directory Listings}

### 2017-04-01 paper/sections/7_conclusion.tex
paper: wordsmith section 5 and add conclusion/discussion
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Discussion}

% Could we have implemented this on something other than CephFS?
One question that comes up often in our work with Ceph is: ``Why do we always
choose Ceph and can we implement this on another storage system?" Ceph is
popular, robust, and open source -- getting it merged into mainline has a
better chance of getting used than if we had built the system `clean-slate'
from the ground up. Using Ceph, we can also leverage its robustness using an
approach called ``programmability".

% How awesome programmability is
``Programmability" is a way of designing storage systems that encourages code
re-use and composition.  A programmable storage system exposes internal
subsystem as building blocks for higher level services. This `dirty-slate'
approach limits redundant code and leverages the robustness of the underlying
storage system. Cudele uses this approach and re-uses some of the building
blocks from the Malacology programmable storage system~\cite{sevilla:eurosys17}
to great success and requires only:

\begin{itemize}

  \item 354 lines of library code

  \item 219 lines of non-destructive metadata server code, which is not used
  unless it is turned on

  \item 4 lines of destructive client/server code to check whether a namespace
  is decoupled

\end{itemize}

\section{Conclusion and Future Work}

Relaxing consistency and durability semantics in the file system is a
double-edged sword. While it achieves better performance and scalability, it
alienates applications that rely on strong consistency and durability. CudeleFS
lets administrators assign consistency and durability guarantees to subtrees in
the global namespace, giving administrators a wide range of policies
optimizations that can be custom fit to the application. Using CudeleFS, we
compare related work on the same file system and conclude that strong
consistency can cause a 104\(\times\) slow down while merging updates and
maintaining durability only have between a \(7-10\times\) slow down.

% Implied Namespaces

% Intermediate Update Bursts

% See if this helps load balancing

### 2017-04-01 paper/sections/4_methodology.tex
Uses times as main font

@@
 \label{sec:methodology-decoupled-namespaces}
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=90mm]{figures/fig-decouple.png}
-\caption{Applications decouple the namespace, write updates to a local journal,
-and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
-\end{figure}
 
@@
 
+\begin{figure}[tb]
+\centering
+\includegraphics[width=90mm]{figures/fig-decouple.png}
+\caption{Applications decouple the namespace, write updates to a local journal,
+and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
+\end{figure}
+
 % - how it makes no gaurantees
@@
 
-\begin{table}
-\begin{tabular}{ r | l }
-  Mechanism         & Description \\\hline
-  RPCs              & round trip remote procedure calls \\
-  Stream            & stream journal into object store \\
-  Create            & events appended to in-memory journal \\
-  Volatile Apply    & apply to metadata store in obj store \\
-  Nonvolatile Apply & apply to metadata store in memory \\
-  Local Persist     & journal saved to client's disk \\
-  Global Persist    & journal saved in object store \\
-\end{tabular}
-\caption{Cudele's mechanisms are composed together to form consistency and
-durability semantics.\label{table:mechanisms}} 
-\end{table}
-
 \subsection{Defining Policies in Cudele}
 \label{sec:setting-policies-with-cudele}
-
-\begin{table}[t]
-\begin{center}
-\caption{Future programmers can explore the consistency (C) and
-durability (D) spectrums by composing Cudele mechanisms. 
-\label{table:spectrum}}
-\begin{tabular}{ l | l | l | l }
-  C \(\rightarrow\) &&& \\  
-  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
-  none                       & create            & create          & RPCs    \\
-                             &                   & +volatile apply &         \\\hdashline
-  local                      & create            & create          & RPCs    \\
-                             & +local persist    & +local persist  & +local  \\
-                             &                   & +volatile apply &  persist\\\hdashline
-  global                     & create            & create          & RPCs    \\
-                             & +global persist   & +global persist & +stream \\
-                             &                   & +volatile apply &         \\
-\end{tabular}
-\end{center}
-\end{table}
 
@@
 complete.
+
+\begin{table}
+\begin{tabular}{ r | l }
+  Mechanism         & Description \\\hline
+  RPCs              & round trip remote procedure calls \\
+  Stream            & stream journal into object store \\
+  Create            & events appended to in-memory journal \\
+  Volatile Apply    & apply to metadata store in obj store \\
+  Nonvolatile Apply & apply to metadata store in memory \\
+  Local Persist     & journal saved to client's disk \\
+  Global Persist    & journal saved in object store \\
+\end{tabular}
+\caption{Cudele's mechanisms are composed together to form consistency and
+durability semantics.\label{table:mechanisms}} 
+\end{table}
+
+\begin{table}[t]
+\begin{center}
+\caption{Future programmers can explore the consistency (C) and
+durability (D) spectrums by composing Cudele mechanisms. 
+\label{table:spectrum}}
+\begin{tabular}{ l | l | l | l }
+  C \(\rightarrow\) &&& \\  
+  D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
+  none                       & create            & create          & RPCs    \\
+                             &                   & +volatile apply &         \\\hdashline
+  local                      & create            & create          & RPCs    \\
+                             & +local persist    & +local persist  & +local  \\
+                             &                   & +volatile apply &  persist\\\hdashline
+  global                     & create            & create          & RPCs    \\
+                             & +global persist   & +global persist & +stream \\
+                             &                   & +volatile apply &         \\
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 
-\begin{listing}
-\begin{minted}[frame=single,
-               framesep=3mm,
-               xleftmargin=21pt,
-               tabsize=4]{js}
-{     
-    "allocated_inodes": "100000"
-    "interfere_policy": "block"
-    "consistency": "create"
-    "durability": "local_persist"
-}
-\end{minted}
-\caption{Implementing DeltaFS with CudeleFS.}
-\label{src:deltafs}
-\end{listing}
-
-\begin{listing}
-\begin{minted}[frame=single,
-               framesep=3mm,
-               xleftmargin=21pt,
-               tabsize=4]{js}
-{     
-    "allocated_inodes": "100000"
-    "interfere_policy": "block"
-    "consistency": "create+volatile_apply"
-    "durability": "local_persist"
-}
-\end{minted}
-\caption{Implementing BatchFS with CudeleFS.}
-\label{src:batchfs}
-\end{listing}
-
 %\begin{listing}

### 2017-04-01 paper/build.sh

@@ -4,13 +4,12 @@ rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.sync
 set -ex
 
 # Build the paper
-docker rm -f latex || true
-docker run \
+docker run --rm \
   --name latex \
   --entrypoint=/bin/bash \
   -v `pwd`/:/mnt \
-  ivotron/texlive_v2:latest -c \
-    "cd /mnt ; \
+  ivotron/texlive:latest -c \
+    "cd /mnt && \
      pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper && \
      bibtex paper && \
      pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper && \
@@ -27,4 +26,4 @@ set +e
 rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.synctex.gz build.log >> /dev/null 2>&1
 
 echo "SUCCESS"
-exit 0
+

### 2017-04-01 paper/build.sh
popper: fix paper build
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -9,7 +9,7 @@ docker run \
   --name latex \
   --entrypoint=/bin/bash \
   -v `pwd`/:/mnt \
-  ivotron/texlive_v2:latest -c \
+  michaelsevilla/texlive:latest -c \
     "cd /mnt ; \
      pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper && \
      bibtex paper && \

### 2017-04-01 paper/sections/0_abstract.tex
Fixes typos and minor grammar edits

@@
 deploy them on a different system.  We present a file system and API that
-lets clients to specify their consistency/durability requirements and assign
-them to subtrees in the namespace, allowing administrators to optimize subtrees
+lets clients specify their consistency/durability requirements and assign
+them to subtrees in the namespace, allowing users to optimize subtrees
 within the same namespace for different workloads.  We draw conclusions about

### 2017-04-01 paper/sections/1_introduction.tex
Fixes typos and minor grammar edits

@@
 BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
-do more client side processing and merge updates when the job is done. HPC
-developers are turning to this non-POSIX solution because their applications
+do more client-side processing and merge updates when the job is done. HPC
+developers are turning to these non-POSIX solutions because their applications
 are well-understood ({\it e.g.}, well-defined read/write phases,
@@
 synchronization, false sharing, and serialization.  The consistency and
-durability semantics for these systems is shown in the table in
-Figure~\ref{fig:subtree-policies} . While the performance benefits are obvious
+durability semantics for these systems is shown in
+Figure~\ref{fig:subtree-policies} (table) . While the performance benefits are obvious
 for these users, applications that rely on the file system's guarantees must be
@@
 % What did we do
-We propose subtree policies, an interface that lets future programmers control
+We propose a subtree policy API that lets future programmers control
 the consistency and durability for subtrees in the file system namespace. For performance, one
@@
 
-  \item a comparison of the strategies used in recently proposed research systems against
-  previously unexplored metadata designs.
+  \item an apples-to-apples comparison of the strategies used in recently proposed research systems against
+  previously unexplored metadata designs, all implemented using Cudele.
 
@@
 Our results confirm the assertions of ``clean-state" research systems that
-decouple namespaces; specifically that the technique drastically improve
+decouple namespaces; specifically that the technique drastically improves
 performance (104\(\times\) speed up) but we go a step further by quantifying
@@
 file format in systems that are based on in-memory data structures.
-Section~\ref{sec:related-work} places Cudele in the context of other related
-work. Section~\ref{sec:posix-overheads} quantifies the cost of POSIX
+Section~\ref{sec:posix-overheads} quantifies the cost of POSIX
 consistency and system-defined durability and
@@
 explored and previously unexplored metadata designs.
+Section~\ref{sec:related-work} places Cudele in the context of other related
+work.

### 2017-04-01 paper/sections/2_related_work.tex
Fixes typos and minor grammar edits

@@
 file operations only require 1 RPC and for weak scaling because requests will
-never incurr extra RPCs due to a full cache.  CephFS employs both techniques to
+never incur extra RPCs due to a full cache.  CephFS employs both techniques to
 a lesser extent; directories can be replicated or sharded but the caching and
@@
 scatter-gather process; this halts updates on the directories and blocks until
-the authoratitive metadata server responds.  These protocols are discusssed in
+the authoritative metadata server responds.  These protocols are discussed in
 more detail in the next section but their inclusion here is a testament to the
@@
 accesses to the namespace, so the clients can batch local operations and merge
-with a global namespace image lazily. Similarily, DeltaFS eliminates RPC
+with a global namespace image lazily. Similarly, DeltaFS eliminates RPC
 traffic using subtree snapshots for non-conflicting workloads and middleware
@@
 merges, applications needing some degree of ground truth can either manage
-consistency thesmelves on a read or add a bolt-on service to manage the
+consistency themselves on a read or add a bolt-on service to manage the
 consistency.

### 2017-04-01 paper/sections/3_posix_overheads.tex
Uses times as main font
Fixes typos and minor grammar edits

@@
 \includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
-\caption{Create-heavy workloads (untar) incurr the highest disk, network, and
+\caption{Create-heavy workloads (such as \texttt{untar}) incur the highest disk, network, and
 CPU utilization because of the consistency and durability demands of
@@
 
-In our examiniation of the overheads of POSIX we benchmark and analyze CephFS,
+In our examination of the overheads of POSIX we benchmark and analyze CephFS,
 the file system that uses the Ceph's object store ({\it i.e.} RADOS) to store
@@
 research because of the checkpoint/restart use case but they also happen to
-stress the underyling storage system the most.
+stress the underlying storage system the most.
 Figure~\ref{fig:overhead-creates} shows the resource utilization of compiling
-the Linux kernel.  The untar phase, which is characterized by many creates, has
+the Linux kernel.  The \texttt{untar} phase, which is characterized by many creates, has
 the highest resource usage which suggests that it is stressing the consistency
@@
 the kernel client so that we can find the true create speed of the server; our
-experiments show a low CPU utilziation for the clients which indicates that we
+experiments show a low CPU utilization for the clients which indicates that we
 are stressing the servers more.
@@
 \noindent\textbf{CephFS Design}: a journal of metadata updates that streams into
-the resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
-WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to grow to large which
-ensures (1) sequential writes into the object store and (2) the ability for daemons to
-trim redundant or irrelevent journal entries.  The journal is striped over
+the resilient object store, similar to LFS~\cite{rosenblum:acm1992-LFS} and
+WAFL~\cite{hitz:wtec1994-WAFL}, grows to large extents which
+allows to (1) sequentially write into the object store and (2) provide the ability for daemons to
+trim redundant or irrelevant journal entries.  The journal is striped over
 objects where multiple journal updates can reside on the same object. There are
@@
 As shown in Figure~\ref{fig:journal}, in addition to the metadata journal,
-CephFS also represents metadata in RADOS as a metadata store, where directories
+CephFS also stores metadata in RADOS, where directories
 and their file inodes are stored as objects.  The metadata server applies the
-updates in the journal to the metadata store when the journal reaches a certain
+updates in the journal to this metadata store when the journal reaches a certain
 size. The metadata store is optimized for recovery ({\it i.e.} reading) while
@@
 size the bigger that the writes into the object store are. The trade-off for
-better performance is memory consumption because larger segment sizes take up
-more space with their buffers. Figure~\ref{fig:slowdown-journal-b} shows how
+better performance is memory consumption because larger segments take up
+more space for buffering. Figure~\ref{fig:slowdown-journal-b} shows how
 the metadata server periodically stops serving requests to flush ({\it i.e.}
-apply journal updates to) the metadata store.  The journal overhead is
+apply journal updates to) to the metadata store.  The journal overhead is
 sufficient enough to slow down metadata throughput but not so much as to
-overwhelm the bandidth of the object store. We measured our peak bandwidth to
+overwhelm the bandwidth of the object store. We measured our peak bandwidth to
 be 100MB/s, which is the speed of our network link.\\
@@
 \noindent\textbf{Comparison to decoupled namespaces}: In BatchFS and DeltaFS,
-as far as we can tell, when a client or server fails there is no recovery
+to the best of our knowledge, when a client or server fails there is no recovery
 scheme. For BatchFS, if a client fails when it is writing to the local
 log-structured merged tree (implemented as an SSTable) then those batched
-metadata operations are lost. For DeltaFS, if the client fails then on restart
+metadata operations are lost. For DeltaFS, if the client fails then, on restart,
 the computation does the work again -- since the snapshots of the namespace are
@@
 BatchFS and DeltaFS use IndexFS. IndexFS writes metadata to SSTables, which
-initially reside in memory but are later to flushed to the underlying
+initially reside in memory but are later flushed to the underlying
 distributed file system.
@@
 consistent. To reduce the number of RPCs needed for consistency, clients can
-obtain capabilities for reading, reading and updating, caching reads, writing,
+obtain capabilities for reading, creating and updating inodes, as well as caching reads,
 buffering writes, changing the file size, and performing lazy IO.
@@
   \end{subfigure}
-  \caption{When a client create stream is ``isolated" then lookups resolve
-  locally but when a second client ``interferes" by creating in the same
+  \caption{When a client create stream is ``isolated" lookups resolve
+  locally; but when a second client ``interferes" by creating in the same
   directory, the directory inode capability is revoked forcing all clients to
@@
 with 7 clients in Figure~\ref{fig:interfere-c} we see how different
-interferring operations affect performance, again when compared to the
+interfering operations affect performance, again when compared to the
 performance of a single isolate client.  ``create" has little effect but
@@
 completes.  In BatchFS the merge is delayed by the application using an API to
-switch between asynchronous to sychronous mode. The merge itself is explicitly
+switch between asynchronous to synchronous mode. The merge itself is explicitly
 managed by the application but future work looks at more automated
-methdologies. In DeltaFS snapshots of the metadata subtrees stays on the client
+methodologies. In DeltaFS snapshots of the metadata subtrees stays on the client
 machines; there is no ground truth and consistent namespaces are constructed

### 2017-04-01 paper/sections/4_methodology.tex
Fixes typos and minor grammar edits

--- 2017-04-01 paper/sections/4_methodology.tex
@@
 % describe table
-The spectrum of consistency and durability guarantees that adminstrators can
+The spectrum of consistency and durability guarantees that administrators can
 construct is shown in Table~\ref{table:spectrum}. The columns are the different
@@
 they are safely saved in the object store. If a failure occurs during ``global
-persist" or if we inject a new policy that chanages a subtree from ``local
+persist" or if we inject a new policy that changes a subtree from ``local
 persist" to ``global persist", CudeleFS no guarantee until the mechanisms are

### 2017-04-01 paper/sections/5_implementation.tex
Fixes typos and minor grammar edits

@@
 
-We use a ``programmable storage" approach~\cite{sevilla:eurosys17} to design
+We use a programmable storage approach~\cite{sevilla:eurosys17} to design
 CudeleFS; namely, we try to leverage components inside Ceph to inherit the

### 2017-04-01 paper/build.sh

@@ -9,7 +9,7 @@ docker run \
   --name latex \
   --entrypoint=/bin/bash \
   -v `pwd`/:/mnt \
-  ivotron/texlive_v2:latest -c \
+  ivotron/texlive:latest -c \
     "cd /mnt ; \
      pdflatex -synctex=1 -interaction=nonstopmode -shell-escape paper && \
      bibtex paper && \

### 2017-04-01 paper/sections/5_implementation.tex
Uses times as main font (#16)

--- 2017-04-01 paper/sections/5_implementation.tex
@@
 save their updates to disk or to the object store.
+
+\begin{listing}
+\begin{minted}[frame=single,
+               framesep=3mm,
+               xleftmargin=21pt,fontsize=\footnotesize,
+               tabsize=4]{js}
+{     
+    "allocated_inodes": "100000"
+    "interfere_policy": "block"
+    "consistency": "create"
+    "durability": "local_persist"
+}
+\end{minted}
+\caption{Implementing DeltaFS with CudeleFS.}
+\label{src:deltafs}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[frame=single,
+               framesep=3mm,
+               xleftmargin=21pt,fontsize=\footnotesize,
+               tabsize=4]{js}
+{     
+    "allocated_inodes": "100000"
+    "interfere_policy": "block"
+    "consistency": "create+volatile_apply"
+    "durability": "local_persist"
+}
+\end{minted}
+\caption{Implementing BatchFS with CudeleFS.}
+\label{src:batchfs}
+\end{listing}
 

### 2017-04-01 paper/build.sh

@@ -4,8 +4,7 @@ rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.sync
 set -ex
 
 # Build the paper
-docker rm -f latex || true
-docker run \
+docker run --rm \
   --name latex \
   --entrypoint=/bin/bash \
   -v `pwd`/:/mnt \
@@ -27,4 +26,4 @@ set +e
 rm -r _minted-paper paper.aux paper.bbl paper.blg paper.log paper.out paper.synctex.gz build.log >> /dev/null 2>&1
 
 echo "SUCCESS"
-exit 0
+

### 2017-04-01 paper/sections/3_posix_overheads.tex

@@
 
+\begin{figure}[tb]
+\centering
+\includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
+\caption{Create-heavy workloads (untar) incurr the highest disk, network, and
+CPU utilization because of the consistency and durability demands of
+CephFS.}\label{fig:overhead-creates}
+\end{figure}
+
 In this section, we quantify the costs of strong consistency and global
@@
 \label{sec:durability}
+
+% what is durability
+While durability is not specified by POSIX, users expect that files they create
+or modify survive failures.  We define three types of durability: global,
+local, and none.  Global
+durability means that the client or server can fail at any time and metadata
+will not be lost. Local durability means that metadata can be lost if the
+client or server stays down after a failure. None means that metadata is volatile
+and that the system provides no guarantees when clients or servers fail. 
+None is different than local durability because regardless of the type failure,
+metadata will be lost when components die in a None configuration.\\
+
+% - sequential IO, trim redundant operations
+\noindent\textbf{CephFS Design}: a journal of metadata updates that streams into
+the resilient object store. Similar to LFS~\cite{rosenblum:acm1992-LFS} and
+WAFL~\cite{hitz:wtec1994-WAFL} the metadata journal is designed to grow to large which
+ensures (1) sequential writes into the object store and (2) the ability for daemons to
+trim redundant or irrelevent journal entries.  The journal is striped over
+objects where multiple journal updates can reside on the same object. There are
+two tunables for controlling the journal: the segment size and the number of
+parallel segments that can be written in parallel. Unless the journal saturates 
+memory or CPU resources, larger values for these tunables results in better
+performance.
+
+% purpose of the journal
+As shown in Figure~\ref{fig:journal}, in addition to the metadata journal,
+CephFS also represents metadata in RADOS as a metadata store, where directories
+and their file inodes are stored as objects.  The metadata server applies the
+updates in the journal to the metadata store when the journal reaches a certain
+size. The metadata store is optimized for recovery ({\it i.e.} reading) while
+the metadata journal is write-optimized.
 

### 2017-04-02 experiments/fuse/scale-clients-isolated/run.sh
microbenchmarks: add isolation experiments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,17 +1,13 @@
 #!/bin/bash
 
 set -ex
+NFILES=98000
 
 # setup the docker command
 RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root"
 ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
 CEPH_ANSIBLE="$RUN -v `pwd`/site/roles/ceph-ansible:/root $ANSIBLE"
 SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
-
-#for nclients in 2 3 4 5 6 7 8; do
-#  site="nojournal-cache"
-#  for run in `seq 0 2`; do
-#    for job in "creates" "creates-touchstream" "creates-stat" "creates-touch"; do
 
 for nclients in 2 3 4 5 6 7; do
   site="nojournal-cache"
@@ -35,17 +31,12 @@ for nclients in 2 3 4 5 6 7; do
         $CEPH_ANSIBLE ceph.yml #cephfs.yml
         $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml
   
-        ## warmup and get baseline
-        #for i in `seq 0 2`; do
-        #  ./ansible-playbook.sh -e site=$results -e nfiles=100000 ../workloads/warmup.yml
-        #done
-  
         # run job
-        ./ansible-playbook.sh -e site=$results -e nfiles=10 ../workloads/setup_mdtest.yml
+        ./ansible-playbook.sh -e site=$results -e nfiles=$NFILES ../workloads/setup_mdtest.yml
         if [ $decoupled -eq 1 ]; then
-          ./ansible-playbook.sh -e site=$results -e nfiles=10 ../workloads/decouple.yml
+          ./ansible-playbook.sh -e site=$results -e nfiles=$NFILES ../workloads/decouple.yml
         fi
-        ./ansible-playbook.sh -e site=$results -e nfiles=10 -e drop_delay=30 ../workloads/${job}.yml
+        ./ansible-playbook.sh -e site=$results -e nfiles=$NFILES -e drop_delay=30 ../workloads/${job}.yml
         ./ansible-playbook.sh -e site=$results collect.yml
       done
     done

### 2017-04-02 experiments/run.sh

@@ -1,7 +1,8 @@
 #!/bin/bash
 set -x
 #for w in "droplease-cache" "droplease-nocache"; do
-for w in "droplease-cache-creates" "droplease-nocache-creates"; do
+#for w in "droplease-cache-creates" "droplease-nocache-creates"; do
+for w in "fuse/scale-clients-isolated" "fuse/scale-merge"; do
   cd $w
   ./run.sh
   cd -

### 2017-04-02 paper/sections/4_methodology.tex
paper: add to evaluation and reorg sections 3/4 - isolated namespace experiment with allow and block - talk about code in implementation and programmability - shift section 3 to focus on what the mechanisms do
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
-description of what each mechanism does.  Of the 6 mechanisms in
-Figure~\ref{fig:decouple} only 4 had to implemented and just 1 required changes to
-the underlying storage system itself.
+description of what each mechanism does.  Decoupled clients use the ``Create"
+mechanism to append metadata updates to a local, in-memory journal. Once the
+job is complete, the system calls CudeleFS mechanisms to achieve the desired
+consistency/durability semantics.  CudeleFS provides a library for clients to
+link into and all operations are performed by the client.
 
-\subsubsection{No Changes} ``RPCs" and ``Stream" are part of the default CephFS
-implementation; they are used to get strong consistency and global durability,
-respectively.  Using existing configuration settings in Ceph we can turn
-``Stream" on and off.  If it is off, then the metadata servers will not save
-journals in the object store and the daemons that apply the journal to the
-metadata store will never run.
-
-\subsubsection{Library} for ``Create", ``Nonvolatile Apply", ``Local/Global
-Persist", CudeleFS provides a library for clients to link into and all
-operations are performed by the client.  Decoupled clients use the ``Create"
-mechanism to append metadata updates to a local, in-memory journal.  For
-``Local Persist", clients write serialized log events to a file on local disk
-and for ``Global Persist", clients push the journal into the objects store. The
-overheads for both ``Local Persist" and ``Global Persist" is the write
-bandwidth of the local disk and object store, respectively.  ``Nonvolatile
+\subsubsection{Mechanisms Used for Consistency} ``RPCs" send remote procedure
+calls for every metadata operation from the client to the metadata server,
+assuming the request cannot be satisified by the inode cache. This mechanism is
+part of the default CephFS implementation and is the strongest form of
+consistency because clients see metadata updates right away.  ``Nonvolatile
 Apply" replays the client's journal onto the metadata cluster's metadata store.
@@
 notice new journal updates in the object store and replay the events onto their
-in-memory metadata stores.  These implementations required no changes to CephFS
-because the metadata servers know how to read the events the library is writing
-into the object store.  By re-using the journal subsystem to implement the
-namespace decoupling, Cudele leverages the write/read optimized data
-structures, the formats for persisting events (similar to TableFS's
-SSTables~\cite{ren:atc2013-tablefs}), and the functions for replaying events
-onto the internal namespace data structures.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=90mm]{figures/fig-decouple.png}
-\caption{Applications decouple the namespace, write updates to a local journal,
-and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
-\end{figure}
-
-% - how it makes no gaurantees
-\subsubsection{Changes to Metadata Server} the ``Volatile Apply" mechanism
-takes an in-memory journal on the client and applies the updates directly to
-the in-memory namespace maintained by the metadata servers. We say volatile
-because -- in exchange for peak performance -- Cudele makes no consistency or
-durability guarantees while ``Volatile Apply" is executing.  If a concurrent
-update from a client occurs there is no rule for resolving conflicts and if the
-client or metadata server crashes there may be no way to recover.
+in-memory metadata stores.  ``Volatile Apply" takes an in-memory journal on the
+client and applies the updates directly to the in-memory namespace maintained
+by the metadata servers. We say volatile because -- in exchange for peak
+performance -- CudeleFS makes no consistency or durability guarantees while
+``Volatile Apply" is executing.  If a concurrent update from a client occurs
+there is no rule for resolving conflicts and if the client or metadata server
+crashes there may be no way to recover.
 
 % difference between apply and volatile apply
-In contrast, ``Nonvolatile Apply" uses the the object store to apply the
-journal of updates from the client onto the metadata servers' metadata store.
-``Nonvolatile Apply" is safer but has a performance overhead because objects in
-the metadata store need to be read from and written back to the object store.
+The biggest difference between ``Volatile Apply" and ``Nonvolatile Apply" is
+the medium they use to communciate. ``Volatile Apply" applies updates directly
+to the metadata servers' metadata store while ``Nonvolatile Apply" uses the
+object store to communicate the journal of updates from the client to the
+metadata servers.  ``Nonvolatile Apply" is safer but has a large performance
+overhead because objects in the metadata store need to be read from and written
+back to the object store.
 
@@
 
+\subsubsection{Mechanisms Used for Durability} ``Stream" is one of the
+mechanisms used by default in CephFS.  Using existing configuration settings in
+Ceph we can turn ``Stream" on and off.  If it is off, then the metadata servers
+will not save journals in the object store and the daemons that apply the
+journal to the metadata store will never run.  For ``Local Persist", clients
+write serialized log events to a file on local disk and for ``Global Persist",
+clients push the journal into the objects store. The overheads for both ``Local
+Persist" and ``Global Persist" is the write bandwidth of the local disk and
+object store, respectively.  These persist mechanisms are part a library that
+links into the client.
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=90mm]{figures/fig-decouple.png}
+\caption{Applications decouple the namespace, write updates to a local journal,
+and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
+\end{figure}
+
 \subsection{Defining Policies in Cudele}
 \label{sec:setting-policies-with-cudele}
+
+\begin{listing}
+\begin{minted}[frame=single,
+               framesep=3mm,
+               xleftmargin=21pt,fontsize=\footnotesize,
+               tabsize=4]{js}
+{     
+    "allocated_inodes": "100000"
+    "interfere_policy": "block"
+    "consistency": "create"
+    "durability": "local_persist"
+}
+\end{minted}
+\caption{Implementing DeltaFS with CudeleFS.}
+\label{src:deltafs}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[frame=single,
+               framesep=3mm,
+               xleftmargin=21pt,fontsize=\footnotesize,
+               tabsize=4]{js}
+{     
+    "allocated_inodes": "100000"
+    "interfere_policy": "block"
+    "consistency": "create+volatile_apply"
+    "durability": "local_persist"
+}
+\end{minted}
+\caption{Implementing BatchFS with CudeleFS.}
+\label{src:batchfs}
+\end{listing}
 
@@
 \begin{center}
-\caption{Future programmers can explore the consistency (C) and
-durability (D) spectrums by composing Cudele mechanisms. 
-\label{table:spectrum}}
 \begin{tabular}{ l | l | l | l }
@@
 \end{tabular}
+
+\caption{Future programmers can explore the consistency (C) and
+durability (D) spectrums by composing Cudele mechanisms. 
+\label{table:spectrum}}
 \end{center}
@@
 %\label{src:batchfs}
+
 %\end{listing}

### 2017-04-02 paper/sections/5_implementation.tex
paper: add to evaluation and reorg sections 3/4 - isolated namespace experiment with allow and block - talk about code in implementation and programmability - shift section 3 to focus on what the mechanisms do
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

--- 2017-04-01 paper/sections/5_implementation.tex
@@
 CudeleFS; namely, we try to leverage components inside Ceph to inherit the
-robustness and correctness of the internal subsystems. In this section, we
-first describe a Ceph internal subsystem or component and then we show how we
-use it in CudeleFS.
+robustness and correctness of the internal subsystems. Using this `dirty-slate'
+approach, we only had to implement 4 of the 6 mechanisms from
+Figure~\ref{fig:decouple} and just 1 required changes to the underlying storage
+system itself.  In this section, we first describe a Ceph internal subsystem or
+component and then we show how we use it in CudeleFS.
+
+%- no change (rpcs, stream)s, library(create, nonvolatile apply, local/global
+%persist),  change to metadata server (nonvolatile apply)
 
@@
 % how i used it
-\noindent\textbf{Cudele}: uses the metadata store format to write objects
+\noindent\textbf{CudeleFS}: uses the metadata store format to write objects
 formatted like the journal into the object store. By writing with the same
 format, the metadata servers can read and use the recovery code to materialize
-the updates from a client's decoupled namespace ({\it i.e.} merge).
+the updates from a client's decoupled namespace ({\it i.e.} merge).  These
+implementations required no changes to CephFS because the metadata servers know
+how to read the events the library is writing into the object store.  By
+re-using the journal subsystem to implement the namespace decoupling, Cudele
+leverages the write/read optimized data structures, the formats for persisting
+events (similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the
+functions for replaying events onto the internal namespace data structures.\\
+
+\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply
 
@@
 
-\noindent\textbf{Cudele}: uses the journal subsystems within the metadata
+\noindent\textbf{CudeleFS}: uses the journal subsystems within the metadata
 server to replay updates onto the in-memory metadata store.  When the clients
@@
 binary file of  metadata updates to the metadata server. The metadata server
-uses its recovery code to ``replay" the updates.
+uses its recovery code to ``replay" the updates.\\
+
+\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply, Local/Persist
 
@@
 
-\noindent\textbf{Cudele}: using the journal tool, clients can materialize the
+\noindent\textbf{CudeleFS}: using the journal tool, clients can materialize the
 journal in memory and append events to their own copy. When they are done they
 export their journal to a binary file. Clients also use the journal tool to
-save their updates to disk or to the object store.
+save their updates to disk or to the object store.\\
+
+\noindent\textbf{Used to Implement}: Create, Volatile/Nonvolatile Apply
 
@@
 to clients that decouple parts of the namespace and to skip inodes used by the
-client at merge time.
+client at merge time.\\
+
+\noindent\textbf{Used to Implement}: Nonvolatile Apply, Create
 

### 2017-04-02 paper/sections/6_results.tex
Fixes typos and minor grammar edits (#15)
partial
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: add to evaluation and reorg sections 3/4 - isolated namespace experiment with allow and block - talk about code in implementation and programmability - shift section 3 to focus on what the mechanisms do
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \subsection{Benefits of Isolated Subtree Policies}
+
+\begin{listing}[tb]
+\begin{minted}[frame=single,
+               framesep=3mm,
+               xleftmargin=21pt,
+               tabsize=4]{js}
+{     
+    "allocated_inodes": "100000"
+    "interfere_policy": "block"
+    "consistency": "RPCs"
+    "durability": "stream"
+}
+\end{minted}
+\caption{Existing CephFS on Cudele with blocking.}
+\label{src:block}
+\end{listing}
+
+\begin{figure}[tb]
+\centering
+\includegraphics[width=1.0\linewidth]{graphs/slowdown-allow-block.png}
+\caption{
+[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-blockallow.ipynb}{source}]
+Using the ``allow" and ``block" API, users can isolate directories from
+interferring clients. Performance with blocking turned is the same as
+``isolated" from Figure~\ref{fig:interfere-b}.
+\label{fig:slowdown-allow-block}}
+\end{figure}
+
 % setup: 2cs sep dirs, 2cs DN, 1c malicious write
-In this section, we show the isolation benefits of subtree policies.  We repeat
-the experiment from Figure~\ref{fig:interfere-b}, where clients write to their
-own private directories and another client interferes at 30 seconds.  We also
-have 2 clients write to decoupled namespaces and merge their updates 90
-seconds. The
+Next we show how CudeleFS can be programmed to block interferring clients. We
+use the same problematic workload from Figure~\ref{fig:interfere-b}, where
+clients write to their own private directories and another client interferes at
+30 seconds.  We also have another client write to a decoupled namespace and
+merge its updates at 90 seconds.  We equip each client directory with the
+configuration in Listing~\ref{src:block}.  Each directory functions with RPCs
+and stream enabled -- which is the default implementation of CephFS. The only
+difference is that we enable blocking on each subtree so while the tree behaves
+like CephFS, all interferring operations will be returned with \texttt{-EBUSY}.
+Note that IndexFS does a similar operation with leases except clients block.
+
+% results
+The benefits of this isolations are shown next to the bars from
+Figure~\ref{fig:interfere-b} in Figure~\ref{fig:slowdown-allow-block}. Because
+\{``allow", ``block"\} has the same performance as \{``interfere",
+``isolated"\}, we draw the following three conclusions: (1) clients that use the
+API to block interferring clients  get the same performance as isolated
+clients, (2) there is a neglible effect on performance for the extra work the
+metadata server does to return \texttt{-EBUSY}, and (3) merging updates from
+the decoupled client has a neglible effect on performance.\\
+
+\noindent\textbf{Takeaway}: the API is sufficient for isolating performance
+when applications require it. This is a way of controlling consistency.
 

### 2017-04-02 paper/sections/1_introduction.tex
paper: fix implementation, address #13 and #18
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 \includegraphics[width=0.5\textwidth]{figures/subtree-policies.png}
-\caption{Administrators can assign weaker consistency and durability policies
+\caption{Users can assign weaker consistency and durability policies
 to subtrees to get the same performance benefits of state-of-the-art HPC
@@
 degrees of consistency and durability by exposing mechanisms used within the
-file system as a client library.  Cudele supports 3 forms of consistency
+file system as a client library.  CudeleFS supports 3 forms of consistency
 (invisible, weak, and strong) and 3 degrees of durability (none, local, and
-global) giving the administrator a wide range of policies and optimizations
+global) giving the user a wide range of policies and optimizations
 that can be custom fit to an application. Our contributions:
@@
 
-  \item a prototype that lets administrators program a range of
+  \item a prototype that lets users program a range of
   consistency and durability semantics (9 permutations), allowing them to custom
@@
 explored and previously unexplored metadata designs.
-Section~\ref{sec:related-work} places Cudele in the context of other related
+Section~\ref{sec:related-work} places CudeleFS in the context of other related
 work.

### 2017-04-02 paper/sections/4_methodology.tex
paper: fix implementation, address #13 and #18
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-In this section we describe CudeleFS, our file system that lets administrators
+\begin{figure}[tb]
+\centering
+\includegraphics[width=90mm]{figures/fig-decouple.png}
+\caption{Applications decouple the namespace, write updates to a local journal,
+and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
+\end{figure}
+
+\begin{table}
+\begin{tabular}{ r | l }
+  Mechanism         & Description \\\hline
+  RPCs              & round trip remote procedure calls \\
+  Stream            & stream journal into object store \\
+  Create            & events appended to in-memory journal \\
+  Volatile Apply    & apply to metadata store in obj store \\
+  Nonvolatile Apply & apply to metadata store in memory \\
+  Local Persist     & journal saved to client's disk \\
+  Global Persist    & journal saved in object store \\
+\end{tabular}
+\caption{Cudele's mechanisms are composed together to form consistency and
+durability semantics.\label{table:mechanisms}} 
+\end{table}
+
+In this section we describe CudeleFS, our file system that lets users
 assign consistency and durability semantics to subtrees in the global
@@
 constructing consistency and durability guarantees. CudeleFS exposes these
-mechanisms and the administrator composes them together to construct
+mechanisms and the user composes them together to construct
 \textbf{policies}. These policies are assigned to subtrees and they dictate how
@@
 % describe the figure
-Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in Cudele and
+Figure~\ref{fig:decouple} shows the mechanisms (labeled arrows) in CudeleFS and
 which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
@@
 notice new journal updates in the object store and replay the events onto their
-in-memory metadata stores.  ``Volatile Apply" takes an in-memory journal on the
-client and applies the updates directly to the in-memory namespace maintained
+in-memory metadata stores.  ``Volatile Apply" takes the client's in-memory journal on the
+client and applies the updates directly to the in-memory metadata store maintained
 by the metadata servers. We say volatile because -- in exchange for peak
@@
 %The metadata store and journal
-%are different ways of representing the namespace.  Cudele presents 6
+%are different ways of representing the namespace.  CudeleFS presents 6
 %mechanisms: RPCs, Stream, Create, Volatile Apply, Local Persist, and Global
@@
 Ceph we can turn ``Stream" on and off.  If it is off, then the metadata servers
-will not save journals in the object store and the daemons that apply the
-journal to the metadata store will never run.  For ``Local Persist", clients
-write serialized log events to a file on local disk and for ``Global Persist",
+will not save journals in the object store. For ``Local Persist", clients write
+serialized log events to a file on local disk and for ``Global Persist",
 clients push the journal into the objects store. The overheads for both ``Local
 Persist" and ``Global Persist" is the write bandwidth of the local disk and
-object store, respectively.  These persist mechanisms are part a library that
+object store, respectively.  These persist mechanisms are part of the library that
 links into the client.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=90mm]{figures/fig-decouple.png}
-\caption{Applications decouple the namespace, write updates to a local journal,
-and delay metadata updates using the CudeleFS Mechanisms }\label{fig:decouple}
-\end{figure}
 
@@
 \label{sec:setting-policies-with-cudele}
-
-\begin{listing}
-\begin{minted}[frame=single,
-               framesep=3mm,
-               xleftmargin=21pt,fontsize=\footnotesize,
-               tabsize=4]{js}
-{     
-    "allocated_inodes": "100000"
-    "interfere_policy": "block"
-    "consistency": "create"
-    "durability": "local_persist"
-}
-\end{minted}
-\caption{Implementing DeltaFS with CudeleFS.}
-\label{src:deltafs}
-\end{listing}
-
-\begin{listing}
-\begin{minted}[frame=single,
-               framesep=3mm,
-               xleftmargin=21pt,fontsize=\footnotesize,
-               tabsize=4]{js}
-{     
-    "allocated_inodes": "100000"
-    "interfere_policy": "block"
-    "consistency": "create+volatile_apply"
-    "durability": "local_persist"
-}
-\end{minted}
-\caption{Implementing BatchFS with CudeleFS.}
-\label{src:batchfs}
-\end{listing}
-
-% describe table
-The spectrum of consistency and durability guarantees that administrators can
-construct is shown in Table~\ref{table:spectrum}. The columns are the different
-consistency semantics and the rows cover the spectrum of durability guarantees.
-For consistency: ``invisible" means the system does not handle merging updates
-into a global namespace and it is assumed that middleware or the application
-manages consistency lazily; ``weak" merges updates at some time in the
-future ({\it e.g.}, when the system has time, when number of updates reaches a
-certain threshold, when the client is done writing, etc.); and updates in
-``strong" consistency are seen immediately by all clients. For durability,
-``none" means that updates are volatile and will be lost on a failure. Stronger
-guarantees are made with ``local", which means updates will be retained if the
-client node recovers, and ``global", where all updates are always recoverable.
-
-% which system they represent and which are impossible
-Existing, state-of-the-art systems in HPC can be represented by the cells in
-Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
-have global consistency and durability; DeltaFS uses ``invisible" consistency
-and ``local" durability and BatchFS uses ``weak" consistency and ``local"
-durability.  To compose the mechanisms administrators inject which mechanisms
-(described in Section~\S\ref{sec:cudeles-mechanisms}) to run and which to use
-in parallel using a domain specific language.
-
-% what is impossible
-Although we can achieve all permutations of the different guarantees in
-Table~\ref{table:spectrum}, not all of them make sense. For example, it
-makes little sense to do \texttt{creates+RPCs} since both mechanisms do the same
-thing or \texttt{stream+save} since ``global" durability is stronger and has
-more overhead than ``local" durability.
-
-% talk of eventual consistency
-The consistency and durability properties in Table~\ref{table:spectrum} are not
-guaranteed until all mechanisms in the cell are complete. The compositions
-should be considered atomic and there are no guarantees while transitioning
-between policies. For example, updates are not deemed globally consistent until
-they are safely saved in the object store. If a failure occurs during ``global
-persist" or if we inject a new policy that changes a subtree from ``local
-persist" to ``global persist", CudeleFS no guarantee until the mechanisms are
-complete.
-
-\begin{table}
-\begin{tabular}{ r | l }
-  Mechanism         & Description \\\hline
-  RPCs              & round trip remote procedure calls \\
-  Stream            & stream journal into object store \\
-  Create            & events appended to in-memory journal \\
-  Volatile Apply    & apply to metadata store in obj store \\
-  Nonvolatile Apply & apply to metadata store in memory \\
-  Local Persist     & journal saved to client's disk \\
-  Global Persist    & journal saved in object store \\
-\end{tabular}
-\caption{Cudele's mechanisms are composed together to form consistency and
-durability semantics.\label{table:mechanisms}} 
-\end{table}
 
@@
 
-\caption{Future programmers can explore the consistency (C) and
-durability (D) spectrums by composing Cudele mechanisms. 
+\caption{Users can explore the consistency (C) and
+durability (D) spectrums by composing CudeleFS mechanisms. 
 \label{table:spectrum}}
@@
 
-\subsection{Cudele Namespace API}
-\label{sec:cudele-namespace-api}
+% describe table
+The spectrum of consistency and durability guarantees that users can
+construct is shown in Table~\ref{table:spectrum}. The columns are the different
+consistency semantics and the rows cover the spectrum of durability guarantees.
+For consistency: ``invisible" means the system does not handle merging updates
+into a global namespace and it is assumed that middleware or the application
+manages consistency lazily; ``weak" merges updates at some time in the
+future ({\it e.g.}, when the system has time, when the number of updates reaches a
+certain threshold, when the client is done writing, etc.); and updates in
+``strong" consistency are seen immediately by all clients. For durability,
+``none" means that updates are volatile and will be lost on a failure. Stronger
+guarantees are made with ``local", which means updates will be retained if the
+client node recovers, and ``global", where all updates are always recoverable.
+
+% which system they represent and which are impossible
+Existing, state-of-the-art systems in HPC can be represented by the cells in
+Table~\ref{table:spectrum}.  POSIX-compliant systems like CephFS and IndexFS
+have global consistency and durability ~\footnote{This is the normal case.
+IndexFS also has bulk merge which would transition the system into ``weak
+consistency"}; DeltaFS uses ``invisible" consistency and ``local" durability
+and BatchFS uses ``weak" consistency and ``local" durability. These systems
+have other features that could push them into different semantics but we assign
+labels here based on the points emphasized in the papers.  To compose the
+mechanisms users inject which mechanisms to run and which to use in parallel
+using a domain specific language.  Although we can achieve all permutations of
+the different guarantees in Table~\ref{table:spectrum}, not all of them make
+sense. For example, it makes little sense to do \texttt{creates+RPCs} since
+both mechanisms do the same thing or \texttt{stream+save} since ``global"
+durability is stronger and has more overhead than ``local" durability.
+
+% talk of eventual consistency
+The consistency and durability properties in Table~\ref{table:spectrum} are not
+guaranteed until all mechanisms in the cell are complete. The compositions
+should be considered atomic and there are no guarantees while transitioning
+between policies. For example, updates are not deemed to have ``global" durability until
+they are safely saved in the object store. If a failure occurs during ``global
+persist" or if we inject a new policy that changes a subtree from ``local
+persist" to ``global persist", CudeleFS makes no guarantee until the mechanisms are
+complete.
+
+\subsection{CudeleFS Namespace API}
+\label{sec:cudelefs-namespace-api}
 
 % the interface
-The interface for setting the subtree policies is with \{path,
-\(<\)block\(|\)overwrite\(>\), pre-allocated inodes\} tuples. For example:
-
-\texttt{(msevilla/mydir, policies.yml)}
-
-would decouple the path \texttt{msevilla/mydir} and would apply the policies in
-\texttt{policies.yml}. The policies file supports the following values:
+Users control consistency and durability for subtrees by contacting the monitor
+daemon with a directory path and presesenting a policies configuration file.
+For example, (msevilla/mydir, policies.yml) would decouple the path
+``msevilla/mydir" and would apply the policies in ``policies.yml".  The
+policies file supports the following parameters (default values are in parenthesis):
 
@@
 
-  \item \texttt{allocated\_inodes}: the number of inodes to allocate to the
-  decoupled namespace (default 100)
+  \item \texttt{consistency}: which consistency model to use (\texttt{RPCs})
 
-  \item \texttt{interfere\_callback}: how to handle a request from another
-  client targeted at the now decoupled subtree (default \texttt{overwrite})
+  \item \texttt{durability}: which durability model to use (\texttt{stream})
 
-  \item \texttt{consistency\_callback}: which consistency model to use (default
-  \texttt{RPCs})
+  \item \texttt{allocated\_inodes}: the number of inodes to provision to the
+  decoupled namespace (100)
 
-  \item \texttt{durability\_callback}: which durability model to use (default
-  \texttt{stream})
+  \item \texttt{interfere\_policy}: how to handle a request from another
+  client targeted at this subtree (\texttt{allow}))
 
@@
 
+The ``Consistency" and ``Durability" parameters are set to the CudeleFS
+mechanisms; they can be composed (\(+\)) or run in parallel (\(|\|\)).
+``Allocated Inodes" is a way for the application to specify how many files it
+intends to create. It is a contract so that the file system can provision
+enough resources for the incumbent merge and so it can give valid inodes to
+other clients. The inodes can be used anywhere within the decoupled namespace
+({i.e.} at any depth in the subtree).
+
 % description
+``Interfere Policy" has two settings: \texttt{block} and \texttt{allow}.
 For \texttt{block}, any requests to this part of the namespace returns with
@@
 updates, for example it wants approximations for results that take too long to
-compute, it can select \texttt{overwrite}. In this case, metadata will be
+compute, it can select \texttt{allow}. In this case, metadata from the interferring client will be
 written and the computation from the decoupled namespace will take priority at
@@
 existing CephFS implementation. To implement DeltaFS on CudeleFS, the user
-would use the configuration from Listing~\ref{src:deltafs}. BatchFS's merging
-back into the global namespace can be achieved with the configuration in
-Listing~\ref{src:batchfs}.
+would use the configuration:
 
-%\begin{listing}
-%\begin{minted}[frame=single,
-%               framesep=3mm,
-%               xleftmargin=21pt,
-%               tabsize=4]{js}
-%{     
-%    "allocated_inodes": "-"
-%    "interfere_policy": "-"
-%    "consistency": "RPCs"
-%    "durability": "stream"
-%}
-%\end{minted}
-%\caption{Existing CephFS on Cudele.}
-%\label{src:batchfs}
+\begin{listing}[h]
+\begin{minted}[frame=single,
+               framesep=3mm,
+               xleftmargin=21pt,fontsize=\footnotesize,
+               tabsize=4]{js}
+{     
+    "allocated_inodes": "100000"
+    "interfere_policy": "block"
+    "consistency": "create"
+    "durability": "local_persist"
+}
+\end{minted}
+\end{listing}
 
-%\end{listing}
+BatchFS merges updates into the global namespace after the job is complete.
+This can be achieved with by composing two mechanisms:
+
+\begin{listing}[h]
+\begin{minted}[frame=single,
+               framesep=3mm,
+               xleftmargin=21pt,fontsize=\footnotesize,
+               tabsize=4]{js}
+{     
+    "allocated_inodes": "100000"
+    "interfere_policy": "block"
+    "consistency": "create+volatile_apply"
+    "durability": "local_persist"
+}
+\end{minted}
+\end{listing}

### 2017-04-02 paper/sections/5_implementation.tex
paper: fix implementation, address #13 and #18
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 % how i used it
-\noindent\textbf{CudeleFS}: uses the metadata store format to write objects
-formatted like the journal into the object store. By writing with the same
-format, the metadata servers can read and use the recovery code to materialize
-the updates from a client's decoupled namespace ({\it i.e.} merge).  These
-implementations required no changes to CephFS because the metadata servers know
-how to read the events the library is writing into the object store.  By
-re-using the journal subsystem to implement the namespace decoupling, Cudele
-leverages the write/read optimized data structures, the formats for persisting
-events (similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the
-functions for replaying events onto the internal namespace data structures.\\
+\noindent\textbf{CudeleFS}: the ``RPCs" mechanism uses the metadata store to
+service requests. Using code designed for recovery, ``Nonvolatile Apply" and
+``Volatile Apply" replay updates onto the metadata store in memory and in the
+object store, respectively.  When the clients are ready to merge their updates
+back into the global namespace, they pass a binary file of  metadata updates to
+the metadata server.
 
-\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply
+%\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply
 
@@
 
-\noindent\textbf{CudeleFS}: uses the journal subsystems within the metadata
-server to replay updates onto the in-memory metadata store.  When the clients
-are ready to merge their updates back into the global namespace, they pass a
-binary file of  metadata updates to the metadata server. The metadata server
-uses its recovery code to ``replay" the updates.\\
-
-\noindent\textbf{Used to Implement}: RPCs, Stream, Volatile Apply, Local/Persist
+\noindent\textbf{CudeleFS}: the journal format is used by ``Stream", ``Create",
+``Local Persist", and ``Global Persist".  ``Stream" is the default
+implementation for achieving global durability in CephFS but the rest of the
+mechanisms are implemeted by writing with the journal format.  By writing with
+the same format, the metadata servers can read and use the recovery code to
+materialize the updates from a client's decoupled namespace ({\it i.e.} merge).
+These implementations required no changes to CephFS because the metadata
+servers know how to read the events the library is writing.  By re-using the
+journal subsystem to implement the namespace decoupling, CudeleFS leverages the
+write/read optimized data structures, the formats for persisting events
+(similar to TableFS's SSTables~\cite{ren:atc2013-tablefs}), and the functions
+for replaying events onto the internal namespace data structures.
 
@@
 
-\noindent\textbf{CudeleFS}: using the journal tool, clients can materialize the
-journal in memory and append events to their own copy. When they are done they
-export their journal to a binary file. Clients also use the journal tool to
-save their updates to disk or to the object store.\\
-
-\noindent\textbf{Used to Implement}: Create, Volatile/Nonvolatile Apply
+\noindent\textbf{CudeleFS}: the external library the clients link into is based
+on the journal tool. It already had functions for importing, exporting, and
+modifying the updates in the journal so we re-purposed that code to implement
+``Create", ``Volatile Apply", and ``Nonvolatile Apply".
 
@@
 metadata servers. Without contention clients can resolve metadata reads locally
-thus reducing operations like \texttt{lookup()}.  For example, if a client or
+thus reducing the number of operations ({\it e.g.}, \texttt{lookup()}s).  For example, if a client or
 metadata server is not caching the directory inode, all creates within that
@@
 
-\noindent\textbf{Cudele}: uses the internal inode cache code to allocate inodes
-to clients that decouple parts of the namespace and to skip inodes used by the
-client at merge time.\\
-
-\noindent\textbf{Used to Implement}: Nonvolatile Apply, Create
+\noindent\textbf{Cudele}: ``Nonvolatile Apply" uses the internal inode cache
+code to allocate inodes to clients that decouple parts of the namespace and to
+skip inodes used by the client at merge time.\\
 
@@
 
-In CephFS, inodes store policies for policies, like how the file is striped
+In CephFS, inodes already store policies, like how the file is striped
 across the object store or for managing subtrees for load balancing. These
@@
 underlying metadata. The underlying implementation stores executable code in
-the inode that calls the different Cudele mechanisms. Of course, there are many
+the inode that calls the different CudeleFS mechanisms. Of course, there are many
 security and access control aspects of this approach but that is beyond the

### 2017-04-02 paper/sections/6_results.tex
paper: fix implementation, address #13 and #18
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-We evaluate Cudele on a 15 node cluster, partitioned into 8 object storage
+We evaluate CudeleFS on a 15 node cluster, partitioned into 8 object storage
 servers, 3 metadata servers, and 2 monitor servers. The object storage servers
@@
 
-\subsection{Cudele Mechanism Performance}
+\subsection{CudeleFS Mechanism Performance}
 
@@
 [\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
-The performance of the Cudele mechanisms normalized to the runtime of the
+The performance of the CudeleFS mechanisms normalized to the runtime of the
 create mechanism. The runtime of the create mechanism is the time it takes to
@@
 apply" is much larger than all the other mechanisms.  That mechanism was not
-implemented as part of Cudele -- it was a debugging and recovery tool packaged
+implemented as part of CudeleFS -- it was a debugging and recovery tool packaged
 with CephFS. It works by iterating over the updates in the journal and pulling
@@
 We compare these semantics because the final metadata states are equivalent.
-Cudele makes no guarantees during execution of the mechanisms or when
+CudeleFS makes no guarantees during execution of the mechanisms or when
 transitioning semantics -- the semantics are guaranteed {\it once the mechanism
@@
 \end{minted}
-\caption{Existing CephFS on Cudele with blocking.}
+\caption{Existing CephFS on CudeleFS with blocking.}
 \label{src:block}

### 2017-04-02 paper/sections/7_conclusion.tex
paper: fix implementation, address #13 and #18
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 approach limits redundant code and leverages the robustness of the underlying
-storage system. Cudele uses this approach and re-uses some of the building
+storage system. CudeleFS uses this approach and re-uses some of the building
 blocks from the Malacology programmable storage system~\cite{sevilla:eurosys17}
@@
 alienates applications that rely on strong consistency and durability. CudeleFS
-lets administrators assign consistency and durability guarantees to subtrees in
-the global namespace, giving administrators a wide range of policies
+lets users assign consistency and durability guarantees to subtrees in
+the global namespace, giving users a wide range of policies
 optimizations that can be custom fit to the application. Using CudeleFS, we

### 2017-04-03 paper/sections/1_introduction.tex
paper: update graphs and scan evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 applications designed for different, state-of-the-art HPC architectures.  We
-present Cudele, a prototype programmable file system that supports different
+present CudeleFS, a prototype programmable file system that supports different
 degrees of consistency and durability by exposing mechanisms used within the
@@
   \item an apples-to-apples comparison of the strategies used in recently proposed research systems against
-  previously unexplored metadata designs, all implemented using Cudele.
+  previously unexplored metadata designs, all implemented using CudeleFS.
 
@@
 consistency and system-defined durability and
-Section~\ref{sec:methodology-decoupled-namespaces} presents the Cudele
-prototype and API. Section~\ref{sec:implementation} describes Cudele's
+Section~\ref{sec:methodology-decoupled-namespaces} presents the CudeleFS
+prototype and API. Section~\ref{sec:implementation} describes the CudeleFS
 mechanisms and shows how re-using internal subsystems results in an

### 2017-04-03 paper/sections/3_posix_overheads.tex
partial
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: update graphs and scan evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 help for create-heavy workloads.
-
-\begin{figure}[tb]
-\centering
-\includegraphics[width=1\linewidth]{./graphs/overhead-creates.png}
-\caption{Create-heavy workloads (untar) incurr the highest disk, network, and
-CPU utilization because of the consistency and durability demands of
-CephFS.}\label{fig:overhead-creates}
-\end{figure}
 
@@
   \end{subfigure}
-  \caption{The overhead of the file system metadata journal. The segment size
+  \caption{The overhead of global durability in CephFS. The segment size
   is the threshold that the metadata server starts
@@
 \end{figure*}
-
-% what is durability
-While durability is not specified by POSIX, users expect that files they create
-or modify survive failures.  We define three types of durability: global,
-local, and none.  Global
-durability means that the client or server can fail at any time and metadata
-will not be lost. Local durability means that metadata can be lost if the
-client or server stays down after a failure. None means that metadata is volatile
-and that the system provides no guarantees when clients or servers fail. 
-None is different than local durability because regardless of the type failure,
-metadata will be lost when components die in a None configuration.\\
-
-% - sequential IO, trim redundant operations
-\noindent\textbf{CephFS Design}: a journal of metadata updates that streams into
-the resilient object store, similar to LFS~\cite{rosenblum:acm1992-LFS} and
-WAFL~\cite{hitz:wtec1994-WAFL}, grows to large extents which
-allows to (1) sequentially write into the object store and (2) provide the ability for daemons to
-trim redundant or irrelevant journal entries.  The journal is striped over
-objects where multiple journal updates can reside on the same object. There are
-two tunables for controlling the journal: the segment size and the number of
-parallel segments that can be written in parallel. Unless the journal saturates 
-memory or CPU resources, larger values for these tunables results in better
-performance.
-
-% purpose of the journal
-As shown in Figure~\ref{fig:journal}, in addition to the metadata journal,
-CephFS also stores metadata in RADOS, where directories
-and their file inodes are stored as objects.  The metadata server applies the
-updates in the journal to this metadata store when the journal reaches a certain
-size. The metadata store is optimized for recovery ({\it i.e.} reading) while
-the metadata journal is write-optimized.
 
@@
   \end{subfigure}
-  \caption{When a client create stream is ``isolated" lookups resolve
-  locally; but when a second client ``interferes" by creating in the same
-  directory, the directory inode capability is revoked forcing all clients to
-  centralize lookups at the metadata server.
+  \caption{The overhead of strong consistency in CephFS. When a second client
+  ``interferes", capabilities are revoked and metadata servers do more work.
   \label{fig:interfere}}

### 2017-04-03 paper/sections/5_implementation.tex
paper: update graphs and scan evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\noindent\textbf{Cudele}: ``Nonvolatile Apply" uses the internal inode cache
+\noindent\textbf{CudeleFS}: ``Nonvolatile Apply" uses the internal inode cache
 code to allocate inodes to clients that decouple parts of the namespace and to
@@
 
-\noindent\textbf{Cudele}: uses the large inodes to store consistency and
+\noindent\textbf{CudeleFS}: uses the large inodes to store consistency and
 durability policies in the directory inode. This approach uses the File Type

### 2017-04-03 paper/sections/6_results.tex
paper: update graphs and scan evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 experiments run in Docker containers; this makes it easier to tear down and
-re-initialize ({\it e.g.}, dropping the kernel cache) for the cluster between
+re-initialize the cluster ({\it e.g.}, dropping the kernel cache) between
 experiments.
@@
 
-Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the Cudele
+Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the CudeleFS
 mechanisms, normalized to the time it takes to write 100K file create updates
-to the client's in-memory journal ({\it i.e.} the create mechanism). Bars above
-\(10^0\) are slower than the create mechanism and bars below are faster.
-``Stream" is an approximation of the overhead and is calculated by subtracting
-the runtime of the job with the journal turned off from the runtime with the
-journal turned on. All the slowdowns and speedups reported are for the 100K
-file create job, the largest workoad we tested.
+to the client's in-memory journal ({\it i.e.} the ``Create" mechanism). Bars
+above \(10^0\) are slower than the ``Create" mechanism and bars below are
+faster.  ``Stream" is an approximation of the overhead and is calculated by
+subtracting the runtime of the job with the journal turned off from the runtime
+with the journal turned on. All the slowdowns and speedups reported are for the
+100K file create job, the largest workoad we tested\footnote{We stopped testing
+at 100K because this is the maximum recommended size of a directory in CephFS;
+preliminary experiemnts with larger directory sizes show memory problems.}.
 
 % RPCs vs. apply: calls to metadata server vs. RADOS
-\subsubsection{Overhead of RPCs} ``RPCs" is \(66\times\) slower than ``volatile
-apply" because sending individual metadata updates over the network is costly.
-While ``RPCs" sends a request for every file create, ``nonvolatile apply"
+\subsubsection{Overhead of RPCs} ``RPCs" is \(66\times\) slower than ``Volatile
+Apply" because sending individual metadata updates over the network is costly.
+While ``RPCs" sends a request for every file create, ``Nonvolatile Apply"
 writes all the updates to the in-memory journal and applies them to the
@@
 decoupled namespace directly to the metadata server is faster, communicating
-through the object store (``nonvolatile apply") is \(10\times\) slower.
+through the object store (``Nonvolatile Apply") is \(10\times\) slower.
 
@@
 % v_apply vs. apply/persist: communicating through RADOS
-\subsubsection{Overhead of ``nonvolatile apply"} The cost of ``nonvolatile
+\subsubsection{Overhead of ``Nonvolatile Apply"} The cost of ``Nonvolatile
 apply" is much larger than all the other mechanisms.  That mechanism was not
@@
 i.e.} \texttt{/}).  The cost of communcating through the object store is shown
-by comparing the runtime of ``volatile apply" + ``global persist" to
-``nonvolatile apply". These two operations end up with the same final metadata
-state but using ``nonvolatile apply" is clearly inferior.
+by comparing the runtime of ``Volatile apply" + ``Global persist" to
+``Nonvolatile Apply". These two operations end up with the same final metadata
+state but using ``Nonvolatile Apply" is clearly inferior.
 
 % persist vs. save: one disk vs. many
-\subsubsection{Parallelism of the Object Store} Comparing ``local" and ``global
+\subsubsection{Parallelism of the Object Store} Comparing ``Local" and ``Global
 persist" demonstrates the bandwidth advantages of storing the journal in a
-distributed object store. As the journal size increases, the ``global persist"
+distributed object store. For 100K file creates, the ``Global Persist"
 performance is \(1.5\times\) faster because the object store is leveraging the
@@
 \(2.5\text{KB}*1\text{ million files} = 2.38\text{GB}\). Transfer times for
-files this large on an HPC network are reasonable.
+payloads of this size on an HPC network are reasonable.\\
+
+\noindent\textbf{Takeaway}: the CudeleFS mechansisms have overheads and costs
+that can differ {\it by order of magnitude}. CudeleFS gives users the ability
+to compose these mechanisms based on their application's correctness
+requirements and performance goals.
 
@@
 
-Figure~\ref{fig:slowdown-strong-weak} shows the runtimes of weak and
-strong consistency based systems implemented on Cudele, normalized to the
-runtime of the create mechanism (again, just creating files in the client's
-in-memory journal). We scaled the number of files up to 100K which is the
-maximum size of a directory by default in CephFS. We use the following
-compositions from the mechanisms in Table~\ref{table:spectrum}:
-
-\begin{itemize}
-  \item Strong Consistency
-  \begin{itemize}
-    \item[]  \texttt{RPCs} \(+\) \texttt{stream}
-  \end{itemize}
-  \item Strong Consistency, No Fault Tolerance
-  \begin{itemize}
-    \item[] \texttt{RPCs}
-  \end{itemize}
-  \item Weak Consistency
-  \begin{itemize}
-    \item[] \texttt{creates} \(+\) \texttt{local persist}
-  \end{itemize}
-  \item Weak Consistency, No Fault Tolerance
-  \begin{itemize}
-    \item[] \texttt{creates} \(+\) \texttt{local persist} \(+\) \texttt{volatile apply}
-  \end{itemize}
-\end{itemize}
+Figure~\ref{fig:slowdown-strong-weak} shows the runtimes of systems employing
+weak and strong consistency, normalized to the runtime of the ``Create"
+mechanism (again, just creating files in the client's in-memory journal).  We
+use the following compositions from the mechanisms in
+Table~\ref{table:spectrum}:  Strong Consistency = ``RPCs" + ``Stream"; Strong
+Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Creates" +
+``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Creates" +
+``Local Persist" + ``Volatile Apply".
 
@@
 \subsubsection{Durability \(<<\) Consistency} The \(1.15\times\) overhead of
-``Strong Consistency" compared to ``Strong Consistency, No Fault Tolerance" for
+Strong Consistency compared to Strong Consistency, No Fault Tolerance for
 100K files is neglible. It suggests that the overhead of consistency is much
@@
 we scale the number of files because the cost of streaming the journal into the
-object store is constant. We omit the same analysis for ``Weak Consistency"
+object store is constant. We omit the same analysis for Weak Consistency
 because the runtimes are so short that the normalized slowdowns are misleading.
@@
 was argued in previous work outlining the benefit of
-SSTables~\cite{ren:atc2013-tablefs, ren:sc2014-indexfs}.
+SSTables~\cite{ren:atc2013-tablefs, ren:sc2014-indexfs}.\\
+
+\noindent\textbf{Takeaway}: CudeleFS shows the true benefit of eventual
+consistency, where we see over a \(100\times\) slowdown for achieving strong
+consistency, in the worst case.
 
@@
 use the same problematic workload from Figure~\ref{fig:interfere-b}, where
-clients write to their own private directories and another client interferes at
+clients write to their own private directories and another client interferes with a stream of creates at
 30 seconds.  We also have another client write to a decoupled namespace and
@@
 % results
-The benefits of this isolations are shown next to the bars from
-Figure~\ref{fig:interfere-b} in Figure~\ref{fig:slowdown-allow-block}. Because
-\{``allow", ``block"\} has the same performance as \{``interfere",
-``isolated"\}, we draw the following three conclusions: (1) clients that use the
-API to block interferring clients  get the same performance as isolated
-clients, (2) there is a neglible effect on performance for the extra work the
-metadata server does to return \texttt{-EBUSY}, and (3) merging updates from
-the decoupled client has a neglible effect on performance.\\
+To show the benefits of this isolations, our results in
+Figure~\ref{fig:slowdown-allow-block} are plotted alongside the bars from
+Figure~\ref{fig:interfere-b}. Because ``allow"/``interfere and
+``block"/``isolated" have the same performance we draw the following three
+conclusions: (1) clients that use the API to block interferring clients  get
+the same performance as isolated clients, (2) there is a neglible effect on
+performance for the extra work the metadata server does to return
+\texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
+neglible effect on performance.\\
 
-\noindent\textbf{Takeaway}: the API is sufficient for isolating performance
-when applications require it. This is a way of controlling consistency.
+\noindent\textbf{Takeaway}: the API lets users isolate directories when
+applications need better and more reliable performance. This is a way of
+controlling consistency.
 
@@
 \subsection{Partial Directory Listings}
-
-\subsection{Cost of Nonvolatile Apply}
-% Why so slow?
-Creating many files in the same directory would touch the same object but the
-existing implementation results in this object being repeatedly pushed/pulled.

### 2017-04-04 paper/sections/3_posix_overheads.tex
paper: carlos evaluation comments, fix graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
+\begin{figure*}[t]
+  \centering
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
+      \caption{Runtime normalized to no journal} \label{fig:overhead-a}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/behavior-interfere.png}
+      \caption{Interference forces \texttt{lookup()}s}
+      \label{fig:overhead-b}
+  \end{subfigure}
+  \begin{subfigure}[b]{.3\linewidth}
+      \centering
+      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere.png}
+      \caption{Interference hurts variability}
+      \label{fig:overhead-c}
+  \end{subfigure}
+  \caption{The overhead of global durability and strong consistency in CephFS.
+  (a) shows the effect of different journal segment sizes, which are streamed
+  into the object store for fault tolerance. (b) and (c) show that when a second
+  client ``interferes", capabilities are revoked and metadata servers do more
+  work.  \label{fig:overhead}}
+\end{figure*}
+
 % what is durability
@@
 \includegraphics[width=1\linewidth]{./figures/journal.png} 
-\caption{CephFS uses a journal to stage updates and tracks dirty metadata in
-the collective memory of the metadata servers. Each metadata server maintains its own journal, which is
-broken up into 4MB segments. These segments are pushed into RADOS and deleted
-when that particular segment is trimmed from the end of the log. In addition to
-journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
+\caption{CephFS has two views of the file system namespace: a journal of
+metadata updates and a metadata store. For fault tolerance, they are stored in
+the object store as segments and fragments, respectively.
+\label{fig:journal}}
 \end{figure}
-\begin{figure*}[t]
-  \centering
-  \begin{subfigure}[b]{.3\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/slowdown-journal.png}
-      \caption{Journal Segment Size} \label{fig:slowdown-journal-a}
-  \end{subfigure}
-  \begin{subfigure}[b]{.3\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/behavior-journal.png}
-      \caption{180MB Segment vs. No Journal}
-      \label{fig:slowdown-journal-b}
-  \end{subfigure}
-  \begin{subfigure}[b]{.3\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/slowdown-cache.png}
-      \caption{Inode Cache}
-      \label{fig:slowdown-journal-c}
-  \end{subfigure}
-  \caption{The overhead of global durability in CephFS. The segment size
-  is the threshold that the metadata server starts
-  trimming.\label{fig:slowdown-journal}}
-\end{figure*}
 
 % Effects on performance
-Figure~\ref{fig:slowdown-journal} shows that journaling metadata updates into
-the object store has an overhead. Figure~\ref{fig:slowdown-journal-a} shows the
+Figure~\ref{fig:overhead-a} shows the
 effect of journaling of different journal segment sizes; the larger the segment
@@
 better performance is memory consumption because larger segments take up
-more space for buffering. Figure~\ref{fig:slowdown-journal-b} shows how
+more space for buffering. When journaling is on, 
 the metadata server periodically stops serving requests to flush ({\it i.e.}
-apply journal updates to) to the metadata store.  The journal overhead is
+apply journal updates) to the metadata store.  The journal overhead is
 sufficient enough to slow down metadata throughput but not so much as to
@@
 
-\begin{figure*}[t]
-  \centering
-  \begin{subfigure}[b]{.3\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/behavior-interfere.png}
-      \caption{Inteference forces \texttt{lookup()}s}
-      \label{fig:interfere-a}
-  \end{subfigure}
-  \begin{subfigure}[b]{.3\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere.png}
-      \caption{Interference hurts performance}
-      \label{fig:interfere-b}
-  \end{subfigure}
-  \begin{subfigure}[b]{.3\linewidth}
-      \centering
-      \includegraphics[width=1.0\linewidth]{graphs/slowdown-interfere-types.png}
-      \caption{Different interferring requests}
-      \label{fig:interfere-c}
-  \end{subfigure}
-  \caption{The overhead of strong consistency in CephFS. When a second client
-  ``interferes", capabilities are revoked and metadata servers do more work.
-  \label{fig:interfere}}
-\end{figure*}
-
 % inode cache - reduces RPCs (lookups for create, readdirs for stats)
@@
 The benefits of caching the directory inode when creating files is shown in
-Figure~\ref{fig:interfere-a}.  If only one client is creating files in a
+Figure~\ref{fig:overhead-b}.  If only one client is creating files in a
 directory (``isolated" curve on \(y1\) axis) then that client can lookup the
@@
 extra load but the overall performance decreases.  This degradation is shown in
-Figure~\ref{fig:interfere-b}, where we scale the number of clients and show
-increased runtime and variability. The performance is compared to a single
-isolated client and the error bar is the standard deviation of the runtime of
+Figure~\ref{fig:overhead-c}, where we scale the number of clients and show
+increased variability. The performance is compared to a single
+isolated client and the \(y\) axis the standard deviation of the runtime of
 all clients.  For the ``interfere" bars, each client creates files in private
@@
 when clients interfere. At 7 clients, the metadata server is overloaded so the
-directory caching in the ``isolated" bars has no benefit. Zooming in on runs
-with 7 clients in Figure~\ref{fig:interfere-c} we see how different
-interfering operations affect performance, again when compared to the
-performance of a single isolate client.  ``create" has little effect but
-``stat" and ``create many" cause noticeable slowdowns because of the extra work
-they impose on the metadata server.\\
+directory caching in the ``isolated" bars has no benefit.\\
 

### 2017-04-04 paper/sections/4_methodology.tex
paper: update graphs and scan evaluation
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: carlos evaluation comments, fix graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
   Stream            & stream journal into object store \\
-  Create            & events appended to in-memory journal \\
+  Append Client Journal & events appended to in-memory journal \\
   Volatile Apply    & apply to metadata store in obj store \\
@@
 \end{tabular}
-\caption{Cudele's mechanisms are composed together to form consistency and
+\caption{The CudeleFS mechanisms are composed together to form consistency and
 durability semantics.\label{table:mechanisms}} 
@@
 
-\subsection{Cudele's Mechanisms}
-\label{sec:cudeles-mechanisms}
+\subsection{The CudeleFS Mechanisms}
+\label{sec:the-cudelesfs-mechanisms}
 
@@
 which daemon(s) they are performed by.  Table~\ref{table:mechanisms} has a
-description of what each mechanism does.  Decoupled clients use the ``Create"
+description of what each mechanism does.  Decoupled clients use the ``Append Client Journal""
 mechanism to append metadata updates to a local, in-memory journal. Once the
@@
 %CephFS. ``Stream" has the metadata servers stream a journal of metadata updates
-%into the object store. ``Create" allows clients to append metadata events to an
+%into the object store. ``Append Client Journal"" allows clients to append metadata events to an
 %in-memory journal. ``Volatile apply" 
@@
 
-\subsection{Defining Policies in Cudele}
+\subsection{Defining Policies in CudeleFS}
 \label{sec:setting-policies-with-cudele}
@@
   D \(\downarrow\)  	     & invisible         & weak        & strong  \\\hline
-  none                       & create            & create          & RPCs    \\
+  none                       & append client journal            & append client journal          & RPCs    \\
                              &                   & +volatile apply &         \\\hdashline
-  local                      & create            & create          & RPCs    \\
+  local                      & append client journal            & append client journal          & RPCs    \\
                              & +local persist    & +local persist  & +local  \\
                              &                   & +volatile apply &  persist\\\hdashline
-  global                     & create            & create          & RPCs    \\
+  global                     & append client journal            & append client journal          & RPCs    \\
                              & +global persist   & +global persist & +stream \\
@@
 the different guarantees in Table~\ref{table:spectrum}, not all of them make
-sense. For example, it makes little sense to do \texttt{creates+RPCs} since
+sense. For example, it makes little sense to do \texttt{append client journal+RPCs} since
 both mechanisms do the same thing or \texttt{stream+save} since ``global"
@@
     "interfere_policy": "block"
-    "consistency": "create"
+    "consistency": "append_client_journal"
     "durability": "local_persist"
@@
     "interfere_policy": "block"
-    "consistency": "create+volatile_apply"
+    "consistency": "append_client_journal+volatile_apply"
     "durability": "local_persist"

### 2017-04-04 paper/sections/5_implementation.tex
paper: carlos evaluation comments, fix graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 
-\noindent\textbf{CudeleFS}: the journal format is used by ``Stream", ``Create",
+\noindent\textbf{CudeleFS}: the journal format is used by ``Stream", ``Append Client Journal"",
 ``Local Persist", and ``Global Persist".  ``Stream" is the default
@@
 modifying the updates in the journal so we re-purposed that code to implement
-``Create", ``Volatile Apply", and ``Nonvolatile Apply".
+``Append Client Journal"", ``Volatile Apply", and ``Nonvolatile Apply".
 

### 2017-04-04 paper/sections/6_results.tex
paper: carlos evaluation comments, fix graphs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 The performance of the CudeleFS mechanisms normalized to the runtime of the
-create mechanism. The runtime of the create mechanism is the time it takes to
-write 100 file creates to the client's in-memory journal of metadata
-updates.\label{fig:slowdown-mechanisms}}
+``Append Client Journal" mechanism (the runtime of writing \(n\) file creates
+to the client's in-memory journal).  \label{fig:slowdown-mechanisms}}
 \end{figure}
@@
 Figure~\ref{fig:slowdown-mechanisms} shows the runtime of the CudeleFS
-mechanisms, normalized to the time it takes to write 100K file create updates
-to the client's in-memory journal ({\it i.e.} the ``Create" mechanism). Bars
-above \(10^0\) are slower than the ``Create" mechanism and bars below are
-faster.  ``Stream" is an approximation of the overhead and is calculated by
-subtracting the runtime of the job with the journal turned off from the runtime
-with the journal turned on. All the slowdowns and speedups reported are for the
-100K file create job, the largest workoad we tested\footnote{We stopped testing
-at 100K because this is the maximum recommended size of a directory in CephFS;
-preliminary experiemnts with larger directory sizes show memory problems.}.
+mechanisms, normalized to the time it takes to write \(n\) file create updates
+to the client's in-memory journal ({\it i.e.} the ``Append Client Journal""
+mechanism). Bars above \(10^0\) are slower than the ``Append Client Journal""
+mechanism and bars below are faster.  ``Stream" is an approximation of the
+overhead and is calculated by subtracting the runtime of the job with the
+journal turned off from the runtime with the journal turned on.  The largest
+workoad we tested is 100K file creates in the same directory.  This is the
+maximum recommended size of a directory in CephFS; preliminary experiemnts with
+larger directory sizes show memory problems.
+
+\subsubsection{Poorly Scaling Data Structures} Despite doing the same amount of
+work, mechanisms that rely on poorly scaling data structures have large
+slowdowns for the larger number of creates. For example, ``RPCs" which relies
+on the internal CephFS directory structures, goes from a \(40\times\) slowdown
+for 10K files to a \(90\times\) slowdown for 100K files. It is a well-known
+problem that directory data structures do not scale when creating files in the
+same directory~\cite{ren:sc2014-indexfs} and any mechanism that uses these data
+structures will experience similar slowdowns. Other mechanisms that write
+events to a journal ({\it e.g.} the persists, ``Volatile Apply") experience a
+much less drastic slowdown because the journal data structure does not need to
+be scanned for every operation.
 
@@
 Figure~\ref{fig:slowdown-strong-weak} shows the runtimes of systems employing
-weak and strong consistency, normalized to the runtime of the ``Create"
+weak and strong consistency, normalized to the runtime of the ``Append Client Journal""
 mechanism (again, just creating files in the client's in-memory journal).  We
@@
 Table~\ref{table:spectrum}:  Strong Consistency = ``RPCs" + ``Stream"; Strong
-Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Creates" +
-``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Creates" +
+Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Append Client Journal"s" +
+``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Append Client Journal"s" +
 ``Local Persist" + ``Volatile Apply".

### 2017-04-04 paper/sections/3_posix_overheads.tex
paper: jeff/carlos intro comments, fix figs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 ensures (1) sequential writes into the object store and (2) the ability for daemons to
-trim redundant or irrelevent journal entries.  The journal is striped over
+trim redundant or irrelevant journal entries.  The journal is striped over
 objects where multiple journal updates can reside on the same object. There are
@@
 throughput of the ``interfere" curve because the metadata server can handle the
-extra load but the overall performance decreases.  This degradation is shown in
+extra load but performance suffers.  This degradation is shown in
 Figure~\ref{fig:overhead-c}, where we scale the number of clients and show
-increased variability. The performance is compared to a single
-isolated client and the \(y\) axis the standard deviation of the runtime of
+increased variability. The results are normalized to a single
+isolated client and the \(y\) axis is the standard deviation of the runtime of
 all clients.  For the ``interfere" bars, each client creates files in private

### 2017-04-04 paper/sections/6_results.tex
paper: jeff/carlos intro comments, fix figs
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 mechanisms, normalized to the time it takes to write \(n\) file create updates
-to the client's in-memory journal ({\it i.e.} the ``Append Client Journal""
-mechanism). Bars above \(10^0\) are slower than the ``Append Client Journal""
+to the client's in-memory journal ({\it i.e.} the ``Append Client Journal"
+mechanism). Bars above \(10^0\) are slower than the ``Append Client Journal"
 mechanism and bars below are faster.  ``Stream" is an approximation of the
@@
 journal turned off from the runtime with the journal turned on.  The largest
-workoad we tested is 100K file creates in the same directory.  This is the
-maximum recommended size of a directory in CephFS; preliminary experiemnts with
+workload we tested is 100K file creates in the same directory.  This is the
+maximum recommended size of a directory in CephFS; preliminary experiments with
 larger directory sizes show memory problems.
@@
 experiment directory and the object that contains the root directory ({\it
-i.e.} \texttt{/}).  The cost of communcating through the object store is shown
+i.e.} \texttt{/}).  The cost of communicating through the object store is shown
 by comparing the runtime of ``Volatile apply" + ``Global persist" to
@@
 
-\noindent\textbf{Takeaway}: the CudeleFS mechansisms have overheads and costs
+\noindent\textbf{Takeaway}: the CudeleFS mechanisms have overheads and costs
 that can differ {\it by order of magnitude}. CudeleFS gives users the ability
@@
 
-Figure~\ref{fig:slowdown-strong-weak} shows the runtimes of systems employing
-weak and strong consistency, normalized to the runtime of the ``Append Client Journal""
+Figure~\ref{fig:slowdown-strong-weak} shows the runtime of systems employing
+weak and strong consistency, normalized to the runtime of the ``Append Client Journal"
 mechanism (again, just creating files in the client's in-memory journal).  We
@@
 Table~\ref{table:spectrum}:  Strong Consistency = ``RPCs" + ``Stream"; Strong
-Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Append Client Journal"s" +
-``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Append Client Journal"s" +
+Consistency, No Fault Tolerance = ``RPCs"; Weak Consistency = ``Append Client Journal" +
+``Local Persist"; and Weak Consistency, No Fault Tolerance = ``Append Client Journal" +
 ``Local Persist" + ``Volatile Apply".
@@
 Strong Consistency compared to Strong Consistency, No Fault Tolerance for
-100K files is neglible. It suggests that the overhead of consistency is much
+100K files is negligible. It suggests that the overhead of consistency is much
 larger than the overhead of durability. This conclusion should be stronger as
@@
 
+\begin{figure}[tb]
+\centering
+\includegraphics[width=1.0\linewidth]{graphs/slowdown-allow-block.png}
+\caption{
+[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-blockallow.ipynb}{source}]
+Using the ``allow" and ``block" API, users can isolate directories from
+interfering clients. Performance with blocking turned is the same as
+``isolated" from Figure~\ref{fig:overhead-b}.
+\label{fig:slowdown-allow-block}}
+\end{figure}
+
+% setup: 2cs sep dirs, 2cs DN, 1c malicious write
+Next we show how CudeleFS can be programmed to block interfering clients. We
+use the same problematic workload from Figure~\ref{fig:overhead-b}, where
+clients write to their own private directories and another client interferes with a stream of creates at
+30 seconds.  We also have another client write to a decoupled namespace and
+merge its updates at 90 seconds.  We equip each client directory with the
+configuration:
+
 \begin{listing}[tb]
-\begin{minted}[frame=single,
-               framesep=3mm,
-               xleftmargin=21pt,
+\begin{minted}[xleftmargin=21pt,
                tabsize=4]{js}
@@
 \end{minted}
-\caption{Existing CephFS on CudeleFS with blocking.}
-\label{src:block}
 \end{listing}
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width=1.0\linewidth]{graphs/slowdown-allow-block.png}
-\caption{
-[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-blockallow.ipynb}{source}]
-Using the ``allow" and ``block" API, users can isolate directories from
-interferring clients. Performance with blocking turned is the same as
-``isolated" from Figure~\ref{fig:interfere-b}.
-\label{fig:slowdown-allow-block}}
-\end{figure}
-
-% setup: 2cs sep dirs, 2cs DN, 1c malicious write
-Next we show how CudeleFS can be programmed to block interferring clients. We
-use the same problematic workload from Figure~\ref{fig:interfere-b}, where
-clients write to their own private directories and another client interferes with a stream of creates at
-30 seconds.  We also have another client write to a decoupled namespace and
-merge its updates at 90 seconds.  We equip each client directory with the
-configuration in Listing~\ref{src:block}.  Each directory functions with RPCs
-and stream enabled -- which is the default implementation of CephFS. The only
+Each directory functions with ``RPCs"
+and ``Stream" enabled -- which is the default implementation of CephFS. The only
 difference is that we enable blocking on each subtree so while the tree behaves
-like CephFS, all interferring operations will be returned with \texttt{-EBUSY}.
+like CephFS, all interfering operations will be returned with \texttt{-EBUSY}.
 Note that IndexFS does a similar operation with leases except clients block.
@@
 To show the benefits of this isolations, our results in
-Figure~\ref{fig:slowdown-allow-block} are plotted alongside the bars from
+Figure~\ref{fig:slowdown-allow-block} are plotted alongside the variance bars from
 Figure~\ref{fig:interfere-b}. Because ``allow"/``interfere and
-``block"/``isolated" have the same performance we draw the following three
-conclusions: (1) clients that use the API to block interferring clients  get
-the same performance as isolated clients, (2) there is a neglible effect on
+``block"/``isolated" have the same variability we draw the following three
+conclusions: (1) clients that use the API to block interfering clients  get
+the same performance as isolated clients, (2) there is a negligible effect on
 performance for the extra work the metadata server does to return
 \texttt{-EBUSY}, and (3) merging updates from the decoupled client has a
-neglible effect on performance.\\
+negligible effect on performance.\\
 

### 2017-04-04 paper/sections/6_results.tex
paper: anonymize and add ivo feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@
 experiments presented here are available in the repository for this
-article\footnote{\url{https://github.com/michaelsevilla/cudele-popper/}}.
+article\footnote{Links have been removed for double-blind submission}.
 Experiments can be examined in more detail, or even re-run, by visiting the
@@
 \caption{
-[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
+[\href{https://...}{source}]
 The performance of the CudeleFS mechanisms normalized to the runtime of the
@@
 \caption{
-[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-decoupled.ipynb}{source}]
+[\href{https://...}{source}]
 In-memory client journal scales with the number of
@@
 mechanism). Bars above \(10^0\) are slower than the ``Append Client Journal"
-mechanism and bars below are faster.  ``Stream" is an approximation of the
+mechanism.  ``Stream" is an approximation of the
 overhead and is calculated by subtracting the runtime of the job with the
@@
 much less drastic slowdown because the journal data structure does not need to
-be scanned for every operation.
+be scanned for every operation. Events are written to the end of the journal without
+even checking the validity ({e.g.}, if the file already exists for a create),
+which is another form of relaxed consistency because the file system assumes the
+application has resolved conflicting updates in a different way.
 
@@
 \noindent\textbf{Takeaway}: the CudeleFS mechanisms have overheads and costs
-that can differ {\it by order of magnitude}. CudeleFS gives users the ability
+that can differ {\it by orders of magnitude}. CudeleFS gives users the ability
 to compose these mechanisms based on their application's correctness
@@
 \includegraphics[width=1.0\linewidth]{graphs/slowdown-strong-v-weak.png}
-\caption{The RPC per metadata update of ``Strong Consistency" has a large
+\caption{
+[\href{https://...}{source}]
+The RPC per metadata update of ``Strong Consistency" has a large
 overhead compared to the decoupled namespace strategy of ``Weak
@@
 \caption{
-[\href{https://github.com/michaelsevilla/cudele-popper/blob/master/viz-blockallow.ipynb}{source}]
+[\href{https://...}{source}]
 Using the ``allow" and ``block" API, users can isolate directories from
@@
 
+Figure~\ref{boner} scales the number of merges and shows linear performance.
+The performance is linear for two reasons: (1) the merges are serialized and
+(2) applying the updates to the in-memory metadata store does not touch data
+structures that scale poorly. CudeleFS serializes merge requests
+
 \subsection{Partial Directory Listings}

### 2017-04-04 experiments/fuse/scale-clients/run.sh
macrobenchmarks: merge experiments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -10,7 +10,7 @@ SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
 
 for nclients in 7; do
   site="nojournal-cache"
-  for run in `seq 3 5`; do
+  for run in `seq 0 6`; do
     for job in "creates" "creates-touchstream" "creates-stat" "creates-touch"; do
       # configure ceph and setup results directory
       results="$job-$nclients/run$run"
@@ -22,9 +22,9 @@ for nclients in 7; do
       cp inventory/${nclients}client site/roles/ceph-ansible/hosts
 
       # cleanup and start ceph
-      for k in `seq 0 3`; do
-        $SRL_ANSIBLE cleanup.yml || true
-      done
+      #for k in `seq 0 3`; do
+      #  $SRL_ANSIBLE cleanup.yml || true
+      #done
       $SRL_ANSIBLE cleanup.yml
       $CEPH_ANSIBLE ceph.yml cephfs.yml
       $SRL_ANSIBLE ceph_pgs.yml ceph_monitor.yml ceph_wait.yml

### 2017-04-04 experiments/fuse/scale-merge/run.sh
macrobenchmarks: merge experiments
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

@@ -1,0 +1,47 @@
+#!/bin/bash
+
+set -ex
+#NFILES=98000
+NFILES=98000
+
+# setup the docker command
+RUN="docker run -it --rm --net host -v $HOME/.ssh:/root/.ssh -w /root"
+ANSIBLE="michaelsevilla/ansible --forks 50 --skip-tags package-install,with_pkg"
+CEPH_ANSIBLE="$RUN -v `pwd`/site/roles/ceph-ansible:/root $ANSIBLE"
+SRL_ANSIBLE="$RUN -v `pwd`/site:/root $ANSIBLE"
+
+#    for job in "creates" "creates-touchstream" "creates-stat" "creates-touch"; do
+#for nclients in 2 3 4 5 6 7; do
+  #for run in 0 1 2; do
+      #for decoupled in 0 1; do
+for nclients in 7; do
+  site="journal-nocache"
+  for run in 0; do
+    for job in "scale-merge"; do
+      for decoupled in 0; do
+        # configure ceph and setup results directory
+        results="$job-$nclients-decoupled-$decoupled/run$run"
+        mkdir -p results/$results/logs || true
+        cp site/* site/roles/ceph-ansible || true
+        cp site_confs/${site}.yml site/group_vars/all
+        cp -r site/group_vars site/roles/ceph-ansible/
+        cp inventory/${nclients}client site/hosts
+        cp inventory/${nclients}client site/roles/ceph-ansible/hosts
+  
+        # cleanup and start ceph
+        $SRL_ANSIBLE cleanup.yml
+        $CEPH_ANSIBLE ceph.yml #cephfs.yml
+        $SRL_ANSIBLE ceph_pgs.yml ceph_wait.yml #ceph_monitor.yml
+  
+        # run job
+#        ./ansible-playbook.sh -e site=$results -e nfiles=$NFILES -e drop_delay=30 ../workloads/scale-merge.yml
+#        ./ansible-playbook.sh -e site=$results -e nfiles=$NFILES ../workloads/setup_mdtest.yml
+#        if [ $decoupled -eq 1 ]; then
+#          ./ansible-playbook.sh -e site=$results -e nfiles=$NFILES ../workloads/decouple.yml
+#        fi
+#        ./ansible-playbook.sh -e site=$results -e nfiles=$NFILES -e drop_delay=30 ../workloads/${job}.yml
+#        ./ansible-playbook.sh -e site=$results collect.yml
+      done
+    done
+  done
+done

### 2017-04-04 experiments/fuse/scale-merge/workloads/docreates.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+set -ex
+INO=$1
+docker exec cephfs cephfs-journal-tool event create list --nfiles 10 --persist true --decoupled_dir dir-$HOSTNAME --file /etc/ceph/events-$HOSTNAME.bin --start_ino $INO

### 2017-04-04 experiments/transfertime/createfile.sh

@@ -1,0 +1,12 @@
+#! /bin/bash
+
+set -ex
+
+OTHERHOST="issdm-0"
+
+mkdir results
+dd if=/dev/zero of=dummy10k.bin bs=23247507 count=1
+dd if=/dev/zero of=dummy100k.bin bs=232497507 count=1
+
+/usr/bin/time --format="%e" scp dummy10k.bin $OTHERHOST:/tmp/blah > results/dummy10k.txt 2>&1
+/usr/bin/time --format="%e" scp dummy100k.bin $OTHERHOST:/tmp/blah > results/dummy100k.txt 2>&1

### 2017-04-04 experiments/transfertime/test.sh

@@ -1,0 +1,39 @@
+#!/bin/bash
+
+
+set -ex
+
+ino=1099511628877
+
+# generate
+#for i in `seq 1 24`; do
+#  ino=$((ino+98000))
+#  docker exec ceph-issdm-12-mds cephfs-journal-tool event create summary --nfiles 98000 --persist true --decoupled_dir dir$i --file /etc/ceph/$i-100k.bin --start_ino $ino >> log.txt 2>&1
+#done
+
+#for i in `seq 1 24`; do
+#  ino=$((ino+50000))
+#  docker exec ceph-issdm-12-mds cephfs-journal-tool event create summary --nfiles 50000 --persist true --decoupled_dir dir$i --file /etc/ceph/$i-50k.bin --start_ino $ino >> log.txt 2>&1
+#done
+#
+#for i in `seq 1 24`; do
+#  ino=$((ino+1000))
+#  docker exec ceph-issdm-12-mds cephfs-journal-tool event create summary --nfiles 1000 --persist true --decoupled_dir dir$i --file /etc/ceph/$i-1k.bin --start_ino $ino >> log.txt 2>&1
+#done
+
+#rm -r results || true
+mkdir results 
+#for i in `seq 1 24`; do
+#  /usr/bin/time --format="%e" docker exec ceph-issdm-12-mds ceph daemon mds.issdm-12 merge /etc/ceph/$i-1k.bin >> results/log-1k 2>> results/summary-1k
+#  sleep 10
+#done
+
+#for i in `seq 1 24`; do
+#  /usr/bin/time --format="%e" docker exec ceph-issdm-12-mds ceph daemon mds.issdm-12 merge /etc/ceph/$i-50k.bin >> results/log-50k 2>> results/summary-50k
+#  sleep 10
+#done
+
+for i in `seq 1 24`; do
+  /usr/bin/time --format="%e" docker exec ceph-issdm-12-mds ceph daemon mds.issdm-12 merge /etc/ceph/$i-100k.bin >> results/log-100k 2>> results/summary-100k
+  sleep 20
+done
